---
title: 'MOFA: Downstream analysis'
author: "Qian-Wu Liao"
output:
  workflowr::wflow_html:
    toc: true
    toc_float: true
    code_folding: hide
---

<font size='4'> Description: Do enrichment analysis of MOFA model trained on DIA
Plasma Proteomics (DIAPP), Untargeted Plasma Metabolomics (Unt. PM), Untargeted
Plasma Lipidomics (Unt. PL), and DIA Normal Tissue Proteomics (DIANP) for biological
interpretations. Note that DIAPP in this script indicates latest data from AG Krijgsveld
and DIANP indicates AG Klingmüller one. </font>\

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 8, fig.width = 10, warning = F)
knitr::opts_knit$set(root.dir = '/Users/qianwu/Desktop/SMART-CARE_LungCancer_MethodDev')
```

Load libraries
```{r message = F}
library('MOFA2')
library('reticulate')
# use_python('/Users/qianwu/opt/anaconda3/bin/python')
library('AnnotationDbi')
library('org.Hs.eg.db')
# This annotation object is accessed using AnnotationDbi:select()
hs <- org.Hs.eg.db
library('msigdbr')
library('clusterProfiler')
library('ggrepel')
library('SummarizedExperiment')
library('tidyverse')
# Load user-defined functions
source('./code/misc.R')
source('./code/mofa_funcs.R')

# Set plot theme
th <- theme_bw(base_size = 15) +
  theme(axis.title = element_text(face = 'bold'),
        axis.text = element_text(face = 'bold'),
        axis.ticks = element_line(linewidth = 0.8),
        legend.text = element_text(size = 15))
```

# DIA PP + Unt. PM + Unt. PL + DIANP

Overview model setup
```{r}
# Load trained MOFA model
mofaObj <- readRDS('./data/mofa/newDIAPP_UntPM_UntPL_DIANP_9Fac.rds')
# Overview model
MOFA2::plot_data_overview(mofaObj)
```

Show data variance explained by factors
```{r}
# Show data variance explained by learned factors
MOFA2::plot_variance_explained(mofaObj, max_r2 = 10) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1, size = 13),
        axis.text.y = element_text(size = 15, face = 'bold'),
        strip.text = element_text(size = 15),
        legend.title = element_text(size = 15), legend.text = element_text(size = 11))
# ggsave('./output/Latest/mofa_PP+UntPM_PL+NP_varExplained.png', device = 'png', dpi = 400, height = 8, width = 10)
```

Show significant factor
```{r}
# Show significant recurrence-related factors
sigFactor <- mofa_vizSigFactor(mofaObj, smpGroup = 'Baseline Group')
```

Show features with highest weights, which is most associated with recurrence patients
```{r}
# Show top features of significant factor with highest absolute weights
# Prepare feature annotation table
# DIA Tissue Proteomics (Klingmüller)
pro_gene_DIA_Klin <- rowData(readRDS('./data/AG_Klingmuller/proTissueNorm_DIA.rds')) %>%
  tibble::as_tibble(rownames = 'Proteins') %>%
  tibble::column_to_rownames('Proteins')
# Modify feature names
tmp_mofaObj <- mofa_rmFeatSuffix(mofaObj, view = 'DIA Normal Proteomics',
                                 feat_anno = list(`DIA Normal Proteomics` = pro_gene_DIA_Klin))

MOFA2::plot_top_weights(tmp_mofaObj, view = 'DIA Normal Proteomics', factor = 5,
                        nfeatures = 10, scale = T, sign = 'negative') +
  labs(title = 'DIA Normal Tissue Proteomics') +
  theme(axis.title = element_text(size = 20, face = 'bold'),
        axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 20, face = 'bold'),
        strip.text = element_text(size = 20, face = 'bold'))
# ggsave('./output/Latest/mofa_PP+UntPM_PL+NP_topWeightedFeats_NP.png', device = 'png', dpi = 400, height = 8, width = 5)


# Untargeted Plasma Lipidomics
mzrt <- rowData(readRDS('./data/AG_Hopf/lipPlasmaVsn.rds')) %>%
  tibble::as_tibble(rownames = 'Feature') %>%
  tibble::column_to_rownames('Feature')
# Modify feature names
tmp_mofaObj <- mofa_rmFeatSuffix(mofaObj, view = 'Untargeted Plasma Lipidomics',
                                 feat_anno = list(`Untargeted Plasma Lipidomics` = mzrt))

MOFA2::plot_top_weights(tmp_mofaObj, view = 'Untargeted Plasma Lipidomics', factor = 5,
                        nfeatures = 10, scale = T, sign = 'negative') +
  labs(title = 'Untargeted Plasma Lipidomics') +
  theme(axis.title = element_text(size = 20, face = 'bold'),
        axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 20, face = 'bold'),
        strip.text = element_text(size = 20, face = 'bold'))
# ggsave('./output/Latest/mofa_PP+UntPM_PL+NP_topWeightedFeats_PL.png', device = 'png', dpi = 400, height = 8, width = 5)
```

```{r eval=F, include=F}
# Show molecular signature of significant factor in data
# Prepare feature and sample annotation tables
# DIA Tissue Proteomics (Klingmüller)
pro_gene_DIA_Klin <- rowData(readRDS('./data/AG_Klingmuller/proTissueNorm_DIA.rds')) %>%
  tibble::as_tibble(rownames = 'Proteins') %>%
  tibble::column_to_rownames('Proteins')
smpAnno <- dplyr::select(samples_metadata(mofaObj), sample, Recurrence) %>%
  tibble::column_to_rownames('sample')

mofa_plotDataHeatmap(mofaObj, factor = 5, view = 'DIA Normal Proteomics',
                     group = 'Baseline Group', num_feats = 20, feat_anno = pro_gene_DIA_Klin,
                     smp_anno = smpAnno, smp_cluster = 'Recurrence', rm_smpSuffix = '_P_B',
                     cluster_rows = F, cluster_cols = F, show_rownames = T, show_colnames = T,
                     color = colorRampPalette(c('navy', 'white', 'red'))(100),
                     scale = 'row', main = 'Baseline DIA Normal Tissue Proteomics - Fac5')
```

## Enrichment analysis

```{r}
# Tidy up weight matrix and feature metadata to reduce information loss
# Prepare feature annotation table to extract proteins' gene annotations
# DIA Tissue Proteomics (Klingmüller)
pro_gene_DIA_Klin <- rowData(readRDS('./data/AG_Klingmuller/proTissueNorm_DIA.rds')) %>%
  tibble::as_tibble(rownames = 'Proteins') %>%
  tibble::column_to_rownames('Proteins')
proNormalWeight <- mofa_keepUniFeats(mofaObj, factor = 5, view = 'DIA Normal Proteomics',
                                     feat_anno = pro_gene_DIA_Klin, to_genes = T)

# Rank feature weights
rankedGeneList <- mofa_rankFeatList(proNormalWeight, factor = 5)
# Make Recurrence group positive weights
rankedGeneList <- sort(-rankedGeneList, decreasing = T)
```

### Curated gene sets
MSigDB - C2

```{r message=F}
# Curated gene sets - DIA Normal Proteomics
# View supported species and collections in MSigDB
# msigdbr_species()
# msigdbr_collections()
# Retrieve annotated gene sets
msigTabC2 <- msigdbr(species = 'Homo sapiens', category = 'C2', subcategory = NULL) %>%
  dplyr::select(gs_name, human_gene_symbol)

# Run GSEA
gseaResC2 <- clusterProfiler::GSEA(geneList = rankedGeneList, TERM2GENE = msigTabC2,
                                   minGSSize = 10, maxGSSize = 500,
                                   pvalueCutoff = 0.05, pAdjustMethod = 'BH',
                                   by = 'fgsea', eps = 0)

# Plot enrichment analysis results
gseaRes4Plot <- gseaResC2@result %>%
  dplyr::select(Description, setSize, NES, pvalue, p.adjust, leading_edge) %>%
  dplyr::arrange(p.adjust) %>%
  dplyr::slice_head(n = 8) %>%
  dplyr::mutate(leading_edge = stringr::str_extract(leading_edge, '\\d\\d'),
                leading_edge = as.numeric(leading_edge),
                # Description = stringr::str_remove(Description, '^REACTOME_'),
                Recurrence = ifelse(test = NES > 0, yes = 'Yes', no = 'No')) %>%
  dplyr::rename(gene_ratio = leading_edge)
midGrad <- mean(gseaRes4Plot$p.adjust)

ggplot(gseaRes4Plot, aes(x=NES, y=reorder(Description, NES), fill=p.adjust, col=Recurrence)) +
  geom_bar(stat = 'identity', width = 0.7, size = 2) +
  scale_fill_gradient2(low='#5B1C00', high='#FFFCA5', mid = '#E86900', midpoint = midGrad, name = 'Adjusted pVal') +
  scale_color_manual(values = c(Yes = 'red', No = 'darkblue')) +
  labs(x = 'Normalized Enrichment Score', y = 'Pathway') +
  theme_minimal() +
  theme(axis.title = element_text(size = 16, face = 'bold'),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 10, face = 'bold'),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12))
  # theme(axis.title = element_text(size = 20, face = 'bold'),
  #       axis.text.x = element_text(size = 18),
  #       axis.text.y = element_text(size = 18, face = 'bold'),
  #       legend.title = element_text(size = 20),
  #       legend.text = element_text(size = 16),
  #       panel.grid.minor = element_blank())
# ggsave('./output/Latest/mofa_PP+UntPM_PL+NP_EA_C2.png', device = 'png', dpi = 400, height = 8, width = 13)

# ggplot(gseaRes4Plot, aes(x=NES, y=reorder(Description, NES), fill=p.adjust)) +
#   geom_point(aes(size = gene_ratio), shape = 21) +
#   scale_fill_gradient2(low='#5B1C00', high='#FFFCA5', mid = '#E86900',
#                        midpoint = midGrad, name = 'Adjusted p-value') +
#   scale_size_continuous(name = 'Core gene ratio') +
#   geom_vline(xintercept = 0) +
#   labs(x = 'Normalized Enrichment Score (NES)', y = 'Pathway') +
#   theme_minimal()

# enrichplot::dotplot(gseaResC2, x = 'GeneRatio', color = 'p.adjust', showCategory = 10,
#                     font.size = 12, title = 'DIA Normal Proteomics (Curated)')
```
=> Increased uptake, synthesis, and storage of lipids has been demonstrated to contribute
to growth of many cancer types, including lung cancer.

### Oncogenic signature gene sets
MSigDB - C6

```{r message=F}
# Oncogenic signatures gene sets - DIA Normal Proteomics
# Retrieve annotated gene sets
msigTabC6 <- msigdbr(species = 'Homo sapiens', category = 'C6', subcategory = NULL) %>%
  dplyr::select(gs_name, human_gene_symbol)
# Retrieve descriptions of gene sets for interpretation
descripC6 <- msigdbr(species = 'Homo sapiens', category = 'C6', subcategory = NULL) %>%
  dplyr::select(gs_name, gs_description) %>%
  dplyr::filter(!duplicated(gs_name))

# Run GSEA
gseaResC6 <- clusterProfiler::GSEA(geneList = rankedGeneList, TERM2GENE = msigTabC6,
                                   minGSSize = 10, maxGSSize = 500,
                                   pvalueCutoff = 0.05, pAdjustMethod = 'BH',
                                   by = 'fgsea', eps = 0)

# Plot enrichment analysis results
gseaRes4Plot <- gseaResC6@result %>%
  dplyr::select(Description, setSize, NES, pvalue, p.adjust, leading_edge) %>%
  dplyr::arrange(p.adjust) %>%
  dplyr::slice_head(n = 8) %>%
  dplyr::mutate(leading_edge = stringr::str_extract(leading_edge, '\\d\\d'),
                leading_edge = as.numeric(leading_edge),
                Recurrence = ifelse(test = NES > 0, yes = 'Yes', no = 'No')) %>%
  dplyr::rename(gene_ratio = leading_edge)
midGrad <- mean(gseaRes4Plot$p.adjust)

ggplot(gseaRes4Plot, aes(x=NES, y=reorder(Description, NES), fill=p.adjust, col=Recurrence)) +
  geom_bar(stat = 'identity', width = 0.6, size = 2) +
  scale_fill_gradient2(low='#5B1C00', high='#FFFCA5', mid = '#E86900', midpoint = midGrad, name = 'Adjusted pVal') +
  scale_color_manual(values = c(Yes = 'red', No = 'darkblue')) +
  labs(x = 'Normalized Enrichment Score', y = 'Pathway') +
  theme_minimal() +
  theme(axis.title = element_text(size = 16, face = 'bold'),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 10, face = 'bold'),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12))
  # theme(axis.title = element_text(size = 20, face = 'bold'),
  #       axis.text.x = element_text(size = 18),
  #       axis.text.y = element_text(size = 18, face = 'bold'),
  #       legend.title = element_text(size = 20),
  #       legend.text = element_text(size = 16),
  #       panel.grid.minor = element_blank())
# ggsave('./output/Latest/mofa_PP+UntPM_PL+NP_EA_C6.png', device = 'png', dpi = 400, height = 8, width = 10)
```
=> Aberrant activation of KRAS and mTOR and suppression of p53 has been reported
to be associated with lung cancer. mTOR signaling pathway regulates many fundamental
cellular activities, including lipid metabolism.

### Hallmark gene sets
MSigDB - H

```{r message=F}
# Hallmark gene sets - DIA Normal Proteomics
# Retrieve annotated gene sets
msigTabH <- msigdbr(species = 'Homo sapiens', category = 'H', subcategory = NULL) %>%
  dplyr::select(gs_name, human_gene_symbol)

# Run GSEA
gseaResH <- clusterProfiler::GSEA(geneList = rankedGeneList, TERM2GENE = msigTabH,
                                  minGSSize = 10, maxGSSize = 500,
                                  pvalueCutoff = 0.05, pAdjustMethod = 'BH',
                                  by = 'fgsea', eps = 0)

# Plot enrichment analysis results
gseaRes4Plot <- gseaResH@result %>%
  dplyr::select(Description, setSize, NES, pvalue, p.adjust, leading_edge) %>%
  dplyr::arrange(p.adjust) %>%
  dplyr::slice_head(n = 8) %>%
  dplyr::mutate(leading_edge = stringr::str_extract(leading_edge, '\\d\\d'),
                leading_edge = as.numeric(leading_edge),
                Recurrence = ifelse(test = NES > 0, yes = 'Yes', no = 'No')) %>%
  dplyr::rename(gene_ratio = leading_edge)
midGrad <- mean(gseaRes4Plot$p.adjust)

ggplot(gseaRes4Plot, aes(x=NES, y=reorder(Description, NES), fill=p.adjust, col=Recurrence)) +
  geom_bar(stat = 'identity', width = 0.6, size = 2) +
  scale_fill_gradient2(low='#5B1C00', high='#FFFCA5', mid = '#E86900', midpoint = midGrad, name = 'Adjusted pVal') +
  scale_color_manual(values = c(Yes = 'red', No = 'darkblue')) +
  labs(x = 'Normalized Enrichment Score', y = 'Pathway') +
  theme_minimal() +
  theme(axis.title = element_text(size = 16, face = 'bold'),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 10, face = 'bold'),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12))
```




```{r eval=F, include=F}
# MSigDB - DIA Normal Proteomics
# Tidy up weight matrix and feature metadata to reduce information loss
# Prepare feature annotation table to extract proteins' gene annotations
# DIA Tissue Proteomics (Klingmüller)
pro_gene_DIA_Klin <- rowData(readRDS('./data/AG_Klingmuller/proTissueNorm_DIA.rds')) %>%
  tibble::as_tibble(rownames = 'Proteins') %>%
  tibble::column_to_rownames('Proteins')
proNormalWeight <- mofa_keepUniFeats(mofaObj, factor = 5, view = 'DIA Normal Proteomics',
                                     feat_anno = pro_gene_DIA_Klin, to_genes = T)
# Rank feature weights
rankedGeneList <- mofa_rankFeatList(proNormalWeight, factor = 5)

# Ontology gene sets
# Retrieve annotated gene sets
msigTabC5BP <- msigdbr(species = 'Homo sapiens', category = 'C5', subcategory = 'GO:BP') %>%
  dplyr::select(gs_name, human_gene_symbol)

# Run GSEA
gseaResC5BP <- clusterProfiler::GSEA(geneList = rankedGeneList, TERM2GENE = msigTabC5BP,
                                     minGSSize = 10, maxGSSize = 500,
                                     pvalueCutoff = 0.05, pAdjustMethod = 'BH',
                                     by = 'fgsea', eps = 0)
# Plot enrichment analysis result
enrichplot::dotplot(gseaResC5BP, x = 'GeneRatio', color = 'p.adjust', showCategory = 10,
                    font.size = 12, title = 'DIA Normal Proteomics (GO:BP)')




# GO:BP - DIA Normal Proteomics (Another way)
# Map gene symbols to Entrez IDs
proNormalWeight2 <- SYMBOL2ENTREZID(proNormalWeight)
# Rank feature weights
rankedGeneList2 <- mofa_rankFeatList(proNormalWeight2, factor = 5)

# Run GSEA
# @result$setSize: Set size of intersection between (ranked) gene list and gene set
# Leading edge analysis: 'tags' indicates percentage of leading edge subset (core
# of genes, @result$core_enrichment) in intersected set between ranked gene list
# (@geneList) and GO gene set (@geneSets) contributing to ES.
# 'Count' in plot indicates length of leading edge subset and 'GeneRatio' indicates 'tags'.
# Parameter 'scoreType' is set to 'pos' or 'neg' for one-tailed test. Use case:
# For ranked list based on absolute values of log(FC), only positive side is of
# interest -> 'pos'
gseaResBP <- clusterProfiler::gseGO(geneList = rankedGeneList2, ont = 'BP', OrgDb = hs,
                                    keyType = 'ENTREZID', minGSSize = 10, maxGSSize = 500,
                                    pvalueCutoff = 0.05, pAdjustMethod = 'BH', by = 'fgsea',
                                    eps = 0, scoreType = 'std')
enrichplot::dotplot(gseaResBP, x = 'GeneRatio', color = 'p.adjust', showCategory = 15,
                    font.size = 12, title = 'DIA Normal Proteomics (GO:BP)')

# Show leading edge analysis of an example
# gseaResBP@result$leading_edge[1]
# # Retrieve size of ranked gene list
# geneListSize <- length(names(gseaResBP@geneList))
# # Retrieve size of example GO gene set
# goGeneSetSize <- length(gseaResBP@geneSets[[gseaResBP@result$ID[1]]])
# # Compute number of intersected genes between gene list and GO gene set
# setSize <- length(intersect(names(gseaResBP@geneList), gseaResBP@geneSets[[gseaResBP@result$ID[1]]]))
# # Retrieve size of leading edge subset
# leadEdgeSubset <- length(unlist(stringr::str_split(gseaResBP@result$core_enrichment[1], '/')))
# # Sanity check
# print(paste('setSize = intersection of ranked gene list and gene set:',
#             identical(setSize, gseaResBP@result$setSize[1])))
# print(paste('tags = leading edge subset size / setSize:', leadEdgeSubset / setSize))
# # Make ES plot
# enrichplot::gseaplot2(gseaResBP, geneSetID = gseaResBP@result$ID[1],title = gseaResBP@result$Description[1])
```
