---
title: 'Preprocessing: Discovery lung cancer patient cohort'
author: "Qian-Wu Liao"
output:
  workflowr::wflow_html:
    toc: yes
    toc_float: yes
    code_folding: hide
---

<font size='4'> Description: Preprocess Tissue and Plasma DIA Proteomics generated
by Karim Aljakouch from AG Krijgsveld, including data cleansing, filtering, and
normalization (VSN). All needed information was then stored in SummarizedExperiment
objects for further analyses.</font>

Q: There is no '1HTNWJ_TU' and got extra Tumor sample '6EY5JO_II_TU'?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 8, fig.width = 10, warning = F)
knitr::opts_knit$set(root.dir = '/Users/qianwu/Desktop/SMART-CARE_LungCancer_Discovery')
```

Load libraries
```{r load libraries, message=F}
library(readr)
library(readxl)
library(vsn)
library(limma)
library(visdat)
library(ggrepel)
library(SummarizedExperiment)
library(tidyverse)
# Load user-defined functions
source('./code/analysis_pipeline.R')
source('./code/misc.R')

# Set plot theme
th <- theme_bw(base_size = 15) +
  theme(axis.title = element_text(face = 'bold'),
        axis.text = element_text(face = 'bold'),
        axis.ticks = element_line(linewidth = 0.8),
        legend.text = element_text(size = 15))
```

```{r prepare global info, message=F}
# Load sample metadata
smpMetadat <- readxl::read_excel('./data/sample_metadata.xlsx')
colnames(smpMetadat) <- smpMetadat[3,, drop = F]
smpMetadat <- dplyr::slice(smpMetadat, -c(1:3)) %>%
  dplyr::select(c(Code, Parents, Visit, `Material submitted`, `Date and time of collection or surgery`,
                  `SMART-CARE cohort identifier`)) %>%
  dplyr::rename(Sample = Code, Patient = Parents, TimePoint = Visit, SmpType = `Material submitted`,
                Date = `Date and time of collection or surgery`, Cohort = `SMART-CARE cohort identifier`) %>%
  dplyr::mutate(Sample = stringr::str_remove_all(Sample, '^SC_T_S_|^SC_DIS_S_|^SC_S_|_P'),
                Patient = stringr::str_remove(Patient, '^/THRX_SPACE/THRX_DB/'),
                TimePoint = dplyr::case_when(TimePoint == 'PRETHERAPEUTIC' ~ 'Baseline',
                                             TimePoint == 'FOLLOW-UP' ~ 'Follow-up',
                                             TimePoint == 'RECURRENCE' ~ 'Recurrence'),
                SmpType = dplyr::case_when(SmpType == 'EDTA_PLASMA' ~ 'Plasma',
                                           SmpType == 'FRESH_FROZEN_TISSUE' ~ 'Tissue'),
                Condition = dplyr::case_when(grepl('_TU', Sample) ~ 'Tumor',
                                             grepl('_NG', Sample) ~ 'Normal',
                                             !grepl('_TU|_NG', Sample) ~ TimePoint),
                Date = stringr::str_extract(Date, '^\\d+-\\d+-\\d+'),
                Cohort = dplyr::case_when(Cohort == 'DISCOVERY_COHORT' ~ 'Discovery',
                                          Cohort == 'METHOD_DEVELOPMENT_COHORT' ~ 'MethodDev'),
                Date = as.Date(Date, format = '%Y-%m-%d'))

# Load patient metadata
patientMetadat <- readxl::read_excel('./data/patient_metadata.xlsx')
colnames(patientMetadat) <- patientMetadat[3,, drop = F]
patientMetadat <- dplyr::slice(patientMetadat, -c(1:3)) %>%
  dplyr::select(c(Code, Gender, `Age at diagnosis`, `Pathological stage`, `Smoking status`,
                  `Adjuvant chemotherapy`,)) %>%
  dplyr::rename(Patient = Code, Age = `Age at diagnosis`, Stage = `Pathological stage`,
                Smoking = `Smoking status`, Adjuvant = `Adjuvant chemotherapy`) %>%
  dplyr::mutate(Gender = dplyr::case_when(Gender == 'MALE' ~ 'Male',
                                          Gender == 'FEMALE' ~ 'Female'),
                Smoking = dplyr::case_when(Smoking == 'SMOKER' ~ 'Smoker',
                                           Smoking == 'EX-SMOKER' ~ 'Ex-smoker',
                                           Smoking == 'NON-SMOKER' ~ 'Non-smoker'),
                Adjuvant = dplyr::case_when(Adjuvant == 'true' ~ 'True',
                                            Adjuvant == 'false' ~ 'False'),
                Age = as.numeric(Age))
# Include patient recurrence information in patient metadata
recurPats <- dplyr::filter(smpMetadat, TimePoint == 'Recurrence') %>%
  dplyr::pull(Patient)
patientMetadat <- dplyr::mutate(patientMetadat, Recurrence = dplyr::case_when(Patient %in% recurPats ~ 'Yes',
                                                                              !Patient %in% recurPats ~ 'No'))

# Load aliquot metadata
aliquotMetadat <- readxl::read_excel('./data/aliquot_metadata.xlsx')
colnames(aliquotMetadat) <- aliquotMetadat[3,, drop = F]
aliquotMetadat <- dplyr::slice(aliquotMetadat, -c(1:3)) %>%
  dplyr::select(Code, Parents, `Tumor Cell Content (%)`, `Cohort Identifier`, `Submission to`,
                `Delivered?`) %>%
  dplyr::rename(Aliquot = Code, Sample = Parents, TumorPurity = `Tumor Cell Content (%)`,
                Cohort = `Cohort Identifier`, To = `Submission to`, Delivered = `Delivered?`) %>%
  dplyr::mutate(Sample = stringr::str_remove_all(Sample, '^/THRX_SPACE/THRX_DB/SC_T_S_|_P'),
                Sample = stringr::str_remove(Sample, '^/THRX_SPACE/THRX_DB/SC_DIS_S_'),
                Sample = stringr::str_remove(Sample, '^/THRX_SPACE/THRX_DB/SC_S_'),
                Sample = stringr::str_remove(Sample, '^/THRX_SPACE/THRX_DB/SC_DIS_'),
                Sample = dplyr::case_when(Sample == 'SDIR55_TU' ~ 'PDIR55_TU',
                                          Sample == 'SDIR55_NG' ~ 'PDIR55_NG',
                                          !Sample %in% c('SDIR55_TU', 'SDIR55_NG') ~ Sample),
                Cohort = dplyr::case_when(Cohort == 'method establishment' ~ 'MethodDev',
                                          Cohort == 'DISCOVERY_COHORT' ~ 'Discovery'),
                TumorPurity = as.numeric(TumorPurity),
                Sample = dplyr::case_when(grepl('^SC_DIS_A_Y4W4L7_P_V2', Aliquot) ~ 'Y4W4L7_V2',
                                          !grepl('^SC_DIS_A_Y4W4L7_P_V2', Aliquot) ~ Sample)) %>%
  dplyr::filter(To == 'KRIJGSVELD')
# Prepare tumor purity information
tumorPurityInfo <- dplyr::select(aliquotMetadat, Sample, TumorPurity)

# Combine all needed metadata
summMetadat <- dplyr::left_join(smpMetadat, patientMetadat, by = 'Patient') %>%
  dplyr::left_join(tumorPurityInfo)

# List sample annotations to keep in SE objects
smpAnno <- c('Patient', 'SmpType', 'TimePoint', 'Date', 'Cohort', 'Condition',
             'Recurrence', 'Gender', 'Age', 'Smoking', 'Stage', 'Adjuvant')
```

# Tissue Proteomics
Tumor and Normal samples were first preprocessed together to assure they are on
same scale, so that they can be compared to check data quality. Then, Tumor and
Normal samples were preprocessed separately to preserve maximal amounts of reliable features.

```{r message=F}
# Load unprocessed data
proTissueTab <- readr::read_tsv('./data/AG_Krijgsveld/20230815_MarcS_Discovery_Tissue.pg_matrix.tsv') %>%
  dplyr::select(-c(Protein.Ids, First.Protein.Description)) %>%
  tidyr::pivot_longer(cols = -c('Protein.Group', 'Protein.Names', 'Genes'),
                      names_to = 'Sample', values_to = 'Abundance') %>%
  dplyr::mutate(Protein.Names = stringr::str_remove_all(Protein.Names, '_HUMAN'),
                Sample = stringr::str_remove_all(Sample, '^.+SC_A_|^.+SC_DIS_A_|4\\.raw$'))
# Include summarized metadata
# Add an extra sample to summarized metadata
tmp_summMetadat <- summMetadat
smp <- summMetadat[summMetadat$Sample == '6EY5JO_TU',]
smp[1, 1] <- '6EY5JO_II_TU'
tmp_summMetadat <- rbind(tmp_summMetadat, smp)
proTissueTab <- dplyr::left_join(proTissueTab, tmp_summMetadat, by = 'Sample')
```

Show dimensions of original data (86 samples and 7986 features)
```{r}
# Preprocess data
proTissue <- doPreprocessing(proTissueTab, feat = 'Protein.Group', smp = 'Sample',
                             val = 'Abundance', featAnno = c('Protein.Names', 'Genes'),
                             smpAnno = c(smpAnno, 'TumorPurity'), do_featFilt = T,
                             cutoff = 0.5, viz_miss = T, save_path = './data/AG_Krijgsveld/proTissue')
dim(proTissue$ori.data)
```

Show missingness of original data
```{r}
proTissue$ori.data.miss +
  theme(axis.text.x = element_text(size = 10))
```

Show distribution of original data
```{r}
proTissue$ori.data.dist +
  theme(axis.text.x = element_text(size = 10))
```

**Remove features quantified in less than 1/2 of samples**\
Show dimensions of filtered data (86 samples and 6077 features)
```{r}
dim(proTissue$filt.data)
```

Show missingness of filtered data
```{r}
proTissue$filt.data.miss +
  theme(axis.text.x = element_text(size = 10))
```

**Normalize filtered data by VSN**\
Show distribution of vsn normalized data
```{r}
proTissue$vsn.data.dist +
  theme(axis.text.x = element_text(size = 10))
```

Show feature mean-variance relationship of vsn normalized data
```{r}
proTissue$vsn.feat.mean.var
```

**Metadata-assisted quality control**
```{r}
# Do single-omics analysis
proTissueSOA <- doSOA(proTissue$vsn.data, meta_var = 'Condition', pca_method = 'ppca',
                      do_onlyPCA = T)
# Prepare labels for samples of interest
pcTab <- proTissueSOA$pcTab %>%
  dplyr::mutate(Label = dplyr::case_when(Sample %in% c('I0HVOL_TU', '7EAOX7_TU',
                                                       'MJMTYR_TU', 'XFKHGP_TU',
                                                       'XFKHGP_NG', 'MJMTYR_NG',
                                                       'I0HVOL_NG', '7EAOX7_NG') ~ Sample))
ggplot(pcTab, aes(x=Condition, y=`PC1 (31.5%)`, col=Condition, fill=Condition, label=Label)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(position = position_jitter(0.3), size = 3, show.legend = F) +
  geom_text_repel(show.legend = F, size = 5) +
  scale_color_brewer(palette = 'Dark2') +
  scale_fill_brewer(palette = 'Dark2') +
  labs(x = 'Sample condition') +
  th #+ theme(axis.title = element_text(size = 28),
             # axis.text.x = element_text(size = 24), axis.text.y = element_text(size = 18),
             # legend.title = element_text(size = 26), legend.text = element_text(size = 24))
# ggsave('./output/group_meeting/quality_control_proTissue_Krij.png', device = 'png', dpi = 400, height = 8, width = 10)
```
=> Tumor purity: I0HVOL_TU - 87% / 7EAOX7_TU - 70% / MJMTYR_TU - 40% / XFKHGP_TU - 100%\
(Mean of sample tumor cell contents is 74%, ranging from 40% to 100%.)

## Tumor samples

**Remove features quantified in less than 1/2 of samples**\
Show dimensions of filtered data (43 samples and 6350 features)
```{r}
# Subset certain samples
proTumorTab <- dplyr::filter(proTissueTab, Condition == 'Tumor')
# Preprocess data
proTumor <- doPreprocessing(proTumorTab, feat = 'Protein.Group', smp = 'Sample',
                            val = 'Abundance', featAnno = c('Protein.Names', 'Genes'),
                            smpAnno = c(smpAnno, 'TumorPurity'), do_featFilt = T,
                            cutoff = 0.5, viz_miss = T, save_path = './data/AG_Krijgsveld/proTumor')
dim(proTumor$filt.data)
```

Show missingness of filtered data
```{r}
proTumor$filt.data.miss
```

**Normalize filtered data by VSN**\
Show distribution of vsn normalized data
```{r}
proTumor$vsn.data.dist
```

Show feature mean-variance relationship of vsn normalized data
```{r}
proTumor$vsn.feat.mean.var
```

## Normal samples

**Remove features quantified in less than 1/2 of samples**\
Show dimensions of filtered data (43 samples and 5778 features)
```{r}
# Subset certain samples
proNormalTab <- dplyr::filter(proTissueTab, Condition == 'Normal')
# Preprocess data
proNormal <- doPreprocessing(proNormalTab, feat = 'Protein.Group', smp = 'Sample',
                             val = 'Abundance', featAnno = c('Protein.Names', 'Genes'),
                             smpAnno = c(smpAnno, 'TumorPurity'), do_featFilt = T,
                             cutoff = 0.5, viz_miss = T, save_path = './data/AG_Krijgsveld/proNormal')
dim(proNormal$filt.data)
```

Show missingness of filtered data
```{r}
proNormal$filt.data.miss
```

**Normalize filtered data by VSN**\
Show distribution of vsn normalized data
```{r}
proNormal$vsn.data.dist
```

Show feature mean-variance relationship of vsn normalized data
```{r}
proNormal$vsn.feat.mean.var
```

```{r message=F, eval=F}
# Fix patient recurrence information in Tissue Proteomics from AG Klingmüller that
# contains Method Development and part of Discovery cohort samples
# Load normalized data
proTissue_Klin <- readRDS('./data/AG_Klingmueller/proTissueVsn.rds')
# Prepare patient recurrence information for both Method Development and Discovery cohorts
# Retrieve Method Development cohort samples
methodDev_smpMetadat <- readxl::read_excel('./data/sample_metadata.xlsx')
colnames(methodDev_smpMetadat) <- methodDev_smpMetadat[3,, drop = F]
methodDev_smps <- dplyr::slice(methodDev_smpMetadat, -c(1:3)) %>%
  dplyr::filter(`SMART-CARE cohort identifier` == 'METHOD_DEVELOPMENT_COHORT',
                grepl('_TG|_NG', Code)) %>%
  dplyr::pull(Code) %>%
  stringr::str_remove('^SC_T_S_')
# Prepare recurrence information of Method Development cohort patients
proTissueSub <- proTissue_Klin[, colnames(proTissue_Klin) %in% methodDev_smps, drop = F]
methodDev_recurInfo <- tibble::as_tibble(colData(proTissueSub)) %>%
  dplyr::select(Patient, Recurrence) %>%
  dplyr::filter(!duplicated(Patient))
# Prepare recurrence information of Discovery cohort patients
methodDev_pats <- unique(stringr::str_remove(methodDev_smps, '_.+'))
discovery_recurInfo <- dplyr::select(patientMetadat, Patient, Recurrence) %>%
  dplyr::filter(!Patient %in% methodDev_pats)
# Combine all recurrence information
recurInfo <- dplyr::bind_rows(methodDev_recurInfo, discovery_recurInfo)
# Update sample annotations
smpAnno <- tibble::as_tibble(colData(proTissue_Klin), rownames = 'Sample') %>%
  dplyr::select(-Recurrence) %>%
  dplyr::left_join(recurInfo, by = 'Patient') %>%
  tibble::column_to_rownames('Sample') %>%
  dplyr::relocate(Recurrence, .after = Patient) %>%
  dplyr::relocate(Identifier, .after = Batch)
# Create SE object with updated sample annotations
featAnno <- tibble::as_tibble(rowData(proTissue_Klin), rownames = 'Feature') %>%
  tibble::column_to_rownames('Feature')
proTissue_Klin <- SummarizedExperiment(assays = list(Abundance = assay(proTissue_Klin)),
                                       rowData = featAnno, colData = smpAnno)
# saveRDS(proTissue_Klin, './data/AG_Klingmueller/proTissueVsn.rds')
```




# Plasma Proteomics
Due to complexity of plasma samples, we will be focusing on analysis of Baseline
samples. 63 out of 81 patients got Baseline samples and they were extracted and
preprocessed to preserve maximal amounts of reliable features.

```{r message=F}
# Load unprocessed data
proPlasmaTab <- readr::read_tsv('./data/AG_Krijgsveld/20230811_MarcS_Discovery_Plasma.pg_matrix.tsv') %>%
  dplyr::select(-c(Protein.Ids, First.Protein.Description)) %>%
  tidyr::pivot_longer(cols = -c('Protein.Group', 'Protein.Names', 'Genes'),
                      names_to = 'Sample', values_to = 'Abundance') %>%
  dplyr::mutate(Protein.Names = stringr::str_remove_all(Protein.Names, '_HUMAN'),
                Sample = stringr::str_remove_all(Sample, '^.+SC_DIS_A_|_4\\.raw$|\\.raw$|^.+KA_|_P'),
                Sample = stringr::str_replace(Sample, 'P_', '_'),
                Sample = stringr::str_remove(Sample, '^SC_A_'),
                Sample = dplyr::case_when(Sample == 'FD01075852' ~ 'R1WTOP_V1',
                                          Sample == 'FD05926483' ~ 'R1WTOP_V2',
                                          Sample == 'FD01282286' ~ 'R1WTOP_V3',
                                          Sample == 'FD07174643' ~ 'R1WTOP_V4',
                                          !Sample %in% c('FD01075852', 'FD05926483',
                                                         'FD01282286', 'FD07174643') ~ Sample),
                Sample = stringr::str_remove_all(Sample, '_FD\\d+|_LV\\d+'))
# Include summarized metadata
proPlasmaTab <- dplyr::left_join(proPlasmaTab, summMetadat, by = 'Sample')
```

Explore time points and exact dates of plasma samples
```{r fig.height=10}
# Explore time points of plasma samples
dplyr::select(proPlasmaTab, Sample, TimePoint, Date) %>%
  dplyr::filter(!duplicated(Sample)) %>%
  DT::datatable()

# Visualize exact dates of plasma samples
tmp_proPlasmaTab <- dplyr::mutate(proPlasmaTab,
                                  Label = dplyr::case_when(TimePoint == 'Baseline' ~ 'Baseline',
                                                           TimePoint != 'Baseline' ~ 'Follow-up'))
ggplot(tmp_proPlasmaTab, aes(x=Date, y=Patient, color=Label)) +
  geom_point(size = 2) +
  labs(title = 'Exact Dates') +
  scale_x_date(date_breaks = 'year', date_labels = '%Y') +
  scale_color_manual(name = 'Time Point', values = c('red2', 'grey40')) +
  theme_minimal() +
  theme(axis.title = element_text(size = 12, face = 'bold'),
        axis.text.x = element_text(size = 10, angle = 60, vjust = 1, hjust = 1),
        plot.title = element_text(size = 14, face = 'bold', hjust = 0.5),
        legend.title = element_text(size = 12), legend.text = element_text(size = 10))
#   theme(axis.title = element_text(size = 24, face = 'bold'),
#         axis.text.x = element_text(size = 12, angle = 60, vjust = 1, hjust = 1),
#         legend.title = element_text(size = 22), legend.text = element_text(size = 20))
# ggsave('./output/group_meeting/smp_dates_proPlasma_Krij.png', device = 'png', dpi = 400, height = 10, width = 12)
```

```{r include=F, eval=F}
# Preprocess data
proPlasma <- doPreprocessing(proPlasmaTab, feat = 'Protein.Group', smp = 'Sample',
                             val = 'Abundance', featAnno = c('Protein.Names', 'Genes'),
                             smpAnno = smpAnno, do_featFilt = T, cutoff = 0.5,
                             viz_miss = T, bins = 20, save_path = './data/AG_Krijgsveld/proPlasma')
# Show dimensions of original data (289 samples and 969 features)
dim(proPlasma$ori.data)
# Show missingness of original data
proPlasma$ori.data.miss +
  labs(x = 'Samples') +
  theme(axis.text.x = element_blank(), axis.title = element_text(size = 13, face = 'bold'),
        panel.grid.major.y = element_blank())
# Show sample mean-variance relationship of original data
proPlasma$ori.smp.mean.var

# Remove features quantified in less than 1/2 of samples
# Show dimensions of filtered data (289 samples and 607 features)
dim(proPlasma$filt.data)
# Show missingness of filtered data
proPlasma$filt.data.miss +
  labs(x = 'Samples') +
  theme(axis.text.x = element_blank(), axis.title = element_text(size = 13, face = 'bold'))

# Normalize filtered data by VSN
# Show sample mean-variance relationship of vsn normalized data
proPlasma$vsn.smp.mean.var +
  ylim(2.4, 3.6)
# Show feature mean-variance relationship of vsn normalized data
proPlasma$vsn.feat.mean.var
```

## Baseline samples

Show dimensions of original data (63 samples and 969 features)
```{r}
# Subset certain samples
proBaseTab <- dplyr::filter(proPlasmaTab, Condition == 'Baseline')
# Preprocess data
proBase <- doPreprocessing(proBaseTab, feat = 'Protein.Group', smp = 'Sample',
                           val = 'Abundance', featAnno = c('Protein.Names', 'Genes'),
                           smpAnno = smpAnno, do_featFilt = T, cutoff = 0.5,
                           viz_miss = T, bins = 20, save_path = './data/AG_Krijgsveld/proBase')
dim(proBase$ori.data)
```

Show missingness of original data
```{r}
proBase$ori.data.miss +
  theme(panel.grid.minor.y = element_blank())
```

Show distribution of original data
```{r}
proBase$ori.data.dist
```

**Remove features quantified in less than 1/2 of samples**\
Show dimensions of filtered data (63 samples and 613 features)
```{r}
dim(proBase$filt.data)
```

Show missingness of filtered data
```{r}
proBase$filt.data.miss
```

**Normalize filtered data by VSN**\
Show distribution of vsn normalized data
```{r}
proBase$vsn.data.dist
```

Show feature mean-variance relationship of vsn normalized data
```{r}
proBase$vsn.feat.mean.var
```
