---
title: 'Preprocessing: Discovery lung cancer patient cohort'
author: "Qian-Wu Liao"
output:
  workflowr::wflow_html:
    toc: yes
    toc_float: yes
    code_folding: hide
---

<font size='4'> Description: Preprocess Tissue and Plasma DIA Proteomics generated
by Karim Aljakouch from AG Krijgsveld, including data cleansing, filtering, and
normalization (VSN). All needed information was then stored in SummarizedExperiment
objects for further analyses. </font>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 8, fig.width = 10, warning = F)
knitr::opts_knit$set(root.dir = '/Users/qianwu/Desktop/SMART-CARE_LungCancer_Discovery')
```

Load libraries
```{r load libraries, message=F}
library(readr)
library(readxl)
library(vsn)
library(limma)
library(visdat)
library(ggrepel)
library(SummarizedExperiment)
library(tidyverse)
# Load user-defined functions
source('./code/analysis_pipeline.R')
source('./code/misc.R')

# Set plot theme
th <- theme_bw(base_size = 15) +
  theme(axis.title = element_text(face = 'bold'),
        axis.text = element_text(face = 'bold'),
        axis.ticks = element_line(linewidth = 0.8),
        legend.text = element_text(size = 15))
```

```{r prepare global info, message=F}
# Load sample metadata
smpMetadat <- readxl::read_excel('./data/sample_metadata.xlsx')
colnames(smpMetadat) <- smpMetadat[3,, drop = F]
smpMetadat <- dplyr::slice(smpMetadat, -c(1:3)) %>%
  dplyr::filter(`SMART-CARE cohort identifier` == 'DISCOVERY_COHORT') %>%
  dplyr::select(c(Code, Visit, `Material submitted`, `Date and time of collection or surgery`, Parents)) %>%
  dplyr::rename(Sample = Code, Status = Visit, SmpType = `Material submitted`,
                Date = `Date and time of collection or surgery`, Patient = Parents) %>%
  dplyr::mutate(Sample = stringr::str_remove_all(Sample, '^SC_DIS_S_|^SC_S_|_P'),
                Status = dplyr::case_when(Status == 'PRETHERAPEUTIC' ~ 'Baseline',
                                          Status == 'FOLLOW-UP' ~ 'Follow-up',
                                          Status == 'RECURRENCE' ~ 'Recurrence'),
                SmpType = dplyr::case_when(SmpType == 'EDTA_PLASMA' ~ 'Plasma',
                                           SmpType == 'FRESH_FROZEN_TISSUE' ~ 'Tissue'),
                Date = stringr::str_extract(Date, '\\d+-\\d+'),
                Patient = stringr::str_remove(Patient, '/THRX_SPACE/THRX_DB/')) %>%
  dplyr::arrange(Patient)

# Load patient metadata
patientMetadat <- readxl::read_excel('./data/patient_metadata.xlsx')
colnames(patientMetadat) <- patientMetadat[3,, drop = F]
patientMetadat <- dplyr::slice(patientMetadat, -c(1:3)) %>%
  dplyr::select(-c(Identifier, Diagnosis, Nutrition)) %>%
  dplyr::rename(Patient = Code, Age = `Age at diagnosis`, Stage = `Pathological stage`,
                Smoking = `Smoking status`, Adjuvant = `Adjuvant chemotherapy`) %>%
  dplyr::mutate(Gender = dplyr::case_when(Gender == 'MALE' ~ 'Male',
                                          Gender == 'FEMALE' ~ 'Female'),
                Smoking = dplyr::case_when(Smoking == 'SMOKER' ~ 'Smoker',
                                           Smoking == 'EX-SMOKER' ~ 'Ex-smoker',
                                           Smoking == 'NON-SMOKER' ~ 'Non-smoker'),
                Adjuvant = dplyr::case_when(Adjuvant == 'true' ~ 'True',
                                            Adjuvant == 'false' ~ 'False'),
                Age = as.numeric(Age))
# Include patient recurrence information
recurPats <- dplyr::filter(smpMetadat, Status == 'Recurrence') %>%
  dplyr::pull(Patient)
patientMetadat <- dplyr::mutate(patientMetadat,
                                Recurrence = dplyr::case_when(Patient %in% recurPats ~ 'Yes',
                                                              !Patient %in% recurPats ~ 'No')) %>%
  dplyr::relocate(Recurrence, .after = Patient)

# List sample annotations to keep in SE objects
smpAnno <- c('Condition', 'Patient', 'Recurrence', 'Gender', 'Age', 'Smoking',
             'Stage', 'Adjuvant')
```

# Tissue Proteomics
Tumor and Normal samples were first preprocessed together to assure they are on
same scale, so that they can be compared to check data quality. Then, Tumor and
Normal samples were preprocessed separately to preserve maximal amounts of reliable features.

```{r message=F}
# Load unprocessed data
proTissueTab <- readr::read_tsv('./data/AG_Krijgsveld/20230815_MarcS_Discovery_Tissue.pg_matrix.tsv') %>%
  dplyr::select(-c(Protein.Ids, First.Protein.Description)) %>%
  tidyr::pivot_longer(cols = -c('Protein.Group', 'Protein.Names', 'Genes'),
                      names_to = 'Sample', values_to = 'Abundance') %>%
  dplyr::mutate(Protein.Names = stringr::str_remove_all(Protein.Names, '_HUMAN'),
                Sample = stringr::str_remove_all(Sample, '^.+SC_A_|^.+SC_DIS_A_|4\\.raw$'),
                Condition = dplyr::case_when(grepl('_NG', Sample) ~ 'Normal',
                                             grepl('TU', Sample) ~ 'Tumor'),
                Patient = stringr::str_remove(Sample, '_.+$'))
# Include patient metadata
proTissueTab <- dplyr::left_join(proTissueTab, patientMetadat, by = 'Patient')
```

Show dimensions of original data (86 samples and 7986 features)
```{r}
# Preprocess data
proTissue <- doPreprocessing(proTissueTab, feat = 'Protein.Group', smp = 'Sample',
                             val = 'Abundance', featAnno = c('Protein.Names', 'Genes'),
                             smpAnno = smpAnno, do_featFilt = T, cutoff = 0.67,
                             viz_miss = T) #save_path = './data/AG_Krijgsveld/proTissue'
dim(proTissue$ori.data)
```

Show missingness of original data
```{r}
proTissue$ori.data.miss +
  theme(axis.text.x = element_text(size = 10))
```

Show distribution of original data
```{r}
proTissue$ori.data.dist +
  theme(axis.text.x = element_text(size = 10))
```

**Remove features quantified in less than 2/3 of samples**\
Show dimensions of filtered data (86 samples and 5316 features)
```{r}
dim(proTissue$filt.data)
```

Show missingness of filtered data
```{r}
proTissue$filt.data.miss +
  theme(axis.text.x = element_text(size = 10))
```

**Normalize filtered data by VSN**\
Show distribution of vsn normalized data
```{r}
proTissue$vsn.data.dist +
  theme(axis.text.x = element_text(size = 10))
```

Show feature mean-variance relationship of vsn normalized data
```{r}
proTissue$vsn.feat.mean.var
```

**Check data quality**
```{r}
# Do single-omics analysis
proTissueSOA <- doSOA(proTissue$vsn.data, meta_var = 'Condition', pca_method = 'ppca',
                      do_onlyPCA = T)
# Prepare labels for samples of interest
pcTab <- proTissueSOA$pcTab %>%
  dplyr::mutate(Label = dplyr::case_when(Sample %in% c('I0HVOL_TU', '7EAOX7_TU',
                                                       'MJMTYR_TU', 'XFKHGP_TU',
                                                       'XFKHGP_NG', 'MJMTYR_NG',
                                                       'I0HVOL_NG', '7EAOX7_NG') ~ Sample))
ggplot(pcTab, aes(x=Condition, y=`PC1 (36.3%)`, col=Condition, fill=Condition, label=Label)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(position = position_jitter(0.3), size = 3, show.legend = F) +
  geom_text_repel(show.legend = F) +
  scale_color_brewer(palette = 'Dark2') +
  scale_fill_brewer(palette = 'Dark2') +
  labs(x = 'Sample condition') +
  th
```
=> I0HVOL_TU: 87%, 14mg / 7EAOX7_TU: 70%, 19mg / MJMTYR_TU: 40%, 31mg / XFKHGP_TU: 100%, 4mg\
(Tumor cell contents ranging from 40 - 100%, and Tissue weight from 4 - 54mg)

## Tumor samples

**Remove features quantified in less than 1/2 of samples**\
Show dimensions of filtered data (43 samples and 6350 features)
```{r}
# Subset certain samples
proTumorTab <- dplyr::filter(proTissueTab, Condition == 'Tumor')
# Preprocess data
proTumor <- doPreprocessing(proTumorTab, feat = 'Protein.Group', smp = 'Sample',
                            val = 'Abundance', featAnno = c('Protein.Names', 'Genes'),
                            smpAnno = smpAnno, do_featFilt = T, cutoff = 0.5,
                            viz_miss = T) #save_path = './data/AG_Krijgsveld/proTumor'
dim(proTumor$filt.data)
```

Show missingness of filtered data
```{r}
proTumor$filt.data.miss
```

**Normalize filtered data by VSN**\
Show distribution of vsn normalized data
```{r}
proTumor$vsn.data.dist
```

Show feature mean-variance relationship of vsn normalized data
```{r}
proTumor$vsn.feat.mean.var
```

## Normal samples

**Remove features quantified in less than 1/2 of samples**\
Show dimensions of filtered data (43 samples and 5778 features)
```{r}
# Subset certain samples
proNormalTab <- dplyr::filter(proTissueTab, Condition == 'Normal')
# Preprocess data
proNormal <- doPreprocessing(proNormalTab, feat = 'Protein.Group', smp = 'Sample',
                             val = 'Abundance', featAnno = c('Protein.Names', 'Genes'),
                             smpAnno = smpAnno, do_featFilt = T, cutoff = 0.5,
                             viz_miss = T) #save_path = './data/AG_Krijgsveld/proNormal'
dim(proNormal$filt.data)
```

Show missingness of filtered data
```{r}
proNormal$filt.data.miss
```

**Normalize filtered data by VSN**\
Show distribution of vsn normalized data
```{r}
proNormal$vsn.data.dist
```

Show feature mean-variance relationship of vsn normalized data
```{r}
proNormal$vsn.feat.mean.var
```

```{r message=F, eval=F}
# Fix patient recurrence information in Tissue Proteomics from AG Klingmüller that
# contains Method Development and part of Discovery cohort samples
# Load normalized data
proTissue_Klin <- readRDS('./data/AG_Klingmueller/proTissueVsn.rds')
# Prepare patient recurrence information for both Method Development and Discovery cohorts
# Retrieve Method Development cohort samples
methodDev_smpMetadat <- readxl::read_excel('./data/sample_metadata.xlsx')
colnames(methodDev_smpMetadat) <- methodDev_smpMetadat[3,, drop = F]
methodDev_smps <- dplyr::slice(methodDev_smpMetadat, -c(1:3)) %>%
  dplyr::filter(`SMART-CARE cohort identifier` == 'METHOD_DEVELOPMENT_COHORT',
                grepl('_TG|_NG', Code)) %>%
  dplyr::pull(Code) %>%
  stringr::str_remove('^SC_T_S_')
# Prepare recurrence information of Method Development cohort patients
proTissueSub <- proTissue_Klin[, colnames(proTissue_Klin) %in% methodDev_smps, drop = F]
methodDev_recurInfo <- tibble::as_tibble(colData(proTissueSub)) %>%
  dplyr::select(Patient, Recurrence) %>%
  dplyr::filter(!duplicated(Patient))
# Prepare recurrence information of Discovery cohort patients
methodDev_pats <- unique(stringr::str_remove(methodDev_smps, '_.+'))
discovery_recurInfo <- dplyr::select(patientMetadat, Patient, Recurrence) %>%
  dplyr::filter(!Patient %in% methodDev_pats)
# Combine all recurrence information
recurInfo <- dplyr::bind_rows(methodDev_recurInfo, discovery_recurInfo)
# Update sample annotations
smpAnno <- tibble::as_tibble(colData(proTissue_Klin), rownames = 'Sample') %>%
  dplyr::select(-Recurrence) %>%
  dplyr::left_join(recurInfo, by = 'Patient') %>%
  tibble::column_to_rownames('Sample') %>%
  dplyr::relocate(Recurrence, .after = Patient) %>%
  dplyr::relocate(Identifier, .after = Batch)
# Create SE object with updated sample annotations
featAnno <- tibble::as_tibble(rowData(proTissue_Klin), rownames = 'Feature') %>%
  tibble::column_to_rownames('Feature')
proTissue_Klin <- SummarizedExperiment(assays = list(Abundance = assay(proTissue_Klin)),
                                       rowData = featAnno, colData = smpAnno)
# saveRDS(proTissue_Klin, './data/AG_Klingmueller/proTissueVsn.rds')
```




# Plasma Proteomics
Baseline samples were extracted and preprocessed to preserve maximal amounts of reliable features.

```{r message=F}
# Load unprocessed data
proPlasmaTab <- readr::read_tsv('./data/AG_Krijgsveld/20230811_MarcS_Discovery_Plasma.pg_matrix.tsv') %>%
  dplyr::select(-c(Protein.Ids, First.Protein.Description)) %>%
  tidyr::pivot_longer(cols = -c('Protein.Group', 'Protein.Names', 'Genes'),
                      names_to = 'Sample', values_to = 'Abundance') %>%
  dplyr::mutate(Protein.Names = stringr::str_remove_all(Protein.Names, '_HUMAN'),
                Sample = stringr::str_remove_all(Sample, '^.+SC_DIS_A_|_4\\.raw$|\\.raw$|^.+KA_|_P'),
                Sample = stringr::str_replace(Sample, 'P_', '_'),
                Sample = stringr::str_remove(Sample, '^SC_A_'),
                Sample = dplyr::case_when(Sample == 'FD01075852' ~ 'R1WTOP_V1',
                                          Sample == 'FD05926483' ~ 'R1WTOP_V2',
                                          Sample == 'FD01282286' ~ 'R1WTOP_V3',
                                          Sample == 'FD07174643' ~ 'R1WTOP_V4',
                                          !Sample %in% c('FD01075852', 'FD05926483',
                                                         'FD01282286', 'FD07174643') ~ Sample),
                Sample = stringr::str_remove_all(Sample, '_FD\\d+|_LV\\d+'),
                Patient = stringr::str_remove(Sample, '_.+$'),
                TimePoint = stringr::str_remove(Sample, '^.+_'),
                TimePoint = factor(TimePoint, levels = paste0('V', seq_along(unique(TimePoint)))))
# Include sample and patient metadata
smpPlasmaMetadat <- dplyr::filter(smpMetadat, SmpType == 'Plasma') %>%
  dplyr::select(Sample, Status, Date) %>%
  dplyr::rename(Condition = Status)
proPlasmaTab <- dplyr::left_join(proPlasmaTab, smpPlasmaMetadat, by = 'Sample')
proPlasmaTab <- dplyr::left_join(proPlasmaTab, patientMetadat, by = 'Patient')
```

Explore time points and exact dates of plasma samples
```{r fig.height=10}
# Explore time points of plasma samples
dplyr::select(proPlasmaTab, Sample, Condition, Date) %>%
  dplyr::filter(!duplicated(Sample)) %>%
  DT::datatable()

# Visualize time points of plasma samples
ggplot(proPlasmaTab, aes(x=TimePoint, y=Patient)) +
  geom_point(shape = 4, size = 2) +
  labs(x = 'Time point', title = 'Time Points') +
  theme_minimal() +
  theme(axis.title = element_text(size = 12, face = 'bold'), axis.text.x = element_text(size = 10),
        plot.title = element_text(size = 14, face = 'bold', hjust = 0.5),)

# Visualize exact dates of plasma samples
tmp_proPlasmaTab <- dplyr::mutate(proPlasmaTab,
                                  Date = paste0(Date, '-01'),
                                  Date = as.Date(Date),
                                  Label = dplyr::case_when(Condition == 'Baseline' ~ 'Baseline',
                                                           Condition != 'Baseline' ~ 'Follow-up'))
ggplot(tmp_proPlasmaTab, aes(x=Date, y=Patient, color=Label)) +
  geom_point(size = 2) +
  labs(title = 'Exact Dates') +
  scale_x_date(date_breaks = 'year', date_labels = '%Y') +
  scale_color_manual(name = 'Time point', values = c('red2', 'grey40')) +
  theme_minimal() +
  theme(axis.title = element_text(size = 12, face = 'bold'),
        axis.text.x = element_text(size = 10, angle = 60, vjust = 1, hjust = 1),
        plot.title = element_text(size = 14, face = 'bold', hjust = 0.5),
        legend.title = element_text(size = 12), legend.text = element_text(size = 10))
```

```{r include=F, eval=F}
# Preprocess data
proPlasma <- doPreprocessing(proPlasmaTab, feat = 'Protein.Group', smp = 'Sample',
                             val = 'Abundance', featAnno = c('Protein.Names', 'Genes'),
                             smpAnno = c(smpAnno, 'TimePoint', 'Date'), do_featFilt = T,
                             cutoff = 0.67, viz_miss = T, bins = 20) #save_path = './data/AG_Krijgsveld/proPlasma'
# Show dimensions of original data (289 samples and 969 features)
dim(proPlasma$ori.data)
# Show missingness of original data
proPlasma$ori.data.miss +
  labs(x = 'Samples') +
  theme(axis.text.x = element_blank(), axis.title = element_text(size = 13, face = 'bold'),
        panel.grid.major.y = element_blank())
# Show sample mean-variance relationship of original data
proPlasma$ori.smp.mean.var

# Remove features quantified in less than 2/3 of samples
# Show dimensions of filtered data (289 samples and 542 features)
dim(proPlasma$filt.data)
# Show missingness of filtered data
proPlasma$filt.data.miss +
  labs(x = 'Samples') +
  theme(axis.text.x = element_blank(), axis.title = element_text(size = 13, face = 'bold'))

# Normalize filtered data by VSN
# Show sample mean-variance relationship of vsn normalized data
proPlasma$vsn.smp.mean.var +
  ylim(2.4, 3.6)
# Show feature mean-variance relationship of vsn normalized data
proPlasma$vsn.feat.mean.var
```

## Baseline samples

Show dimensions of original data (42 samples and 969 features)
```{r}
# Subset certain samples
proBaseTab <- dplyr::filter(proPlasmaTab, Condition == 'Baseline')
# Preprocess data
proBase <- doPreprocessing(proBaseTab, feat = 'Protein.Group', smp = 'Sample',
                           val = 'Abundance', featAnno = c('Protein.Names', 'Genes'),
                           smpAnno = c(smpAnno, 'TimePoint', 'Date'), do_featFilt = T,
                           cutoff = 0.5, viz_miss = T, bins = 20) #save_path = './data/AG_Krijgsveld/proBase'
dim(proBase$ori.data)
```

Show missingness of original data
```{r}
proBase$ori.data.miss +
  theme(panel.grid.minor.y = element_blank())
```

Show distribution of original data
```{r}
proBase$ori.data.dist
```

**Remove features quantified in less than 1/2 of samples**\
Show dimensions of filtered data (42 samples and 665 features)
```{r}
dim(proBase$filt.data)
```

Show missingness of filtered data
```{r}
proBase$filt.data.miss
```

**Normalize filtered data by VSN**\
Show distribution of vsn normalized data
```{r}
proBase$vsn.data.dist
```

Show feature mean-variance relationship of vsn normalized data
```{r}
proBase$vsn.feat.mean.var
```
