---
title: "Single-omics analysis: Discovery lung cancer patient cohort"
author: "Qian-Wu Liao"
output:
  workflowr::wflow_html:
    toc: yes
    toc_float: yes
    code_folding: hide
---

<font size='4'> Description: </font>

**Metadata variables**\
Patients whose Tissue and Baseline Plasma samples were collected:\
Note that Baseline Plasma sample of Patient '1JV272' (Female, 64 y/o, Ex-smoker,
Stage IIB, Adjuvant chemotherapy used, and Cancer recurrence) is missing.\
Recurrence -> Cancer recurrences, Yes:No = 19:24\
Gender -> Female:Male = 18:25\
Age -> Diagnosis ages ranging from 51 to 76\
Smoking -> Smoker:Ex-smoker:Non-Smoker = 11:26:6\
Stage -> Pathological stages, IB:IIA:IIB = 25:6:12\
Adjuvant -> Adjuvant chemotherapy, True:False = 14:29\
Samples:\
Condition -> Tumor:Normal = 1:1

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 8, fig.width = 10, warning = F)
knitr::opts_knit$set(root.dir = '/Users/qianwu/Desktop/SMART-CARE_LungCancer_Discovery')
```

Load libraries
```{r library loading, message = F}
library('pcaMethods')
library('limma')
library('proDA')
library('pheatmap')
library('ggplotify')
library('ggrepel')
library('SummarizedExperiment')
library('tidyverse')
# Load user-defined functions
source('./code/analysis_pipeline.R')
source('./code/misc.R')

# Set plot theme
th <- theme_bw(base_size = 15) +
    theme(axis.title = element_text(face = 'bold'),
          axis.text = element_text(face = 'bold'),
          axis.ticks = element_line(linewidth = 0.8),
          legend.text = element_text(size = 15))
```

Display associations between cancer recurrence and the other patient metadata variables
to identify potential predictors (or confounders)
```{r}
# Test associations between cancer recurrence and the other patient metadata variables
# Prepare patient metadata
proTissue <- readRDS('./data/AG_Krijgsveld/proTissueVsn.rds')
smpMetadat <- tibble::as_tibble(colData(proTissue)) %>%
  dplyr::filter(!duplicated(Patient)) %>%
  dplyr::select(-Condition)
smpRecur <- dplyr::select(smpMetadat, Patient, Recurrence)
smpMetadat <- dplyr::select(smpMetadat, -Recurrence)
testAsso(smpMetadat, smpRecur, cmn_col = 'Patient') %>%
  dplyr::select(-Stat)
```

# Tissue Proteomics

```{r}
# Load preprocessed data
proTissue <- readRDS('./data/AG_Krijgsveld/proTissueVsn.rds')
# Prepare feature annotation table
featAnnoTab <- tibble::as_tibble(rowData(proTissue), rownames = 'Protein') %>%
  dplyr::select(Protein, Genes) %>%
  dplyr::rename(Gene = Genes) %>%
  tibble::column_to_rownames('Protein')

# Perform single-omics analysis
proTissueRes <- doSOA(proTissue, meta_var = c('Condition', 'Recurrence', 'Gender',
                                              'Age', 'Smoking', 'Stage', 'Adjuvant'),
                      do_onlyPCA = T, pca_method = 'ppca')
```

Display significant associations between PCs (Var1) and sample metadata variables
(Var2). Note that each patient has two samples that have same patient annotations,
so association results may have additive effects.
```{r}
proTissueRes$pcSigAssoRes
```

PC1 significantly separate Tumor and Normal samples, which indicates decent data
quality. Yet, there are 4 Tumor samples misclassified, reported as follows:\
I0HVOL_TU: 87%, 14mg / 7EAOX7_TU: 70%, 19mg / MJMTYR_TU: 40%, 31mg / XFKHGP_TU: 100%, 4mg\
(Tumor cell contents ranging from 40 - 100%, and Tissue weight from 4 - 54mg)
```{r}
pcTab <- proTissueRes$pcTab %>%
  dplyr::mutate(Label = dplyr::case_when(Sample %in% c('I0HVOL_TU', '7EAOX7_TU',
                                                       'MJMTYR_TU', 'XFKHGP_TU',
                                                       'XFKHGP_NG', 'MJMTYR_NG',
                                                       'I0HVOL_NG', '7EAOX7_NG') ~ Sample))
ggplot(pcTab, aes(x=Condition, y=`PC1 (31.5%)`, color=Condition, fill=Condition, label=Label)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(position = position_jitter(0.3), size = 3, show.legend = F) +
  geom_text_repel(show.legend = F) +
  scale_color_brewer(palette = 'Dark2') +
  scale_fill_brewer(palette = 'Dark2') +
  labs(x = 'Sample condition') +
  th
```
=> Tumor and Normal tissue samples will be analyzed separately for two reasons:
(1) They are characteristically different and might be offsets to each other, and
(2) Normal tissues is always obtainable in case that Tumor tissues were not preserved
after surgery.

## Tumor samples

```{r}
# Subset certain samples
proTumor <- proTissue[, which(colData(proTissue)$Condition == 'Tumor')]
# Perform single-omics analysis
proTumorRes <- doSingleOmicsAnalysis(proTumor, soa_metaVar = 'Recurrence', num_sigFeats = 6,
                                     pca_method = 'ppca', num_PCs = 40, num_PCfeats = 30,
                                     use_limma = F, use_proDA = T, feat_anno = featAnnoTab,
                                     feat_conv = T, plot_title = NULL, show_rownames = F,
                                     pca_metaVar = c('Recurrence', 'Gender', 'Age',
                                                     'Smoking', 'Stage', 'Adjuvant'))
```

Display significant associations between PCs (Var1) and sample metadata variables
(Var2) to overview sources of variance in data
```{r}
proTumorRes$data.var.source
```

Display significant associations between features (Var1) and cancer recurrence (Var2).
Owing to missing values, there may be situation that only few observations in sample
groups, resulting in unreliable t-statistics. To this end, we used probabilistic
dropout model (proDA) to account for data missingness.
```{r}
proTumorRes$sig.feat.tab
```

Visualize molecular signatures of significant features in data. Note that features
in rows are ordered by t-statistics.
```{r}
proTumorRes$sig.feat.heat
```

Visualize top significant features identified in Tumor samples. Note that abundances
in Normal samples are also provided to have more comprehensive analysis.
```{r}
# Prepare data matrix and metadata including both Tumor and Normal samples
datMat <- proTissueRes$data
smpAnnoTab <- proTissueRes$smpMetadata
# Prepare significant feature table
featSigAssoTab <- proTumorRes$sig.feat.tab
num_sigFeats <- 6
gpLevel = c('No_Tumor', 'Yes_Tumor', 'No_Normal', 'Yes_Normal')
condiCol = c('grey50', 'firebrick') #c(Normal, Tumor)
# Extract top significant features and prepare needed information
topSigFeats <- featSigAssoTab[1:num_sigFeats,] %>%
  dplyr::select(Var1, pVal, Gene) %>%
  dplyr::mutate(newVar1 = paste0(Gene, '\n(p = ', round(pVal, 4), ')'),
                newVar1 = factor(newVar1, levels = unique(newVar1)))
topSigFeatDat <- tibble::as_tibble(datMat[topSigFeats$Var1,], rownames = 'Var1') %>%
  tidyr::pivot_longer(cols = -'Var1', names_to = 'Sample', values_to = 'Abundance') %>%
  dplyr::left_join(topSigFeats, by = 'Var1') %>%
  dplyr::left_join(smpAnnoTab, by = 'Sample') %>%
  dplyr::mutate(Recur_Condi = dplyr::case_when(Condition == 'Tumor' ~ paste0(Recurrence, '_Tumor'),
                                               Condition == 'Normal' ~ paste0(Recurrence, '_Normal')),
                Recur_Condi = factor(Recur_Condi, levels = gpLevel))
ggplot(topSigFeatDat, aes(x=Recur_Condi, y=Abundance, col=Condition, fill=Recurrence)) +
  geom_boxplot(alpha = 1, outlier.shape = NA) +
  geom_jitter(position = position_jitter(0.2), size = 1, show.legend = F) +
  labs(x = 'Recurrence') +
  scale_color_manual(values = condiCol) +
  scale_fill_manual(values=c('#00BFC4', '#F8766D')) +
  facet_wrap(vars(newVar1), scales = 'free') +
  th + theme(strip.text = element_text(size = 13, face = 'bold'),
             axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

Assess data power by p-value histogram
```{r}
proTumorRes$pval.hist
```

Display significant associations between PCs (Var1) and cancer recurrence (Var2)
```{r}
proTumorRes$sig.pc.tab
```

Visualize significant PCs and display their top features with greatest absolute loadings
```{r}
proTumorRes$sig.pc.dist$PC20
proTumorRes$pc.top.feat.tab$PC20
```

## Normal samples

```{r}
# Subset certain samples
proNormal <- proTissue[, which(colData(proTissue)$Condition == 'Normal')]
# Perform single-omics analysis
proNormalRes <- doSingleOmicsAnalysis(proNormal, soa_metaVar = 'Recurrence', num_sigFeats = 6,
                                      pca_method = 'ppca', num_PCs = 40, num_PCfeats = 30,
                                      use_limma = F, use_proDA = T, feat_anno = featAnnoTab,
                                      feat_conv = T, plot_title = NULL, show_rownames = F,
                                      pca_metaVar = c('Recurrence', 'Gender', 'Age',
                                                      'Smoking', 'Stage', 'Adjuvant'))
```

Display significant associations between PCs (Var1) and sample metadata variables
(Var2) to overview sources of variance in data
```{r}
proNormalRes$data.var.source
```

Display significant associations between features (Var1) and cancer recurrence (Var2)
```{r}
proNormalRes$sig.feat.tab
```

Visualize molecular signatures of significant features in data. Note that features
in rows are ordered by t-statistics.
```{r}
proNormalRes$sig.feat.heat
```

Visualize top significant features identified in Normal samples. Note that abundances
in Tumor samples are also provided to have more comprehensive analysis.
```{r}
# Prepare data matrix and metadata including both Tumor and Normal samples
datMat <- proTissueRes$data
smpAnnoTab <- proTissueRes$smpMetadata
# Prepare significant feature table
featSigAssoTab <- proNormalRes$sig.feat.tab
num_sigFeats <- 6
gpLevel = c('No_Normal', 'Yes_Normal', 'No_Tumor', 'Yes_Tumor')
condiCol = c('firebrick', 'grey50') #c(Normal, Tumor)
# Extract top significant features and prepare needed information
topSigFeats <- featSigAssoTab[1:num_sigFeats,] %>%
  dplyr::select(Var1, pVal, Gene) %>%
  dplyr::mutate(newVar1 = paste0(Gene, '\n(p = ', round(pVal, 4), ')'),
                newVar1 = factor(newVar1, levels = unique(newVar1)))
topSigFeatDat <- tibble::as_tibble(datMat[topSigFeats$Var1,], rownames = 'Var1') %>%
  tidyr::pivot_longer(cols = -'Var1', names_to = 'Sample', values_to = 'Abundance') %>%
  dplyr::left_join(topSigFeats, by = 'Var1') %>%
  dplyr::left_join(smpAnnoTab, by = 'Sample') %>%
  dplyr::mutate(Recur_Condi = dplyr::case_when(Condition == 'Tumor' ~ paste0(Recurrence, '_Tumor'),
                                               Condition == 'Normal' ~ paste0(Recurrence, '_Normal')),
                Recur_Condi = factor(Recur_Condi, levels = gpLevel))
ggplot(topSigFeatDat, aes(x=Recur_Condi, y=Abundance, col=Condition, fill=Recurrence)) +
  geom_boxplot(alpha = 1, outlier.shape = NA) +
  geom_jitter(position = position_jitter(0.2), size = 1, show.legend = F) +
  labs(x = 'Recurrence') +
  scale_color_manual(values = condiCol) +
  scale_fill_manual(values=c('#00BFC4', '#F8766D')) +
  facet_wrap(vars(newVar1), scales = 'free') +
  th + theme(strip.text = element_text(size = 13, face = 'bold'),
             axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

Assess data power by p-value histogram
```{r}
proNormalRes$pval.hist
```

Display significant associations between PCs (Var1) and cancer recurrence (Var2)
```{r}
proNormalRes$sig.pc.tab
```

Visualize significant PCs and display their top features with greatest absolute loadings
```{r}
proNormalRes$sig.pc.dist$PC7
proNormalRes$pc.top.feat.tab$PC7

proNormalRes$sig.pc.dist$PC38
proNormalRes$pc.top.feat.tab$PC38
```




# Plasma Proteomics
We first focus on analysis of Baseline samples because (1) what we are most interested
in is predicting recurrence, and (2) Baseline samples are less complex compared to Follow-up samples.

```{r include=F, eval=F}
# Load preprocessed data
proPlasma <- readRDS('./data/AG_Krijgsveld/proPlasmaVsn.rds')
# Perform single-omics analysis
proPlasmaRes <- doSOA(proPlasma, meta_var = c('Condition', 'Recurrence', 'Gender', 'Adjuvant'),
                      do_onlyPCA = T, pca_method = 'ppca', alpha = 0.01)

# Display significant associations between PCs (Var1) and sample metadata variables
# (Var2). Note that all patients got multiple samples that have same patient annotations,
# so association results may have additive effects. Furthermore, patient annotations,
# Age, Smoking, and Stage, were excluded because they are Baseline time point-related.
proPlasmaRes$pcSigAssoRes

# Visualize PC1 that is significantly associated with sample conditions
pcTab <- proPlasmaRes$pcTab
ggplot(pcTab, aes(x=Condition, y=`PC1 (35.3%)`, color=Condition, fill=Condition)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(position = position_jitter(0.2), size = 2, show.legend = F) +
  scale_color_brewer(palette = 'Dark2') +
  scale_fill_brewer(palette = 'Dark2') +
  labs(x = 'Sample condition') +
  th
```

## Baseline samples

```{r}
# Load preprocessed data
proBase <- readRDS('./data/AG_Krijgsveld/proBaseVsn.rds')
# Prepare feature annotation table
featAnnoTab <- tibble::as_tibble(rowData(proBase), rownames = 'Protein') %>%
  dplyr::select(Protein, Genes) %>%
  dplyr::rename(Gene = Genes) %>%
  tibble::column_to_rownames('Protein')

# Perform single-omics analysis
proBaseRes <- doSingleOmicsAnalysis(proBase, soa_metaVar = 'Recurrence', num_sigFeats = 6,
                                    pca_method = 'ppca', num_PCs = 40, num_PCfeats = 30,
                                    use_limma = F, use_proDA = T, feat_anno = featAnnoTab,
                                    feat_conv = T, plot_title = NULL,
                                    pca_metaVar = c('Recurrence', 'Gender', 'Age',
                                                    'Smoking', 'Stage', 'Adjuvant'))
```

Display significant associations between PCs (Var1) and sample metadata variables
(Var2) to overview sources of variance in data
```{r}
proBaseRes$data.var.source
```

Display significant associations between features (Var1) and cancer recurrence (Var2).
Owing to missing values, there may be situation that only few observations in sample
groups, resulting in unreliable t-statistics. To this end, we used probabilistic
dropout model (proDA) to account for data missingness.
```{r}
proBaseRes$sig.feat.tab

# Immunoglobulins (antibodies) are membrane-bound or secreted glycoproteins produced
# by B lymphocytes. In recognition phase of humoral immunity, membrane-bound immunoglobulins
# serve as receptors which, upon binding of a specific antigen, trigger clonal expansion
# and differentiation of B lymphocytes into immunoglobulins-secreting plasma cells.
# Secreted immunoglobulins mediate effector phase of humoral immunity, which results
# in elimination of bound antigens.
```

```{r eval=F}
# Check Gender (PC1) and Smoking (PC6) to avoid potential confounders
# Identify significant associations between features and patient genders
proBaseSex <- doSingleOmicsAnalysis(proBase, soa_metaVar = 'Gender', pca_method = 'ppca',
                                    num_PCs = 40, use_limma = F, use_proDA = T,
                                    pca_metaVar = c('Recurrence', 'Gender', 'Age',
                                                    'Smoking', 'Stage', 'Adjuvant'))
sigFeatsSex <- proBaseSex$sig.feat.tab$Var1
sigFeatsRecur <- proBaseRes$sig.feat.tab$Var1
sigFeatsRecur[sigFeatsRecur %in% sigFeatsSex]
# Protein 'P20742', 'Q0VDD7', and 'P02768' are also significantly associated with patient genders.

# Protein encoded by PZP is highly expressed in late-pregnancy serum, which can
# inhibit proteinases by its unique trapping mechanism. Appears to be elevated in
# sera of presymptomatic Alzheimer's disease patients. HCG is present during pregnancy,
# yet it relates to testicular carcinoma (tumor marker) when present in sera of males.

# Protein encoded by BRME1 (Break Repair MEiotic recombinase recruit factor 1) is
# predicted to be involved in meiosis 1 (meiotic double-strand break repair) and
# spermatogenesis (male fertility).

# ALB encodes most abundant protein albumin in human blood, which functions in regulation
# of blood plasma colloid osmotic pressure and acts as carrier protein for endogenous
# molecules hormones, fatty acids, and metabolites, and as well as exogenous drugs.
```

Visualize molecular signatures of significant features in data. Note that features
in rows are ordered by t-statistics.
```{r}
proBaseRes$sig.feat.heat
```

Visualize top significant features
```{r}
proBaseRes$top.sig.feat.dist
```

Assess data power by p-value histogram
```{r}
proBaseRes$pval.hist
```

Display significant associations between PCs (Var1) and cancer recurrence (Var2)
```{r}
proBaseRes$sig.pc.tab
```

Visualize significant PCs and display their top features with greatest absolute loadings
```{r}
proBaseRes$sig.pc.dist$PC7
proBaseRes$pc.top.feat.tab$PC7

proBaseRes$sig.pc.dist$PC5
proBaseRes$pc.top.feat.tab$PC5

proBaseRes$sig.pc.dist$PC20
proBaseRes$pc.top.feat.tab$PC20
```
