---
title: 'Cross-cohort analysis'
author: "Qian-Wu Liao"
output:
  workflowr::wflow_html:
    toc: yes
    toc_float: yes
    code_folding: hide
---

<font size='4'> Description: Investigate reproducibility of data and results between
Method Development and Discovery lung cancer patient cohorts. </font>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 8, fig.width = 10, warning = F)
knitr::opts_knit$set(root.dir = '/Users/qianwu/Desktop/SMART-CARE_LungCancer')
```

Load libraries
```{r load libraries, message=F}
library(ggrepel)
library(ggvenn)
library(SummarizedExperiment)
library(tidyverse)
# Load user-defined functions
source('./code/misc.R')

# Set plot theme
th <- theme_bw(base_size = 15) +
  theme(axis.title = element_text(face = 'bold'),
        axis.text = element_text(face = 'bold'),
        axis.ticks = element_line(linewidth = 0.8),
        legend.text = element_text(size = 15))
```

```{r load user-defined functions}
doFeatAssoTest <- function(data, meta_var, subset = NULL, prot_gene_tbl = NULL,
                           method = 'proDA') {
  #' Conduct univariate association test (t-test for now) between each feature and
  #' sample metadata variable of interest
  #' 
  #' Parameters
  #' data: A SummarizedExperiment object that contains the data matrix and metadata
  #' meta_var: A character specifying the sample metadata variable that has two
  #' levels for conducting t-test
  #' subset: A vector of length 2 containing characters specifying the sample metadata
  #' variable and one of its level to split the dataset. Default is NULL, meaning
  #' the whole dataset is used
  #' prot_gene_tbl: A table containing proteins and their encoding genes for providing
  #' gene names in a resulting table. Default is NULL. The column for proteins should
  #' be named 'Proteins' and for genes should be named 'Genes'. Furthermore, only
  #' the first protein and gene should be kept if there are more than one
  #' method: A character specifying the method used to conduct t-test, which should
  #' be one of 'proDA' (default) and 'limma'
  #' 
  #' Return
  #' assoTestRes: A table of feature significance test and statistical results
  
  # Check arguments
  # meta_var should be length of 1 and in sample metadata
  # subset should be length of 2 and in sample metadata
  # method should be one of 'proDA' and limma''
  
  # Subset samples if specified
  if (!is.null(subset)) {
    subsetIdx <- which(colData(data)[[subset[1]]] == subset[2])
    data <- data[, subsetIdx]
  }
  datMat <- SummarizedExperiment::assay(data)
  smpAnno <- tibble::as_tibble(colData(data), rownames = 'Sample')
  
  # Perform analysis
  if (method == 'proDA') {
    # Fit linear probabilistic dropout model to normalized data
    fit <- proDA(datMat, design = ~ smpAnno[[meta_var]])
    diffTerm <- proDA::result_names(fit)[2]
    assoTestRes <- proDA::test_diff(fit, contrast = diffTerm, sort_by = 'pval') %>%
      dplyr::select(name, pval, adj_pval, diff) %>%
      dplyr::rename(Feature = name, pVal = pval, pValAdj = adj_pval, logFC = diff)
  } else if (method == 'limma') {
    design <- model.matrix(~ smpAnno[[meta_var]])
    fit <- limma::lmFit(datMat, design = design)
    fit <- limma::eBayes(fit)
    assoTestRes <- limma::topTable(fit, coef = 2, number = Inf) %>%
      tibble::as_tibble(rownames = 'Feature') %>%
      dplyr::select(Feature, P.Value, adj.P.Val, logFC) %>%
      dplyr::rename(pVal = P.Value, pValAdj = adj.P.Val)
  }
  assoTestRes <- dplyr::mutate(assoTestRes,
                               pVal = as.numeric(formatC(pVal, format = 'g', digits = 3)),
                               pValAdj = as.numeric(formatC(pValAdj, format = 'g', digits = 3)))
  # Include feature 
  if (!is.null(prot_gene_tbl)) {
    assoTestRes <- dplyr::mutate(assoTestRes, Gene = plyr::mapvalues(Feature,
                                                                     from = prot_gene_tbl$Proteins,
                                                                     to = prot_gene_tbl$Genes,
                                                                     warn_missing = F)) %>%
      dplyr::relocate(Gene, .after = Feature)
  }
  
  return(assoTestRes)
}

# data1 <- proTissue_Discovery
# data2 <- proTissue_MethodDev
# name1 <- 'Discovery'
# name2 <- 'MethodDev'
# meta_var <- 'Recurrence'
# subset <- c('Condition', 'Tumor')
# method <- 'proDA'
# prot_gene_tbl <- featInfo
# alpha <- 0.05
# num_topSigFeats <- 6
# plot_pVal <- T
# plot_topSigFeats <- T
# plot_logFC <- T
summarizeFeatAssoRes <- function(data1, data2, name1, name2, meta_var, subset = NULL,
                                 prot_gene_tbl = NULL, method = 'proDA', alpha = 0.05,
                                 num_topSigFeats = 6, plot_pVal = T, plot_topSigFeats = T,
                                 plot_logFC = T) {
  #' Conduct univariate analysis (t-test for now) between features and sample metadata
  #' variable of interest, and combine and tidy up significance test and statistical
  #' results (p-values, top significant features, and log fold changes) into tables for making plots
  #' 
  #' Parameters
  #' data1, data2: SummarizedExperiment objects that contain the data matrices and metadata
  #' name1, name2: Characters indicating the dataset names in the plot legend
  #' meta_var: A character specifying the sample metadata variable that has two
  #' levels for grouping samples to conduct t-test
  #' subset: A vector of length 2 containing characters specifying the sample metadata
  #' variable and one of its level to split the dataset. Default is NULL, meaning
  #' the whole dataset is used
  #' prot_gene_tbl: A table containing proteins and their encoding genes for providing
  #' and labeling gene names in plots. Default is NULL. The column for proteins
  #' should be named 'Proteins' and for genes should be named 'Genes'. Furthermore,
  #' only the first protein and gene should be kept if there are more than one
  #' method: A character specifying the method used to conduct t-test, which should
  #' be one of 'proDA' (default) and 'limma'
  #' alpha: A numeric value specifying the significance level for summarizing p-value
  #' information. Default is 0.05
  #' num_topSigFeats: A numeric value specifying the number of top significant features
  #' to visualize. Default is 6
  #' plot_pVal: A logical variable indicating whether to generate the summarized
  #' p-value table for plotting. Default is TRUE
  #' plot_topSigFeats: A logical variable indicating whether to generate the summarized
  #' table of top significant features based on 'data1' for plotting. Default is TRUE
  #' plot_logFC: A logical variable indicating whether to generate the summarized
  #' log(FC) table for plotting. Default is TRUE
  #' 
  #' Return
  #' A list containing the following components:
  #' tbl_featAsso1: A table containing association test results of 'data1'
  #' tbl_featAsso2: A table containing association test results of 'data2'
  #' tbl_pVal: A table containing summarized significance test results, labeling
  #' information, etc. for making plots
  #' tbl_topSigFeats: A table containing abundances of top significant features,
  #' grouping information, etc. for making plots
  #' tbl_logFC: A table containing summarized log(FC) results, coloring information,
  #' etc. for making plots
  
  # Extract common features between two datasets
  # Select first protein as representative if there are more than one
  rownames(data1) <- stringr::str_remove(rownames(data1), ';.*')
  rownames(data2) <- stringr::str_remove(rownames(data2), ';.*')
  commonFeats <- intersect(rownames(data1), rownames(data2))
  
  # Compute associations between features and sample metadata variable of interest
  # and keep only results of common features
  assoTestRes1 <- doFeatAssoTest(data1, meta_var = meta_var, subset = subset,
                                 prot_gene_tbl = prot_gene_tbl, method = method)
  assoTestRes_dat1 <- dplyr::filter(assoTestRes1, Feature %in% commonFeats)
  assoTestRes2 <- doFeatAssoTest(data2, meta_var = meta_var, subset = subset,
                                 prot_gene_tbl = prot_gene_tbl, method = method)
  assoTestRes_dat2 <- dplyr::filter(assoTestRes2, Feature %in% commonFeats)
  
  # Summarize resulting p-Values
  if (plot_pVal) {
    # Prepare feature significance information
    tmp_assoTestRes_dat1 <- dplyr::select(assoTestRes_dat1, Feature, pVal) %>%
      dplyr::mutate(Sig_dat1 = dplyr::case_when(pVal <= alpha ~ 'Yes',
                                                pVal > alpha ~ 'No')) %>%
      dplyr::rename(pVal_dat1 = pVal)
    tmp_assoTestRes_dat2 <- dplyr::select(assoTestRes_dat2, Feature, pVal) %>%
      dplyr::mutate(Sig_dat2 = dplyr::case_when(pVal <= alpha ~ 'Yes',
                                                pVal > alpha ~ 'No')) %>%
      dplyr::rename(pVal_dat2 = pVal)
    # Combine association test results and all needed information into a table
    combinedAssoTestRes_pVal <- dplyr::left_join(tmp_assoTestRes_dat1, tmp_assoTestRes_dat2,
                                                 by = 'Feature') %>%
      # Include coloring information
      dplyr::mutate(Significance = dplyr::case_when(
        Sig_dat1 == 'Yes' & Sig_dat2 == 'Yes' ~ 'Sig. in Both',
        Sig_dat1 == 'Yes' & Sig_dat2 == 'No' ~ paste('Sig. in', name1),
        Sig_dat1 == 'No' & Sig_dat2 == 'Yes' ~ paste('Sig. in', name2),
        Sig_dat1 == 'No' & Sig_dat2 == 'No' ~ 'Not Sig.'
      ),
      Significance = factor(Significance, levels = c('Sig. in Both',
                                                     paste('Sig. in', name1),
                                                     paste('Sig. in', name2),
                                                     'Not Sig.')))
    # Include labeling information
    if (is.null(prot_gene_tbl)) {
      combinedAssoTestRes_pVal <- dplyr::mutate(combinedAssoTestRes_pVal,
                                                Label = dplyr::case_when(
                                                  Significance == 'Sig. in Both' ~ Feature
                                                ))
    } else {
      combinedAssoTestRes_pVal <- dplyr::mutate(combinedAssoTestRes_pVal,
                                                Gene = plyr::mapvalues(
                                                  Feature,
                                                  from = prot_gene_tbl$Proteins,
                                                  to = prot_gene_tbl$Genes,
                                                  warn_missing = F
                                                ),
                                                Label = dplyr::case_when(
                                                  Significance == 'Sig. in Both' ~ Gene
                                                ))
    }
    # Reorder points in graph
    significance <- combinedAssoTestRes_pVal$Significance
    combinedAssoTestRes_pVal <- combinedAssoTestRes_pVal[c(which(significance == 'Not Sig.'),
                                                           which(significance == 'Sig. in Both'),
                                                           which(significance == paste('Sig. in', name1)),
                                                           which(significance == paste('Sig. in', name2))),]
  } else {
    combinedAssoTestRes_pVal <- NULL
  }
  
  # Summarize top significant features
  # Check if there is any common significant feature
    sigFeats1 <- dplyr::filter(assoTestRes_dat1, pVal <= alpha) %>%
      dplyr::arrange() %>%
      dplyr::pull(Feature)
    sigFeats2 <- dplyr::filter(assoTestRes_dat2, pVal <= alpha) %>%
      dplyr::arrange() %>%
      dplyr::pull(Feature)
    sigFeats <- intersect(sigFeats1, sigFeats2)
  if (plot_topSigFeats & (length(sigFeats) != 0)) {
    # Retrieve top common significant features based on 'data1'
    topSigFeats <- intersect(sigFeats1, sigFeats2)
    if (length(topSigFeats) > num_topSigFeats) {
      topSigFeats <- topSigFeats[1:num_topSigFeats]
    }
    # Extract abundances of top common significant features and combine all needed
    # information into a table
    # Subset samples if specified
    if (!is.null(subset)) {
      data1 <- data1[, which(colData(data1)[[subset[1]]] == subset[2])]
      data2 <- data2[, which(colData(data2)[[subset[1]]] == subset[2])]
    }
    # Prepare sample grouping information
    smpGpInfo1 <- tibble::as_tibble(colData(data1), rownames = 'Sample') %>%
      dplyr::select(Sample, all_of(meta_var))
    smpGpInfo2 <- tibble::as_tibble(colData(data2), rownames = 'Sample') %>%
      dplyr::select(Sample, all_of(meta_var))
    # Prepare abundance information for top common significant features
    topSigFeatTab1 <- t(SummarizedExperiment::assay(data1)) %>%
      tibble::as_tibble(rownames = 'Sample') %>%
      dplyr::select(c(Sample, topSigFeats)) %>%
      tidyr::pivot_longer(cols = -'Sample', names_to = 'Feature', values_to = 'Abundance') %>%
      dplyr::left_join(smpGpInfo1, by = 'Sample') %>%
      dplyr::mutate(DataName = name1)
    topSigFeatTab2 <- t(SummarizedExperiment::assay(data2)) %>%
      tibble::as_tibble(rownames = 'Sample') %>%
      dplyr::select(c(Sample, topSigFeats)) %>%
      tidyr::pivot_longer(cols = -'Sample', names_to = 'Feature', values_to = 'Abundance') %>%
      dplyr::left_join(smpGpInfo2, by = 'Sample') %>%
      dplyr::mutate(DataName = name2)
    # Combine all information
    combinedTopSigFeatTab <- dplyr::bind_rows(topSigFeatTab1, topSigFeatTab2) %>%
      # Include new grouping information
      dplyr::mutate(newGroup = paste0(.data[[meta_var]], '_', DataName))
    # Create new feature names if feature annotations are provided
    if (!is.null(prot_gene_tbl)) {
      combinedTopSigFeatTab <- dplyr::mutate(combinedTopSigFeatTab,
                                             Gene = plyr::mapvalues(Feature,
                                                                    from = prot_gene_tbl$Proteins,
                                                                    to = prot_gene_tbl$Genes,
                                                                    warn_missing = F),
                                             newFeature = paste0(Feature, ' (', Gene, ')'),
                                             newFeature = factor(newFeature,
                                                                 levels = unique(newFeature)))
    }
  } else {
    combinedTopSigFeatTab <- NULL
  }
  
  # Summarize resulting log(FC)
  if (plot_logFC) {
    # Prepare feature log(FC) information
    tmp_assoTestRes_dat1 <- dplyr::select(assoTestRes_dat1, Feature, logFC) %>%
      dplyr::mutate(Dir_dat1 = dplyr::case_when(logFC > 0 ~ 'Up',
                                                logFC < 0 ~ 'Down')) %>%
      dplyr::rename(logFC_dat1 = logFC)
    tmp_assoTestRes_dat2 <- dplyr::select(assoTestRes_dat2, Feature, logFC) %>%
      dplyr::mutate(Dir_dat2 = dplyr::case_when(logFC > 0 ~ 'Up',
                                                logFC < 0 ~ 'Down')) %>%
      dplyr::rename(logFC_dat2 = logFC)
    # Combine association test results and all needed information into a table
    combinedAssoTestRes_logFC <- dplyr::left_join(tmp_assoTestRes_dat1, tmp_assoTestRes_dat2,
                                                  by = 'Feature') %>%
      # Include coloring information
      dplyr::mutate(Direction = dplyr::case_when(
        Dir_dat1 == 'Up' & Dir_dat2 == 'Up' ~ 'Up in Both',
        Dir_dat1 == 'Down' & Dir_dat2 == 'Down' ~ 'Down in Both',
        Dir_dat1 == 'Up' & Dir_dat2 == 'Down' ~ 'Inconsistent',
        Dir_dat1 == 'Down' & Dir_dat2 == 'Up' ~ 'Inconsistent'
      ),
      Direction = factor(Direction, levels = c('Up in Both', 'Down in Both', 'Inconsistent')))
    # Include gene annotations
    if (!is.null(prot_gene_tbl)) {
      combinedAssoTestRes_logFC <- dplyr::mutate(combinedAssoTestRes_logFC,
                                                 Gene = plyr::mapvalues(Feature,
                                                                        from = prot_gene_tbl$Proteins,
                                                                        to = prot_gene_tbl$Genes,
                                                                        warn_missing = F))
    }
    # Reorder points in graph
    direction <- combinedAssoTestRes_logFC$Direction
    combinedAssoTestRes_logFC <- combinedAssoTestRes_logFC[c(which(direction == 'Inconsistent'),
                                                             which(direction == 'Down in Both'),
                                                             which(direction == 'Up in Both')),]
  } else {
    combinedAssoTestRes_logFC <- NULL
  }
  
  return(list(tbl_featAsso1 = assoTestRes1, tbl_featAsso2 = assoTestRes2,
              tbl_pVal = combinedAssoTestRes_pVal, tbl_topSigFeats = combinedTopSigFeatTab,
              tbl_logFC = combinedAssoTestRes_logFC))
}
```

# Tissue DIA Proteomics
Metadata-assisted quality assessment: Check association of tissue type-related feature
significance between Discovery and Method Development datasets

## AG Krijgsveld

Display association of tissue type-related feature significance between Discovery
and Method Development DIA Proteomics generated by AG Krijgsveld\
**p-values (significance level = 0.05)**
```{r}
# Load normalized data
proTissue_Discovery <- readRDS('./data/Discovery/AG_Krijgsveld/proTissueVsn.rds')
proTissue_MethodDev <- readRDS('./data/MethodDev/AG_Krijgsveld/new_proTissueVsn.rds')
# Prepare protein-gene table for providing gene names
featInfo <- tibble::as_tibble(rowData(proTissue_Discovery), rownames = 'Proteins') %>%
  dplyr::select(Proteins, Genes) %>%
  dplyr::mutate(Proteins = stringr::str_remove(Proteins, ';.*'),
                Genes = stringr::str_remove(Genes, ';.*'))

# Prepare summarized tables of comparison between Tumor and Normal tissue samples
condiFeatAssoTabs_Krij <- summarizeFeatAssoRes(proTissue_Discovery, proTissue_MethodDev,
                                               'Discovery', 'MethodDev', meta_var = 'Condition',
                                               subset = NULL, method = 'proDA', alpha = 0.05,
                                               prot_gene_tbl = featInfo)
featAssoTabs <- condiFeatAssoTabs_Krij

# Plot p-values
ggplot(featAssoTabs$tbl_pVal, aes(x=-log10(pVal_dat1), y=-log10(pVal_dat2), col=Significance)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.8) +
  scale_color_manual(values = c('red', 'darkgreen', 'navy', 'grey')) +
  labs(x = '-log10(pVal) of Discovery Cohort', y = '-log10(pVal) of MethodDev Cohort',
       title = 'Tumor vs Normal: Tissue DIA Proteomics (AG Krijgsveld)') +
  th #+ theme(axis.title = element_text(size = 22),
             # axis.text = element_text(size = 16),
             # legend.title = element_text(size = 20), legend.text = element_text(size = 18))
# ggsave('./output/SC_meeting/pVals_Dis_vs_MetDev_proTissue_condi_Krij.png', device = 'png', dpi = 400, height = 8, width = 11)

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_pVal$Significance == 'Sig. in Both') /
                   sum(featAssoTabs$tbl_pVal$Significance != 'Not Sig.'), 3) * 100
cat(paste0('=> ', percent, '% of significant results are consistent'))
```

**log(FC)**
```{r}
# Plot log(FC)
ggplot(featAssoTabs$tbl_logFC, aes(x=logFC_dat1, y=logFC_dat2, col=Direction)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.8) +
  scale_color_manual(values = c('red', 'navy', 'grey')) +
  labs(x = 'log(FC) of Discovery Cohort', y = 'log(FC) of MethodDev Cohort',
       title = 'Tumor vs Normal: Tissue DIA Proteomics (AG Krijgsveld)') +
  th #+ theme(axis.title = element_text(size = 22),
             # axis.text = element_text(size = 16),
             # legend.title = element_text(size = 20), legend.text = element_text(size = 18))
# ggsave('./output/SC_meeting/log(FC)_Dis_vs_MetDev_proTissue_condi_Krij.png', device = 'png', dpi = 400, height = 8, width = 11)

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_logFC$Direction != 'Inconsistent') /
                   nrow(featAssoTabs$tbl_logFC), 3) * 100
cat(paste0('=> ', percent, '% of results are consistent'))
```

## AG Klingmüller

Display association of tissue type-related feature significance between Discovery
and Method Development DIA Proteomics generated by AG Klingmüller\
**p-values (significance level = 0.05)**
```{r}
# Load normalized data
proTissue_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsn.rds')
proTissue_Discovery <- proTissue_Klin[, colData(proTissue_Klin)$Cohort == 'Discovery']
proTissue_MethodDev <- proTissue_Klin[, colData(proTissue_Klin)$Cohort == 'MethodDev']
# Prepare protein-gene table for providing gene names
featInfo <- tibble::as_tibble(rowData(proTissue_Discovery), rownames = 'Proteins') %>%
  dplyr::rename(Genes = PG.Genes) %>%
  dplyr::mutate(Proteins = stringr::str_remove(Proteins, ';.*'),
                Genes = stringr::str_remove(Genes, ';.*'))

# Prepare summarized tables of comparison between Tumor and Normal tissue samples
condiFeatAssoTabs_Klin <- summarizeFeatAssoRes(proTissue_Discovery, proTissue_MethodDev,
                                               'Discovery', 'MethodDev', meta_var = 'Condition',
                                               subset = NULL, method = 'proDA', alpha = 0.05,
                                               prot_gene_tbl = featInfo)
featAssoTabs <- condiFeatAssoTabs_Klin

# Plot p-values
ggplot(featAssoTabs$tbl_pVal, aes(x=-log10(pVal_dat1), y=-log10(pVal_dat2), col=Significance)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.8) +
  scale_color_manual(values = c('red', 'darkgreen', 'navy', 'grey')) +
  labs(x = '-log10(pVal) of Discovery Cohort', y = '-log10(pVal) of MethodDev Cohort',
       title = 'Tumor vs Normal: Tissue DIA Proteomics (AG Klingmüller)') +
  th #+ theme(axis.title = element_text(size = 22),
             # axis.text = element_text(size = 16),
             # legend.title = element_text(size = 20), legend.text = element_text(size = 18))
# ggsave('./output/SC_meeting/pVals_Dis_vs_MetDev_proTissue_condi_Klin.png', device = 'png', dpi = 400, height = 8, width = 11)

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_pVal$Significance == 'Sig. in Both') /
                   sum(featAssoTabs$tbl_pVal$Significance != 'Not Sig.'), 3) * 100
cat(paste0('=> ', percent, '% of significant results are consistent'))
```

**log(FC)**
```{r}
# Plot log(FC)
ggplot(featAssoTabs$tbl_logFC, aes(x=logFC_dat1, y=logFC_dat2, col=Direction)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.8) +
  scale_color_manual(values = c('red', 'navy', 'grey')) +
  labs(x = 'log(FC) of Discovery Cohort', y = 'log(FC) of MethodDev Cohort',
       title = 'Tumor vs Normal: Tissue DIA Proteomics (AG Klingmüller)') +
  th #+ theme(axis.title = element_text(size = 22),
#              axis.text = element_text(size = 16),
#              legend.title = element_text(size = 20), legend.text = element_text(size = 18))
# ggsave('./output/SC_meeting/log(FC)_Dis_vs_MetDev_proTissue_condi_Klin.png', device = 'png', dpi = 400, height = 8, width = 11)

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_logFC$Direction != 'Inconsistent') /
                   nrow(featAssoTabs$tbl_logFC), 3) * 100
cat(paste0('=> ', percent, '% of results are consistent'))
```

```{r include=F, eval=F}
## Method Development
# Load normalized data
proTissue_Krij <- readRDS('./data/MethodDev/AG_Krijgsveld/new_proTissueVsn.rds')
proTissue_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsn.rds')
proTissue_Klin <- proTissue_Klin[, colData(proTissue_Klin)$Cohort == 'MethodDev']
# Prepare protein-gene table for providing gene names
featInfo <- tibble::as_tibble(rowData(proTissue_Klin), rownames = 'Proteins') %>%
  dplyr::rename(Genes = PG.Genes) %>%
  dplyr::mutate(Proteins = stringr::str_remove(Proteins, ';.*'),
                Genes = stringr::str_remove(Genes, ';.*'))

# Prepare summarized tables of comparison between Tumor and Normal tissue samples
condiFeatAssoTabs_MethodDev <- summarizeFeatAssoRes(proTissue_Krij, proTissue_Klin,
                                                    'Krij.', 'Klin.', meta_var = 'Condition',
                                                    subset = NULL, method = 'proDA', alpha = 0.05,
                                                    prot_gene_tbl = featInfo)
featAssoTabs <- condiFeatAssoTabs_MethodDev

# Plot p-values
ggplot(featAssoTabs$tbl_pVal, aes(x=-log10(pVal_dat1), y=-log10(pVal_dat2), col=Significance)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.8) +
  scale_color_manual(values = c('red', 'darkgreen', 'navy', 'grey')) +
  labs(x = '-log10(pVal) of Krij.', y = '-log10(pVal) of Klin.',
       title = 'Tumor vs Normal: Tissue DIA Proteomics (MethodDev)') +
  th
# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_pVal$Significance == 'Sig. in Both') /
                   sum(featAssoTabs$tbl_pVal$Significance != 'Not Sig.'), 3) * 100
cat(paste0('=> ', percent, '% of significant results are consistent'))

# Plot log(FC)
ggplot(featAssoTabs$tbl_logFC, aes(x=logFC_dat1, y=logFC_dat2, col=Direction)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.8) +
  scale_color_manual(values = c('red', 'navy', 'grey')) +
  labs(x = 'log(FC) of Krij.', y = 'log(FC) of Klin.',
       title = 'Tumor vs Normal: Tissue DIA Proteomics (MethodDev)') +
  th
# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_logFC$Direction != 'Inconsistent') /
                   nrow(featAssoTabs$tbl_logFC), 3) * 100
cat(paste0('=> ', percent, '% of results are consistent'))

## Discovery
# Load normalized data
proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/proTissueVsn.rds')
proTissue_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsn.rds')
proTissue_Klin <- proTissue_Klin[, colData(proTissue_Klin)$Cohort == 'Discovery']
# Prepare protein-gene table for providing gene names
featInfo <- tibble::as_tibble(rowData(proTissue_Klin), rownames = 'Proteins') %>%
  dplyr::rename(Genes = PG.Genes) %>%
  dplyr::mutate(Proteins = stringr::str_remove(Proteins, ';.*'),
                Genes = stringr::str_remove(Genes, ';.*'))

# Prepare summarized tables of comparison between Tumor and Normal tissue samples
condiFeatAssoTabs_Discovery <- summarizeFeatAssoRes(proTissue_Krij, proTissue_Klin,
                                                    'Krij.', 'Klin.', meta_var = 'Condition',
                                                    subset = NULL, method = 'proDA', alpha = 0.05,
                                                    prot_gene_tbl = featInfo)
featAssoTabs <- condiFeatAssoTabs_Discovery

# Plot p-values
ggplot(featAssoTabs$tbl_pVal, aes(x=-log10(pVal_dat1), y=-log10(pVal_dat2), col=Significance)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.8) +
  scale_color_manual(values = c('red', 'darkgreen', 'navy', 'grey')) +
  labs(x = '-log10(pVal) of Krij.', y = '-log10(pVal) of Klin.',
       title = 'Tumor vs Normal: Tissue DIA Proteomics (Discovery)') +
  th
# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_pVal$Significance == 'Sig. in Both') /
                   sum(featAssoTabs$tbl_pVal$Significance != 'Not Sig.'), 3) * 100
cat(paste0('=> ', percent, '% of significant results are consistent'))

# Plot log(FC)
ggplot(featAssoTabs$tbl_logFC, aes(x=logFC_dat1, y=logFC_dat2, col=Direction)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.8) +
  scale_color_manual(values = c('red', 'navy', 'grey')) +
  labs(x = 'log(FC) of Krij.', y = 'log(FC) of Klin.',
       title = 'Tumor vs Normal: Tissue DIA Proteomics (Discovery)') +
  th
# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_logFC$Direction != 'Inconsistent') /
                   nrow(featAssoTabs$tbl_logFC), 3) * 100
cat(paste0('=> ', percent, '% of results are consistent'))
```


# Normal Tissue Proteomics
Check association of recurrence-related feature significance between Discovery and
Method Development datasets from same working group, so that working group differences
can be excluded

## AG Krijgsveld

Display association of feature recurrence predictive power between Discovery and
Method Development Normal Tissue DIA Proteomics generated by AG Krijgsveld\
**p-values (significance level = 0.05)**
```{r}
# Load normalized data
proTissue_Discovery <- readRDS('./data/Discovery/AG_Krijgsveld/proTissueVsn.rds')
proTissue_MethodDev <- readRDS('./data/MethodDev/AG_Krijgsveld/new_proTissueVsn.rds')
# Prepare protein-gene table for providing gene names
featInfo <- tibble::as_tibble(rowData(proTissue_Discovery), rownames = 'Proteins') %>%
  dplyr::select(Proteins, Genes) %>%
  dplyr::mutate(Proteins = stringr::str_remove(Proteins, ';.*'),
                Genes = stringr::str_remove(Genes, ';.*'))

# Prepare summarized tables of comparison between Recurrence and Non-recurrence patient groups
recurNormalFeatAssoTabs_Krij <- summarizeFeatAssoRes(proTissue_Discovery, proTissue_MethodDev,
                                                     'Discovery', 'MethodDev',
                                                     meta_var = 'Recurrence',
                                                     method = 'proDA',
                                                     subset = c('Condition', 'Normal'),
                                                     prot_gene_tbl = featInfo,
                                                     alpha = 0.05)
featAssoTabs <- recurNormalFeatAssoTabs_Krij

# Plot p-values
ggplot(featAssoTabs$tbl_pVal, aes(x=-log10(pVal_dat1), y=-log10(pVal_dat2),
                                  col=Significance, label=Label)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.5) +
  scale_color_manual(values = c('red', 'darkgreen', 'navy', 'grey')) +
  geom_text_repel(size = 4, show.legend = F) +
  labs(x = '-log10(pVal) of Discovery Cohort', y = '-log10(pVal) of MethodDev Cohort',
       title = 'Recurrence: Normal Tissue DIA Proteomics (AG Krijgsveld)') +
  th #+ theme(axis.title = element_text(size = 22),
             # axis.text = element_text(size = 16),
             # legend.title = element_text(size = 20), legend.text = element_text(size = 18))
# ggsave('./output/SC_meeting/pVals_Dis_vs_MetDev_proNormal_recur_Krij.png', device = 'png', dpi = 400, height = 8, width = 11)

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_pVal$Significance == 'Sig. in Both') /
                   sum(featAssoTabs$tbl_pVal$Significance != 'Not Sig.'), 3) * 100
cat(paste0('=> ', percent, '% of significant results are consistent'))
```

**log(FC)**
```{r}
# Plot log(FC)
ggplot(featAssoTabs$tbl_logFC, aes(x=logFC_dat1, y=logFC_dat2, col=Direction)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.5) +
  scale_color_manual(values = c('red', 'navy', 'grey')) +
  labs(x = 'log(FC) of Discovery Cohort', y = 'log(FC) of MethodDev Cohort',
       title = 'Recurrence: Normal Tissue DIA Proteomics (AG Krijgsveld)') +
  th #+ theme(axis.title = element_text(size = 22),
             # axis.text = element_text(size = 16),
             # legend.title = element_text(size = 20), legend.text = element_text(size = 18))
# ggsave('./output/SC_meeting/log(FC)_Dis_vs_MetDev_proNormal_recur_Krij.png', device = 'png', dpi = 400, height = 8, width = 11)

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_logFC$Direction != 'Inconsistent') /
                   nrow(featAssoTabs$tbl_logFC), 3) * 100
cat(paste0('=> ', percent, '% of results are consistent'))
```

Visualize top significant proteins identified in both Discovery and Method Development
datasets and based on Discovery one. Note that Method Development data is provided
to have more comprehensive analysis and it may not be on same scale as Discovery data.\
ns: p > 0.05, ∗: p <= 0.05, ∗∗: p <= 0.01, ***: p <= 0.001
```{r fig.height=12}
# Plot top significant features based on 'data1'
topSigFeatTab <- dplyr::mutate(featAssoTabs$tbl_topSigFeats,
                               newGroup = factor(newGroup, levels = c('Yes_Discovery',
                                                                      'No_Discovery',
                                                                      'Yes_MethodDev',
                                                                      'No_MethodDev')))
ggplot(topSigFeatTab, aes(x=newGroup, y=Abundance, col=DataName, fill=Recurrence)) +
  geom_boxplot(alpha = 1, outlier.shape = NA, linewidth = 1) +
  geom_jitter(position = position_jitter(0.2), size = 3, show.legend = F) +
  ggpubr::stat_compare_means(method = 't.test', paired = F, method.args = list(var.equal = T),
                             comparisons = list(c('Yes_Discovery', 'No_Discovery'),
                                                c('Yes_MethodDev', 'No_MethodDev')),
                             label = 'p.signif', tip.length = 0.015, bracket.size = 0.7, size = 4) +
  labs(x = 'Recurrence', title = 'Recurrence: Normal Tissue DIA Proteomics (AG Krijgsveld)') +
  scale_color_manual(name = 'Cohort', values = c('firebrick', 'grey50')) +
  scale_fill_manual(values=c('#00BFC4', '#F8766D')) +
  facet_wrap(vars(newFeature), scales = 'free') +
  th +
  theme(strip.text = element_text(size = 13, face = 'bold'),
        axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1))
#   theme(strip.text = element_text(size = 14, face = 'bold'),
#         axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1, size = 20),
#         axis.title = element_text(size = 26),
#         legend.title = element_text(size = 24), legend.text = element_text(size = 22))
# ggsave('./output/group_meeting/topSigFeats_Discovery_vs_MethodDev_proNormal_Krij.png', device = 'png', dpi = 400, height = 8, width = 16)
```

## AG Klingmüller

Display association of feature recurrence predictive power between Discovery and
Method Development Normal Tissue DIA Proteomics generated by AG Klingmüller\
**p-values (significance level = 0.05)**
```{r}
# Load normalized data
proTissue_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsn.rds')
proTissue_Discovery <- proTissue_Klin[, colData(proTissue_Klin)$Cohort == 'Discovery']
proTissue_MethodDev <- proTissue_Klin[, colData(proTissue_Klin)$Cohort == 'MethodDev']
# Prepare protein-gene table for providing gene names
featInfo <- tibble::as_tibble(rowData(proTissue_Discovery), rownames = 'Proteins') %>%
  dplyr::rename(Genes = PG.Genes) %>%
  dplyr::mutate(Proteins = stringr::str_remove(Proteins, ';.*'),
                Genes = stringr::str_remove(Genes, ';.*'))

# Prepare summarized tables of comparison between Recurrence and Non-recurrence patient groups
recurNormalFeatAssoTabs_Klin <- summarizeFeatAssoRes(proTissue_Discovery, proTissue_MethodDev,
                                                     'Discovery', 'MethodDev',
                                                     meta_var = 'Recurrence',
                                                     subset = c('Condition', 'Normal'),
                                                     prot_gene_tbl = featInfo,
                                                     method = 'proDA',
                                                     alpha = 0.05)
featAssoTabs <- recurNormalFeatAssoTabs_Klin

# Plot p-values
ggplot(featAssoTabs$tbl_pVal, aes(x=-log10(pVal_dat1), y=-log10(pVal_dat2),
                                  col=Significance, label=Label)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.5) +
  scale_color_manual(values = c('red', 'darkgreen', 'navy', 'grey')) +
  geom_text_repel(size = 4, show.legend = F) +
  labs(x = '-log10(pVal) of Discovery Cohort', y = '-log10(pVal) of MethodDev Cohort',
       title = 'Recurrence: Normal Tissue DIA Proteomics (AG Klingmüller)') +
  th #+ theme(axis.title = element_text(size = 22),
             # axis.text = element_text(size = 16),
             # legend.title = element_text(size = 20), legend.text = element_text(size = 18))
# ggsave('./output/SC_meeting/pVals_Dis_vs_MetDev_proNormal_recur_Klin.png', device = 'png', dpi = 400, height = 8, width = 11)

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_pVal$Significance == 'Sig. in Both') /
                   sum(featAssoTabs$tbl_pVal$Significance != 'Not Sig.'), 3) * 100
cat(paste0('=> ', percent, '% of significant results are consistent'))
```

**log(FC)**
```{r}
# Plot log(FC)
ggplot(featAssoTabs$tbl_logFC, aes(x=logFC_dat1, y=logFC_dat2, col=Direction)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.5) +
  scale_color_manual(values = c('red', 'navy', 'grey')) +
  labs(x = 'log(FC) of Discovery Cohort', y = 'log(FC) of MethodDev Cohort',
       title = 'Recurrence: Normal Tissue DIA Proteomics (AG Klingmüller)') +
  th #+ theme(axis.title = element_text(size = 22),
             # axis.text = element_text(size = 16),
             # legend.title = element_text(size = 20), legend.text = element_text(size = 18))
# ggsave('./output/SC_meeting/log(FC)_Dis_vs_MetDev_proNormal_recur_Klin.png', device = 'png', dpi = 400, height = 8, width = 11)

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_logFC$Direction != 'Inconsistent') /
                   nrow(featAssoTabs$tbl_logFC), 3) * 100
cat(paste0('=> ', percent, '% of results are consistent'))
```

Visualize top significant proteins identified in both Discovery and Method Development
datasets and based on Discovery one. Note that Method Development data is provided
to have more comprehensive analysis and it may not be on same scale as Discovery data.\
ns: p > 0.05, ∗: p <= 0.05, ∗∗: p <= 0.01, ***: p <= 0.001
```{r fig.height=12}
# Plot top significant features based on 'data1'
topSigFeatTab <- dplyr::mutate(featAssoTabs$tbl_topSigFeats,
                               newGroup = factor(newGroup, levels = c('Yes_Discovery',
                                                                      'No_Discovery',
                                                                      'Yes_MethodDev',
                                                                      'No_MethodDev')))
ggplot(topSigFeatTab, aes(x=newGroup, y=Abundance, col=DataName, fill=Recurrence)) +
  geom_boxplot(alpha = 1, outlier.shape = NA, linewidth = 1) +
  geom_jitter(position = position_jitter(0.2), size = 3, show.legend = F) +
  ggpubr::stat_compare_means(method = 't.test', paired = F, method.args = list(var.equal = T),
                             comparisons = list(c('Yes_Discovery', 'No_Discovery'),
                                                c('Yes_MethodDev', 'No_MethodDev')),
                             label = 'p.signif', tip.length = 0.015, bracket.size = 0.7, size = 4) +
  labs(x = 'Recurrence', title = 'Recurrence: Normal Tissue DIA Proteomics (AG Klingmüller)') +
  scale_color_manual(name = 'Cohort', values = c('firebrick', 'grey50')) +
  scale_fill_manual(values=c('#00BFC4', '#F8766D')) +
  facet_wrap(vars(newFeature), scales = 'free') +
  th +
  theme(strip.text = element_text(size = 13, face = 'bold'),
        axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1))
```

## Integrative analysis

Show intersections of potential Normal proteins identified in different datasets
(significance level = 0.05)
```{r}
# Prepare recurrence-related significant features
alpha <- 0.05
# AG Klingmüller
sigFeats_Discovery_Klin <- recurNormalFeatAssoTabs_Klin$tbl_featAsso1 %>%
  dplyr::filter(pVal <= alpha) %>%
  dplyr::pull(Feature)
sigFeats_MethodDev_Klin <- recurNormalFeatAssoTabs_Klin$tbl_featAsso2 %>%
  dplyr::filter(pVal <= alpha) %>%
  dplyr::pull(Feature)
# AG Krijgsveld
sigFeats_Discovery_Krij <- recurNormalFeatAssoTabs_Krij$tbl_featAsso1 %>%
  dplyr::filter(pVal <= alpha) %>%
  dplyr::pull(Feature)
sigFeats_MethodDev_Krij <- recurNormalFeatAssoTabs_Krij$tbl_featAsso2 %>%
  dplyr::filter(pVal <= alpha) %>%
  dplyr::pull(Feature)

# Make venn graphs
ggvenn(list(Discovery_Krij. = sigFeats_Discovery_Krij, MethodDev_Krij. = sigFeats_MethodDev_Krij,
            MethodDev_Klin. = sigFeats_MethodDev_Klin),
       fill_color = c('darkred', 'darkblue', 'darkgreen'), fill_alpha = 0.6,
       set_name_color = c('darkred', 'darkblue', 'darkgreen'), set_name_size = 9,
       text_size = 5, stroke_size = 1.5)
# ggsave('./output/SC_meeting/venn_sigFeats_Krij_dis.png', device = 'png', dpi = 400, height = 8, width = 11)
# Identify fully overlapped features
# ol <- intersect(intersect(sigFeats_Discovery_Krij, sigFeats_MethodDev_Krij), sigFeats_MethodDev_Klin)
# recurNormalFeatAssoTabs_Krij$tbl_featAsso1[recurNormalFeatAssoTabs_Krij$tbl_featAsso1$Feature %in% ol,]
```
=> That one protein identified in all datasets is 'P37235 (HPCAL1)'.

```{r}
ggvenn(list(Discovery_Klin. = sigFeats_Discovery_Klin, MethodDev_Klin. = sigFeats_MethodDev_Klin,
            MethodDev_Krij. = sigFeats_MethodDev_Krij),
       fill_color = c('darkorange', 'darkgreen', 'darkblue'), fill_alpha = 0.6,
       set_name_color = c('darkorange', 'darkgreen', 'darkblue'), set_name_size = 9,
       text_size = 5, stroke_size = 1.5)
# ggsave('./output/SC_meeting/venn_sigFeats_Klin_dis.png', device = 'png', dpi = 400, height = 8, width = 11)
# Identify fully overlapped features
# ol <- intersect(intersect(sigFeats_Discovery_Klin, sigFeats_MethodDev_Klin), sigFeats_MethodDev_Krij)
# recurNormalFeatAssoTabs_Klin$tbl_featAsso1[recurNormalFeatAssoTabs_Klin$tbl_featAsso1$Feature %in% ol,]
```
=> Those proteins identified in all datasets are 'Q5S007 (LRRK2)', 'Q99417 (MYCBP)',
'Q96PE3 (INPP4A)', 'Q8NBQ5 (HSD17B11)', 'Q05823 (RNASEL)', 'Q9Y3E5 (PTRH2)', 'Q9H8L6 (MMRN2)'

```{r}
ggvenn(list(Discovery_Krij. = sigFeats_Discovery_Krij, Discovery_Klin. = sigFeats_Discovery_Klin,
            MethodDev_Klin. = sigFeats_MethodDev_Klin),
       fill_color = c('darkred', 'darkorange', 'darkgreen'), fill_alpha = 0.6,
       set_name_color = c('darkred', 'darkorange', 'darkgreen'), set_name_size = 9,
       text_size = 5, stroke_size = 1.5)
# ggsave('./output/SC_meeting/venn_sigFeats_Krij_Klin_dis.png', device = 'png', dpi = 400, height = 8, width = 11)
# Identify fully overlapped features
# ol <- intersect(intersect(sigFeats_Discovery_Krij, sigFeats_Discovery_Klin), sigFeats_MethodDev_Klin)
# recurNormalFeatAssoTabs_Krij$tbl_featAsso1[recurNormalFeatAssoTabs_Krij$tbl_featAsso1$Feature %in% ol,]
```
=> That one protein identified in all datasets is 'O94855 (SEC24D)'.

```{r eval=F, include=F}
# Total 1158 potential Normal Tissue Proteins were identified in Method Development
# Normal Tissue DIA Proteomics from AG Klingmüller and AG Krijgsveld, and 892 of
# them are captured in Discovery dataset of AG Krijgsveld.

# Show predictive ability of captured potential proteins in Discovery dataset
featAssoRes_Discovery <- recurNormalFeatAssoTabs_Krij$tbl_featAsso1
# Extract only results of captured potential proteins
sigFeats_MethodDev <- unique(c(sigFeats_MethodDev_Klin, sigFeats_MethodDev_Krij))
sigFeats_MethodDev <- intersect(sigFeats_MethodDev, featAssoRes_Discovery$Feature)
featAssoRes_Discovery %>%
  dplyr::filter(Feature %in% sigFeats_MethodDev) %>%
  dplyr::select(Feature, pVal, pValAdj, Gene) %>%
  DT::datatable()
```

```{r eval=F, include=F}
# Focus on AG Klingmüller's dataset to exclude working group differences.

# Show predictive ability of potential proteins identified in MethodDev dataset in Discovery one
featAssoRes_Discovery <- recurNormalFeatAssoTabs_Klin$tbl_featAsso1
# Extract only results of potential proteins identified in MethodDev dataset
common_sigFeats <- unique(c(sigFeats_Discovery_Klin, sigFeats_MethodDev_Klin))
featAssoRes_Discovery %>%
  dplyr::filter(Feature %in% common_sigFeats) %>%
  dplyr::select(Feature, pVal, pValAdj, Gene) %>%
  DT::datatable()
```




# Tumor Tissue Proteomics
Check association of recurrence-related feature significance between Discovery and
Method Development datasets from same working group, so that working group differences
can be excluded

## AG Krijgsveld

Display association of feature recurrence predictive power between Discovery and
Method Development Tumor Tissue DIA Proteomics generated by AG Krijgsveld\
**p-values (significance level = 0.05)**
```{r}
# Load normalized data
proTissue_Discovery <- readRDS('./data/Discovery/AG_Krijgsveld/proTissueVsn.rds')
proTissue_MethodDev <- readRDS('./data/MethodDev/AG_Krijgsveld/new_proTissueVsn.rds')
# Prepare protein-gene table for providing gene names
featInfo <- tibble::as_tibble(rowData(proTissue_Discovery), rownames = 'Proteins') %>%
  dplyr::select(Proteins, Genes) %>%
  dplyr::mutate(Proteins = stringr::str_remove(Proteins, ';.*'),
                Genes = stringr::str_remove(Genes, ';.*'))

# Prepare summarized tables of comparison between Recurrence and Non-recurrence patient groups
recurTumorFeatAssoTabs_Krij <- summarizeFeatAssoRes(proTissue_Discovery, proTissue_MethodDev,
                                                    'Discovery', 'MethodDev',
                                                    meta_var = 'Recurrence',
                                                    subset = c('Condition', 'Tumor'),
                                                    prot_gene_tbl = featInfo,
                                                    method = 'proDA',
                                                    alpha = 0.05)
featAssoTabs <- recurTumorFeatAssoTabs_Krij

# Plot p-values
ggplot(featAssoTabs$tbl_pVal, aes(x=-log10(pVal_dat1), y=-log10(pVal_dat2),
                                  col=Significance, label=Label)) +
  geom_point(size = 2) +
  scale_color_manual(values = c('red', 'darkgreen', 'navy', 'grey')) +
  geom_text_repel(size = 4, show.legend = F) +
  labs(x = '-log10(pVal) of Discovery Cohort', y = '-log10(pVal) of MethodDev Cohort',
       title = 'Recurrence: Tumor Tissue DIA Proteomics (AG Krijgsveld)') +
  th

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_pVal$Significance == 'Sig. in Both') /
                   sum(featAssoTabs$tbl_pVal$Significance != 'Not Sig.'), 3) * 100
cat(paste0('=> ', percent, '% of significant results are consistent'))
```

**log(FC)**
```{r}
# Plot log(FC)
ggplot(featAssoTabs$tbl_logFC, aes(x=logFC_dat1, y=logFC_dat2, col=Direction)) +
  geom_point(size = 2) +
  scale_color_manual(values = c('red', 'navy', 'grey')) +
  labs(x = 'log(FC) of Discovery Cohort', y = 'log(FC) of MethodDev Cohort',
       title = 'Recurrence: Tumor Tissue DIA Proteomics (AG Krijgsveld)') +
  th

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_logFC$Direction != 'Inconsistent') /
                   nrow(featAssoTabs$tbl_logFC), 3) * 100
cat(paste0('=> ', percent, '% of results are consistent'))
```

Visualize top significant proteins identified in both Discovery and Method Development
datasets and based on Discovery one. Note that Method Development data is provided
to have more comprehensive analysis and it may not be on same scale as Discovery data.\
ns: p > 0.05, ∗: p <= 0.05, ∗∗: p <= 0.01, ***: p <= 0.001
```{r fig.height=12}
# Plot top significant features based on 'data1'
topSigFeatTab <- dplyr::mutate(featAssoTabs$tbl_topSigFeats,
                               newGroup = factor(newGroup, levels = c('Yes_Discovery',
                                                                      'No_Discovery',
                                                                      'Yes_MethodDev',
                                                                      'No_MethodDev')))
ggplot(topSigFeatTab, aes(x=newGroup, y=Abundance, col=DataName, fill=Recurrence)) +
  geom_boxplot(alpha = 1, outlier.shape = NA, linewidth = 1) +
  geom_jitter(position = position_jitter(0.2), size = 3, show.legend = F) +
  ggpubr::stat_compare_means(method = 't.test', paired = F, method.args = list(var.equal = T),
                             comparisons = list(c('Yes_Discovery', 'No_Discovery'),
                                                c('Yes_MethodDev', 'No_MethodDev')),
                             label = 'p.signif', tip.length = 0.015, bracket.size = 0.7, size = 4) +
  labs(x = 'Recurrence', title = 'Recurrence: Tumor Tissue DIA Proteomics (AG Krijgsveld)') +
  scale_color_manual(name = 'Cohort', values = c('firebrick', 'grey50')) +
  scale_fill_manual(values=c('#00BFC4', '#F8766D')) +
  facet_wrap(vars(newFeature), scales = 'free') +
  th +
  theme(strip.text = element_text(size = 13, face = 'bold'),
        axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1))
```

## AG Klingmüller

Display association of feature recurrence predictive power between Discovery and
Method Development Tumor Tissue DIA Proteomics generated by AG Klingmüller\
**p-values (significance level = 0.05)**
```{r}
# Load normalized data
proTissue_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsn.rds')
proTissue_Discovery <- proTissue_Klin[, colData(proTissue_Klin)$Cohort == 'Discovery']
proTissue_MethodDev <- proTissue_Klin[, colData(proTissue_Klin)$Cohort == 'MethodDev']
# Prepare protein-gene table for providing gene names
featInfo <- tibble::as_tibble(rowData(proTissue_Discovery), rownames = 'Proteins') %>%
  dplyr::rename(Genes = PG.Genes) %>%
  dplyr::mutate(Proteins = stringr::str_remove(Proteins, ';.*'),
                Genes = stringr::str_remove(Genes, ';.*'))

# Prepare summarized tables of comparison between Recurrence and Non-recurrence patient groups
recurTumorFeatAssoTabs_Klin <- summarizeFeatAssoRes(proTissue_Discovery, proTissue_MethodDev,
                                                    'Discovery', 'MethodDev',
                                                    meta_var = 'Recurrence',
                                                    subset = c('Condition', 'Tumor'),
                                                    prot_gene_tbl = featInfo,
                                                    method = 'proDA',
                                                    alpha = 0.05)
featAssoTabs <- recurTumorFeatAssoTabs_Klin

# Plot p-values
ggplot(featAssoTabs$tbl_pVal, aes(x=-log10(pVal_dat1), y=-log10(pVal_dat2),
                                  col=Significance, label=Label)) +
  geom_point(size = 2) +
  scale_color_manual(values = c('red', 'darkgreen', 'navy', 'grey')) +
  geom_text_repel(size = 4, show.legend = F) +
  labs(x = '-log10(pVal) of Discovery Cohort', y = '-log10(pVal) of MethodDev Cohort',
       title = 'Recurrence: Tumor Tissue DIA Proteomics (AG Klingmüller)') +
  th

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_pVal$Significance == 'Sig. in Both') /
                   sum(featAssoTabs$tbl_pVal$Significance != 'Not Sig.'), 3) * 100
cat(paste0('=> ', percent, '% of significant results are consistent'))
```

**log(FC)**
```{r}
# Plot log(FC)
ggplot(featAssoTabs$tbl_logFC, aes(x=logFC_dat1, y=logFC_dat2, col=Direction)) +
  geom_point(size = 2) +
  scale_color_manual(values = c('red', 'navy', 'grey')) +
  labs(x = 'log(FC) of Discovery Cohort', y = 'log(FC) of MethodDev Cohort',
       title = 'Recurrence: Tumor Tissue DIA Proteomics (AG Klingmüller)') +
  th

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_logFC$Direction != 'Inconsistent') /
                   nrow(featAssoTabs$tbl_logFC), 3) * 100
cat(paste0('=> ', percent, '% of results are consistent'))
```

Visualize top significant proteins identified in both Discovery and Method Development
datasets and based on Discovery one. Note that Method Development data is provided
to have more comprehensive analysis and it may not be on same scale as Discovery data.\
ns: p > 0.05, ∗: p <= 0.05, ∗∗: p <= 0.01, ***: p <= 0.001
```{r}
# Plot top significant features based on 'data1'
topSigFeatTab <- dplyr::mutate(featAssoTabs$tbl_topSigFeats,
                               newGroup = factor(newGroup, levels = c('Yes_Discovery',
                                                                      'No_Discovery',
                                                                      'Yes_MethodDev',
                                                                      'No_MethodDev')))
ggplot(topSigFeatTab, aes(x=newGroup, y=Abundance, col=DataName, fill=Recurrence)) +
  geom_boxplot(alpha = 1, outlier.shape = NA, linewidth = 1) +
  geom_jitter(position = position_jitter(0.2), size = 3, show.legend = F) +
  ggpubr::stat_compare_means(method = 't.test', paired = F, method.args = list(var.equal = T),
                             comparisons = list(c('Yes_Discovery', 'No_Discovery'),
                                                c('Yes_MethodDev', 'No_MethodDev')),
                             label = 'p.signif', tip.length = 0.015, bracket.size = 0.7, size = 4) +
  labs(x = 'Recurrence', title = 'Recurrence: Tumor Tissue DIA Proteomics (AG Klingmüller)') +
  scale_color_manual(name = 'Cohort', values = c('firebrick', 'grey50')) +
  scale_fill_manual(values=c('#00BFC4', '#F8766D')) +
  facet_wrap(vars(newFeature), scales = 'free') +
  th +
  theme(strip.text = element_text(size = 13, face = 'bold'),
        axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1))
```

## Integrative analysis

Show intersections of potential Tumor proteins identified in different datasets
(significance level = 0.05)
```{r}
# Prepare recurrence-related significant features
alpha <- 0.05
# AG Klingmüller
sigFeats_Discovery_Klin <- recurTumorFeatAssoTabs_Klin$tbl_featAsso1 %>%
  dplyr::filter(pVal <= alpha) %>%
  dplyr::pull(Feature)
sigFeats_MethodDev_Klin <- recurTumorFeatAssoTabs_Klin$tbl_featAsso2 %>%
  dplyr::filter(pVal <= alpha) %>%
  dplyr::pull(Feature)
# AG Krijgsveld
sigFeats_Discovery_Krij <- recurTumorFeatAssoTabs_Krij$tbl_featAsso1 %>%
  dplyr::filter(pVal <= alpha) %>%
  dplyr::pull(Feature)
sigFeats_MethodDev_Krij <- recurTumorFeatAssoTabs_Krij$tbl_featAsso2 %>%
  dplyr::filter(pVal <= alpha) %>%
  dplyr::pull(Feature)

# Make venn graphs
ggvenn(list(Discovery_Krij. = sigFeats_Discovery_Krij, MethodDev_Krij. = sigFeats_MethodDev_Krij,
            MethodDev_Klin. = sigFeats_MethodDev_Klin),
       fill_color = c('darkred', 'darkblue', 'darkgreen'), fill_alpha = 0.6,
       set_name_color = c('darkred', 'darkblue', 'darkgreen'), set_name_size = 9,
       text_size = 5, stroke_size = 1.5)
```

```{r}
ggvenn(list(Discovery_Klin. = sigFeats_Discovery_Klin, MethodDev_Klin. = sigFeats_MethodDev_Klin,
            MethodDev_Krij. = sigFeats_MethodDev_Krij),
       fill_color = c('darkorange', 'darkgreen', 'darkblue'), fill_alpha = 0.6,
       set_name_color = c('darkorange', 'darkgreen', 'darkblue'), set_name_size = 9,
       text_size = 5, stroke_size = 1.5)
```

```{r}
ggvenn(list(Discovery_Krij. = sigFeats_Discovery_Krij, Discovery_Klin. = sigFeats_Discovery_Klin,
            MethodDev_Klin. = sigFeats_MethodDev_Klin),
       fill_color = c('darkred', 'darkorange', 'darkgreen'), fill_alpha = 0.6,
       set_name_color = c('darkred', 'darkorange', 'darkgreen'), set_name_size = 9,
       text_size = 5, stroke_size = 1.5)
```
