---
title: 'Cross-cohort analysis'
author: "Qian-Wu Liao"
output:
  workflowr::wflow_html:
    toc: yes
    toc_float: yes
    code_folding: hide
---

<font size='4'> Description: We currently got Method Development and Discovery Tissue
Proteomics from AG Krijgsveld and AG Klingmüller. This script is to investigate
data quality and result reproducibility between different cohorts and different
working groups. Comparisons between cohorts can exclude differences in working
groups, but different patient cohorts may introduce a lot more noise, and comparisons
between working groups can exclude huge variation in patients. Results imply that
biological variation (individual patients and tumor heterogeneity) is larger than
technical differences. </font>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 8, fig.width = 10, warning = F)
knitr::opts_knit$set(root.dir = '/Users/qianwu/Desktop/SMART-CARE_LungCancer')
```

Load libraries
```{r load libraries, message=F}
library(ggrepel)
library(ggvenn)
library('AnnotationDbi')
library('org.Hs.eg.db')
# This annotation object is accessed using AnnotationDbi:select()
hs <- org.Hs.eg.db
library('msigdbr')
library('clusterProfiler')
library('enrichplot')
library(SummarizedExperiment)
library(tidyverse)
# Load user-defined functions
source('./code/misc.R')

# Set plot theme
th <- theme_bw(base_size = 15) +
  theme(axis.title = element_text(face = 'bold'),
        axis.text = element_text(face = 'bold'),
        axis.ticks = element_line(linewidth = 0.8),
        legend.text = element_text(size = 15))
```

```{r load user-defined functions}
doFeatAssoTest <- function(data, meta_var, subset = NULL, prot_gene_tbl = NULL,
                           method = 'proDA', unwantVar = NULL) {
  #' Conduct univariate association test (t-test for now) between each feature and
  #' sample metadata variable of interest
  #' 
  #' Parameters
  #' data: A SummarizedExperiment object that contains the data matrix and metadata
  #' meta_var: A character specifying the sample metadata variable that has two
  #' levels for conducting t-test
  #' subset: A vector of length 2 containing characters specifying the sample metadata
  #' variable and one of its level to split the dataset. Default is NULL, meaning
  #' the whole dataset is used
  #' prot_gene_tbl: A table containing proteins and their encoding genes for providing
  #' gene names in a resulting table. Default is NULL. The column for proteins should
  #' be named 'Proteins' and for genes should be named 'Genes'. Furthermore, only
  #' the first protein and gene should be kept if there are more than one
  #' method: A character specifying the method used to conduct t-test, which should
  #' be one of 'proDA' (default) and 'limma'
  #' unwantVar: A character or a vector of characters specifying the metadata variables
  #' that will be accounted for by linear models (regressing out). Default is NULL
  #' 
  #' Return
  #' assoTestRes: A table of feature significance test and statistical results
  
  # Check arguments
  # meta_var should be length of 1 and in sample metadata
  # subset should be length of 2 and in sample metadata
  # method should be one of 'proDA' and limma''
  
  # Subset samples if specified
  if (!is.null(subset)) {
    subsetIdx <- which(colData(data)[[subset[1]]] == subset[2])
    data <- data[, subsetIdx]
  }
  datMat <- SummarizedExperiment::assay(data)
  smpAnno <- tibble::as_tibble(colData(data), rownames = 'Sample')
  
  # Perform analysis
  # Prepare model design matrix
  if (is.null(unwantVar)) {
      design <- model.matrix(~ smpAnno[[meta_var]])
    } else {
      formu <- paste0('~', paste0(c(meta_var, unwantVar), collapse = '+')) %>%
        as.formula()
      design <- model.matrix(formu, data = smpAnno)
    }
  if (method == 'proDA') {
    # Fit linear probabilistic dropout model to normalized data
    fit <- proDA(datMat, design = design)
    diffTerm <- proDA::result_names(fit)[2]
    assoTestRes <- proDA::test_diff(fit, contrast = diffTerm, sort_by = 'pval') %>%
      dplyr::select(name, pval, adj_pval, diff) %>%
      dplyr::rename(Feature = name, pVal = pval, pValAdj = adj_pval, logFC = diff)
  } else if (method == 'limma') {
    fit <- limma::lmFit(datMat, design = design)
    fit <- limma::eBayes(fit)
    assoTestRes <- limma::topTable(fit, coef = 2, number = Inf) %>%
      tibble::as_tibble(rownames = 'Feature') %>%
      dplyr::select(Feature, P.Value, adj.P.Val, logFC) %>%
      dplyr::rename(pVal = P.Value, pValAdj = adj.P.Val)
  }
  assoTestRes <- dplyr::mutate(assoTestRes,
                               pVal = as.numeric(formatC(pVal, format = 'g', digits = 3)),
                               pValAdj = as.numeric(formatC(pValAdj, format = 'g', digits = 3)))
  # Include feature 
  if (!is.null(prot_gene_tbl)) {
    assoTestRes <- dplyr::mutate(assoTestRes, Gene = plyr::mapvalues(Feature,
                                                                     from = prot_gene_tbl$Proteins,
                                                                     to = prot_gene_tbl$Genes,
                                                                     warn_missing = F)) %>%
      dplyr::relocate(Gene, .after = Feature)
  }
  
  return(assoTestRes)
}


summarizeFeatAssoRes <- function(data1, data2, name1, name2, meta_var, subset = NULL,
                                 prot_gene_tbl = NULL, method = 'proDA',
                                 unwantVar1 = NULL, unwantVar2 = NULL,
                                 alpha = 0.05, num_topSigFeats = 6, plot_pVal = T,
                                 plot_topSigFeats = T, plot_logFC = T) {
  #' Conduct univariate analysis (t-test for now) between features and sample metadata
  #' variable of interest, and combine and tidy up significance test and statistical
  #' results (p-values, top significant features, and log fold changes) into tables for making plots
  #' 
  #' Parameters
  #' data1, data2: SummarizedExperiment objects that contain the data matrices and metadata
  #' name1, name2: A character indicating the names of datasets (data1 and data2)
  #' in the plot legend
  #' meta_var: A character specifying the sample metadata variable that has two
  #' levels for grouping samples to conduct t-test
  #' subset: A vector of length 2 containing characters specifying the sample metadata
  #' variable and one of its level to split the dataset. Default is NULL, meaning
  #' the whole dataset is used
  #' prot_gene_tbl: A table containing proteins and their encoding genes for providing
  #' and labeling gene names in plots. Default is NULL. The column for proteins
  #' should be named 'Proteins' and for genes should be named 'Genes'. Furthermore,
  #' only the first protein and gene should be kept if there are more than one
  #' method: A character specifying the method used to conduct t-test, which should
  #' be one of 'proDA' (default) and 'limma'
  #' unwantVar1, unwantVar2: A character or a vector of characters specifying the
  #' metadata variables (data1 and data2) that will be accounted for by linear models
  #' (regressing out). Default is NULL
  #' alpha: A numeric value specifying the significance level for summarizing p-value
  #' information. Default is 0.05
  #' num_topSigFeats: A numeric value specifying the number of top significant features
  #' to visualize. Default is 6
  #' plot_pVal: A logical variable indicating whether to generate the summarized
  #' p-value table for plotting. Default is TRUE
  #' plot_topSigFeats: A logical variable indicating whether to generate the summarized
  #' table of top significant features based on 'data1' for plotting. Default is TRUE
  #' plot_logFC: A logical variable indicating whether to generate the summarized
  #' log(FC) table for plotting. Default is TRUE
  #' 
  #' Return
  #' A list containing the following components:
  #' tbl_featAsso1: A table containing association test results of 'data1'
  #' tbl_featAsso2: A table containing association test results of 'data2'
  #' tbl_pVal: A table containing summarized significance test results, labeling
  #' information, etc. for making plots
  #' tbl_topSigFeats: A table containing abundances of top significant features,
  #' grouping information, etc. for making plots
  #' tbl_logFC: A table containing summarized log(FC) results, coloring information,
  #' etc. for making plots
  
  # Extract common features between two datasets
  # Select first protein as representative if there are more than one
  rownames(data1) <- stringr::str_remove(rownames(data1), ';.*')
  rownames(data2) <- stringr::str_remove(rownames(data2), ';.*')
  commonFeats <- intersect(rownames(data1), rownames(data2))
  
  # Compute associations between features and sample metadata variable of interest
  # and keep only results of common features
  assoTestRes1 <- doFeatAssoTest(data1, meta_var = meta_var, subset = subset,
                                 prot_gene_tbl = prot_gene_tbl, method = method,
                                 unwantVar = unwantVar1)
  assoTestRes_dat1 <- dplyr::filter(assoTestRes1, Feature %in% commonFeats)
  assoTestRes2 <- doFeatAssoTest(data2, meta_var = meta_var, subset = subset,
                                 prot_gene_tbl = prot_gene_tbl, method = method,
                                 unwantVar = unwantVar2)
  assoTestRes_dat2 <- dplyr::filter(assoTestRes2, Feature %in% commonFeats)
  
  # Summarize resulting p-Values
  if (plot_pVal) {
    # Prepare feature significance information
    tmp_assoTestRes_dat1 <- dplyr::select(assoTestRes_dat1, Feature, pVal) %>%
      dplyr::mutate(Sig_dat1 = dplyr::case_when(pVal <= alpha ~ 'Yes',
                                                pVal > alpha ~ 'No')) %>%
      dplyr::rename(pVal_dat1 = pVal)
    tmp_assoTestRes_dat2 <- dplyr::select(assoTestRes_dat2, Feature, pVal) %>%
      dplyr::mutate(Sig_dat2 = dplyr::case_when(pVal <= alpha ~ 'Yes',
                                                pVal > alpha ~ 'No')) %>%
      dplyr::rename(pVal_dat2 = pVal)
    # Combine association test results and all needed information into a table
    combinedAssoTestRes_pVal <- dplyr::left_join(tmp_assoTestRes_dat1, tmp_assoTestRes_dat2,
                                                 by = 'Feature') %>%
      # Include coloring information
      dplyr::mutate(Significance = dplyr::case_when(
        Sig_dat1 == 'Yes' & Sig_dat2 == 'Yes' ~ 'Sig. in Both',
        Sig_dat1 == 'Yes' & Sig_dat2 == 'No' ~ paste('Sig. in', name1),
        Sig_dat1 == 'No' & Sig_dat2 == 'Yes' ~ paste('Sig. in', name2),
        Sig_dat1 == 'No' & Sig_dat2 == 'No' ~ 'Not Sig.'
      ),
      Significance = factor(Significance, levels = c('Sig. in Both',
                                                     paste('Sig. in', name1),
                                                     paste('Sig. in', name2),
                                                     'Not Sig.')))
    # Include labeling information
    if (is.null(prot_gene_tbl)) {
      combinedAssoTestRes_pVal <- dplyr::mutate(combinedAssoTestRes_pVal,
                                                Label = dplyr::case_when(
                                                  Significance == 'Sig. in Both' ~ Feature
                                                ))
    } else {
      combinedAssoTestRes_pVal <- dplyr::mutate(combinedAssoTestRes_pVal,
                                                Gene = plyr::mapvalues(
                                                  Feature,
                                                  from = prot_gene_tbl$Proteins,
                                                  to = prot_gene_tbl$Genes,
                                                  warn_missing = F
                                                ),
                                                Label = dplyr::case_when(
                                                  Significance == 'Sig. in Both' ~ Gene
                                                ))
    }
    # Reorder points in graph
    significance <- combinedAssoTestRes_pVal$Significance
    combinedAssoTestRes_pVal <- combinedAssoTestRes_pVal[c(which(significance == 'Not Sig.'),
                                                           which(significance == 'Sig. in Both'),
                                                           which(significance == paste('Sig. in', name1)),
                                                           which(significance == paste('Sig. in', name2))),]
  } else {
    combinedAssoTestRes_pVal <- NULL
  }
  
  # Summarize top significant features
  # Check if there is any common significant feature
    sigFeats1 <- dplyr::filter(assoTestRes_dat1, pVal <= alpha) %>%
      dplyr::arrange() %>%
      dplyr::pull(Feature)
    sigFeats2 <- dplyr::filter(assoTestRes_dat2, pVal <= alpha) %>%
      dplyr::arrange() %>%
      dplyr::pull(Feature)
    sigFeats <- intersect(sigFeats1, sigFeats2)
  if (plot_topSigFeats & (length(sigFeats) != 0)) {
    # Retrieve top common significant features based on 'data1'
    topSigFeats <- intersect(sigFeats1, sigFeats2)
    if (length(topSigFeats) > num_topSigFeats) {
      topSigFeats <- topSigFeats[1:num_topSigFeats]
    }
    # Extract abundances of top common significant features and combine all needed
    # information into a table
    # Subset samples if specified
    if (!is.null(subset)) {
      data1 <- data1[, which(colData(data1)[[subset[1]]] == subset[2])]
      data2 <- data2[, which(colData(data2)[[subset[1]]] == subset[2])]
    }
    # Prepare sample grouping information
    smpGpInfo1 <- tibble::as_tibble(colData(data1), rownames = 'Sample') %>%
      dplyr::select(Sample, all_of(meta_var))
    smpGpInfo2 <- tibble::as_tibble(colData(data2), rownames = 'Sample') %>%
      dplyr::select(Sample, all_of(meta_var))
    # Prepare abundance information for top common significant features
    topSigFeatTab1 <- t(SummarizedExperiment::assay(data1)) %>%
      tibble::as_tibble(rownames = 'Sample') %>%
      dplyr::select(c(Sample, topSigFeats)) %>%
      tidyr::pivot_longer(cols = -'Sample', names_to = 'Feature', values_to = 'Abundance') %>%
      dplyr::left_join(smpGpInfo1, by = 'Sample') %>%
      dplyr::mutate(DataName = name1)
    topSigFeatTab2 <- t(SummarizedExperiment::assay(data2)) %>%
      tibble::as_tibble(rownames = 'Sample') %>%
      dplyr::select(c(Sample, topSigFeats)) %>%
      tidyr::pivot_longer(cols = -'Sample', names_to = 'Feature', values_to = 'Abundance') %>%
      dplyr::left_join(smpGpInfo2, by = 'Sample') %>%
      dplyr::mutate(DataName = name2)
    # Combine all information
    combinedTopSigFeatTab <- dplyr::bind_rows(topSigFeatTab1, topSigFeatTab2) %>%
      # Include new grouping information
      dplyr::mutate(newGroup = paste0(.data[[meta_var]], '_', DataName))
    # Create new feature names if feature annotations are provided
    if (!is.null(prot_gene_tbl)) {
      combinedTopSigFeatTab <- dplyr::mutate(combinedTopSigFeatTab,
                                             Gene = plyr::mapvalues(Feature,
                                                                    from = prot_gene_tbl$Proteins,
                                                                    to = prot_gene_tbl$Genes,
                                                                    warn_missing = F),
                                             newFeature = paste0(Feature, ' (', Gene, ')'),
                                             newFeature = factor(newFeature,
                                                                 levels = unique(newFeature)))
    }
  } else {
    combinedTopSigFeatTab <- NULL
  }
  
  # Summarize resulting log(FC)
  if (plot_logFC) {
    # Prepare feature log(FC) information
    tmp_assoTestRes_dat1 <- dplyr::select(assoTestRes_dat1, Feature, logFC) %>%
      dplyr::mutate(Dir_dat1 = dplyr::case_when(logFC > 0 ~ 'Up',
                                                logFC < 0 ~ 'Down')) %>%
      dplyr::rename(logFC_dat1 = logFC)
    tmp_assoTestRes_dat2 <- dplyr::select(assoTestRes_dat2, Feature, logFC) %>%
      dplyr::mutate(Dir_dat2 = dplyr::case_when(logFC > 0 ~ 'Up',
                                                logFC < 0 ~ 'Down')) %>%
      dplyr::rename(logFC_dat2 = logFC)
    # Combine association test results and all needed information into a table
    combinedAssoTestRes_logFC <- dplyr::left_join(tmp_assoTestRes_dat1, tmp_assoTestRes_dat2,
                                                  by = 'Feature') %>%
      # Include coloring information
      dplyr::mutate(Direction = dplyr::case_when(
        Dir_dat1 == 'Up' & Dir_dat2 == 'Up' ~ 'Up in Both',
        Dir_dat1 == 'Down' & Dir_dat2 == 'Down' ~ 'Down in Both',
        Dir_dat1 == 'Up' & Dir_dat2 == 'Down' ~ 'Inconsistent',
        Dir_dat1 == 'Down' & Dir_dat2 == 'Up' ~ 'Inconsistent'
      ),
      Direction = factor(Direction, levels = c('Up in Both', 'Down in Both', 'Inconsistent')))
    # Include gene annotations
    if (!is.null(prot_gene_tbl)) {
      combinedAssoTestRes_logFC <- dplyr::mutate(combinedAssoTestRes_logFC,
                                                 Gene = plyr::mapvalues(Feature,
                                                                        from = prot_gene_tbl$Proteins,
                                                                        to = prot_gene_tbl$Genes,
                                                                        warn_missing = F))
    }
    # Reorder points in graph
    direction <- combinedAssoTestRes_logFC$Direction
    combinedAssoTestRes_logFC <- combinedAssoTestRes_logFC[c(which(direction == 'Inconsistent'),
                                                             which(direction == 'Down in Both'),
                                                             which(direction == 'Up in Both')),]
  } else {
    combinedAssoTestRes_logFC <- NULL
  }
  
  return(list(tbl_featAsso1 = assoTestRes1, tbl_featAsso2 = assoTestRes2,
              tbl_pVal = combinedAssoTestRes_pVal, tbl_topSigFeats = combinedTopSigFeatTab,
              tbl_logFC = combinedAssoTestRes_logFC))
}


calcFeatCorr <- function(data1, data2, show_nFeats = F) {
  #' Systematically compute feature correlations between two PROTEOMICS (for now) datasets
  #' 
  #' Parameters
  #' data1, data2: SummarizedExperiment objects that contain data matrices
  #' 
  #' Return
  #' corrRes: A table of correlations of overlapped features between two datasets
  
  # Extract data matrix
  dat1 <- SummarizedExperiment::assay(data1)
  dat2 <- SummarizedExperiment::assay(data2)
  
  # Make two matrices share same sample and feature spaces and in same row and column order
  # Keep only common samples
  commonSmps <- intersect(colnames(dat1), colnames(dat2))
  dat1 <- dat1[, commonSmps]
  dat2 <- dat2[, commonSmps]
  # Keep only common features
  # Select first protein as representative if there are more than one
  rownames(dat1) <- stringr::str_remove(rownames(dat1), ';.*')
  rownames(dat2) <- stringr::str_remove(rownames(dat2), ';.*')
  commonFeats <- intersect(rownames(dat1), rownames(dat2))
  dat1 <- dat1[commonFeats,]
  dat2 <- dat2[commonFeats,]
  if (show_nFeats) {
    cat('Features in data1: ', nrow(data1), '\n', 'Features in data2: ', nrow(data2), '\n',
        'Features in common: ', length(commonFeats), sep = '')
  }
  
  # Compute feature correlations between two proteomics datasets
  # Do sanity check
  if (identical(rownames(dat1), rownames(dat2)) &
      identical(colnames(dat1), colnames(dat2))) {
    corrs <- sapply(seq_len(nrow(dat1)), function(i) {
      corr <- cor.test(dat1[i,], dat2[i,], method = 'pearson', use = "pairwise.complete.obs")
      corr$estimate
    })
    corrRes <- data.frame(Feature = rownames(dat1), Correlation = corrs) %>%
      dplyr::arrange(Correlation)
  }
  
  return(corrRes)
}
```

# Tissue DIA Proteomics
Metadata-assisted quality assessment: Check association of tissue type-related feature
significance between Discovery and Method Development datasets or between Krij.'s
and Klin.'s datasets

## AG Krijgsveld

Display association of tissue type-related feature significance between Discovery
and Method Development DIA Proteomics generated by AG Krijgsveld\
**p-values (significance level = 0.05)**
```{r}
# Load normalized data
proTissue_Discovery <- readRDS('./data/Discovery/AG_Krijgsveld/proTissueVsn.rds')
proTissue_MethodDev <- readRDS('./data/MethodDev/AG_Krijgsveld/new_proTissueVsn.rds')
# Prepare protein-gene table for providing gene names
featInfo <- tibble::as_tibble(rowData(proTissue_Discovery), rownames = 'Proteins') %>%
  dplyr::select(Proteins, Genes) %>%
  dplyr::mutate(Proteins = stringr::str_remove(Proteins, ';.*'),
                Genes = stringr::str_remove(Genes, ';.*'))

# Prepare summarized tables of comparison between Tumor and Normal tissue samples
condiFeatAssoTabs_Krij <- summarizeFeatAssoRes(proTissue_Discovery, proTissue_MethodDev,
                                               'Discovery', 'MethodDev', meta_var = 'Condition',
                                               subset = NULL, method = 'proDA', alpha = 0.05,
                                               prot_gene_tbl = featInfo)
featAssoTabs <- condiFeatAssoTabs_Krij

# Plot p-values
ggplot(featAssoTabs$tbl_pVal, aes(x=-log10(pVal_dat1), y=-log10(pVal_dat2), col=Significance)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.8) +
  scale_color_manual(values = c('red', 'darkgreen', 'navy', 'grey')) +
  labs(x = '-log10(pVal) of Discovery Cohort', y = '-log10(pVal) of MethodDev Cohort',
       title = 'Tumor vs Normal: Tissue DIA Proteomics (AG Krijgsveld)') +
  th #+ theme(axis.title = element_text(size = 22),
             # axis.text = element_text(size = 16),
             # legend.title = element_text(size = 20), legend.text = element_text(size = 18))
# ggsave('./output/SC_meeting/pVals_Dis_vs_MetDev_proTissue_condi_Krij.png', device = 'png', dpi = 400, height = 8, width = 11)

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_pVal$Significance == 'Sig. in Both') /
                   sum(featAssoTabs$tbl_pVal$Significance != 'Not Sig.'), 3) * 100
cat(paste0('=> ', percent, '% of significant results are consistent'))
```
=> Trend is towards Discovery data, which means same features got lower p-values
in Discovery data compared to Method Development one. Yet, it probably does not
mean Discovery data is better, but just got larger sample size.

**log(FC)**
```{r}
# Plot log(FC)
ggplot(featAssoTabs$tbl_logFC, aes(x=logFC_dat1, y=logFC_dat2, col=Direction)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.8) +
  scale_color_manual(values = c('red', 'navy', 'grey')) +
  labs(x = 'log(FC) of Discovery Cohort', y = 'log(FC) of MethodDev Cohort',
       title = 'Tumor vs Normal: Tissue DIA Proteomics (AG Krijgsveld)') +
  th #+ theme(axis.title = element_text(size = 22),
             # axis.text = element_text(size = 16),
             # legend.title = element_text(size = 20), legend.text = element_text(size = 18))
# ggsave('./output/SC_meeting/log(FC)_Dis_vs_MetDev_proTissue_condi_Krij.png', device = 'png', dpi = 400, height = 8, width = 11)

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_logFC$Direction != 'Inconsistent') /
                   nrow(featAssoTabs$tbl_logFC), 3) * 100
cat(paste0('=> ', percent, '% of results are consistent'))
```
=> Factor of sample size is excluded.

## AG Klingmüller

Display association of tissue type-related feature significance between Discovery
and Method Development DIA Proteomics generated by AG Klingmüller\
**p-values (significance level = 0.05)**
```{r}
# Load normalized data
proTissue_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsn.rds')
proTissue_Discovery <- proTissue_Klin[, colData(proTissue_Klin)$Cohort == 'Discovery']
proTissue_MethodDev <- proTissue_Klin[, colData(proTissue_Klin)$Cohort == 'MethodDev']
# Prepare protein-gene table for providing gene names
featInfo <- tibble::as_tibble(rowData(proTissue_Discovery), rownames = 'Proteins') %>%
  dplyr::rename(Genes = PG.Genes) %>%
  dplyr::mutate(Proteins = stringr::str_remove(Proteins, ';.*'),
                Genes = stringr::str_remove(Genes, ';.*'))

# Prepare summarized tables of comparison between Tumor and Normal tissue samples
condiFeatAssoTabs_Klin <- summarizeFeatAssoRes(proTissue_Discovery, proTissue_MethodDev,
                                               'Discovery', 'MethodDev', meta_var = 'Condition',
                                               subset = NULL, method = 'proDA', alpha = 0.05,
                                               prot_gene_tbl = featInfo)
featAssoTabs <- condiFeatAssoTabs_Klin

# Plot p-values
ggplot(featAssoTabs$tbl_pVal, aes(x=-log10(pVal_dat1), y=-log10(pVal_dat2), col=Significance)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.8) +
  scale_color_manual(values = c('red', 'darkgreen', 'navy', 'grey')) +
  labs(x = '-log10(pVal) of Discovery Cohort', y = '-log10(pVal) of MethodDev Cohort',
       title = 'Tumor vs Normal: Tissue DIA Proteomics (AG Klingmüller)') +
  th #+ theme(axis.title = element_text(size = 22),
             # axis.text = element_text(size = 16),
             # legend.title = element_text(size = 20), legend.text = element_text(size = 18))
# ggsave('./output/SC_meeting/pVals_Dis_vs_MetDev_proTissue_condi_Klin.png', device = 'png', dpi = 400, height = 8, width = 11)

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_pVal$Significance == 'Sig. in Both') /
                   sum(featAssoTabs$tbl_pVal$Significance != 'Not Sig.'), 3) * 100
cat(paste0('=> ', percent, '% of significant results are consistent'))
```

**log(FC)**
```{r}
# Plot log(FC)
ggplot(featAssoTabs$tbl_logFC, aes(x=logFC_dat1, y=logFC_dat2, col=Direction)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.8) +
  scale_color_manual(values = c('red', 'navy', 'grey')) +
  labs(x = 'log(FC) of Discovery Cohort', y = 'log(FC) of MethodDev Cohort',
       title = 'Tumor vs Normal: Tissue DIA Proteomics (AG Klingmüller)') +
  th #+ theme(axis.title = element_text(size = 22),
#              axis.text = element_text(size = 16),
#              legend.title = element_text(size = 20), legend.text = element_text(size = 18))
# ggsave('./output/SC_meeting/log(FC)_Dis_vs_MetDev_proTissue_condi_Klin.png', device = 'png', dpi = 400, height = 8, width = 11)

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_logFC$Direction != 'Inconsistent') /
                   nrow(featAssoTabs$tbl_logFC), 3) * 100
cat(paste0('=> ', percent, '% of results are consistent'))
```

## Method Development
Compare data from same patient cohorts (only common samples used) between different
working groups to see if better results can be obtained due to large variation among patients

**p-values (significance level = 0.05)**
```{r}
# Load normalized data
proTissue_Krij <- readRDS('./data/MethodDev/AG_Krijgsveld/new_proTissueVsn.rds')
proTissue_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsn.rds')
proTissue_Klin <- proTissue_Klin[, colData(proTissue_Klin)$Cohort == 'MethodDev']
# Keep only common samples shared between Discovery and Method Development data
cmnSmps <- intersect(colnames(proTissue_Krij), colnames(proTissue_Klin))
proTissue_Krij <- proTissue_Krij[, cmnSmps]
proTissue_Klin <- proTissue_Klin[, cmnSmps]
# Prepare protein-gene table for providing gene names
featInfo <- tibble::as_tibble(rowData(proTissue_Klin), rownames = 'Proteins') %>%
  dplyr::rename(Genes = PG.Genes) %>%
  dplyr::mutate(Proteins = stringr::str_remove(Proteins, ';.*'),
                Genes = stringr::str_remove(Genes, ';.*'))

# Prepare summarized tables of comparison between Tumor and Normal tissue samples
condiFeatAssoTabs_MethodDev <- summarizeFeatAssoRes(proTissue_Krij, proTissue_Klin,
                                                    'Krij.', 'Klin.', meta_var = 'Condition',
                                                    subset = NULL, method = 'proDA', alpha = 0.05,
                                                    prot_gene_tbl = featInfo)
featAssoTabs <- condiFeatAssoTabs_MethodDev

# Plot p-values
ggplot(featAssoTabs$tbl_pVal, aes(x=-log10(pVal_dat1), y=-log10(pVal_dat2), col=Significance)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.8) +
  scale_color_manual(values = c('red', 'darkgreen', 'navy', 'grey')) +
  labs(x = '-log10(pVal) of Krij.', y = '-log10(pVal) of Klin.',
       title = 'Tumor vs Normal: Tissue DIA Proteomics (MethodDev)') +
  th
# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_pVal$Significance == 'Sig. in Both') /
                   sum(featAssoTabs$tbl_pVal$Significance != 'Not Sig.'), 3) * 100
cat(paste0('=> ', percent, '% of significant results are consistent'))
```

**log(FC)**
```{r}
# Plot log(FC)
ggplot(featAssoTabs$tbl_logFC, aes(x=logFC_dat1, y=logFC_dat2, col=Direction)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.8) +
  scale_color_manual(values = c('red', 'navy', 'grey')) +
  labs(x = 'log(FC) of Krij.', y = 'log(FC) of Klin.',
       title = 'Tumor vs Normal: Tissue DIA Proteomics (MethodDev)') +
  th
# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_logFC$Direction != 'Inconsistent') /
                   nrow(featAssoTabs$tbl_logFC), 3) * 100
cat(paste0('=> ', percent, '% of results are consistent'))
```

**Feature correlations**
```{r}
# Compute correlations of common features between two datasets
corrRes <- calcFeatCorr(proTissue_Krij, proTissue_Klin, show_nFeats = F)
ggplot(corrRes, aes(x=Correlation)) +
  geom_histogram(binwidth = 0.1) +
  scale_x_continuous(breaks = c(seq(-0.5, 1, 0.25))) +
  labs(y = 'Frequency', title = 'Feature correlations: Tissue DIA Proteomics (MethodDev)') +
  th
```

## Discovery
Compare data from same patient cohorts (only common samples used) between different
working groups to see if better results can be obtained due to large variation among patients

**p-values (significance level = 0.05)**
```{r}
# Load normalized data
proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/proTissueVsn.rds')
proTissue_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsn.rds')
proTissue_Klin <- proTissue_Klin[, colData(proTissue_Klin)$Cohort == 'Discovery']
# Keep only common samples shared between Discovery and Method Development data
cmnSmps <- intersect(colnames(proTissue_Krij), colnames(proTissue_Klin))
proTissue_Krij <- proTissue_Krij[, cmnSmps]
proTissue_Klin <- proTissue_Klin[, cmnSmps]
# Prepare protein-gene table for providing gene names
featInfo <- tibble::as_tibble(rowData(proTissue_Klin), rownames = 'Proteins') %>%
  dplyr::rename(Genes = PG.Genes) %>%
  dplyr::mutate(Proteins = stringr::str_remove(Proteins, ';.*'),
                Genes = stringr::str_remove(Genes, ';.*'))

# Prepare summarized tables of comparison between Tumor and Normal tissue samples
condiFeatAssoTabs_Discovery <- summarizeFeatAssoRes(proTissue_Krij, proTissue_Klin,
                                                    'Krij.', 'Klin.', meta_var = 'Condition',
                                                    subset = NULL, method = 'proDA', alpha = 0.05,
                                                    prot_gene_tbl = featInfo)
featAssoTabs <- condiFeatAssoTabs_Discovery

# Plot p-values
ggplot(featAssoTabs$tbl_pVal, aes(x=-log10(pVal_dat1), y=-log10(pVal_dat2), col=Significance)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.8) +
  scale_color_manual(values = c('red', 'darkgreen', 'navy', 'grey')) +
  labs(x = '-log10(pVal) of Krij.', y = '-log10(pVal) of Klin.',
       title = 'Tumor vs Normal: Tissue DIA Proteomics (Discovery)') +
  th
# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_pVal$Significance == 'Sig. in Both') /
                   sum(featAssoTabs$tbl_pVal$Significance != 'Not Sig.'), 3) * 100
cat(paste0('=> ', percent, '% of significant results are consistent'))
```

**log(FC)**
```{r}
# Plot log(FC)
ggplot(featAssoTabs$tbl_logFC, aes(x=logFC_dat1, y=logFC_dat2, col=Direction)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.8) +
  scale_color_manual(values = c('red', 'navy', 'grey')) +
  labs(x = 'log(FC) of Krij.', y = 'log(FC) of Klin.',
       title = 'Tumor vs Normal: Tissue DIA Proteomics (Discovery)') +
  th
# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_logFC$Direction != 'Inconsistent') /
                   nrow(featAssoTabs$tbl_logFC), 3) * 100
cat(paste0('=> ', percent, '% of results are consistent'))
```

**Feature correlations**
```{r}
# Compute correlations of common features between two datasets
corrRes <- calcFeatCorr(proTissue_Krij, proTissue_Klin, show_nFeats = F)
ggplot(corrRes, aes(x=Correlation)) +
  geom_histogram(binwidth = 0.1) +
  scale_x_continuous(breaks = c(seq(-0.5, 1, 0.25))) +
  labs(y = 'Frequency', title = 'Feature correlations: Tissue DIA Proteomics (Discovery)') +
  th
```

## Combined data
Combined Discovery and Method Development data (only common samples used)

**p-values (significance level = 0.05)**
```{r}
# Load normalized data
proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds')
proTissue_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsnBC.rds')
# Keep only common samples shared between Discovery and Method Development data
cmnSmps <- intersect(colnames(proTissue_Krij), colnames(proTissue_Klin))
proTissue_Krij <- proTissue_Krij[, cmnSmps]
proTissue_Klin <- proTissue_Klin[, cmnSmps]
# Prepare protein-gene table for providing gene names
featInfo <- tibble::as_tibble(rowData(proTissue_Klin), rownames = 'Proteins') %>%
  dplyr::rename(Genes = PG.Genes) %>%
  dplyr::mutate(Proteins = stringr::str_remove(Proteins, ';.*'),
                Genes = stringr::str_remove(Genes, ';.*'))

# Prepare summarized tables of comparison between Tumor and Normal tissue samples
condiFeatAssoTabs_Combined <- summarizeFeatAssoRes(proTissue_Krij, proTissue_Klin,
                                                   'Krij.', 'Klin.', meta_var = 'Condition',
                                                   subset = NULL, method = 'proDA',
                                                   unwantVar1 = 'Cohort', unwantVar2 = 'Cohort',
                                                   alpha = 0.05, prot_gene_tbl = featInfo)
featAssoTabs <- condiFeatAssoTabs_Combined

# Plot p-values
ggplot(featAssoTabs$tbl_pVal, aes(x=-log10(pVal_dat1), y=-log10(pVal_dat2), col=Significance)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.8) +
  scale_color_manual(values = c('red', 'darkgreen', 'navy', 'grey')) +
  labs(x = '-log10(pVal) of Krij.', y = '-log10(pVal) of Klin.',
       title = 'Tumor vs Normal: Tissue DIA Proteomics (Combined data)') +
  th

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_pVal$Significance == 'Sig. in Both') /
                   sum(featAssoTabs$tbl_pVal$Significance != 'Not Sig.'), 3) * 100
cat(paste0('=> ', percent, '% of significant results are consistent'))
```

**log(FC)**
```{r}
# Plot log(FC)
ggplot(featAssoTabs$tbl_logFC, aes(x=logFC_dat1, y=logFC_dat2, col=Direction)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.8) +
  scale_color_manual(values = c('red', 'navy', 'grey')) +
  labs(x = 'log(FC) of Krij.', y = 'log(FC) of Klin.',
       title = 'Tumor vs Normal: Tissue DIA Proteomics (Combined data)') +
  th

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_logFC$Direction != 'Inconsistent') /
                   nrow(featAssoTabs$tbl_logFC), 3) * 100
cat(paste0('=> ', percent, '% of results are consistent'))
```

**Feature correlations**
```{r}
# Compute correlations of common features between two datasets
cat('Feature sizes: (data1 is Krij. and data 2 is Klin.)\n')
corrRes <- calcFeatCorr(proTissue_Krij, proTissue_Klin, show_nFeats = T)
ggplot(corrRes, aes(x=Correlation)) +
  geom_histogram(binwidth = 0.1) +
  scale_x_continuous(breaks = c(seq(-0.5, 1, 0.25))) +
  labs(y = 'Frequency', title = 'Feature correlations: Tissue DIA Proteomics (Combined)') +
  th
```

```{r eval=F}
# Check whether highly correlated features are highly abundant proteins
# Compute feature means
meanFeats_Krij <- apply(assay(combined_proTissue_Krij), 1, function(featVec) {
  mean(featVec, na.rm = T)
})
meanFeats_Klin <- apply(assay(combined_proTissue_Klin), 1, function(featVec) {
  mean(featVec, na.rm = T)
})
# Keep only first protein as representative if there are more than one
names(meanFeats_Krij) <- stringr::str_remove(names(meanFeats_Krij), ';.+')
names(meanFeats_Klin) <- stringr::str_remove(names(meanFeats_Klin), ';.+')
# Extract common features
cmnFeats <- corrRes$Feature
meanFeats_Krij <- meanFeats_Krij[names(meanFeats_Krij) %in% cmnFeats]
meanFeats_Klin <- meanFeats_Klin[names(meanFeats_Klin) %in% cmnFeats]
# Order features based on correlation results
meanFeats_Krij <- meanFeats_Krij[cmnFeats]
meanFeats_Klin <- meanFeats_Klin[cmnFeats]
# Integrate feature means into correlation results
corrMeanFeats <- corrRes
corrMeanFeats['Mean (Krij.)'] <- meanFeats_Krij
corrMeanFeats['Mean (Klin.)'] <- meanFeats_Klin
ggplot(corrMeanFeats, aes(x=Correlation, y=`Mean (Krij.)`)) +
  geom_point() +
  ggpubr::stat_cor(aes(label=after_stat(r.label)), method = 'pearson', size = 5) +
  th
ggplot(corrMeanFeats, aes(x=Correlation, y=`Mean (Klin.)`)) +
  geom_point() +
  ggpubr::stat_cor(aes(label=after_stat(r.label)), method = 'pearson', size = 5) +
  th
# => Not really....
```

### Enrichment analysis

**QC: t-values of unique or common features in Klin. and Krij. data between Tumor and Normal samples**\
(1) Unique features captured in Klin. data\
(2) Unique features captured in Krij. data\
(3) Common features captured in both Klin. and Krij. data\
=> Use GO:CC, GO:BP, and MSigBD:C8 as reference gene sets to check data uniqueness and similarity

<font size='4'> **Unique features captured in Klin. or Krij. data** </font>
```{r}
# Load preprocessed data
combined_proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds')
combined_proTissue_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsnBC.rds')
# Keep only first protein as representative if there are more than one
rownames(combined_proTissue_Krij) <- stringr::str_remove(rownames(combined_proTissue_Krij), ';.+')
rownames(combined_proTissue_Klin) <- stringr::str_remove(rownames(combined_proTissue_Klin), ';.+')
# Subset data, where features are captured in either Klin. or Krij. data or in both datasets
feats_Krij <- rownames(combined_proTissue_Krij)
feats_Klin <- rownames(combined_proTissue_Klin)
combined_proTissue_uniq_Klin <- combined_proTissue_Klin[!feats_Klin %in% feats_Krij,]
combined_proTissue_uniq_Krij <- combined_proTissue_Krij[!feats_Krij %in% feats_Klin,]
combined_proTissue_cmn_Klin <- combined_proTissue_Klin[feats_Klin %in% feats_Krij]
combined_proTissue_cmn_Krij <- combined_proTissue_Krij[feats_Krij %in% feats_Klin]
#### UNIQUE FEATURES ####
# Do association tests between unique features and sample conditions
sigRes_uniq_Klin <- doSOA(combined_proTissue_uniq_Klin, meta_var = 'Condition', use_proDA = T)
sigRes_uniq_Krij <- doSOA(combined_proTissue_uniq_Krij, meta_var = 'Condition', use_proDA = T)
# Prepare ranked gene list
statRes_Klin <- sigRes_uniq_Klin$featAssoRes %>%
  dplyr::select(Var1, Stat) %>%
  tibble::column_to_rownames('Var1')
featAnno_Klin <- tibble::as_tibble(rowData(combined_proTissue_uniq_Klin), rownames = 'Protein') %>%
  dplyr::rename(Gene = PG.Genes) %>%
  tibble::column_to_rownames('Protein')
rankedGeneList_Klin <- prepRankedFeatList(statRes = statRes_Klin, featAnno = featAnno_Klin, to_gene = T)

statRes_Krij <- sigRes_uniq_Krij$featAssoRes %>%
  dplyr::select(Var1, Stat) %>%
  tibble::column_to_rownames('Var1')
featAnno_Krij <- tibble::as_tibble(rowData(combined_proTissue_uniq_Krij), rownames = 'Protein') %>%
  dplyr::select(Protein, Genes) %>%
  dplyr::rename(Gene = Genes) %>%
  tibble::column_to_rownames('Protein')
rankedGeneList_Krij <- prepRankedFeatList(statRes = statRes_Krij, featAnno = featAnno_Krij, to_gene = T)
```

---- GO:Cellular Component ----
```{r}
# GO:CC
# Retrieve annotated gene sets
msigTabC5CC <- msigdbr(species = 'Homo sapiens', category = 'C5', subcategory = 'GO:CC') %>%
  dplyr::select(gs_name, human_gene_symbol)
# Run GSEA
gseaResC5CC_Klin <- clusterProfiler::GSEA(geneList = rankedGeneList_Klin, TERM2GENE = msigTabC5CC,
                                          minGSSize = 10, maxGSSize = 500,
                                          pvalueCutoff = 0.05, pAdjustMethod = 'BH',
                                          by = 'fgsea', eps = 0)
gseaResC5CC_Krij <- clusterProfiler::GSEA(geneList = rankedGeneList_Krij, TERM2GENE = msigTabC5CC,
                                          minGSSize = 10, maxGSSize = 500,
                                          pvalueCutoff = 0.05, pAdjustMethod = 'BH',
                                          by = 'fgsea', eps = 0)

# Plot enrichment analysis results (enrichment map)
gseaResC5CC <- enrichplot::pairwise_termsim(gseaResC5CC_Klin)
clusterProfiler::emapplot(gseaResC5CC, showCategory = 18, node_label = 'category', repel = T,
                          cex.params = list(category_node = 1.2, category_label = 0.8),
                          highlight.params = list()) +
  labs(title = 'GO:CC - Unique features in Klin. data') +
  theme(title = element_text(face = 'bold', size = 14))
# 77 overrepresented when cutoff for adjusted p-value is 0.05

gseaResC5CC <- enrichplot::pairwise_termsim(gseaResC5CC_Krij)
clusterProfiler::emapplot(gseaResC5CC) +
  labs(title = 'GO:CC - Unique features in Krij. data') +
  theme(title = element_text(face = 'bold', size = 14))
# 4
```

---- GO:Biological Process ----
```{r}
# GO:BP
# Retrieve annotated gene sets
msigTabC5BP <- msigdbr(species = 'Homo sapiens', category = 'C5', subcategory = 'BP') %>%
  dplyr::select(gs_name, human_gene_symbol)
# Run GSEA
gseaResC5BP_Klin <- clusterProfiler::GSEA(geneList = rankedGeneList_Klin, TERM2GENE = msigTabC5BP,
                                          minGSSize = 10, maxGSSize = 500,
                                          pvalueCutoff = 0.05, pAdjustMethod = 'BH',
                                          by = 'fgsea', eps = 0)
gseaResC5BP_Krij <- clusterProfiler::GSEA(geneList = rankedGeneList_Krij, TERM2GENE = msigTabC5BP,
                                          minGSSize = 10, maxGSSize = 500,
                                          pvalueCutoff = 0.05, pAdjustMethod = 'BH',
                                          by = 'fgsea', eps = 0)

gseaRes4Plot <- gseaResC5BP_Klin@result %>%
  dplyr::select(Description, setSize, NES, pvalue, p.adjust, leading_edge) %>%
  dplyr::arrange(p.adjust) %>%
  dplyr::slice_head(n = 20) %>%
  dplyr::mutate(leading_edge = stringr::str_extract(leading_edge, '\\d\\d'),
                leading_edge = as.numeric(leading_edge),
                Recurrence = ifelse(test = NES > 0, yes = 'Yes', no = 'No')) %>%
  dplyr::rename(gene_ratio = leading_edge)
midGrad <- mean(gseaRes4Plot$p.adjust)

ggplot(gseaRes4Plot, aes(x=NES, y=reorder(Description, NES), fill=p.adjust, col=Recurrence)) +
  geom_bar(stat = 'identity', width = 0.7, size = 2) +
  scale_fill_gradient2(low='#5B1C00', high='#FFFCA5', mid = '#E86900', midpoint = midGrad, name = 'Adjusted pVal') +
  scale_color_manual(values = c(Yes = 'red', No = 'darkblue')) +
  labs(x = 'Normalized Enrichment Score', y = 'Pathway',
       title = 'GO:BP - Unique features in Klin. data') +
  theme_minimal() +
  theme(axis.title = element_text(size = 16, face = 'bold'),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 10, face = 'bold'),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        title = element_text(face = 'bold', size = 12))
# 193
```
=> No overrepresented pathways in Krij. data
<br/>
<br/>

<font size='4'> **Common features captured in both Klin. and Krij. data** </font>\
Sample space is also identical!

---- GO:Cellular Component ----\
Overlaps of overrepresented cellular components between datasets
```{r}
#### COMMON FEATURES ####
# Keep only common samples for investigating data similarity at pathway level
smps_Klin <- colnames(combined_proTissue_cmn_Klin)
smps_Krij <- colnames(combined_proTissue_cmn_Krij)
cmnSmps <- intersect(smps_Klin, smps_Krij)
combined_proTissue_cmn_Klin <- combined_proTissue_cmn_Klin[, smps_Klin %in% cmnSmps]
combined_proTissue_cmn_Krij <- combined_proTissue_cmn_Krij[, smps_Krij %in% cmnSmps]
# Do association tests between common features and sample conditions
sigRes_cmn_Klin <- doSOA(combined_proTissue_cmn_Klin, meta_var = 'Condition', use_proDA = T)
sigRes_cmn_Krij <- doSOA(combined_proTissue_cmn_Krij, meta_var = 'Condition', use_proDA = T)
# Prepare ranked gene list
statRes_Klin <- sigRes_cmn_Klin$featAssoRes %>%
  dplyr::select(Var1, Stat) %>%
  tibble::column_to_rownames('Var1')
featAnno_Klin <- tibble::as_tibble(rowData(combined_proTissue_cmn_Klin), rownames = 'Protein') %>%
  dplyr::rename(Gene = PG.Genes) %>%
  tibble::column_to_rownames('Protein')
rankedGeneList_Klin <- prepRankedFeatList(statRes = statRes_Klin, featAnno = featAnno_Klin, to_gene = T)

statRes_Krij <- sigRes_cmn_Krij$featAssoRes %>%
  dplyr::select(Var1, Stat) %>%
  tibble::column_to_rownames('Var1')
featAnno_Krij <- tibble::as_tibble(rowData(combined_proTissue_cmn_Krij), rownames = 'Protein') %>%
  dplyr::select(Protein, Genes) %>%
  dplyr::rename(Gene = Genes) %>%
  tibble::column_to_rownames('Protein')
rankedGeneList_Krij <- prepRankedFeatList(statRes = statRes_Krij, featAnno = featAnno_Krij, to_gene = T)

# GO:CC
# Retrieve annotated gene sets
msigTabC5CC <- msigdbr(species = 'Homo sapiens', category = 'C5', subcategory = 'GO:CC') %>%
  dplyr::select(gs_name, human_gene_symbol)
# Run GSEA
gseaResC5CC_Klin <- clusterProfiler::GSEA(geneList = rankedGeneList_Klin, TERM2GENE = msigTabC5CC,
                                          minGSSize = 10, maxGSSize = 500,
                                          pvalueCutoff = 0.05, pAdjustMethod = 'BH',
                                          by = 'fgsea', eps = 0)
gseaResC5CC_Krij <- clusterProfiler::GSEA(geneList = rankedGeneList_Krij, TERM2GENE = msigTabC5CC,
                                          minGSSize = 10, maxGSSize = 500,
                                          pvalueCutoff = 0.05, pAdjustMethod = 'BH',
                                          by = 'fgsea', eps = 0)
# Visualize overlaps of overrepresented pathways between datasets
ggvenn(list(Combined_Klin. = gseaResC5CC_Klin@result$Description,
            Combined_Krij. = gseaResC5CC_Krij@result$Description),
       fill_color = c('firebrick1', 'dodgerblue1'), fill_alpha = 0.6,
       set_name_color = c('firebrick1', 'dodgerblue1'), set_name_size = 9,
       text_size = 5, stroke_size = 1.5)
```

```{r, eval=F, include=F}
# Plot enrichment analysis results (enrichment map)
gseaResC5CC <- enrichplot::pairwise_termsim(gseaResC5CC_Klin)
clusterProfiler::emapplot(gseaResC5CC)

gseaResC5CC <- enrichplot::pairwise_termsim(gseaResC5CC_Krij)
clusterProfiler::emapplot(gseaResC5CC)
```

---- MSigDB:Cell Type ----\
Overlaps of overrepresented cell types between datasets
```{r}
# Cell type gene sets
# Retrieve annotated gene sets
msigTabC8 <- msigdbr(species = 'Homo sapiens', category = 'C8', subcategory = NULL) %>%
  dplyr::select(gs_name, human_gene_symbol)
# Run GSEA
gseaResC8_Klin <- clusterProfiler::GSEA(geneList = rankedGeneList_Klin, TERM2GENE = msigTabC8,
                                        minGSSize = 10, maxGSSize = 500,
                                        pvalueCutoff = 0.05, pAdjustMethod = 'BH',
                                        by = 'fgsea', eps = 0)
gseaResC8_Krij <- clusterProfiler::GSEA(geneList = rankedGeneList_Krij, TERM2GENE = msigTabC8,
                                        minGSSize = 10, maxGSSize = 500,
                                        pvalueCutoff = 0.05, pAdjustMethod = 'BH',
                                        by = 'fgsea', eps = 0)
# Visualize overlaps of overrepresented pathways between datasets
ggvenn(list(Combined_Klin. = gseaResC8_Klin@result$Description,
            Combined_Krij. = gseaResC8_Krij@result$Description),
       fill_color = c('firebrick1', 'dodgerblue1'), fill_alpha = 0.6,
       set_name_color = c('firebrick1', 'dodgerblue1'), set_name_size = 9,
       text_size = 5, stroke_size = 1.5)
```

Klin. where overrepresented cell types are also observed in Krij.
```{r}
gseaRes4Plot <- gseaResC8_Klin@result %>%
  dplyr::select(Description, setSize, NES, pvalue, p.adjust, leading_edge) %>%
  # Extract overrepresented pathways captured also in Krij. data
  dplyr::filter(Description %in% gseaResC8_Krij@result$Description) %>%
  dplyr::arrange(p.adjust) %>%
  dplyr::slice_head(n = 20) %>%
  dplyr::mutate(leading_edge = stringr::str_extract(leading_edge, '\\d\\d'),
                leading_edge = as.numeric(leading_edge),
                Recurrence = ifelse(test = NES > 0, yes = 'Yes', no = 'No')) %>%
  dplyr::rename(gene_ratio = leading_edge)
midGrad <- mean(gseaRes4Plot$p.adjust)

ggplot(gseaRes4Plot, aes(x=NES, y=reorder(Description, NES), fill=p.adjust, col=Recurrence)) +
  geom_bar(stat = 'identity', width = 0.7, size = 2) +
  scale_fill_gradient2(low='#5B1C00', high='#FFFCA5', mid = '#E86900', midpoint = midGrad, name = 'Adjusted pVal') +
  scale_color_manual(values = c(Yes = 'red', No = 'darkblue')) +
  labs(x = 'Normalized Enrichment Score', y = 'Pathway',
       title = 'MSigDB:C8 - Common features in Klin. data') +
  theme_minimal() +
  theme(axis.title = element_text(size = 16, face = 'bold'),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 10, face = 'bold'),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        title = element_text(face = 'bold', size = 12))
```

Krij. where overrepresented cell types are also observed in Klin.
```{r}
gseaRes4Plot <- gseaResC8_Krij@result %>%
  dplyr::select(Description, setSize, NES, pvalue, p.adjust, leading_edge) %>%
  # Extract overrepresented pathways captured also in Klin. data
  dplyr::filter(Description %in% gseaResC8_Klin@result$Description) %>%
  dplyr::arrange(p.adjust) %>%
  dplyr::slice_head(n = 20) %>%
  dplyr::mutate(leading_edge = stringr::str_extract(leading_edge, '\\d\\d'),
                leading_edge = as.numeric(leading_edge),
                Recurrence = ifelse(test = NES > 0, yes = 'Yes', no = 'No')) %>%
  dplyr::rename(gene_ratio = leading_edge)
midGrad <- mean(gseaRes4Plot$p.adjust)

ggplot(gseaRes4Plot, aes(x=NES, y=reorder(Description, NES), fill=p.adjust, col=Recurrence)) +
  geom_bar(stat = 'identity', width = 0.7, size = 2) +
  scale_fill_gradient2(low='#5B1C00', high='#FFFCA5', mid = '#E86900', midpoint = midGrad, name = 'Adjusted pVal') +
  scale_color_manual(values = c(Yes = 'red', No = 'darkblue')) +
  labs(x = 'Normalized Enrichment Score', y = 'Pathway',
       title = 'MSigDB:C8 - Common features in Krij. data') +
  theme_minimal() +
  theme(axis.title = element_text(size = 16, face = 'bold'),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 10, face = 'bold'),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        title = element_text(face = 'bold', size = 12))
```

**t-values of Normal proteins in Krij. and Klin. data between Recurrence and Nonrecurrence patient groups**\
All features are used!

---- MSigDB:Curated gene sets ----\
Overlaps of overrepresented pathways between datasets
```{r}
# Load preprocessed data
combined_proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds')
combined_proTissue_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsnBC.rds')
# Subset certain sample condition
combined_proNormal_Klin <- combined_proTissue_Klin[, colData(combined_proTissue_Klin)$Condition == 'Normal']
combined_proNormal_Krij <- combined_proTissue_Krij[, colData(combined_proTissue_Krij)$Condition == 'Normal']
# Do association tests between unique features and patient recurrence
sigRes_Klin <- doSOA(combined_proNormal_Klin, meta_var = 'Recurrence', use_proDA = T)
sigRes_Krij <- doSOA(combined_proNormal_Krij, meta_var = 'Recurrence', use_proDA = T)
# Prepare ranked gene list
statRes_Klin <- sigRes_Klin$featAssoRes %>%
  dplyr::select(Var1, Stat) %>%
  tibble::column_to_rownames('Var1')
featAnno_Klin <- tibble::as_tibble(rowData(combined_proNormal_Klin), rownames = 'Protein') %>%
  dplyr::rename(Gene = PG.Genes) %>%
  tibble::column_to_rownames('Protein')
rankedGeneList_Klin <- prepRankedFeatList(statRes = statRes_Klin, featAnno = featAnno_Klin, to_gene = T)

statRes_Krij <- sigRes_Krij$featAssoRes %>%
  dplyr::select(Var1, Stat) %>%
  tibble::column_to_rownames('Var1')
featAnno_Krij <- tibble::as_tibble(rowData(combined_proNormal_Krij), rownames = 'Protein') %>%
  dplyr::select(Protein, Genes) %>%
  dplyr::rename(Gene = Genes) %>%
  tibble::column_to_rownames('Protein')
rankedGeneList_Krij <- prepRankedFeatList(statRes = statRes_Krij, featAnno = featAnno_Krij, to_gene = T)

# Curated gene sets
# Retrieve annotated gene sets
msigTabC2 <- msigdbr(species = 'Homo sapiens', category = 'C2', subcategory = NULL) %>%
  dplyr::select(gs_name, human_gene_symbol)

# Run GSEA
gseaResC2_Klin <- clusterProfiler::GSEA(geneList = rankedGeneList_Klin, TERM2GENE = msigTabC2,
                                        minGSSize = 10, maxGSSize = 500,
                                        pvalueCutoff = 0.05, pAdjustMethod = 'BH',
                                        by = 'fgsea', eps = 0)
gseaResC2_Krij <- clusterProfiler::GSEA(geneList = rankedGeneList_Krij, TERM2GENE = msigTabC2,
                                        minGSSize = 10, maxGSSize = 500,
                                        pvalueCutoff = 0.05, pAdjustMethod = 'BH',
                                        by = 'fgsea', eps = 0)

# Visualize overlaps of overrepresented pathways between datasets
orPathways_Krij <- gseaResC2_Krij@result$Description
orPathways_Klin <- gseaResC2_Klin@result$Description
ggvenn(list(Combined_Krij. = orPathways_Krij,
            Combined_Klin. = orPathways_Klin),
       fill_color = c('firebrick1', 'dodgerblue1'), fill_alpha = 0.6,
       set_name_color = c('firebrick1', 'dodgerblue1'), set_name_size = 9,
       text_size = 5, stroke_size = 1.5)
```

Krij. where represented pathways are also observed in Klin.
```{r}
# Plot enrichment analysis results
gseaRes4Plot <- gseaResC2_Krij@result %>%
  dplyr::select(Description, setSize, NES, pvalue, p.adjust, leading_edge) %>%
  # Extract overrepresented pathways captured also in Klin. data
  dplyr::filter(Description %in% orPathways_Klin) %>%
  dplyr::arrange(p.adjust) %>%
  dplyr::slice_head(n = 8) %>%
  dplyr::mutate(leading_edge = stringr::str_extract(leading_edge, '\\d\\d'),
                leading_edge = as.numeric(leading_edge),
                # Description = stringr::str_remove(Description, '^REACTOME_'),
                Recurrence = ifelse(test = NES > 0, yes = 'Yes', no = 'No')) %>%
  dplyr::rename(gene_ratio = leading_edge)
midGrad <- mean(gseaRes4Plot$p.adjust)

ggplot(gseaRes4Plot, aes(x=NES, y=reorder(Description, NES), fill=p.adjust, col=Recurrence)) +
  geom_bar(stat = 'identity', width = 0.7, size = 2) +
  scale_fill_gradient2(low='#5B1C00', high='#FFFCA5', mid = '#E86900', midpoint = midGrad, name = 'Adjusted pVal') +
  scale_color_manual(values = c(Yes = 'red', No = 'darkblue')) +
  labs(x = 'Normalized Enrichment Score', y = 'Pathway',
       title = 'MSigDB:C2 - Normal proteins in Krij. data') +
  theme_minimal() +
  theme(axis.title = element_text(size = 16, face = 'bold'),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 10, face = 'bold'),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        title = element_text(face = 'bold', size = 12))
```

Klin. where represented pathways are also observed in Krij.
```{r}
gseaRes4Plot <- gseaResC2_Klin@result %>%
  dplyr::select(Description, setSize, NES, pvalue, p.adjust, leading_edge) %>%
  # Extract overrepresented pathways captured also in Krij. data
  dplyr::filter(Description %in% orPathways_Krij) %>%
  dplyr::arrange(p.adjust) %>%
  dplyr::slice_head(n = 8) %>%
  dplyr::mutate(leading_edge = stringr::str_extract(leading_edge, '\\d\\d'),
                leading_edge = as.numeric(leading_edge),
                # Description = stringr::str_remove(Description, '^REACTOME_'),
                Recurrence = ifelse(test = NES > 0, yes = 'Yes', no = 'No')) %>%
  dplyr::rename(gene_ratio = leading_edge)
midGrad <- mean(gseaRes4Plot$p.adjust)

ggplot(gseaRes4Plot, aes(x=NES, y=reorder(Description, NES), fill=p.adjust, col=Recurrence)) +
  geom_bar(stat = 'identity', width = 0.7, size = 2) +
  scale_fill_gradient2(low='#5B1C00', high='#FFFCA5', mid = '#E86900', midpoint = midGrad, name = 'Adjusted pVal') +
  scale_color_manual(values = c(Yes = 'red', No = 'darkblue')) +
  labs(x = 'Normalized Enrichment Score', y = 'Pathway',
       title = 'MSigDB:C2 - Normal proteins in Klin. data') +
  theme_minimal() +
  theme(axis.title = element_text(size = 16, face = 'bold'),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 10, face = 'bold'),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        title = element_text(face = 'bold', size = 12))
```

Relationships between overrepresented pathways and genes of interest (captured in MethodDev, Discovery, and Combined Krij.)
```{r}
# Load SOA results
Combined_proNormalRes_Krij <- readRDS('./data/Discovery/potential_proteins/Combined_proNormalRes_Krij.rds')
Combined_proNormalRes_Klin <- readRDS('./data/Discovery/potential_proteins/Combined_proNormalRes_Klin.rds')
MethodDev_proNormalRes_Krij <- readRDS('./data/Discovery/potential_proteins/MethodDev_proNormalRes_Krij.rds')
Discovery_proNormalRes_Krij <- readRDS('./data/Discovery/potential_proteins/Discovery_proNormalRes_Krij.rds')
# Retrieve significant features from SOA results
Combined_proNormalSigFeats_Krij <- Combined_proNormalRes_Krij$sig.feat.tab$Var1 %>%
  stringr::str_remove_all(';.+')
MethodDev_proNormalSigFeats_Krij <- MethodDev_proNormalRes_Krij$sig.feat.tab$Var1 %>%
  stringr::str_remove_all(';.+')
Discovery_proNormalSigFeats_Krij <- Discovery_proNormalRes_Krij$sig.feat.tab$Var1 %>%
  stringr::str_remove_all(';.+')
Combined_proNormalSigFeats_Klin <- Combined_proNormalRes_Klin$sig.feat.tab$Var1 %>%
  stringr::str_remove_all(';.+')

# Extract common significant features across datasets
# (MethodDev and Discovery Normal Tissue Proteomics shares few significant features)
cmnProNormalSigFeats_Krij <- intersect(Combined_proNormalSigFeats_Krij, MethodDev_proNormalSigFeats_Krij)
cmnProNormalSigFeats_Krij <- intersect(cmnProNormalSigFeats_Krij, Discovery_proNormalSigFeats_Krij)
# intersect(cmnProNormalSigFeats_Krij, Combined_proNormalSigFeats_Klin)
# Change proteins to genes
cmnProNormalSigGenes_Krij <- tibble::as_tibble(rowData(combined_proNormal_Krij), rownames = 'Feature') %>%
  dplyr::filter(Feature %in% cmnProNormalSigFeats_Krij) %>%
  dplyr::pull(Genes)

# Extract common overrepresented pathways between two datasets
cmnORPathways <- intersect(orPathways_Krij, orPathways_Klin)
cmnORGeneSets <- gseaResC2_Krij@geneSets[names(gseaResC2_Krij@geneSets) %in% cmnORPathways]
# Remove overrepresented pathways that do not include any feature of interest
keptGeneSets <- sapply(seq_len(length(cmnORGeneSets)), function(i) {
  any(cmnProNormalSigGenes_Krij %in% cmnORGeneSets[[i]])
})
cmnORGeneSets <- cmnORGeneSets[keptGeneSets]
# Create binary matrix indicating presence of features of interest in overrepresented pathways
orGeneBinary <- matrix(nrow = length(cmnORGeneSets), ncol = length(cmnProNormalSigGenes_Krij),
                        dimnames = list(names(cmnORGeneSets), cmnProNormalSigGenes_Krij))
for (i in seq_len(nrow(orGeneBinary))) {
  for (j in seq_len(ncol(orGeneBinary))) {
    if (colnames(orGeneBinary)[j] %in% cmnORGeneSets[[rownames(orGeneBinary)[i]]]) {
      orGeneBinary[i, j] <- T
    } else {
      orGeneBinary[i, j] <- F
    }
  }
}
# Remove features of interest not included in any represented pathway
keptGenes <- apply(orGeneBinary, 2, any)
orGeneBinary <- orGeneBinary[, keptGenes]

# Plot heatmap
pw <- 'REACTOME_ACTIVATION_OF_THE_MRNA_UPON_BINDING_OF_THE_CAP_BINDING_COMPLEX_AND_EIFS_AND_SUBSEQUENT_BINDING_TO_43S'
orGeneHeatmap <- tibble::as_tibble(orGeneBinary, rownames = 'Pathway') %>%
  tidyr::pivot_longer(cols = -'Pathway', names_to = 'Gene', values_to = 'tStat') %>%
  dplyr::mutate(tStat = dplyr::case_when(tStat ~ rankedGeneList_Krij[Gene],
                                         !tStat ~ 0),
                Pathway = dplyr::case_when(Pathway == pw ~ 'REACTOME_ACTIVATION_OF_THE_MRNA_AND_BINDING_TO_43S',
                                           Pathway != pw ~ Pathway))

ggplot(orGeneHeatmap, aes(x=Gene, y=Pathway, fill=tStat)) +
  geom_tile() +
  scale_fill_gradient2(high = 'darkred', low = 'darkblue', mid = 'white', midpoint = 0) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

---- MSigDB:Oncogenic gene sets ----\
Overlaps of overrepresented pathways between datasets
```{r}
# Oncogenic gene sets
# Retrieve annotated gene sets
msigTabC6 <- msigdbr(species = 'Homo sapiens', category = 'C6', subcategory = NULL) %>%
  dplyr::select(gs_name, human_gene_symbol)

# Run GSEA
gseaResC6_Klin <- clusterProfiler::GSEA(geneList = rankedGeneList_Klin, TERM2GENE = msigTabC6,
                                        minGSSize = 10, maxGSSize = 500,
                                        pvalueCutoff = 0.05, pAdjustMethod = 'BH',
                                        by = 'fgsea', eps = 0)
gseaResC6_Krij <- clusterProfiler::GSEA(geneList = rankedGeneList_Krij, TERM2GENE = msigTabC6,
                                        minGSSize = 10, maxGSSize = 500,
                                        pvalueCutoff = 0.05, pAdjustMethod = 'BH',
                                        by = 'fgsea', eps = 0)

# Visualize overlaps of overrepresented pathways between datasets
orPathways_Krij <- gseaResC6_Krij@result$Description
orPathways_Klin <- gseaResC6_Klin@result$Description
ggvenn(list(Combined_Krij. = orPathways_Krij,
            Combined_Klin. = orPathways_Klin),
       fill_color = c('firebrick1', 'dodgerblue1'), fill_alpha = 0.6,
       set_name_color = c('firebrick1', 'dodgerblue1'), set_name_size = 9,
       text_size = 5, stroke_size = 1.5)

# Plot enrichment analysis results
gseaRes4Plot <- gseaResC6_Klin@result %>%
  dplyr::select(Description, setSize, NES, pvalue, p.adjust, leading_edge) %>%
  dplyr::arrange(p.adjust) %>%
  dplyr::slice_head(n = 8) %>%
  dplyr::mutate(leading_edge = stringr::str_extract(leading_edge, '\\d\\d'),
                leading_edge = as.numeric(leading_edge),
                Recurrence = ifelse(test = NES > 0, yes = 'Yes', no = 'No')) %>%
  dplyr::rename(gene_ratio = leading_edge)
midGrad <- mean(gseaRes4Plot$p.adjust)

ggplot(gseaRes4Plot, aes(x=NES, y=reorder(Description, NES), fill=p.adjust, col=Recurrence)) +
  geom_bar(stat = 'identity', width = 0.7, size = 2) +
  scale_fill_gradient2(low='#5B1C00', high='#FFFCA5', mid = '#E86900', midpoint = midGrad, name = 'Adjusted pVal') +
  scale_color_manual(values = c(Yes = 'red', No = 'darkblue')) +
  labs(x = 'Normalized Enrichment Score', y = 'Pathway',
       title = 'MSigDB:C6 - Normal proteins in Klin. data') +
  theme_minimal() +
  theme(axis.title = element_text(size = 16, face = 'bold'),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 10, face = 'bold'),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        title = element_text(face = 'bold', size = 12))
```
=> Only overrepresented pathway in Krij. is 'BMI1_DN.V1_UP'.

<font size='4'> **Unique features in Krij. or Klin. data** </font>\
---- MSigDB:Curated gene sets ----
```{r}
# Keep only first protein as representative if there are more than one
rownames(combined_proNormal_Krij) <- stringr::str_remove(rownames(combined_proNormal_Krij), ';.+')
rownames(combined_proNormal_Klin) <- stringr::str_remove(rownames(combined_proNormal_Klin), ';.+')
# Prepare protein-gene map table
featAnno_Krij <- tibble::as_tibble(rowData(combined_proNormal_Krij), rownames = 'Protein') %>%
  dplyr::select(Protein, Genes) %>%
  dplyr::rename(Gene = Genes) %>%
  tibble::column_to_rownames('Protein')
featAnno_Klin <- tibble::as_tibble(rowData(combined_proNormal_Klin), rownames = 'Protein') %>%
  dplyr::rename(Gene = PG.Genes) %>%
  tibble::column_to_rownames('Protein')
# Extract unique features in Krij. and Klin. data
feats_Krij <- rownames(combined_proNormal_Krij)
uniqFeats_Krij <- feats_Krij[!feats_Krij %in% feats_Klin]
uniqFeats_Krij <- featAnno_Krij[uniqFeats_Krij,]
rankedGeneList_Krij <- rankedGeneList_Krij[names(rankedGeneList_Krij) %in% uniqFeats_Krij]

feats_Klin <- rownames(combined_proNormal_Klin)
uniqFeats_Klin <- feats_Klin[!feats_Klin %in% feats_Krij]
uniqFeats_Klin <- featAnno_Klin[uniqFeats_Klin,]
rankedGeneList_Klin <- rankedGeneList_Klin[names(rankedGeneList_Klin) %in% uniqFeats_Klin]

# Curated gene sets
# Retrieve annotated gene sets
msigTabC2 <- msigdbr(species = 'Homo sapiens', category = 'C2', subcategory = NULL) %>%
  dplyr::select(gs_name, human_gene_symbol)

# Run GSEA
gseaResC2_Krij <- clusterProfiler::GSEA(geneList = rankedGeneList_Krij, TERM2GENE = msigTabC2,
                                        minGSSize = 10, maxGSSize = 500,
                                        pvalueCutoff = 0.05, pAdjustMethod = 'BH',
                                        by = 'fgsea', eps = 0)
gseaResC2_Klin <- clusterProfiler::GSEA(geneList = rankedGeneList_Klin, TERM2GENE = msigTabC2,
                                        minGSSize = 10, maxGSSize = 500,
                                        pvalueCutoff = 0.05, pAdjustMethod = 'BH',
                                        by = 'fgsea', eps = 0)

# Plot enrichment analysis results
gseaRes4Plot <- gseaResC2_Klin@result %>%
  dplyr::select(Description, setSize, NES, pvalue, p.adjust, leading_edge) %>%
  dplyr::arrange(p.adjust) %>%
  dplyr::slice_head(n = 8) %>%
  dplyr::mutate(leading_edge = stringr::str_extract(leading_edge, '\\d\\d'),
                leading_edge = as.numeric(leading_edge),
                # Description = stringr::str_remove(Description, '^REACTOME_'),
                Recurrence = ifelse(test = NES > 0, yes = 'Yes', no = 'No')) %>%
  dplyr::rename(gene_ratio = leading_edge)
midGrad <- mean(gseaRes4Plot$p.adjust)

ggplot(gseaRes4Plot, aes(x=NES, y=reorder(Description, NES), fill=p.adjust, col=Recurrence)) +
  geom_bar(stat = 'identity', width = 0.7, size = 2) +
  scale_fill_gradient2(low='#5B1C00', high='#FFFCA5', mid = '#E86900', midpoint = midGrad, name = 'Adjusted pVal') +
  scale_color_manual(values = c(Yes = 'red', No = 'darkblue')) +
  labs(x = 'Normalized Enrichment Score', y = 'Pathway',
       title = 'MSigDB:C2 - Unique normal proteins in Klin. data') +
  theme_minimal() +
  theme(axis.title = element_text(size = 16, face = 'bold'),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 10, face = 'bold'),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        title = element_text(face = 'bold', size = 12))
```
=> No overrepresented pathways in Krij. data

---- GO:Biological Process ----
```{r}
# GO:BP
# Retrieve annotated gene sets
msigTabC5BP <- msigdbr(species = 'Homo sapiens', category = 'C5', subcategory = 'BP') %>%
  dplyr::select(gs_name, human_gene_symbol)
# Run GSEA
gseaResC5BP_Krij <- clusterProfiler::GSEA(geneList = rankedGeneList_Krij, TERM2GENE = msigTabC5BP,
                                          minGSSize = 10, maxGSSize = 500,
                                          pvalueCutoff = 0.05, pAdjustMethod = 'BH',
                                          by = 'fgsea', eps = 0)
gseaResC5BP_Klin <- clusterProfiler::GSEA(geneList = rankedGeneList_Klin, TERM2GENE = msigTabC5BP,
                                          minGSSize = 10, maxGSSize = 500,
                                          pvalueCutoff = 0.05, pAdjustMethod = 'BH',
                                          by = 'fgsea', eps = 0)

gseaRes4Plot <- gseaResC5BP_Krij@result %>%
  dplyr::select(Description, setSize, NES, pvalue, p.adjust, leading_edge) %>%
  dplyr::arrange(p.adjust) %>%
  dplyr::slice_head(n = 8) %>%
  dplyr::mutate(leading_edge = stringr::str_extract(leading_edge, '\\d\\d'),
                leading_edge = as.numeric(leading_edge),
                Recurrence = ifelse(test = NES > 0, yes = 'Yes', no = 'No')) %>%
  dplyr::rename(gene_ratio = leading_edge)
midGrad <- mean(gseaRes4Plot$p.adjust)

ggplot(gseaRes4Plot, aes(x=NES, y=reorder(Description, NES), fill=p.adjust, col=Recurrence)) +
  geom_bar(stat = 'identity', width = 0.7, size = 2) +
  scale_fill_gradient2(low='#5B1C00', high='#FFFCA5', mid = '#E86900', midpoint = midGrad, name = 'Adjusted pVal') +
  scale_color_manual(values = c(Yes = 'red', No = 'darkblue')) +
  labs(x = 'Normalized Enrichment Score', y = 'Pathway',
       title = 'GO:BP - Unique normal proteins in Krij. data') +
  theme_minimal() +
  theme(axis.title = element_text(size = 16, face = 'bold'),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 10, face = 'bold'),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        title = element_text(face = 'bold', size = 12))


gseaRes4Plot <- gseaResC5BP_Klin@result %>%
  dplyr::select(Description, setSize, NES, pvalue, p.adjust, leading_edge) %>%
  dplyr::arrange(p.adjust) %>%
  dplyr::slice_head(n = 8) %>%
  dplyr::mutate(leading_edge = stringr::str_extract(leading_edge, '\\d\\d'),
                leading_edge = as.numeric(leading_edge),
                Recurrence = ifelse(test = NES > 0, yes = 'Yes', no = 'No')) %>%
  dplyr::rename(gene_ratio = leading_edge)
midGrad <- mean(gseaRes4Plot$p.adjust)

ggplot(gseaRes4Plot, aes(x=NES, y=reorder(Description, NES), fill=p.adjust, col=Recurrence)) +
  geom_bar(stat = 'identity', width = 0.7, size = 2) +
  scale_fill_gradient2(low='#5B1C00', high='#FFFCA5', mid = '#E86900', midpoint = midGrad, name = 'Adjusted pVal') +
  scale_color_manual(values = c(Yes = 'red', No = 'darkblue')) +
  labs(x = 'Normalized Enrichment Score', y = 'Pathway',
       title = 'GO:BP - Unique normal proteins in Klin. data') +
  theme_minimal() +
  theme(axis.title = element_text(size = 16, face = 'bold'),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 10, face = 'bold'),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        title = element_text(face = 'bold', size = 12))
```
<br/>
<br/>

**Correlation coefficients of common features between two datasets**
```{r}
# Load preprocessed data
combined_proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds')
combined_proTissue_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsnBC.rds')
# Compute correlations of common features between two datasets
corrRes <- calcFeatCorr(combined_proTissue_Krij, combined_proTissue_Klin)
# Prepare ranked gene list
statRes <- tibble::column_to_rownames(corrRes, 'Feature')
featAnno <- tibble::as_tibble(rowData(combined_proTissue_Krij), rownames = 'Protein') %>%
  dplyr::select(Protein, Genes) %>%
  dplyr::rename(Gene = Genes) %>%
  tibble::column_to_rownames('Protein')
rankedGeneList <- prepRankedFeatList(statRes = statRes, featAnno = featAnno, to_gene = T)

# Cell type sets
# View supported species and collections in MSigDB
# msigdbr_species()
# msigdbr_collections()
# Retrieve annotated gene sets
msigTabC8 <- msigdbr(species = 'Homo sapiens', category = 'C8', subcategory = NULL) %>%
  dplyr::select(gs_name, human_gene_symbol)

# Run GSEA
gseaResC8 <- clusterProfiler::GSEA(geneList = rankedGeneList, TERM2GENE = msigTabC8,
                                   minGSSize = 10, maxGSSize = 500,
                                   pvalueCutoff = 0.05, pAdjustMethod = 'BH',
                                   by = 'fgsea', eps = 0)

# Plot enrichment analysis results
gseaRes4Plot <- gseaResC8@result %>%
  dplyr::select(Description, setSize, NES, pvalue, p.adjust, leading_edge) %>%
  dplyr::arrange(p.adjust) %>%
  dplyr::slice_head(n = 20) %>%
  dplyr::mutate(leading_edge = stringr::str_extract(leading_edge, '\\d\\d'),
                leading_edge = as.numeric(leading_edge),
                Relationship = ifelse(test = NES > 0, yes = 'Positive', no = 'Negative')) %>%
  dplyr::rename(gene_ratio = leading_edge)
midGrad <- mean(gseaRes4Plot$p.adjust)

ggplot(gseaRes4Plot, aes(x=NES, y=reorder(Description, NES), fill=p.adjust, col=Relationship)) +
  geom_bar(stat = 'identity', width = 0.7, size = 2) +
  scale_fill_gradient2(low='#5B1C00', high='#FFFCA5', mid = '#E86900', midpoint = midGrad, name = 'Adjusted pVal') +
  scale_color_manual(values = c(Positive = 'red', Negative = 'darkblue')) +
  labs(x = 'Normalized Enrichment Score', y = 'Pathway',
       title = 'MSigDB:C8 - Relationships of features between Krij. and Klin. data') +
  theme_minimal() +
  theme(axis.title = element_text(size = 16, face = 'bold'),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 10, face = 'bold'),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        title = element_text(face = 'bold', size = 12))

# Hallmark gene sets
# Retrieve annotated gene sets
msigTabH <- msigdbr(species = 'Homo sapiens', category = 'H', subcategory = NULL) %>%
  dplyr::select(gs_name, human_gene_symbol)

# Run GSEA
gseaResH <- clusterProfiler::GSEA(geneList = rankedGeneList, TERM2GENE = msigTabH,
                                  minGSSize = 10, maxGSSize = 500,
                                  pvalueCutoff = 0.05, pAdjustMethod = 'BH',
                                  by = 'fgsea', eps = 0)

# Plot enrichment analysis results
gseaRes4Plot <- gseaResH@result %>%
  dplyr::select(Description, setSize, NES, pvalue, p.adjust, leading_edge) %>%
  dplyr::arrange(p.adjust) %>%
  dplyr::slice_head(n = 8) %>%
  dplyr::mutate(leading_edge = stringr::str_extract(leading_edge, '\\d\\d'),
                leading_edge = as.numeric(leading_edge),
                Relationship = ifelse(test = NES > 0, yes = 'Positive', no = 'Negative')) %>%
  dplyr::rename(gene_ratio = leading_edge)
midGrad <- mean(gseaRes4Plot$p.adjust)

ggplot(gseaRes4Plot, aes(x=NES, y=reorder(Description, NES), fill=p.adjust, col=Relationship)) +
  geom_bar(stat = 'identity', width = 0.6, size = 2) +
  scale_fill_gradient2(low='#5B1C00', high='#FFFCA5', mid = '#E86900', midpoint = midGrad, name = 'Adjusted pVal') +
  scale_color_manual(values = c(Positive = 'red', Negative = 'darkblue')) +
  labs(x = 'Normalized Enrichment Score', y = 'Pathway',
       title = 'MSigDB:H - Relationships of features between Krij. and Klin. data') +
  theme_minimal() +
  theme(axis.title = element_text(size = 16, face = 'bold'),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 10, face = 'bold'),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        title = element_text(face = 'bold', size = 12))
```




# Normal Tissue Proteomics
Check association of recurrence-related feature significance between Discovery and
Method Development datasets or between Krij.'s and Klin.'s datasets

## AG Krijgsveld

Display association of feature recurrence predictive power between Discovery and
Method Development Normal Tissue DIA Proteomics generated by AG Krijgsveld\
**p-values (significance level = 0.05)**
```{r}
# Load normalized data
proTissue_Discovery <- readRDS('./data/Discovery/AG_Krijgsveld/proTissueVsn.rds')
proTissue_MethodDev <- readRDS('./data/MethodDev/AG_Krijgsveld/new_proTissueVsn.rds')
# Prepare protein-gene table for providing gene names
featInfo <- tibble::as_tibble(rowData(proTissue_Discovery), rownames = 'Proteins') %>%
  dplyr::select(Proteins, Genes) %>%
  dplyr::mutate(Proteins = stringr::str_remove(Proteins, ';.*'),
                Genes = stringr::str_remove(Genes, ';.*'))

# Prepare summarized tables of comparison between Recurrence and Non-recurrence patient groups
recurNormalFeatAssoTabs_Krij <- summarizeFeatAssoRes(proTissue_Discovery, proTissue_MethodDev,
                                                     'Discovery', 'MethodDev',
                                                     meta_var = 'Recurrence',
                                                     method = 'proDA',
                                                     subset = c('Condition', 'Normal'),
                                                     prot_gene_tbl = featInfo,
                                                     alpha = 0.05)
featAssoTabs <- recurNormalFeatAssoTabs_Krij

# Plot p-values
ggplot(featAssoTabs$tbl_pVal, aes(x=-log10(pVal_dat1), y=-log10(pVal_dat2),
                                  col=Significance, label=Label)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.5) +
  scale_color_manual(values = c('red', 'darkgreen', 'navy', 'grey')) +
  geom_text_repel(size = 4, show.legend = F) +
  labs(x = '-log10(pVal) of Discovery Cohort', y = '-log10(pVal) of MethodDev Cohort',
       title = 'Recurrence: Normal Tissue DIA Proteomics (AG Krijgsveld)') +
  th #+ theme(axis.title = element_text(size = 22),
             # axis.text = element_text(size = 16),
             # legend.title = element_text(size = 20), legend.text = element_text(size = 18))
# ggsave('./output/SC_meeting/pVals_Dis_vs_MetDev_proNormal_recur_Krij.png', device = 'png', dpi = 400, height = 8, width = 11)

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_pVal$Significance == 'Sig. in Both') /
                   sum(featAssoTabs$tbl_pVal$Significance != 'Not Sig.'), 3) * 100
cat(paste0('=> ', percent, '% of significant results are consistent'))
```

**log(FC)**
```{r}
# Plot log(FC)
ggplot(featAssoTabs$tbl_logFC, aes(x=logFC_dat1, y=logFC_dat2, col=Direction)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.5) +
  scale_color_manual(values = c('red', 'navy', 'grey')) +
  labs(x = 'log(FC) of Discovery Cohort', y = 'log(FC) of MethodDev Cohort',
       title = 'Recurrence: Normal Tissue DIA Proteomics (AG Krijgsveld)') +
  th #+ theme(axis.title = element_text(size = 22),
             # axis.text = element_text(size = 16),
             # legend.title = element_text(size = 20), legend.text = element_text(size = 18))
# ggsave('./output/SC_meeting/log(FC)_Dis_vs_MetDev_proNormal_recur_Krij.png', device = 'png', dpi = 400, height = 8, width = 11)

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_logFC$Direction != 'Inconsistent') /
                   nrow(featAssoTabs$tbl_logFC), 3) * 100
cat(paste0('=> ', percent, '% of results are consistent'))
```

Visualize top significant proteins identified in both Discovery and Method Development
datasets and based on Discovery one. Note that Method Development data is provided
to have more comprehensive analysis and it may not be on same scale as Discovery data.\
ns: p > 0.05, ∗: p <= 0.05, ∗∗: p <= 0.01, ***: p <= 0.001
```{r fig.height=12}
# Plot top significant features based on 'data1'
topSigFeatTab <- dplyr::mutate(featAssoTabs$tbl_topSigFeats,
                               newGroup = factor(newGroup, levels = c('Yes_Discovery',
                                                                      'No_Discovery',
                                                                      'Yes_MethodDev',
                                                                      'No_MethodDev')))
ggplot(topSigFeatTab, aes(x=newGroup, y=Abundance, col=DataName, fill=Recurrence)) +
  geom_boxplot(alpha = 1, outlier.shape = NA, linewidth = 1) +
  geom_jitter(position = position_jitter(0.2), size = 3, show.legend = F) +
  ggpubr::stat_compare_means(method = 't.test', paired = F, method.args = list(var.equal = T),
                             comparisons = list(c('Yes_Discovery', 'No_Discovery'),
                                                c('Yes_MethodDev', 'No_MethodDev')),
                             label = 'p.signif', tip.length = 0.015, bracket.size = 0.7, size = 4) +
  labs(x = 'Recurrence', title = 'Recurrence: Normal Tissue DIA Proteomics (AG Krijgsveld)') +
  scale_color_manual(name = 'Cohort', values = c('firebrick', 'grey50')) +
  scale_fill_manual(values=c('#00BFC4', '#F8766D')) +
  facet_wrap(vars(newFeature), scales = 'free') +
  th +
  theme(strip.text = element_text(size = 13, face = 'bold'),
        axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1))
#   theme(strip.text = element_text(size = 14, face = 'bold'),
#         axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1, size = 20),
#         axis.title = element_text(size = 26),
#         legend.title = element_text(size = 24), legend.text = element_text(size = 22))
# ggsave('./output/group_meeting/topSigFeats_Discovery_vs_MethodDev_proNormal_Krij.png', device = 'png', dpi = 400, height = 8, width = 16)
```

## AG Klingmüller

Display association of feature recurrence predictive power between Discovery and
Method Development Normal Tissue DIA Proteomics generated by AG Klingmüller\
**p-values (significance level = 0.05)**
```{r}
# Load normalized data
proTissue_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsn.rds')
proTissue_Discovery <- proTissue_Klin[, colData(proTissue_Klin)$Cohort == 'Discovery']
proTissue_MethodDev <- proTissue_Klin[, colData(proTissue_Klin)$Cohort == 'MethodDev']
# Prepare protein-gene table for providing gene names
featInfo <- tibble::as_tibble(rowData(proTissue_Discovery), rownames = 'Proteins') %>%
  dplyr::rename(Genes = PG.Genes) %>%
  dplyr::mutate(Proteins = stringr::str_remove(Proteins, ';.*'),
                Genes = stringr::str_remove(Genes, ';.*'))

# Prepare summarized tables of comparison between Recurrence and Non-recurrence patient groups
recurNormalFeatAssoTabs_Klin <- summarizeFeatAssoRes(proTissue_Discovery, proTissue_MethodDev,
                                                     'Discovery', 'MethodDev',
                                                     meta_var = 'Recurrence',
                                                     subset = c('Condition', 'Normal'),
                                                     prot_gene_tbl = featInfo,
                                                     method = 'proDA',
                                                     alpha = 0.05)
featAssoTabs <- recurNormalFeatAssoTabs_Klin

# Plot p-values
ggplot(featAssoTabs$tbl_pVal, aes(x=-log10(pVal_dat1), y=-log10(pVal_dat2),
                                  col=Significance, label=Label)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.5) +
  scale_color_manual(values = c('red', 'darkgreen', 'navy', 'grey')) +
  geom_text_repel(size = 4, show.legend = F) +
  labs(x = '-log10(pVal) of Discovery Cohort', y = '-log10(pVal) of MethodDev Cohort',
       title = 'Recurrence: Normal Tissue DIA Proteomics (AG Klingmüller)') +
  th #+ theme(axis.title = element_text(size = 22),
             # axis.text = element_text(size = 16),
             # legend.title = element_text(size = 20), legend.text = element_text(size = 18))
# ggsave('./output/SC_meeting/pVals_Dis_vs_MetDev_proNormal_recur_Klin.png', device = 'png', dpi = 400, height = 8, width = 11)

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_pVal$Significance == 'Sig. in Both') /
                   sum(featAssoTabs$tbl_pVal$Significance != 'Not Sig.'), 3) * 100
cat(paste0('=> ', percent, '% of significant results are consistent'))
```

**log(FC)**
```{r}
# Plot log(FC)
ggplot(featAssoTabs$tbl_logFC, aes(x=logFC_dat1, y=logFC_dat2, col=Direction)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.5) +
  scale_color_manual(values = c('red', 'navy', 'grey')) +
  labs(x = 'log(FC) of Discovery Cohort', y = 'log(FC) of MethodDev Cohort',
       title = 'Recurrence: Normal Tissue DIA Proteomics (AG Klingmüller)') +
  th #+ theme(axis.title = element_text(size = 22),
             # axis.text = element_text(size = 16),
             # legend.title = element_text(size = 20), legend.text = element_text(size = 18))
# ggsave('./output/SC_meeting/log(FC)_Dis_vs_MetDev_proNormal_recur_Klin.png', device = 'png', dpi = 400, height = 8, width = 11)

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_logFC$Direction != 'Inconsistent') /
                   nrow(featAssoTabs$tbl_logFC), 3) * 100
cat(paste0('=> ', percent, '% of results are consistent'))
```

Visualize top significant proteins identified in both Discovery and Method Development
datasets and based on Discovery one. Note that Method Development data is provided
to have more comprehensive analysis and it may not be on same scale as Discovery data.\
ns: p > 0.05, ∗: p <= 0.05, ∗∗: p <= 0.01, ***: p <= 0.001
```{r fig.height=12}
# Plot top significant features based on 'data1'
topSigFeatTab <- dplyr::mutate(featAssoTabs$tbl_topSigFeats,
                               newGroup = factor(newGroup, levels = c('Yes_Discovery',
                                                                      'No_Discovery',
                                                                      'Yes_MethodDev',
                                                                      'No_MethodDev')))
ggplot(topSigFeatTab, aes(x=newGroup, y=Abundance, col=DataName, fill=Recurrence)) +
  geom_boxplot(alpha = 1, outlier.shape = NA, linewidth = 1) +
  geom_jitter(position = position_jitter(0.2), size = 3, show.legend = F) +
  ggpubr::stat_compare_means(method = 't.test', paired = F, method.args = list(var.equal = T),
                             comparisons = list(c('Yes_Discovery', 'No_Discovery'),
                                                c('Yes_MethodDev', 'No_MethodDev')),
                             label = 'p.signif', tip.length = 0.015, bracket.size = 0.7, size = 4) +
  labs(x = 'Recurrence', title = 'Recurrence: Normal Tissue DIA Proteomics (AG Klingmüller)') +
  scale_color_manual(name = 'Cohort', values = c('firebrick', 'grey50')) +
  scale_fill_manual(values=c('#00BFC4', '#F8766D')) +
  facet_wrap(vars(newFeature), scales = 'free') +
  th +
  theme(strip.text = element_text(size = 13, face = 'bold'),
        axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1))
```

## Integrative analysis

Show intersections of potential Normal proteins identified in different datasets
(significance level = 0.05)
```{r}
# Prepare recurrence-related significant features
alpha <- 0.05
# AG Klingmüller
sigFeats_Discovery_Klin <- recurNormalFeatAssoTabs_Klin$tbl_featAsso1 %>%
  dplyr::filter(pVal <= alpha) %>%
  dplyr::pull(Feature)
sigFeats_MethodDev_Klin <- recurNormalFeatAssoTabs_Klin$tbl_featAsso2 %>%
  dplyr::filter(pVal <= alpha) %>%
  dplyr::pull(Feature)
# AG Krijgsveld
sigFeats_Discovery_Krij <- recurNormalFeatAssoTabs_Krij$tbl_featAsso1 %>%
  dplyr::filter(pVal <= alpha) %>%
  dplyr::pull(Feature)
sigFeats_MethodDev_Krij <- recurNormalFeatAssoTabs_Krij$tbl_featAsso2 %>%
  dplyr::filter(pVal <= alpha) %>%
  dplyr::pull(Feature)

# Make venn graphs
ggvenn(list(Discovery_Krij. = sigFeats_Discovery_Krij, MethodDev_Krij. = sigFeats_MethodDev_Krij,
            MethodDev_Klin. = sigFeats_MethodDev_Klin),
       fill_color = c('darkred', 'darkblue', 'darkgreen'), fill_alpha = 0.6,
       set_name_color = c('darkred', 'darkblue', 'darkgreen'), set_name_size = 9,
       text_size = 5, stroke_size = 1.5)
# ggsave('./output/SC_meeting/venn_sigFeats_Krij_dis.png', device = 'png', dpi = 400, height = 8, width = 11)
# Identify fully overlapped features
# ol <- intersect(intersect(sigFeats_Discovery_Krij, sigFeats_MethodDev_Krij), sigFeats_MethodDev_Klin)
# recurNormalFeatAssoTabs_Krij$tbl_featAsso1[recurNormalFeatAssoTabs_Krij$tbl_featAsso1$Feature %in% ol,]
```
=> That one protein identified in all datasets is 'P37235 (HPCAL1)'.

```{r}
ggvenn(list(Discovery_Klin. = sigFeats_Discovery_Klin, MethodDev_Klin. = sigFeats_MethodDev_Klin,
            MethodDev_Krij. = sigFeats_MethodDev_Krij),
       fill_color = c('darkorange', 'darkgreen', 'darkblue'), fill_alpha = 0.6,
       set_name_color = c('darkorange', 'darkgreen', 'darkblue'), set_name_size = 9,
       text_size = 5, stroke_size = 1.5)
# ggsave('./output/SC_meeting/venn_sigFeats_Klin_dis.png', device = 'png', dpi = 400, height = 8, width = 11)
# Identify fully overlapped features
# ol <- intersect(intersect(sigFeats_Discovery_Klin, sigFeats_MethodDev_Klin), sigFeats_MethodDev_Krij)
# recurNormalFeatAssoTabs_Klin$tbl_featAsso1[recurNormalFeatAssoTabs_Klin$tbl_featAsso1$Feature %in% ol,]
```
=> Those proteins identified in all datasets are 'Q5S007 (LRRK2)', 'Q99417 (MYCBP)',
'Q96PE3 (INPP4A)', 'Q8NBQ5 (HSD17B11)', 'Q05823 (RNASEL)', 'Q9Y3E5 (PTRH2)', 'Q9H8L6 (MMRN2)'

```{r}
ggvenn(list(Discovery_Krij. = sigFeats_Discovery_Krij, Discovery_Klin. = sigFeats_Discovery_Klin,
            MethodDev_Klin. = sigFeats_MethodDev_Klin),
       fill_color = c('darkred', 'darkorange', 'darkgreen'), fill_alpha = 0.6,
       set_name_color = c('darkred', 'darkorange', 'darkgreen'), set_name_size = 9,
       text_size = 5, stroke_size = 1.5)
# ggsave('./output/SC_meeting/venn_sigFeats_Krij_Klin_dis.png', device = 'png', dpi = 400, height = 8, width = 11)
# Identify fully overlapped features
# ol <- intersect(intersect(sigFeats_Discovery_Krij, sigFeats_Discovery_Klin), sigFeats_MethodDev_Klin)
# recurNormalFeatAssoTabs_Krij$tbl_featAsso1[recurNormalFeatAssoTabs_Krij$tbl_featAsso1$Feature %in% ol,]
```
=> That one protein identified in all datasets is 'O94855 (SEC24D)'.

```{r eval=F, include=F}
# Total 1158 potential Normal Tissue Proteins were identified in Method Development
# Normal Tissue DIA Proteomics from AG Klingmüller and AG Krijgsveld, and 892 of
# them are captured in Discovery dataset of AG Krijgsveld.

# Show predictive ability of captured potential proteins in Discovery dataset
featAssoRes_Discovery <- recurNormalFeatAssoTabs_Krij$tbl_featAsso1
# Extract only results of captured potential proteins
sigFeats_MethodDev <- unique(c(sigFeats_MethodDev_Klin, sigFeats_MethodDev_Krij))
sigFeats_MethodDev <- intersect(sigFeats_MethodDev, featAssoRes_Discovery$Feature)
featAssoRes_Discovery %>%
  dplyr::filter(Feature %in% sigFeats_MethodDev) %>%
  dplyr::select(Feature, pVal, pValAdj, Gene) %>%
  DT::datatable()
```

```{r eval=F, include=F}
# Focus on AG Klingmüller's dataset to exclude working group differences.

# Show predictive ability of potential proteins identified in MethodDev dataset in Discovery one
featAssoRes_Discovery <- recurNormalFeatAssoTabs_Klin$tbl_featAsso1
# Extract only results of potential proteins identified in MethodDev dataset
common_sigFeats <- unique(c(sigFeats_Discovery_Klin, sigFeats_MethodDev_Klin))
featAssoRes_Discovery %>%
  dplyr::filter(Feature %in% common_sigFeats) %>%
  dplyr::select(Feature, pVal, pValAdj, Gene) %>%
  DT::datatable()
```

## Method Development
Compare data from same patient cohorts (only common samples used) between different
working groups to see if better results can be obtained due to large variation among patients

**p-values (significance level = 0.05)**
```{r}
# Load normalized data
proTissue_Krij <- readRDS('./data/MethodDev/AG_Krijgsveld/new_proTissueVsn.rds')
proTissue_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsn.rds')
proTissue_Klin <- proTissue_Klin[, colData(proTissue_Klin)$Cohort == 'MethodDev']
# Keep only common samples shared between Discovery and Method Development data
cmnSmps <- intersect(colnames(proTissue_Krij), colnames(proTissue_Klin))
proTissue_Krij <- proTissue_Krij[, cmnSmps]
proTissue_Klin <- proTissue_Klin[, cmnSmps]
# Prepare protein-gene table for providing gene names
featInfo <- tibble::as_tibble(rowData(proTissue_Klin), rownames = 'Proteins') %>%
  dplyr::rename(Genes = PG.Genes) %>%
  dplyr::mutate(Proteins = stringr::str_remove(Proteins, ';.*'),
                Genes = stringr::str_remove(Genes, ';.*'))

# Prepare summarized tables of comparison between Recurrence and Non-recurrence patient groups
recurNormalFeatAssoTabs_MethodDev <- summarizeFeatAssoRes(proTissue_Krij, proTissue_Klin,
                                                          'Krij.', 'Klin.',
                                                          meta_var = 'Recurrence',
                                                          subset = c('Condition', 'Normal'),
                                                          method = 'proDA',
                                                          prot_gene_tbl = featInfo,
                                                          alpha = 0.05)
featAssoTabs <- recurNormalFeatAssoTabs_MethodDev

# Plot p-values
ggplot(featAssoTabs$tbl_pVal, aes(x=-log10(pVal_dat1), y=-log10(pVal_dat2), col=Significance)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.8) +
  scale_color_manual(values = c('red', 'darkgreen', 'navy', 'grey')) +
  labs(x = '-log10(pVal) of Krij.', y = '-log10(pVal) of Klin.',
       title = 'Recurrence: Normal Tissue DIA Proteomics (MethodDev)') +
  th
# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_pVal$Significance == 'Sig. in Both') /
                   sum(featAssoTabs$tbl_pVal$Significance != 'Not Sig.'), 3) * 100
cat(paste0('=> ', percent, '% of significant results are consistent'))
```

**log(FC)**
```{r}
# Plot log(FC)
ggplot(featAssoTabs$tbl_logFC, aes(x=logFC_dat1, y=logFC_dat2, col=Direction)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.8) +
  scale_color_manual(values = c('red', 'navy', 'grey')) +
  labs(x = 'log(FC) of Krij.', y = 'log(FC) of Klin.',
       title = 'Recurrence: Normal Tissue DIA Proteomics (MethodDev)') +
  th
# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_logFC$Direction != 'Inconsistent') /
                   nrow(featAssoTabs$tbl_logFC), 3) * 100
cat(paste0('=> ', percent, '% of results are consistent'))
```

## Discovery
Compare data from same patient cohorts (only common samples used) between different
working groups to see if better results can be obtained due to large variation among patients

**p-values (significance level = 0.05)**
```{r}
# Load normalized data
proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/proTissueVsn.rds')
proTissue_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsn.rds')
proTissue_Klin <- proTissue_Klin[, colData(proTissue_Klin)$Cohort == 'Discovery']
# Keep only common samples shared between Discovery and Method Development data
cmnSmps <- intersect(colnames(proTissue_Krij), colnames(proTissue_Klin))
proTissue_Krij <- proTissue_Krij[, cmnSmps]
proTissue_Klin <- proTissue_Klin[, cmnSmps]
# Prepare protein-gene table for providing gene names
featInfo <- tibble::as_tibble(rowData(proTissue_Klin), rownames = 'Proteins') %>%
  dplyr::rename(Genes = PG.Genes) %>%
  dplyr::mutate(Proteins = stringr::str_remove(Proteins, ';.*'),
                Genes = stringr::str_remove(Genes, ';.*'))

# Prepare summarized tables of comparison between Recurrence and Non-recurrence patient groups
recurNormalFeatAssoTabs_Discovery <- summarizeFeatAssoRes(proTissue_Krij, proTissue_Klin,
                                                          'Krij.', 'Klin.',
                                                          meta_var = 'Recurrence',
                                                          subset = c('Condition', 'Normal'),
                                                          method = 'proDA',
                                                          prot_gene_tbl = featInfo,
                                                          alpha = 0.05)
featAssoTabs <- recurNormalFeatAssoTabs_Discovery

# Plot p-values
ggplot(featAssoTabs$tbl_pVal, aes(x=-log10(pVal_dat1), y=-log10(pVal_dat2), col=Significance)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.8) +
  scale_color_manual(values = c('red', 'darkgreen', 'navy', 'grey')) +
  labs(x = '-log10(pVal) of Krij.', y = '-log10(pVal) of Klin.',
       title = 'Recurrence: Normal Tissue DIA Proteomics (Discovery)') +
  th
# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_pVal$Significance == 'Sig. in Both') /
                   sum(featAssoTabs$tbl_pVal$Significance != 'Not Sig.'), 3) * 100
cat(paste0('=> ', percent, '% of significant results are consistent'))
```

**log(FC)**
```{r}
# Plot log(FC)
ggplot(featAssoTabs$tbl_logFC, aes(x=logFC_dat1, y=logFC_dat2, col=Direction)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.8) +
  scale_color_manual(values = c('red', 'navy', 'grey')) +
  labs(x = 'log(FC) of Krij.', y = 'log(FC) of Klin.',
       title = 'Recurrence: Normal Tissue DIA Proteomics (Discovery)') +
  th
# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_logFC$Direction != 'Inconsistent') /
                   nrow(featAssoTabs$tbl_logFC), 3) * 100
cat(paste0('=> ', percent, '% of results are consistent'))
```

## Combined data
Combined Discovery and Method Development data (only common samples used)

**p-values (significance level = 0.05)**
```{r}
# Load normalized data
proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsn.rds')
proTissue_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsn.rds')
# Keep only common samples shared between Discovery and Method Development data
cmnSmps <- intersect(colnames(proTissue_Krij), colnames(proTissue_Klin))
proTissue_Krij <- proTissue_Krij[, cmnSmps]
proTissue_Klin <- proTissue_Klin[, cmnSmps]
# Prepare protein-gene table for providing gene names
featInfo <- tibble::as_tibble(rowData(proTissue_Klin), rownames = 'Proteins') %>%
  dplyr::rename(Genes = PG.Genes) %>%
  dplyr::mutate(Proteins = stringr::str_remove(Proteins, ';.*'),
                Genes = stringr::str_remove(Genes, ';.*'))

# Prepare summarized tables of comparison between Recurrence and Non-recurrence patient groups
recurNormalFeatAssoTabs_Combined <- summarizeFeatAssoRes(proTissue_Krij, proTissue_Klin,
                                                         'Krij.', 'Klin.',
                                                         meta_var = 'Recurrence',
                                                         method = 'proDA',
                                                         unwantVar1 = 'Cohort',
                                                         unwantVar2 = 'Cohort',
                                                         subset = c('Condition', 'Normal'),
                                                         prot_gene_tbl = featInfo,
                                                         alpha = 0.05)
featAssoTabs <- recurNormalFeatAssoTabs_Combined

# Plot p-values
ggplot(featAssoTabs$tbl_pVal, aes(x=-log10(pVal_dat1), y=-log10(pVal_dat2),
                                  col=Significance, label=Label)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.5) +
  scale_color_manual(values = c('red', 'darkgreen', 'navy', 'grey')) +
  geom_text_repel(size = 4, show.legend = F) +
  labs(x = '-log10(pVal) of Krij.', y = '-log10(pVal) of Klin.',
       title = 'Recurrence: Normal Tissue DIA Proteomics (Combined data)') +
  th

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_pVal$Significance == 'Sig. in Both') /
                   sum(featAssoTabs$tbl_pVal$Significance != 'Not Sig.'), 3) * 100
cat(paste0('=> ', percent, '% of significant results are consistent'))
# With all measured samples, around 8%
```

**log(FC)**
```{r}
# Plot log(FC)
ggplot(featAssoTabs$tbl_logFC, aes(x=logFC_dat1, y=logFC_dat2, col=Direction)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.5) +
  scale_color_manual(values = c('red', 'navy', 'grey')) +
  labs(x = 'log(FC) of Krij.', y = 'log(FC) of Klin.',
       title = 'Recurrence: Normal Tissue DIA Proteomics (Combined data)') +
  th

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_logFC$Direction != 'Inconsistent') /
                   nrow(featAssoTabs$tbl_logFC), 3) * 100
cat(paste0('=> ', percent, '% of results are consistent'))
```




# Tumor Tissue Proteomics
Check association of recurrence-related feature significance between Discovery and
Method Development datasets or between Krij.'s and Klin.'s datasets

## AG Krijgsveld

Display association of feature recurrence predictive power between Discovery and
Method Development Tumor Tissue DIA Proteomics generated by AG Krijgsveld\
**p-values (significance level = 0.05)**
```{r}
# Load normalized data
proTissue_Discovery <- readRDS('./data/Discovery/AG_Krijgsveld/proTissueVsn.rds')
proTissue_MethodDev <- readRDS('./data/MethodDev/AG_Krijgsveld/new_proTissueVsn.rds')
# Prepare protein-gene table for providing gene names
featInfo <- tibble::as_tibble(rowData(proTissue_Discovery), rownames = 'Proteins') %>%
  dplyr::select(Proteins, Genes) %>%
  dplyr::mutate(Proteins = stringr::str_remove(Proteins, ';.*'),
                Genes = stringr::str_remove(Genes, ';.*'))

# Prepare summarized tables of comparison between Recurrence and Non-recurrence patient groups
recurTumorFeatAssoTabs_Krij <- summarizeFeatAssoRes(proTissue_Discovery, proTissue_MethodDev,
                                                    'Discovery', 'MethodDev',
                                                    meta_var = 'Recurrence',
                                                    subset = c('Condition', 'Tumor'),
                                                    prot_gene_tbl = featInfo,
                                                    method = 'proDA',
                                                    alpha = 0.05)
featAssoTabs <- recurTumorFeatAssoTabs_Krij

# Plot p-values
ggplot(featAssoTabs$tbl_pVal, aes(x=-log10(pVal_dat1), y=-log10(pVal_dat2),
                                  col=Significance, label=Label)) +
  geom_point(size = 2) +
  scale_color_manual(values = c('red', 'darkgreen', 'navy', 'grey')) +
  geom_text_repel(size = 4, show.legend = F) +
  labs(x = '-log10(pVal) of Discovery Cohort', y = '-log10(pVal) of MethodDev Cohort',
       title = 'Recurrence: Tumor Tissue DIA Proteomics (AG Krijgsveld)') +
  th

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_pVal$Significance == 'Sig. in Both') /
                   sum(featAssoTabs$tbl_pVal$Significance != 'Not Sig.'), 3) * 100
cat(paste0('=> ', percent, '% of significant results are consistent'))
```

**log(FC)**
```{r}
# Plot log(FC)
ggplot(featAssoTabs$tbl_logFC, aes(x=logFC_dat1, y=logFC_dat2, col=Direction)) +
  geom_point(size = 2) +
  scale_color_manual(values = c('red', 'navy', 'grey')) +
  labs(x = 'log(FC) of Discovery Cohort', y = 'log(FC) of MethodDev Cohort',
       title = 'Recurrence: Tumor Tissue DIA Proteomics (AG Krijgsveld)') +
  th

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_logFC$Direction != 'Inconsistent') /
                   nrow(featAssoTabs$tbl_logFC), 3) * 100
cat(paste0('=> ', percent, '% of results are consistent'))
```

Visualize top significant proteins identified in both Discovery and Method Development
datasets and based on Discovery one. Note that Method Development data is provided
to have more comprehensive analysis and it may not be on same scale as Discovery data.\
ns: p > 0.05, ∗: p <= 0.05, ∗∗: p <= 0.01, ***: p <= 0.001
```{r fig.height=12}
# Plot top significant features based on 'data1'
topSigFeatTab <- dplyr::mutate(featAssoTabs$tbl_topSigFeats,
                               newGroup = factor(newGroup, levels = c('Yes_Discovery',
                                                                      'No_Discovery',
                                                                      'Yes_MethodDev',
                                                                      'No_MethodDev')))
ggplot(topSigFeatTab, aes(x=newGroup, y=Abundance, col=DataName, fill=Recurrence)) +
  geom_boxplot(alpha = 1, outlier.shape = NA, linewidth = 1) +
  geom_jitter(position = position_jitter(0.2), size = 3, show.legend = F) +
  ggpubr::stat_compare_means(method = 't.test', paired = F, method.args = list(var.equal = T),
                             comparisons = list(c('Yes_Discovery', 'No_Discovery'),
                                                c('Yes_MethodDev', 'No_MethodDev')),
                             label = 'p.signif', tip.length = 0.015, bracket.size = 0.7, size = 4) +
  labs(x = 'Recurrence', title = 'Recurrence: Tumor Tissue DIA Proteomics (AG Krijgsveld)') +
  scale_color_manual(name = 'Cohort', values = c('firebrick', 'grey50')) +
  scale_fill_manual(values=c('#00BFC4', '#F8766D')) +
  facet_wrap(vars(newFeature), scales = 'free') +
  th +
  theme(strip.text = element_text(size = 13, face = 'bold'),
        axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1))
```

## AG Klingmüller

Display association of feature recurrence predictive power between Discovery and
Method Development Tumor Tissue DIA Proteomics generated by AG Klingmüller\
**p-values (significance level = 0.05)**
```{r}
# Load normalized data
proTissue_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsn.rds')
proTissue_Discovery <- proTissue_Klin[, colData(proTissue_Klin)$Cohort == 'Discovery']
proTissue_MethodDev <- proTissue_Klin[, colData(proTissue_Klin)$Cohort == 'MethodDev']
# Prepare protein-gene table for providing gene names
featInfo <- tibble::as_tibble(rowData(proTissue_Discovery), rownames = 'Proteins') %>%
  dplyr::rename(Genes = PG.Genes) %>%
  dplyr::mutate(Proteins = stringr::str_remove(Proteins, ';.*'),
                Genes = stringr::str_remove(Genes, ';.*'))

# Prepare summarized tables of comparison between Recurrence and Non-recurrence patient groups
recurTumorFeatAssoTabs_Klin <- summarizeFeatAssoRes(proTissue_Discovery, proTissue_MethodDev,
                                                    'Discovery', 'MethodDev',
                                                    meta_var = 'Recurrence',
                                                    subset = c('Condition', 'Tumor'),
                                                    prot_gene_tbl = featInfo,
                                                    method = 'proDA',
                                                    alpha = 0.05)
featAssoTabs <- recurTumorFeatAssoTabs_Klin

# Plot p-values
ggplot(featAssoTabs$tbl_pVal, aes(x=-log10(pVal_dat1), y=-log10(pVal_dat2),
                                  col=Significance, label=Label)) +
  geom_point(size = 2) +
  scale_color_manual(values = c('red', 'darkgreen', 'navy', 'grey')) +
  geom_text_repel(size = 4, show.legend = F) +
  labs(x = '-log10(pVal) of Discovery Cohort', y = '-log10(pVal) of MethodDev Cohort',
       title = 'Recurrence: Tumor Tissue DIA Proteomics (AG Klingmüller)') +
  th

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_pVal$Significance == 'Sig. in Both') /
                   sum(featAssoTabs$tbl_pVal$Significance != 'Not Sig.'), 3) * 100
cat(paste0('=> ', percent, '% of significant results are consistent'))
```

**log(FC)**
```{r}
# Plot log(FC)
ggplot(featAssoTabs$tbl_logFC, aes(x=logFC_dat1, y=logFC_dat2, col=Direction)) +
  geom_point(size = 2) +
  scale_color_manual(values = c('red', 'navy', 'grey')) +
  labs(x = 'log(FC) of Discovery Cohort', y = 'log(FC) of MethodDev Cohort',
       title = 'Recurrence: Tumor Tissue DIA Proteomics (AG Klingmüller)') +
  th

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_logFC$Direction != 'Inconsistent') /
                   nrow(featAssoTabs$tbl_logFC), 3) * 100
cat(paste0('=> ', percent, '% of results are consistent'))
```

Visualize top significant proteins identified in both Discovery and Method Development
datasets and based on Discovery one. Note that Method Development data is provided
to have more comprehensive analysis and it may not be on same scale as Discovery data.\
ns: p > 0.05, ∗: p <= 0.05, ∗∗: p <= 0.01, ***: p <= 0.001
```{r}
# Plot top significant features based on 'data1'
topSigFeatTab <- dplyr::mutate(featAssoTabs$tbl_topSigFeats,
                               newGroup = factor(newGroup, levels = c('Yes_Discovery',
                                                                      'No_Discovery',
                                                                      'Yes_MethodDev',
                                                                      'No_MethodDev')))
ggplot(topSigFeatTab, aes(x=newGroup, y=Abundance, col=DataName, fill=Recurrence)) +
  geom_boxplot(alpha = 1, outlier.shape = NA, linewidth = 1) +
  geom_jitter(position = position_jitter(0.2), size = 3, show.legend = F) +
  ggpubr::stat_compare_means(method = 't.test', paired = F, method.args = list(var.equal = T),
                             comparisons = list(c('Yes_Discovery', 'No_Discovery'),
                                                c('Yes_MethodDev', 'No_MethodDev')),
                             label = 'p.signif', tip.length = 0.015, bracket.size = 0.7, size = 4) +
  labs(x = 'Recurrence', title = 'Recurrence: Tumor Tissue DIA Proteomics (AG Klingmüller)') +
  scale_color_manual(name = 'Cohort', values = c('firebrick', 'grey50')) +
  scale_fill_manual(values=c('#00BFC4', '#F8766D')) +
  facet_wrap(vars(newFeature), scales = 'free') +
  th +
  theme(strip.text = element_text(size = 13, face = 'bold'),
        axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1))
```

## Integrative analysis

Show intersections of potential Tumor proteins identified in different datasets
(significance level = 0.05)
```{r}
# Prepare recurrence-related significant features
alpha <- 0.05
# AG Klingmüller
sigFeats_Discovery_Klin <- recurTumorFeatAssoTabs_Klin$tbl_featAsso1 %>%
  dplyr::filter(pVal <= alpha) %>%
  dplyr::pull(Feature)
sigFeats_MethodDev_Klin <- recurTumorFeatAssoTabs_Klin$tbl_featAsso2 %>%
  dplyr::filter(pVal <= alpha) %>%
  dplyr::pull(Feature)
# AG Krijgsveld
sigFeats_Discovery_Krij <- recurTumorFeatAssoTabs_Krij$tbl_featAsso1 %>%
  dplyr::filter(pVal <= alpha) %>%
  dplyr::pull(Feature)
sigFeats_MethodDev_Krij <- recurTumorFeatAssoTabs_Krij$tbl_featAsso2 %>%
  dplyr::filter(pVal <= alpha) %>%
  dplyr::pull(Feature)

# Make venn graphs
ggvenn(list(Discovery_Krij. = sigFeats_Discovery_Krij, MethodDev_Krij. = sigFeats_MethodDev_Krij,
            MethodDev_Klin. = sigFeats_MethodDev_Klin),
       fill_color = c('darkred', 'darkblue', 'darkgreen'), fill_alpha = 0.6,
       set_name_color = c('darkred', 'darkblue', 'darkgreen'), set_name_size = 9,
       text_size = 5, stroke_size = 1.5)
```

```{r}
ggvenn(list(Discovery_Klin. = sigFeats_Discovery_Klin, MethodDev_Klin. = sigFeats_MethodDev_Klin,
            MethodDev_Krij. = sigFeats_MethodDev_Krij),
       fill_color = c('darkorange', 'darkgreen', 'darkblue'), fill_alpha = 0.6,
       set_name_color = c('darkorange', 'darkgreen', 'darkblue'), set_name_size = 9,
       text_size = 5, stroke_size = 1.5)
```

```{r}
ggvenn(list(Discovery_Krij. = sigFeats_Discovery_Krij, Discovery_Klin. = sigFeats_Discovery_Klin,
            MethodDev_Klin. = sigFeats_MethodDev_Klin),
       fill_color = c('darkred', 'darkorange', 'darkgreen'), fill_alpha = 0.6,
       set_name_color = c('darkred', 'darkorange', 'darkgreen'), set_name_size = 9,
       text_size = 5, stroke_size = 1.5)
```

## Method Development
Compare data from same patient cohorts (only common samples used) between different
working groups to see if better results can be obtained due to large variation among patients

**p-values (significance level = 0.05)**
```{r}
# Load normalized data
proTissue_Krij <- readRDS('./data/MethodDev/AG_Krijgsveld/new_proTissueVsn.rds')
proTissue_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsn.rds')
proTissue_Klin <- proTissue_Klin[, colData(proTissue_Klin)$Cohort == 'MethodDev']
# Keep only common samples shared between Discovery and Method Development data
cmnSmps <- intersect(colnames(proTissue_Krij), colnames(proTissue_Klin))
proTissue_Krij <- proTissue_Krij[, cmnSmps]
proTissue_Klin <- proTissue_Klin[, cmnSmps]
# Prepare protein-gene table for providing gene names
featInfo <- tibble::as_tibble(rowData(proTissue_Klin), rownames = 'Proteins') %>%
  dplyr::rename(Genes = PG.Genes) %>%
  dplyr::mutate(Proteins = stringr::str_remove(Proteins, ';.*'),
                Genes = stringr::str_remove(Genes, ';.*'))

# Prepare summarized tables of comparison between Recurrence and Non-recurrence patient groups
recurTumorFeatAssoTabs_MethodDev <- summarizeFeatAssoRes(proTissue_Krij, proTissue_Klin,
                                                         'Krij.', 'Klin.',
                                                         meta_var = 'Recurrence',
                                                         subset = c('Condition', 'Tumor'),
                                                         method = 'proDA',
                                                         prot_gene_tbl = featInfo,
                                                         alpha = 0.05)
featAssoTabs <- recurTumorFeatAssoTabs_MethodDev

# Plot p-values
ggplot(featAssoTabs$tbl_pVal, aes(x=-log10(pVal_dat1), y=-log10(pVal_dat2), col=Significance)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.8) +
  scale_color_manual(values = c('red', 'darkgreen', 'navy', 'grey')) +
  labs(x = '-log10(pVal) of Krij.', y = '-log10(pVal) of Klin.',
       title = 'Recurrence: Tumor Tissue DIA Proteomics (MethodDev)') +
  th
# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_pVal$Significance == 'Sig. in Both') /
                   sum(featAssoTabs$tbl_pVal$Significance != 'Not Sig.'), 3) * 100
cat(paste0('=> ', percent, '% of significant results are consistent'))
```

**log(FC)**
```{r}
# Plot log(FC)
ggplot(featAssoTabs$tbl_logFC, aes(x=logFC_dat1, y=logFC_dat2, col=Direction)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.8) +
  scale_color_manual(values = c('red', 'navy', 'grey')) +
  labs(x = 'log(FC) of Krij.', y = 'log(FC) of Klin.',
       title = 'Recurrence: Tumor Tissue DIA Proteomics (MethodDev)') +
  th
# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_logFC$Direction != 'Inconsistent') /
                   nrow(featAssoTabs$tbl_logFC), 3) * 100
cat(paste0('=> ', percent, '% of results are consistent'))
```

## Discovery
Compare data from same patient cohorts (only common samples used) between different
working groups to see if better results can be obtained due to large variation among patients

**p-values (significance level = 0.05)**
```{r}
# Load normalized data
proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/proTissueVsn.rds')
proTissue_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsn.rds')
proTissue_Klin <- proTissue_Klin[, colData(proTissue_Klin)$Cohort == 'Discovery']
# Keep only common samples shared between Discovery and Method Development data
cmnSmps <- intersect(colnames(proTissue_Krij), colnames(proTissue_Klin))
proTissue_Krij <- proTissue_Krij[, cmnSmps]
proTissue_Klin <- proTissue_Klin[, cmnSmps]
# Prepare protein-gene table for providing gene names
featInfo <- tibble::as_tibble(rowData(proTissue_Klin), rownames = 'Proteins') %>%
  dplyr::rename(Genes = PG.Genes) %>%
  dplyr::mutate(Proteins = stringr::str_remove(Proteins, ';.*'),
                Genes = stringr::str_remove(Genes, ';.*'))

# Prepare summarized tables of comparison between Recurrence and Non-recurrence patient groups
recurTumorFeatAssoTabs_Discovery <- summarizeFeatAssoRes(proTissue_Krij, proTissue_Klin,
                                                         'Krij.', 'Klin.',
                                                         meta_var = 'Recurrence',
                                                         subset = c('Condition', 'Tumor'),
                                                         method = 'proDA',
                                                         prot_gene_tbl = featInfo,
                                                         alpha = 0.05)
featAssoTabs <- recurTumorFeatAssoTabs_Discovery

# Plot p-values
ggplot(featAssoTabs$tbl_pVal, aes(x=-log10(pVal_dat1), y=-log10(pVal_dat2), col=Significance)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.8) +
  scale_color_manual(values = c('red', 'darkgreen', 'navy', 'grey')) +
  labs(x = '-log10(pVal) of Krij.', y = '-log10(pVal) of Klin.',
       title = 'Recurrence: Tumor Tissue DIA Proteomics (Discovery)') +
  th
# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_pVal$Significance == 'Sig. in Both') /
                   sum(featAssoTabs$tbl_pVal$Significance != 'Not Sig.'), 3) * 100
cat(paste0('=> ', percent, '% of significant results are consistent'))
```

**log(FC)**
```{r}
# Plot log(FC)
ggplot(featAssoTabs$tbl_logFC, aes(x=logFC_dat1, y=logFC_dat2, col=Direction)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.8) +
  scale_color_manual(values = c('red', 'navy', 'grey')) +
  labs(x = 'log(FC) of Krij.', y = 'log(FC) of Klin.',
       title = 'Recurrence: Tumor Tissue DIA Proteomics (Discovery)') +
  th
# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_logFC$Direction != 'Inconsistent') /
                   nrow(featAssoTabs$tbl_logFC), 3) * 100
cat(paste0('=> ', percent, '% of results are consistent'))
```

## Combined data
Combined Discovery and Method Development data (only common samples used)

**p-values (significance level = 0.05)**
```{r}
# Load normalized data
proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsn.rds')
proTissue_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsn.rds')
# Keep only common samples shared between Discovery and Method Development data
cmnSmps <- intersect(colnames(proTissue_Krij), colnames(proTissue_Klin))
proTissue_Krij <- proTissue_Krij[, cmnSmps]
proTissue_Klin <- proTissue_Klin[, cmnSmps]
# Prepare protein-gene table for providing gene names
featInfo <- tibble::as_tibble(rowData(proTissue_Klin), rownames = 'Proteins') %>%
  dplyr::rename(Genes = PG.Genes) %>%
  dplyr::mutate(Proteins = stringr::str_remove(Proteins, ';.*'),
                Genes = stringr::str_remove(Genes, ';.*'))

# Prepare summarized tables of comparison between Recurrence and Non-recurrence patient groups
recurTumorFeatAssoTabs_Combined <- summarizeFeatAssoRes(proTissue_Krij, proTissue_Klin,
                                                        'Krij.', 'Klin.',
                                                        meta_var = 'Recurrence',
                                                        method = 'proDA',
                                                        unwantVar1 = 'Cohort',
                                                        unwantVar2 = 'Cohort',
                                                        subset = c('Condition', 'Tumor'),
                                                        prot_gene_tbl = featInfo,
                                                        alpha = 0.05)
featAssoTabs <- recurTumorFeatAssoTabs_Combined

# Plot p-values
ggplot(featAssoTabs$tbl_pVal, aes(x=-log10(pVal_dat1), y=-log10(pVal_dat2),
                                  col=Significance, label=Label)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.5) +
  scale_color_manual(values = c('red', 'darkgreen', 'navy', 'grey')) +
  geom_text_repel(size = 4, show.legend = F) +
  labs(x = '-log10(pVal) of Krij.', y = '-log10(pVal) of Klin.',
       title = 'Recurrence: Tumor Tissue DIA Proteomics (Combined data)') +
  th

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_pVal$Significance == 'Sig. in Both') /
                   sum(featAssoTabs$tbl_pVal$Significance != 'Not Sig.'), 3) * 100
cat(paste0('=> ', percent, '% of significant results are consistent'))
# With all measured samples, around 3%
```

**log(FC)**
```{r}
# Plot log(FC)
ggplot(featAssoTabs$tbl_logFC, aes(x=logFC_dat1, y=logFC_dat2, col=Direction)) +
  geom_point(size = 2) +
  geom_abline(intercept = 0, slope = 1, col = 'black', linewidth = 0.5) +
  scale_color_manual(values = c('red', 'navy', 'grey')) +
  labs(x = 'log(FC) of Krij.', y = 'log(FC) of Klin.',
       title = 'Recurrence: Tumor Tissue DIA Proteomics (Combined data)') +
  th

# Show percentage of consistent results
percent <- round(sum(featAssoTabs$tbl_logFC$Direction != 'Inconsistent') /
                   nrow(featAssoTabs$tbl_logFC), 3) * 100
cat(paste0('=> ', percent, '% of results are consistent'))
```
