---
title: "Identification of potential recurrence-related proteins"
author: "Qian-Wu Liao"
output:
  workflowr::wflow_html:
    toc: yes
    toc_float: yes
    code_folding: hide
---

<font size='4'> Description: </font>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 8, fig.width = 10, warning = F)
knitr::opts_knit$set(root.dir = '/Users/qianwu/Desktop/SMART-CARE_LungCancer')
```

Load libraries
```{r library loading, message = F}
library('pcaMethods')
library('limma')
library('proDA')
library('pheatmap')
library('ggplotify')
library('ggrepel')
library('ggvenn')
library('AnnotationDbi')
library('org.Hs.eg.db')
# This annotation object is accessed using AnnotationDbi:select()
hs <- org.Hs.eg.db
library('msigdbr')
library('clusterProfiler')
library('SummarizedExperiment')
library('tidyverse')
# Load user-defined functions
source('./code/analysis_pipeline.R')
source('./code/misc.R')

# Set plot theme
th <- theme_bw(base_size = 15) +
    theme(axis.title = element_text(face = 'bold'),
          axis.text = element_text(face = 'bold'),
          axis.ticks = element_line(linewidth = 0.8),
          legend.text = element_text(size = 15))
```

# NTP (AG Krij.)
Normal Tissue DIA Proteomics from AG Krijgsveld, including MethodDev, Discovery,
and Combined datasets

## Significant Proteins

Identify significant proteins using t-test
```{r}
####################################
# Combined (MethodDev + Discovery) #
####################################
# # Load preprocessed data
# Combined_proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds')
# # Prepare feature annotation table
# featAnnoTab <- tibble::as_tibble(rowData(Combined_proTissue_Krij), rownames = 'Protein') %>%
#   dplyr::select(Protein, Genes) %>%
#   dplyr::rename(Gene = Genes) %>%
#   tibble::column_to_rownames('Protein')
# 
# # Subset certain samples
# Combined_proNormal_Krij <- Combined_proTissue_Krij[, colData(Combined_proTissue_Krij)$Condition %in% 'Normal']
# # Perform single-omics analysis
# Combined_proNormalRes_Krij <- doSingleOmicsAnalysis(Combined_proNormal_Krij, soa_metaVar = 'Recurrence',
#                                                     num_sigFeats = 6, pca_method = 'ppca',
#                                                     num_PCs = 40, num_PCfeats = 30,
#                                                     use_limma = F, use_proDA = T,
#                                                     feat_anno = featAnnoTab, feat_conv = T,
#                                                     plot_title = NULL, show_rownames = F,
#                                                     pca_metaVar = c('Recurrence', 'Gender', 'Age',
#                                                                     'Smoking', 'Stage', 'Adjuvant'))
# saveRDS(Combined_proNormalRes_Krij, './data/Discovery/potential_proteins/Combined_proNormalRes_Krij.rds')
# 
# # AG Klingmüller
# # Load preprocessed data
# Combined_proTissue_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsnBC.rds')
# # Prepare feature annotation table
# featAnnoTab <- tibble::as_tibble(rowData(Combined_proTissue_Klin), rownames = 'Protein') %>%
#   dplyr::rename(Gene = PG.Genes) %>%
#   tibble::column_to_rownames('Protein')
# 
# # Subset certain samples
# Combined_proNormal_Klin <- Combined_proTissue_Klin[, colData(Combined_proTissue_Klin)$Condition %in% 'Normal']
# # Perform single-omics analysis
# Combined_proNormalRes_Klin <- doSingleOmicsAnalysis(Combined_proNormal_Klin, soa_metaVar = 'Recurrence',
#                                                     num_sigFeats = 6, pca_method = 'ppca',
#                                                     num_PCs = 37, num_PCfeats = 30,
#                                                     use_limma = F, use_proDA = T,
#                                                     feat_anno = featAnnoTab, feat_conv = T,
#                                                     plot_title = NULL, show_rownames = F,
#                                                     pca_metaVar = c('Recurrence', 'Gender', 'Age',
#                                                                     'Smoking', 'Stage', 'Adjuvant'),
#                                                     soa_unwantVar = 'Adjuvant')
# saveRDS(Combined_proNormalRes_Klin, './data/Discovery/potential_proteins/Combined_proNormalRes_Klin.rds')
# 
# #############
# # MethodDev #
# #############
# # Load normalized data
# MethodDev_proTissue_Krij <- readRDS('./data/MethodDev/AG_Krijgsveld/new_proTissueVsn.rds')
# # Prepare feature annotation table
# featAnnoTab <- tibble::as_tibble(rowData(MethodDev_proTissue_Krij), rownames = 'Protein') %>%
#   dplyr::select(Protein, Genes) %>%
#   dplyr::rename(Gene = Genes) %>%
#   tibble::column_to_rownames('Protein')
# 
# # Subset certain samples
# MethodDev_proNormal_Krij <- MethodDev_proTissue_Krij[, colData(MethodDev_proTissue_Krij)$Condition %in% 'Normal']
# # Perform single-omics analysis
# MethodDev_proNormalRes_Krij <- doSingleOmicsAnalysis(MethodDev_proNormal_Krij, soa_metaVar = 'Recurrence',
#                                                      num_sigFeats = 6, pca_method = 'ppca',
#                                                      num_PCs = 18, num_PCfeats = 30,
#                                                      use_limma = F, use_proDA = T,
#                                                      feat_anno = featAnnoTab, feat_conv = T,
#                                                      plot_title = NULL, show_rownames = F,
#                                                      pca_metaVar = c('Recurrence', 'Gender', 'Age',
#                                                                      'Smoking', 'Stage', 'Adjuvant'))
# saveRDS(MethodDev_proNormalRes_Krij, './data/Discovery/potential_proteins/MethodDev_proNormalRes_Krij.rds')
# 
# #############
# # Discovery #
# #############
# # Load normalized data
# Discovery_proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/proTissueVsn.rds')
# # Prepare feature annotation table
# featAnnoTab <- tibble::as_tibble(rowData(Discovery_proTissue_Krij), rownames = 'Protein') %>%
#   dplyr::select(Protein, Genes) %>%
#   dplyr::rename(Gene = Genes) %>%
#   tibble::column_to_rownames('Protein')
# 
# # Subset certain samples
# Discovery_proNormal_Krij <- Discovery_proTissue_Krij[, colData(Discovery_proTissue_Krij)$Condition %in% 'Normal']
# # Perform single-omics analysis
# Discovery_proNormalRes_Krij <- doSingleOmicsAnalysis(Discovery_proNormal_Krij, soa_metaVar = 'Recurrence',
#                                                      num_sigFeats = 6, pca_method = 'ppca',
#                                                      num_PCs = 40, num_PCfeats = 30,
#                                                      use_limma = F, use_proDA = T,
#                                                      feat_anno = featAnnoTab, feat_conv = T,
#                                                      plot_title = NULL, show_rownames = F,
#                                                      pca_metaVar = c('Recurrence', 'Gender', 'Age',
#                                                                      'Smoking', 'Stage', 'Adjuvant'))
# saveRDS(Discovery_proNormalRes_Krij, './data/Discovery/potential_proteins/Discovery_proNormalRes_Krij.rds')

# Load SOA results
Combined_proNormalRes_Krij <- readRDS('./data/Discovery/potential_proteins/Combined_proNormalRes_Krij.rds')
Combined_proNormalRes_Klin <- readRDS('./data/Discovery/potential_proteins/Combined_proNormalRes_Klin.rds')
MethodDev_proNormalRes_Krij <- readRDS('./data/Discovery/potential_proteins/MethodDev_proNormalRes_Krij.rds')
Discovery_proNormalRes_Krij <- readRDS('./data/Discovery/potential_proteins/Discovery_proNormalRes_Krij.rds')

# Retrieve significant features from SOA results
Combined_proNormalSigFeats_Krij <- Combined_proNormalRes_Krij$sig.feat.tab$Var1 %>%
  stringr::str_remove_all(';.+')
MethodDev_proNormalSigFeats_Krij <- MethodDev_proNormalRes_Krij$sig.feat.tab$Var1 %>%
  stringr::str_remove_all(';.+')
Discovery_proNormalSigFeats_Krij <- Discovery_proNormalRes_Krij$sig.feat.tab$Var1 %>%
  stringr::str_remove_all(';.+')
Combined_proNormalSigFeats_Klin <- Combined_proNormalRes_Klin$sig.feat.tab$Var1 %>%
  stringr::str_remove_all(';.+')
```

Visualize common significant proteins among datasets
```{r}
# Visualize common significant features among datasets
ggvenn(list(Combined_Krij. = Combined_proNormalSigFeats_Krij,
            Discovery_Krij. = Discovery_proNormalSigFeats_Krij,
            MethodDev_Krij. = MethodDev_proNormalSigFeats_Krij),
       fill_color = c('darkred', 'darkblue', 'darkgreen'), fill_alpha = 0.6,
       set_name_color = c('darkred', 'darkblue', 'darkgreen'), set_name_size = 9,
       text_size = 5, stroke_size = 1.5)
ggsave('./output/Discovery/group_meeting/venn_sigFeats_Krij_combined_methdev_dis.png',
       device = 'png', dpi = 400, height = 8, width = 11)

# Extract common significant features across datasets
# (MethodDev and Discovery Normal Tissue Proteomics shares few significant features)
cmnProNormalSigFeats_Krij <- intersect(Combined_proNormalSigFeats_Krij, MethodDev_proNormalSigFeats_Krij)
cmnProNormalSigFeats_Krij <- intersect(cmnProNormalSigFeats_Krij, Discovery_proNormalSigFeats_Krij)
# intersect(cmnProNormalSigFeats_Krij, Combined_proNormalSigFeats_Klin)

# Compare SOA results of combined data from AG Krijgsveld and that of AG Klingmüller
# ggvenn(list(Combined_Krij. = Combined_proNormalSigFeats_Krij,
#             Combined_Klin. = Combined_proNormalSigFeats_Klin),
#        fill_color = c('darkred', 'darkblue'), fill_alpha = 0.6,
#        set_name_color = c('darkred', 'darkblue'), set_name_size = 9,
#        text_size = 5, stroke_size = 1.5)

Combined_proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds')
# Prepare data matrix and metadata including both Tumor and Normal samples
datMat <- assay(Combined_proTissue_Krij)
smpAnnoTab <- colData(Combined_proTissue_Krij) %>%
  tibble::as_tibble(rownames = 'Sample')
# Prepare significant feature table
featSigAssoTab <- Combined_proNormalRes_Krij$sig.feat.tab
cmnFeatSigAssoTab <- featSigAssoTab[featSigAssoTab$Var1 %in% cmnProNormalSigFeats_Krij,]

num_sigFeats <- 6
gpLevel = c('Yes_Normal', 'No_Normal', 'Yes_Tumor', 'No_Tumor')
condiCol = c('firebrick', 'grey50') #c(Normal, Tumor)
# Extract top significant features and prepare needed information
topSigFeats <- cmnFeatSigAssoTab[1:num_sigFeats,] %>%
  dplyr::select(Var1, Gene) %>%
  dplyr::mutate(newVar1 = paste0(Var1, ' (', Gene, ')'),
                newVar1 = factor(newVar1, levels = unique(newVar1)))
topSigFeatDat <- tibble::as_tibble(datMat[topSigFeats$Var1,], rownames = 'Var1') %>%
  tidyr::pivot_longer(cols = -'Var1', names_to = 'Sample', values_to = 'Abundance') %>%
  dplyr::left_join(topSigFeats, by = 'Var1') %>%
  dplyr::left_join(smpAnnoTab, by = 'Sample') %>%
  dplyr::mutate(Recur_Condi = paste0(Recurrence, '_', Condition),
                Recur_Condi = factor(Recur_Condi, levels = gpLevel))
ggplot(topSigFeatDat, aes(x=Recur_Condi, y=Abundance, col=Condition, fill=Recurrence)) +
  geom_boxplot(alpha = 1, outlier.shape = NA, linewidth = 1) +
  geom_jitter(position = position_jitter(0.2), size = 2, show.legend = F) +
  ggpubr::stat_compare_means(method = 't.test', paired = F, method.args = list(var.equal = T),
                             comparisons = list(c('Yes_Normal', 'No_Normal'), c('Yes_Tumor', 'No_Tumor')),
                             label = 'p.signif', tip.length = 0.015, bracket.size = 0.7, size = 4) +
  labs(x = 'Recurrence') +
  scale_color_manual(values = condiCol) +
  scale_fill_manual(values=c('#00BFC4', '#F8766D')) +
  facet_wrap(vars(newVar1), scales = 'free') +
  th +
  # theme(strip.text = element_text(size = 13, face = 'bold'),
  #            axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
  theme(strip.text = element_text(size = 14, face = 'bold'),
        axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1, size = 20),
        axis.title = element_text(size = 26),
        legend.title = element_text(size = 24), legend.text = element_text(size = 22))
ggsave('./output/Discovery/group_meeting/combined_common_topSigFeats_proNormal_Krij.png',
       device = 'png', dpi = 400, height = 12, width = 14)
```

## Enrichment analysis
Combined data

```{r}
# Prepare ranked gene list
Combined_proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds')
Combined_proNormalRes_Krij <- readRDS('./data/Discovery/potential_proteins/Combined_proNormalRes_Krij.rds')
sigRes <- Combined_proNormalRes_Krij$SOA.res$featAssoRes %>%
  dplyr::select(Var1, Stat) %>%
  tibble::column_to_rownames('Var1')
featAnnoTab <- tibble::as_tibble(rowData(Combined_proTissue_Krij), rownames = 'Protein') %>%
  dplyr::select(Protein, Genes) %>%
  dplyr::rename(Gene = Genes) %>%
  tibble::column_to_rownames('Protein')

rankedGeneList <- prepRankedFeatList(statRes = sigRes, featAnno = featAnnoTab, to_gene = T)
```

```{r}
# Curated gene sets
# View supported species and collections in MSigDB
# msigdbr_species()
# msigdbr_collections()
# Retrieve annotated gene sets
msigTabC2 <- msigdbr(species = 'Homo sapiens', category = 'C2', subcategory = NULL) %>%
  dplyr::select(gs_name, human_gene_symbol)

# Run GSEA
gseaResC2 <- clusterProfiler::GSEA(geneList = rankedGeneList, TERM2GENE = msigTabC2,
                                   minGSSize = 10, maxGSSize = 500,
                                   pvalueCutoff = 0.05, pAdjustMethod = 'BH',
                                   by = 'fgsea', eps = 0)

# Plot enrichment analysis results
gseaRes4Plot <- gseaResC2@result %>%
  dplyr::select(Description, setSize, NES, pvalue, p.adjust, leading_edge) %>%
  dplyr::arrange(p.adjust) %>%
  dplyr::slice_head(n = 8) %>%
  dplyr::mutate(leading_edge = stringr::str_extract(leading_edge, '\\d\\d'),
                leading_edge = as.numeric(leading_edge),
                # Description = stringr::str_remove(Description, '^REACTOME_'),
                Recurrence = ifelse(test = NES > 0, yes = 'Yes', no = 'No')) %>%
  dplyr::rename(gene_ratio = leading_edge)
midGrad <- mean(gseaRes4Plot$p.adjust)

ggplot(gseaRes4Plot, aes(x=NES, y=reorder(Description, NES), fill=p.adjust, col=Recurrence)) +
  geom_bar(stat = 'identity', width = 0.7, size = 2) +
  scale_fill_gradient2(low='#5B1C00', high='#FFFCA5', mid = '#E86900', midpoint = midGrad, name = 'Adjusted pVal') +
  scale_color_manual(values = c(Yes = 'red', No = 'darkblue')) +
  labs(x = 'Normalized Enrichment Score', y = 'Pathway') +
  theme_minimal() +
  theme(axis.title = element_text(size = 16, face = 'bold'),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 10, face = 'bold'),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12))
```

```{r}
# Oncogenic signatures gene sets
# Retrieve annotated gene sets
msigTabC6 <- msigdbr(species = 'Homo sapiens', category = 'C6', subcategory = NULL) %>%
  dplyr::select(gs_name, human_gene_symbol)
# Retrieve descriptions of gene sets for interpretation
descripC6 <- msigdbr(species = 'Homo sapiens', category = 'C6', subcategory = NULL) %>%
  dplyr::select(gs_name, gs_description) %>%
  dplyr::filter(!duplicated(gs_name))

# Run GSEA
gseaResC6 <- clusterProfiler::GSEA(geneList = rankedGeneList, TERM2GENE = msigTabC6,
                                   minGSSize = 10, maxGSSize = 500,
                                   pvalueCutoff = 0.05, pAdjustMethod = 'BH',
                                   by = 'fgsea', eps = 0)

# Plot enrichment analysis results
gseaRes4Plot <- gseaResC6@result %>%
  dplyr::select(Description, setSize, NES, pvalue, p.adjust, leading_edge) %>%
  dplyr::arrange(p.adjust) %>%
  dplyr::slice_head(n = 8) %>%
  dplyr::mutate(leading_edge = stringr::str_extract(leading_edge, '\\d\\d'),
                leading_edge = as.numeric(leading_edge),
                Recurrence = ifelse(test = NES > 0, yes = 'Yes', no = 'No')) %>%
  dplyr::rename(gene_ratio = leading_edge)
midGrad <- mean(gseaRes4Plot$p.adjust)

# ggplot(gseaRes4Plot, aes(x=NES, y=reorder(Description, NES), fill=p.adjust, col=Recurrence)) +
#   geom_bar(stat = 'identity', width = 0.6, size = 2) +
#   scale_fill_gradient2(low='#5B1C00', high='#FFFCA5', mid = '#E86900', midpoint = midGrad, name = 'Adjusted pVal') +
#   scale_color_manual(values = c(Yes = 'red', No = 'darkblue')) +
#   labs(x = 'Normalized Enrichment Score', y = 'Pathway') +
#   theme_minimal() +
#   theme(axis.title = element_text(size = 16, face = 'bold'),
#         axis.text.x = element_text(size = 12),
#         axis.text.y = element_text(size = 10, face = 'bold'),
#         legend.title = element_text(size = 12),
#         legend.text = element_text(size = 12))
```
=> There is only one overrepresented biological function: BMI1_DN.V1_UP

```{r}
# Hallmark gene sets
# Retrieve annotated gene sets
msigTabH <- msigdbr(species = 'Homo sapiens', category = 'H', subcategory = NULL) %>%
  dplyr::select(gs_name, human_gene_symbol)

# Run GSEA
gseaResH <- clusterProfiler::GSEA(geneList = rankedGeneList, TERM2GENE = msigTabH,
                                  minGSSize = 10, maxGSSize = 500,
                                  pvalueCutoff = 0.05, pAdjustMethod = 'BH',
                                  by = 'fgsea', eps = 0)

# Plot enrichment analysis results
gseaRes4Plot <- gseaResH@result %>%
  dplyr::select(Description, setSize, NES, pvalue, p.adjust, leading_edge) %>%
  dplyr::arrange(p.adjust) %>%
  dplyr::slice_head(n = 8) %>%
  dplyr::mutate(leading_edge = stringr::str_extract(leading_edge, '\\d\\d'),
                leading_edge = as.numeric(leading_edge),
                Recurrence = ifelse(test = NES > 0, yes = 'Yes', no = 'No')) %>%
  dplyr::rename(gene_ratio = leading_edge)
midGrad <- mean(gseaRes4Plot$p.adjust)

ggplot(gseaRes4Plot, aes(x=NES, y=reorder(Description, NES), fill=p.adjust, col=Recurrence)) +
  geom_bar(stat = 'identity', width = 0.6, size = 2) +
  scale_fill_gradient2(low='#5B1C00', high='#FFFCA5', mid = '#E86900', midpoint = midGrad, name = 'Adjusted pVal') +
  scale_color_manual(values = c(Yes = 'red', No = 'darkblue')) +
  labs(x = 'Normalized Enrichment Score', y = 'Pathway') +
  theme_minimal() +
  theme(axis.title = element_text(size = 16, face = 'bold'),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 10, face = 'bold'),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12))
```
