---
title: 'Single-omics analysis: Combined DIA Proteomics (Discovery + MethodDev)'
author: "Qian-Wu Liao"
output:
  workflowr::wflow_html:
    toc: yes
    toc_float: yes
    code_folding: hide
---

<font size='4'> Description: Conduct univariate (t-test) and multivariate (PCA)
analyses on combined Tissue DIA Proteomics from AG Krijgsveld and AG Klingmüller
(2 datasets) to have overview of data and take initial look at data power in terms
of predicting patient cancer recurrence. </font>

**Metadata variables**\
Patients who got Tissue samples (n = 63):\
Recurrence -> Cancer recurrences, Yes:No = 29:34\
Gender -> Female:Male = 25:38\
Age -> Diagnosis ages ranging from 51 to 84\
Smoking -> Smoker:Ex-smoker:Non-Smoker = 14:40:9\
Stage -> Pathological stages, IB:IIA:IIB = 36:14:13\
Adjuvant -> Adjuvant chemotherapy, True:False = 21:42\
Samples:\
Condition -> Tumor:Normal = 1:1

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 8, fig.width = 10, warning = F)
knitr::opts_knit$set(root.dir = '/Users/qianwu/Desktop/SMART-CARE_LungCancer')
```

Load libraries
```{r library loading, message = F}
library('pcaMethods')
library('limma')
library('proDA')
library('pheatmap')
library('ggplotify')
library('ggrepel')
library('SummarizedExperiment')
library('tidyverse')
# Load user-defined functions
source('./code/analysis_pipeline.R')
source('./code/misc.R')

# Set plot theme
th <- theme_bw(base_size = 15) +
    theme(axis.title = element_text(face = 'bold'),
          axis.text = element_text(face = 'bold'),
          axis.ticks = element_line(linewidth = 0.8),
          legend.text = element_text(size = 15))
```

Display associations between cancer recurrence and the other patient metadata variables
to identify potential predictors (or confounders)
```{r}
# Test associations between cancer recurrence and the other patient metadata variables
# Prepare patient metadata
proTissue <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds')
patientMetadat <- tibble::as_tibble(colData(proTissue)) %>%
  dplyr::filter(!duplicated(Patient)) %>%
  dplyr::select(Patient, Recurrence, Gender, Age, Smoking, Stage, Adjuvant)
patientRecur <- dplyr::select(patientMetadat, Patient, Recurrence)
patientMetadat <- dplyr::select(patientMetadat, -Recurrence)
testAsso(patientMetadat, patientRecur, cmn_col = 'Patient') %>%
  dplyr::select(-Stat)
```

# Tissue Proteomics (AG Krijgsveld)
Discovery + MethodDev

```{r}
# Load preprocessed data
proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds')
# Prepare feature annotation table
featAnnoTab <- tibble::as_tibble(rowData(proTissue_Krij), rownames = 'Protein') %>%
  dplyr::select(Protein, Genes) %>%
  dplyr::rename(Gene = Genes) %>%
  tibble::column_to_rownames('Protein')

# Perform single-omics analysis
proTissueRes_Krij <- doSOA(proTissue_Krij, meta_var = c('Condition', 'Cohort'), pca_method = 'ppca', do_onlyPCA = T)
```

**Do metadata-assisted quality control**\
After testing associations between PCs and tissue sample conditions, we found that
PC1 significantly separates Tumor and Normal samples, which indicates decent data
quality. Yet, there are 4 Tumor samples misclassified, reported as follows:\
=> Tumor purity: I0HVOL_TU - 87% / 7EAOX7_TU - 70% / MJMTYR_TU - 40% / XFKHGP_TU - 100%\
(Mean of sample tumor cell contents is 74%, ranging from 40% to 100%.)
```{r}
pcTab <- proTissueRes_Krij$pcTab %>%
  dplyr::mutate(Label = dplyr::case_when(Sample %in% c('I0HVOL_TU', '7EAOX7_TU',
                                                       'MJMTYR_TU', 'XFKHGP_TU',
                                                       'XFKHGP_NG', 'MJMTYR_NG',
                                                       'I0HVOL_NG', '7EAOX7_NG') ~ Sample))
ggplot(pcTab, aes(x=Condition, y=`PC1 (38%)`, color=Condition, fill=Condition, label=Label)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(position = position_jitter(0.3), size = 3, show.legend = F) +
  geom_text_repel(show.legend = F) +
  scale_color_brewer(palette = 'Dark2') +
  scale_fill_brewer(palette = 'Dark2') +
  labs(x = 'Sample condition') +
  th
```

## Tumor samples

Note that **tumor purity** and **pathological state** are also included in linear
model for more reliable t-test results because they are significantly associated with top PCs.
```{r}
# Subset certain samples
proTumor_Krij <- proTissue_Krij[, which(colData(proTissue_Krij)$Condition == 'Tumor')]
# Perform single-omics analysis
proTumorRes_Krij <- doSingleOmicsAnalysis(proTumor_Krij, soa_metaVar = 'Recurrence', num_sigFeats = 6,
                                          pca_method = 'ppca', num_PCs = 40, num_PCfeats = 30,
                                          use_limma = F, use_proDA = T, feat_anno = featAnnoTab,
                                          feat_conv = T, plot_title = NULL, show_rownames = F,
                                          pca_metaVar = c('Recurrence', 'Gender', 'Age',
                                                          'Smoking', 'Stage', 'Adjuvant',
                                                          'TumorPurity', 'Cohort'),
                                          soa_unwantVar = c('TumorPurity', 'Stage'))
```

Display significant associations between PCs (Var1) and sample metadata variables
(Var2) to overview sources of variance in data
```{r}
proTumorRes_Krij$data.var.source
```

Display significant associations between features (Var1) and cancer recurrence (Var2).
Owing to missing values, there may be situation that only few observations in sample
groups, resulting in unreliable t-statistics. To this end, we used probabilistic
dropout model (proDA) to account for data missingness.
```{r}
proTumorRes_Krij$sig.feat.tab
```

Visualize molecular signatures of significant features in data. Note that features
in rows are ordered by t-statistics.
```{r}
proTumorRes_Krij$sig.feat.heat
```

Visualize top significant features identified in Tumor samples (ns: p > 0.05,
$*$: p <= 0.05, $**$: p <= 0.01)
```{r fig.height=10}
# Prepare data matrix and metadata including both Tumor and Normal samples
datMat <- proTissueRes_Krij$data
smpAnnoTab <- proTissueRes_Krij$smpMetadata
# Prepare significant feature table
featSigAssoTab <- proTumorRes_Krij$sig.feat.tab
num_sigFeats <- 6
gpLevel = c('Yes_Tumor', 'No_Tumor', 'Yes_Normal', 'No_Normal')
condiCol = c('grey50', 'firebrick') #c(Normal, Tumor)
# Extract top significant features and prepare needed information
topSigFeats <- featSigAssoTab[1:num_sigFeats,] %>%
  dplyr::select(Var1, Gene) %>%
  dplyr::mutate(newVar1 = paste0(Var1, ' (', Gene, ')'),
                newVar1 = factor(newVar1, levels = unique(newVar1)))
topSigFeatDat <- tibble::as_tibble(datMat[topSigFeats$Var1,], rownames = 'Var1') %>%
  tidyr::pivot_longer(cols = -'Var1', names_to = 'Sample', values_to = 'Abundance') %>%
  dplyr::left_join(topSigFeats, by = 'Var1') %>%
  dplyr::left_join(smpAnnoTab, by = 'Sample') %>%
  dplyr::mutate(Recur_Condi = paste0(Recurrence, '_', Condition),
                Recur_Condi = factor(Recur_Condi, levels = gpLevel))
ggplot(topSigFeatDat, aes(x=Recur_Condi, y=Abundance, col=Condition, fill=Recurrence)) +
  geom_boxplot(alpha = 1, outlier.shape = NA, linewidth = 1) +
  geom_jitter(position = position_jitter(0.2), size = 2, show.legend = F) +
  ggpubr::stat_compare_means(method = 't.test', paired = F, method.args = list(var.equal = T),
                             comparisons = list(c('Yes_Tumor', 'No_Tumor'), c('Yes_Normal', 'No_Normal')),
                             label = 'p.signif', tip.length = 0.015, bracket.size = 0.7, size = 4) +
  labs(x = 'Recurrence') +
  scale_color_manual(values = condiCol) +
  scale_fill_manual(values=c('#00BFC4', '#F8766D')) +
  facet_wrap(vars(newVar1), scales = 'free') +
  th + theme(strip.text = element_text(size = 13, face = 'bold'),
             axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

Assess data power by p-value histogram
```{r}
proTumorRes_Krij$pval.hist
```

Display significant associations between PCs (Var1) and cancer recurrence (Var2)
```{r}
proTumorRes_Krij$sig.pc.tab
```

Visualize significant PCs and display their top features with greatest absolute loadings
```{r}
proTumorRes_Krij$sig.pc.dist$PC33
proTumorRes_Krij$pc.top.feat.tab$PC33
proTumorRes_Krij$sig.pc.dist$PC11
proTumorRes_Krij$pc.top.feat.tab$PC11
```

## Normal samples

```{r message=F}
# Subset certain samples
proNormal_Krij <- proTissue_Krij[, which(colData(proTissue_Krij)$Condition == 'Normal')]
# Perform single-omics analysis
proNormalRes_Krij <- doSingleOmicsAnalysis(proNormal_Krij, soa_metaVar = 'Recurrence', num_sigFeats = 6,
                                           pca_method = 'ppca', num_PCs = 40, num_PCfeats = 30,
                                           use_limma = F, use_proDA = T, feat_anno = featAnnoTab,
                                           feat_conv = T, plot_title = NULL, show_rownames = F,
                                           pca_metaVar = c('Recurrence', 'Gender', 'Age',
                                                           'Smoking', 'Stage', 'Adjuvant'))
```

Display significant associations between PCs (Var1) and sample metadata variables
(Var2) to overview sources of variance in data
```{r}
proNormalRes_Krij$data.var.source
```

Display significant associations between features (Var1) and cancer recurrence (Var2)
```{r}
proNormalRes_Krij$sig.feat.tab
```

Visualize molecular signatures of significant features in data. Note that features
in rows are ordered by t-statistics.
```{r}
proNormalRes_Krij$sig.feat.heat
```

Visualize top significant features identified in Normal samples (ns: p > 0.05,
$*$: p <= 0.05, $**$: p <= 0.01, $***$: p <= 0.001)
```{r fig.height=10}
# Prepare data matrix and metadata including both Tumor and Normal samples
datMat <- proTissueRes_Krij$data
smpAnnoTab <- proTissueRes_Krij$smpMetadata
# Prepare significant feature table
featSigAssoTab <- proNormalRes_Krij$sig.feat.tab
num_sigFeats <- 6
gpLevel = c('Yes_Normal', 'No_Normal', 'Yes_Tumor', 'No_Tumor')
condiCol = c('firebrick', 'grey50') #c(Normal, Tumor)
# Extract top significant features and prepare needed information
topSigFeats <- featSigAssoTab[1:num_sigFeats,] %>%
  dplyr::select(Var1, Gene) %>%
  dplyr::mutate(newVar1 = paste0(Var1, ' (', Gene, ')'),
                newVar1 = factor(newVar1, levels = unique(newVar1)))
topSigFeatDat <- tibble::as_tibble(datMat[topSigFeats$Var1,], rownames = 'Var1') %>%
  tidyr::pivot_longer(cols = -'Var1', names_to = 'Sample', values_to = 'Abundance') %>%
  dplyr::left_join(topSigFeats, by = 'Var1') %>%
  dplyr::left_join(smpAnnoTab, by = 'Sample') %>%
  dplyr::mutate(Recur_Condi = paste0(Recurrence, '_', Condition),
                Recur_Condi = factor(Recur_Condi, levels = gpLevel))
ggplot(topSigFeatDat, aes(x=Recur_Condi, y=Abundance, col=Condition, fill=Recurrence)) +
  geom_boxplot(alpha = 1, outlier.shape = NA, linewidth = 1) +
  geom_jitter(position = position_jitter(0.2), size = 2, show.legend = F) +
  ggpubr::stat_compare_means(method = 't.test', paired = F, method.args = list(var.equal = T),
                             comparisons = list(c('Yes_Normal', 'No_Normal'), c('Yes_Tumor', 'No_Tumor')),
                             label = 'p.signif', tip.length = 0.015, bracket.size = 0.7, size = 4) +
  labs(x = 'Recurrence') +
  scale_color_manual(values = condiCol) +
  scale_fill_manual(values=c('#00BFC4', '#F8766D')) +
  facet_wrap(vars(newVar1), scales = 'free') +
  th +
  theme(strip.text = element_text(size = 13, face = 'bold'),
             axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
# theme(strip.text = element_text(size = 14, face = 'bold'),
#       axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 20),
#       axis.title = element_text(size = 26),
#       legend.title = element_text(size = 24), legend.text = element_text(size = 22))
```

Assess data power by p-value histogram
```{r}
proNormalRes_Krij$pval.hist
```

There is no significant association between PCs (Var1) and cancer recurrence (Var2).
```{r}
proNormalRes_Krij$sig.pc.tab
```




# Tissue Proteomics (AG Klingmüller)
Discovery (partial) + MethodDev

```{r}
# Load preprocessed data
proTissue_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsnBC.rds')
# Prepare feature annotation table
featAnnoTab <- tibble::as_tibble(rowData(proTissue_Klin), rownames = 'Protein') %>%
  dplyr::select(Protein, PG.Genes) %>%
  dplyr::rename(Gene = PG.Genes) %>%
  tibble::column_to_rownames('Protein')

# Perform single-omics analysis
proTissueRes_Klin <- doSOA(proTissue_Klin, meta_var = c('Condition', 'Cohort'), pca_method = 'ppca', do_onlyPCA = T)
```

**Do metadata-assisted quality control**\
PC1 significantly separates Tumor and Normal samples, which indicates decent data quality.
```{r}
pcTab <- proTissueRes_Klin$pcTab %>%
  dplyr::mutate(Label = dplyr::case_when(Sample %in% c('I0HVOL_TU', '7EAOX7_TU',
                                                       'MJMTYR_TU', 'XFKHGP_TU',
                                                       'XFKHGP_NG', 'MJMTYR_NG',
                                                       'I0HVOL_NG', '7EAOX7_NG') ~ Sample))
ggplot(pcTab, aes(x=Condition, y=`PC1 (31.8%)`, color=Condition, fill=Condition, label=Label)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(position = position_jitter(0.3), size = 3, show.legend = F) +
  geom_text_repel(show.legend = F) +
  scale_color_brewer(palette = 'Dark2') +
  scale_fill_brewer(palette = 'Dark2') +
  labs(x = 'Sample condition') +
  th
```

## Tumor samples

Note that **Stage** are also included in linear model for more reliable t-test results
because it is significantly associated with top PCs.
```{r}
# Subset certain samples
proTumor_Klin <- proTissue_Klin[, which(colData(proTissue_Klin)$Condition == 'Tumor')]
# Perform single-omics analysis
proTumorRes_Klin <- doSingleOmicsAnalysis(proTumor_Klin, soa_metaVar = 'Recurrence', num_sigFeats = 6,
                                          pca_method = 'ppca', num_PCs = 37, num_PCfeats = 30,
                                          use_limma = F, use_proDA = T, feat_anno = featAnnoTab,
                                          feat_conv = T, plot_title = NULL, show_rownames = F,
                                          pca_metaVar = c('Recurrence', 'Gender', 'Age',
                                                          'Smoking', 'Stage', 'Adjuvant',
                                                          'TumorPurity', 'Cohort'),
                                          soa_unwantVar = 'Stage')
```

Display significant associations between PCs (Var1) and sample metadata variables
(Var2) to overview sources of variance in data
```{r}
proTumorRes_Klin$data.var.source
```

Display significant associations between features (Var1) and cancer recurrence (Var2).
Owing to missing values, there may be situation that only few observations in sample
groups, resulting in unreliable t-statistics. To this end, we used probabilistic
dropout model (proDA) to account for data missingness.
```{r}
dplyr::select(proTumorRes_Klin$sig.feat.tab, -Test)
```

Visualize molecular signatures of significant features in data. Note that features
in rows are ordered by t-statistics.
```{r}
proTumorRes_Klin$sig.feat.heat
```

Visualize top significant features identified in Tumor samples (ns: p > 0.05,
$*$: p <= 0.05, $**$: p <= 0.01)
```{r fig.height=10}
# Prepare data matrix and metadata including both Tumor and Normal samples
datMat <- proTissueRes_Klin$data
smpAnnoTab <- proTissueRes_Klin$smpMetadata
# Prepare significant feature table
featSigAssoTab <- proTumorRes_Klin$sig.feat.tab
num_sigFeats <- 6
gpLevel = c('Yes_Tumor', 'No_Tumor', 'Yes_Normal', 'No_Normal')
condiCol = c('grey50', 'firebrick') #c(Normal, Tumor)
# Extract top significant features and prepare needed information
topSigFeats <- featSigAssoTab[1:num_sigFeats,] %>%
  dplyr::select(Var1, Gene) %>%
  dplyr::mutate(newVar1 = paste0(Var1, ' (', Gene, ')'),
                newVar1 = factor(newVar1, levels = unique(newVar1)))
topSigFeatDat <- tibble::as_tibble(datMat[topSigFeats$Var1,], rownames = 'Var1') %>%
  tidyr::pivot_longer(cols = -'Var1', names_to = 'Sample', values_to = 'Abundance') %>%
  dplyr::left_join(topSigFeats, by = 'Var1') %>%
  dplyr::left_join(smpAnnoTab, by = 'Sample') %>%
  dplyr::mutate(Recur_Condi = paste0(Recurrence, '_', Condition),
                Recur_Condi = factor(Recur_Condi, levels = gpLevel))
ggplot(topSigFeatDat, aes(x=Recur_Condi, y=Abundance, col=Condition, fill=Recurrence)) +
  geom_boxplot(alpha = 1, outlier.shape = NA, linewidth = 1) +
  geom_jitter(position = position_jitter(0.2), size = 2, show.legend = F) +
  ggpubr::stat_compare_means(method = 't.test', paired = F, method.args = list(var.equal = T),
                             comparisons = list(c('Yes_Tumor', 'No_Tumor'), c('Yes_Normal', 'No_Normal')),
                             label = 'p.signif', tip.length = 0.015, bracket.size = 0.7, size = 4) +
  labs(x = 'Recurrence') +
  scale_color_manual(values = condiCol) +
  scale_fill_manual(values=c('#00BFC4', '#F8766D')) +
  facet_wrap(vars(newVar1), scales = 'free') +
  th + theme(strip.text = element_text(size = 13, face = 'bold'),
             axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

Assess data power by p-value histogram
```{r}
proTumorRes_Klin$pval.hist
```

Display significant associations between PCs (Var1) and cancer recurrence (Var2)
```{r}
proTumorRes_Klin$sig.pc.tab
```

Visualize significant PCs and display their top features with greatest absolute loadings
```{r}
proTumorRes_Klin$sig.pc.dist$PC19
proTumorRes_Klin$pc.top.feat.tab$PC19
proTumorRes_Klin$sig.pc.dist$PC34
proTumorRes_Klin$pc.top.feat.tab$PC34
```

## Normal samples

Note that **Adjuvant** are also included in linear model for more reliable t-test
results because it is significantly associated with top PCs.
```{r message=F}
# Subset certain samples
proNormal_Klin <- proTissue_Klin[, which(colData(proTissue_Klin)$Condition == 'Normal')]
# Perform single-omics analysis
proNormalRes_Klin <- doSingleOmicsAnalysis(proNormal_Klin, soa_metaVar = 'Recurrence', num_sigFeats = 6,
                                           pca_method = 'ppca', num_PCs = 37, num_PCfeats = 30,
                                           use_limma = F, use_proDA = T, feat_anno = featAnnoTab,
                                           feat_conv = T, plot_title = NULL, show_rownames = F,
                                           pca_metaVar = c('Recurrence', 'Gender', 'Age',
                                                           'Smoking', 'Stage', 'Adjuvant'),
                                           soa_unwantVar = 'Adjuvant')
```

Display significant associations between PCs (Var1) and sample metadata variables
(Var2) to overview sources of variance in data
```{r}
proNormalRes_Klin$data.var.source
```

Display significant associations between features (Var1) and cancer recurrence (Var2)
```{r}
dplyr::select(proNormalRes_Klin$sig.feat.tab, -Test)
```

Visualize molecular signatures of significant features in data. Note that features
in rows are ordered by t-statistics.
```{r}
proNormalRes_Klin$sig.feat.heat
```

Visualize top significant features identified in Normal samples (ns: p > 0.05,
$*$: p <= 0.05, $**$: p <= 0.01, $***$: p <= 0.001)
```{r fig.height=10}
# Prepare data matrix and metadata including both Tumor and Normal samples
datMat <- proTissueRes_Klin$data
smpAnnoTab <- proTissueRes_Klin$smpMetadata
# Prepare significant feature table
featSigAssoTab <- proNormalRes_Klin$sig.feat.tab
num_sigFeats <- 6
gpLevel = c('Yes_Normal', 'No_Normal', 'Yes_Tumor', 'No_Tumor')
condiCol = c('firebrick', 'grey50') #c(Normal, Tumor)
# Extract top significant features and prepare needed information
topSigFeats <- featSigAssoTab[1:num_sigFeats,] %>%
  dplyr::select(Var1, Gene) %>%
  dplyr::mutate(newVar1 = paste0(Var1, ' (', Gene, ')'),
                newVar1 = factor(newVar1, levels = unique(newVar1)))
topSigFeatDat <- tibble::as_tibble(datMat[topSigFeats$Var1,], rownames = 'Var1') %>%
  tidyr::pivot_longer(cols = -'Var1', names_to = 'Sample', values_to = 'Abundance') %>%
  dplyr::left_join(topSigFeats, by = 'Var1') %>%
  dplyr::left_join(smpAnnoTab, by = 'Sample') %>%
  dplyr::mutate(Recur_Condi = paste0(Recurrence, '_', Condition),
                Recur_Condi = factor(Recur_Condi, levels = gpLevel))
ggplot(topSigFeatDat, aes(x=Recur_Condi, y=Abundance, col=Condition, fill=Recurrence)) +
  geom_boxplot(alpha = 1, outlier.shape = NA, linewidth = 1) +
  geom_jitter(position = position_jitter(0.2), size = 2, show.legend = F) +
  ggpubr::stat_compare_means(method = 't.test', paired = F, method.args = list(var.equal = T),
                             comparisons = list(c('Yes_Normal', 'No_Normal'), c('Yes_Tumor', 'No_Tumor')),
                             label = 'p.signif', tip.length = 0.015, bracket.size = 0.7, size = 4) +
  labs(x = 'Recurrence') +
  scale_color_manual(values = condiCol) +
  scale_fill_manual(values=c('#00BFC4', '#F8766D')) +
  facet_wrap(vars(newVar1), scales = 'free') +
  th +
  theme(strip.text = element_text(size = 13, face = 'bold'),
             axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
# theme(strip.text = element_text(size = 14, face = 'bold'),
#       axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 20),
#       axis.title = element_text(size = 26),
#       legend.title = element_text(size = 24), legend.text = element_text(size = 22))
```

Assess data power by p-value histogram
```{r}
proNormalRes_Klin$pval.hist
```

Display significant associations between PCs (Var1) and cancer recurrence (Var2)
```{r}
proNormalRes_Klin$sig.pc.tab
```

Visualize significant PCs and display their top features with greatest absolute loadings
```{r}
proNormalRes_Klin$sig.pc.dist$PC1
proNormalRes_Klin$pc.top.feat.tab$PC1
proNormalRes_Klin$sig.pc.dist$PC9
proNormalRes_Klin$pc.top.feat.tab$PC9
```
