---
title: 'MOFA: Downstream analysis'
author: "Qian-Wu Liao"
output:
  workflowr::wflow_html:
    toc: true
    toc_float: true
    code_folding: hide
---

<font size='4'> Description: </font>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 8, fig.width = 10, warning = F)
knitr::opts_knit$set(root.dir = '/Users/qianwu/Desktop/SMART-CARE_LungCancer')
```

Load libraries
```{r library loading, message = F}
library('MOFA2')
library('reticulate')
# use_python('/Users/qianwu/opt/anaconda3/bin/python')
library('AnnotationDbi')
library('org.Hs.eg.db')
# This annotation object is accessed using AnnotationDbi:select()
hs <- org.Hs.eg.db
library('msigdbr')
library('clusterProfiler')
library('ggrepel')
library('SummarizedExperiment')
library('tidyverse')
# Load user-defined functions
source('./code/misc.R')
source('./code/mofa_funcs.R')

# Set plot theme
th <- theme_bw(base_size = 15) +
  theme(axis.title = element_text(face = 'bold'),
        axis.text = element_text(face = 'bold'),
        axis.ticks = element_line(linewidth = 0.8),
        legend.text = element_text(size = 15))
```

```{r message=F}
# Prepare data tables to train MOFA models
# Load normalized datasets
comb_proTissue_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsnBC.rds')
comb_proNormal_Klin <- comb_proTissue_Klin[, colData(comb_proTissue_Klin)$Condition == 'Normal']

comb_proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds')

dis_lipTissue_Hopf <- readRDS('./data/Discovery/AG_Hopf/lipTissueVsn_WBC25.rds')
dis_lipNormal_Hopf <- dis_lipTissue_Hopf[, colData(dis_lipTissue_Hopf)$Condition == 'Normal']

mDev_lipTissue_Hopf <- readRDS('./data/MethodDev/AG_Hopf/lipTissueVsn.rds')
mDev_lipNormal_Hopf <- mDev_lipTissue_Hopf[, colData(mDev_lipTissue_Hopf)$Condition == 'Normal']

dis_lipBase_Hopf <- readRDS('./data/Discovery/AG_Hopf/lipBaseVsn_B1WBC25.rds')

mDev_lipPlasma_Hopf <- readRDS('./data/MethodDev/AG_Hopf/lipPlasmaVsn.rds')
mDev_lipBase_Hopf <- mDev_lipPlasma_Hopf[, colData(mDev_lipPlasma_Hopf)$TimePoint == 'Baseline']

# Sanity check dimensionality of each dataset (Feature x Sample)
dataList <- list(Comb_Tissue_Proteomics_Klin = comb_proTissue_Klin,
                 Comb_Tissue_Proteomics_Krij = comb_proTissue_Krij,
                 Dis_Tissue_Lipidomics_Hopf = dis_lipTissue_Hopf,
                 MDev_Tissue_Lipidomics_Hopf = mDev_lipTissue_Hopf,
                 Dis_Baseline_Lipidomics_Hopf = dis_lipBase_Hopf,
                 MDev_Plasma_Lipidomics_Hopf = mDev_lipPlasma_Hopf)
# lapply(dataList, dim)

# Prepare significant features for feature selection
# soaRes <- readRDS('./data/Discovery/potential_proteins/Combined_proNormalRes_Klin.rds')
# sigFeats_comb_proNormal_Klin <- soaRes$sig.feat.tab$Var1
# soaRes <- doSOA(dis_lipNormal_Hopf, meta_var = 'Recurrence', use_proDA = T)
# sigFeats_dis_lipNormal_Hopf <- soaRes$featSigAssoRes$Var1
# soaRes <- doSOA(mDev_lipNormal_Hopf, meta_var = 'Recurrence', use_proDA = T)
# sigFeats_mDev_lipNormal_Hopf <- soaRes$featSigAssoRes$Var1
# soaRes <- doSOA(dis_lipBase_Hopf, meta_var = 'Recurrence', use_proDA = T)
# sigFeats_dis_lipBase_Hopf <- soaRes$featSigAssoRes$Var1
# soaRes <- doSOA(mDev_lipBase_Hopf, meta_var = 'Recurrence', use_proDA = T)
# sigFeats_mDev_lipBase_Hopf <- soaRes$featSigAssoRes$Var1

# Convert SE objects to long data for creating MOFA object through 'create_mofa_from_df'
# to include metadata
# comb_proNormalTab_Klin <- mofa_summExp2df(comb_proNormal_Klin, smp_type = 'Tissue',
#                                           data_acqui = 'Combined DIA', data_modal = 'Proteomics') %>%
#   dplyr::filter(feature %in% sigFeats_comb_proNormal_Klin)
# dis_lipNormalTab_Hopf <- mofa_summExp2df(dis_lipNormal_Hopf, smp_type = 'Tissue',
#                                          data_acqui = 'Discovery Untargeted', data_modal = 'Lipidomics') %>%
#   dplyr::filter(feature %in% sigFeats_dis_lipNormal_Hopf)
# mDev_lipNormalTab_Hopf <- mofa_summExp2df(mDev_lipNormal_Hopf, smp_type = 'Tissue',
#                                           data_acqui = 'MethodDev Untargeted', data_modal = 'Lipidomics') %>%
#   dplyr::filter(feature %in% sigFeats_mDev_lipNormal_Hopf)
# dis_lipBaseTab_Hopf <- mofa_summExp2df(dis_lipBase_Hopf, smp_type = 'Plasma',
#                                        data_acqui = ' Discovery Untargeted', data_modal = 'Lipidomics') %>%
#   dplyr::filter(feature %in% sigFeats_dis_lipBase_Hopf)
# mDev_lipBaseTab_Hopf <- mofa_summExp2df(mDev_lipBase_Hopf, smp_type = 'Plasma',
#                                         data_acqui = 'MethodDev Untargeted', data_modal = 'Lipidomics') %>%
#   dplyr::filter(feature %in% sigFeats_mDev_lipBase_Hopf)
```

# Comb. DIA NTP (Klin.) + All Unt. NTL
Combined DIA Normal Tissue Proteomics (MethodDev + Discovery) from AG Klingmüller +
Discovery and MethodDev Untargeted Normal Tissue Lipidomics\
(since features defined by RT/CCS/MZ have difficulty being combined together from different cohorts)

```{r}
# Train MOFA model
# mofaObject <- trainMOFA(list(comb_proNormalTab_Klin, dis_lipNormalTab_Hopf, mDev_lipNormalTab_Hopf),
#                         view_data = T, train_mofa = T, num_factors = 10,
#                         save_path = './data/Discovery/mofa/sigFeats_CombDIANTPKlin_DisMDevUntNTL')

# Load trained model
mofaObject <- readRDS('./data/Discovery/mofa/sigFeats_CombDIANTPKlin_DisMDevUntNTL.rds')
```

Overview model
```{r}
# Overview model
MOFA2::plot_data_overview(mofaObject)
```

Quantify amount of variance explained by each factor in each data modality
```{r}
# Do sanity check for factor correlations
# MOFA2::plot_factor_cor(mofaObject)

# Perform variance dedecomposition analysis (coefficient of determination)
MOFA2::plot_variance_explained(mofaObject, max_r2 = 15)
```

Display and visualize significant recurrence-related factors
```{r}
# Display significant associations between learned factors (Var1) and cancer recurrence
# (Var2) and visualize significant recurrence-related factors
sigFactor <- mofa_vizSigFactor(mofaObject, smpGroup = 'Baseline Group')

# Factor1: p-value = 0.000021 without zero-valued samples
```
=> Samples with extremely small scores of Factor1&4 exist only in Untargeted Lipidomics
that includes few features.\
=> Factor1 is highly correlated with PC1 from Klin. NTP and explains variance also
in MDev. NTL, but in Dis. NTL. Note that Dis. NTL got 21 samples and MDev. NTL got
18 samples in common with Klin. NTP, which might imply worse sample or data quality
of Discovery Tissue Lipidomics (which is proved by their biological quality controls).

Visualize Factor1 without samples with extremely small values
```{r}
# Do t-test without samples existing in Discovery Untargeted Lipidomics but Klin. Proteomics
# Collect samples to be removed
smps_CombNTP <- stringr::str_replace(colnames(comb_proNormal_Klin), '_NG', '_B')
smps_DisNTL <- stringr::str_replace(colnames(dis_lipNormal_Hopf), '_NG', '_B')
rmSmps <- smps_DisNTL[!smps_DisNTL %in% smps_CombNTP]
# Prepare factor table
facTab <- get_factors(mofaObject)[['Baseline Group']] %>%
  tibble::as_tibble(rownames = 'sample') %>%
  dplyr::select(sample, Factor1) %>%
  dplyr::filter(!sample %in% rmSmps)
# Prepare metadata table
metaTab <- samples_metadata(mofaObject) %>%
  dplyr::select(sample, Recurrence) %>%
  dplyr::filter(!sample %in% rmSmps)
# Do t-test
# testAsso(facTab, metaTab, cmn_col = 'sample') %>%
#     dplyr::filter(pVal <= 0.05)
# Visualize factor values
tab4Plot <- dplyr::left_join(facTab, metaTab, by = 'sample')
ggplot(tab4Plot, aes(x=Recurrence, y=Factor1, col=Recurrence, fill=Recurrence)) +
        geom_boxplot(alpha = 0.7, outlier.shape = NA) +
        geom_jitter(position = position_jitter(0.2), show.legend = F) +
        scale_color_manual(values=c('#00BFC4', '#F8766D')) +
        scale_fill_manual(values=c('#00BFC4', '#F8766D')) +
        ggpubr::stat_compare_means(method = 't.test', paired = F,
                                   method.args = list(var.equal = T),
                                   show.legend = F) +
        th
```

Compare Factor1 with recurrence-related PC1 learned from Combined Klin. Normal Tissue Proteomics
```{r}
# Visualize correlation between significant factor (MOFA) and significant PC (PCA)
# Prepare significant factor table
facTab <- sigFactor$tab4Plot %>%
  dplyr::select(sample, Factor1, Recurrence) %>%
  dplyr::mutate(sample = stringr::str_replace(sample, '_B', '_NG'))
# Prepare significant PC table
proNormalRes_Klin <- readRDS('./data/Discovery/potential_proteins/Combined_proNormalRes_Klin.rds')
# Show significant PCs
# proNormalRes_Klin$sig.pc.tab
pcTab <- proNormalRes_Klin$SOA.res$pcTab %>%
  dplyr::select(Sample, `PC1 (11.1%)`) %>%
  dplyr::rename(sample = Sample)
# Combine all information
combinedTab <- dplyr::left_join(facTab, pcTab, by = 'sample')
# Make a plot and compute correlation
ggplot(combinedTab, aes(x=Factor1, y=`PC1 (11.1%)`)) +
  geom_point(aes(col=Recurrence), size = 4) +
  scale_color_manual(values=c('#00BFC4', '#F8766D')) +
  scale_fill_manual(values=c('#00BFC4', '#F8766D')) +
  ggpubr::stat_cor(aes(label=after_stat(r.label)), method = 'pearson', size = 7) +
  th
```

Visualize enrichment analysis results
```{r message=F}
# Prepare ranked feature list
# Tidy up weight matrix and feature metadata to reduce information loss
proGeneTab <- rowData(comb_proNormal_Klin) %>%
  tibble::as_tibble(rownames = 'Proteins') %>%
  tibble::column_to_rownames('Proteins')
tidyWeiMat <- mofa_keepUniFeats(mofaObject, factor = 1, view = 'Combined DIA Normal Proteomics',
                                feat_anno = proGeneTab, to_genes = T)
# Rank feature weights
rankedGeneList <- mofa_rankFeatList(tidyWeiMat, factor = 1)
# Make Recurrence group positive weights
rankedGeneList <- sort(-rankedGeneList, decreasing = T)

# Do enrichment analysis - Hallmark gene sets (H)
# Retrieve annotated gene sets
msigTabH <- msigdbr(species = 'Homo sapiens', category = 'H', subcategory = NULL) %>%
  dplyr::select(gs_name, human_gene_symbol)
# Run GSEA
gseaResH <- clusterProfiler::GSEA(geneList = rankedGeneList, TERM2GENE = msigTabH,
                                  minGSSize = 10, maxGSSize = 500,
                                  pvalueCutoff = 0.05, pAdjustMethod = 'BH',
                                  by = 'fgsea', eps = 0)

# Plot enrichment analysis results
gseaRes4Plot <- gseaResH@result %>%
  dplyr::select(Description, setSize, NES, pvalue, p.adjust, leading_edge) %>%
  dplyr::arrange(p.adjust) %>%
  dplyr::slice_head(n = 8) %>%
  dplyr::mutate(leading_edge = stringr::str_extract(leading_edge, '\\d\\d'),
                leading_edge = as.numeric(leading_edge),
                Recurrence = ifelse(test = NES > 0, yes = 'Yes', no = 'No')) %>%
  dplyr::rename(gene_ratio = leading_edge)
midGrad <- mean(gseaRes4Plot$p.adjust)

ggplot(gseaRes4Plot, aes(x=NES, y=reorder(Description, NES), fill=p.adjust, col=Recurrence)) +
  geom_bar(stat = 'identity', width = 0.6, size = 2) +
  scale_fill_gradient2(low='#5B1C00', high='#FFFCA5', mid = '#E86900', midpoint = midGrad, name = 'Adjusted pVal') +
  scale_color_manual(values = c(Yes = 'red', No = 'darkblue')) +
  labs(x = 'Normalized Enrichment Score', y = 'Pathway', title = 'MSigDB:H - Factor1') +
  theme_minimal() +
  theme(axis.title = element_text(size = 16, face = 'bold'),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 10, face = 'bold'),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12))

# => Factor2: HALLMARK_E2F_TARGETS, HALLMARK_MYC_TARGETS_V1, HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION
# => Factor4: HALLMARK_MYC_TARGETS_V1
```

```{r include=F, eval=F}
# Compare factors learned from different MOFA models
mofaObj1 <- readRDS('./data/Discovery/mofa/sigFeats_CombDIANTPKlin_DisMDevUntNTL.rds')
mofaObj2 <- readRDS('./data/Discovery/mofa/sigFeats_CombDIANTPKlin_DisUntNTL.rds')
mofaObj3 <- readRDS('./data/Discovery/mofa/sigFeats_CombDIANTPKlin_MDevUntNTL.rds')

MOFA2::plot_variance_explained(mofaObj2, max_r2 = 15) +
  labs(title = 'DisUntNTL: Factor1&2 Sig.')
MOFA2::plot_variance_explained(mofaObj3, max_r2 = 15) +
  labs(title = 'MDevUntNTL: Factor1&3 Sig.')

MOFA2::compare_factors(list(DisMDevUntNTL = mofaObj1, DisUntNTL = mofaObj2, MDevUntNTL = mofaObj3))
```

## Comb. DIA NTP (Klin.) + All Unt. BPL
Combined DIA Normal Tissue Proteomics (MethodDev + Discovery) from AG Klingmüller +
Discovery and MethodDev Untargeted Baseline Plasma Lipidomics
(since features defined by RT/CCS/MZ have difficulty being combined together from different cohorts)

```{r}
# Train MOFA model
# mofaObject <- trainMOFA(list(comb_proNormalTab_Klin, dis_lipBaseTab_Hopf, mDev_lipBaseTab_Hopf),
#                         view_data = T, train_mofa = T, num_factors = 10,
#                         save_path = './data/Discovery/mofa/sigFeats_CombDIANTPKlin_DisMDevUntBPL')

# Load trained model
mofaObject <- readRDS('./data/Discovery/mofa/sigFeats_CombDIANTPKlin_DisMDevUntBPL.rds')
```

Overview model
```{r}
# Overview model
MOFA2::plot_data_overview(mofaObject)
```

Quantify amount of variance explained by each factor in each data modality
```{r}
# Do sanity check for factor correlations
# MOFA2::plot_factor_cor(mofaObject)

# Perform variance dedecomposition analysis (coefficient of determination)
MOFA2::plot_variance_explained(mofaObject, max_r2 = 15)
```

Display and visualize significant recurrence-related factors
```{r}
# Display significant associations between learned factors (Var1) and cancer recurrence
# (Var2) and visualize significant recurrence-related factors
sigFactor <- mofa_vizSigFactor(mofaObject, smpGroup = 'Baseline Group')
```

Compare Factor1 with recurrence-related PC1 learned from Combined Klin. Normal Tissue
Proteomics and another Factor1 from model built on Klin. NTP and Unt. NTL
```{r}
# Visualize correlation between significant factors (MOFA) and significant PC (PCA)
# Prepare significant factor table
facTab <- sigFactor$tab4Plot %>%
  dplyr::select(sample, Factor1, Recurrence) %>%
  dplyr::mutate(sample = stringr::str_replace(sample, '_B', '_NG'))
# Prepare significant PC table
proNormalRes_Klin <- readRDS('./data/Discovery/potential_proteins/Combined_proNormalRes_Klin.rds')
# Show significant PCs
# proNormalRes_Klin$sig.pc.tab
pcTab <- proNormalRes_Klin$SOA.res$pcTab %>%
  dplyr::select(Sample, `PC1 (11.1%)`) %>%
  dplyr::rename(sample = Sample)
# Combine all information
combinedTab <- dplyr::left_join(facTab, pcTab, by = 'sample')
# Make a plot and compute correlation
ggplot(combinedTab, aes(x=Factor1, y=`PC1 (11.1%)`)) +
  geom_point(aes(col=Recurrence), size = 4) +
  scale_color_manual(values=c('#00BFC4', '#F8766D')) +
  scale_fill_manual(values=c('#00BFC4', '#F8766D')) +
  ggpubr::stat_cor(aes(label=after_stat(r.label)), method = 'pearson', size = 7) +
  th

# Prepare significant factor tables
facTab <- sigFactor$tab4Plot %>%
  dplyr::select(sample, Factor1, Recurrence) %>%
  dplyr::rename(Factor1_BPL = Factor1)
mofaObj <- readRDS('./data/Discovery/mofa/sigFeats_CombDIANTPKlin_DisMDevUntNTL.rds')
facTab2 <- get_factors(mofaObj)[['Baseline Group']] %>%
  tibble::as_tibble(rownames = 'sample') %>%
  dplyr::select(sample, Factor1) %>%
  dplyr::rename(Factor1_NTL = Factor1)
# Collect samples with extremely small Factor1_NTL, which will be removed
smps_CombNTP <- stringr::str_replace(colnames(comb_proNormal_Klin), '_NG', '_B')
smps_DisNTL <- stringr::str_replace(colnames(dis_lipNormal_Hopf), '_NG', '_B')
rmSmps <- smps_DisNTL[!smps_DisNTL %in% smps_CombNTP]
# Combine all information and remove unwanted samples
combinedTab <- dplyr::left_join(facTab, facTab2, by = 'sample') %>%
  dplyr::filter(!sample %in% rmSmps)
# Make a plot and compute correlation
ggplot(combinedTab, aes(x=Factor1_BPL, y=Factor1_NTL)) +
  geom_point(aes(col=Recurrence), size = 4) +
  scale_color_manual(values=c('#00BFC4', '#F8766D')) +
  scale_fill_manual(values=c('#00BFC4', '#F8766D')) +
  ggpubr::stat_cor(aes(label=after_stat(r.label)), method = 'pearson', size = 7) +
  th
```
=> Samples with extremely small Factor1_NTL were removed.

Visualize enrichment analysis results
```{r message=F}
# Prepare ranked feature list
# Tidy up weight matrix and feature metadata to reduce information loss
proGeneTab <- rowData(comb_proNormal_Klin) %>%
  tibble::as_tibble(rownames = 'Proteins') %>%
  tibble::column_to_rownames('Proteins')
tidyWeiMat <- mofa_keepUniFeats(mofaObject, factor = 1, view = 'Combined DIA Normal Proteomics',
                                feat_anno = proGeneTab, to_genes = T)
# Rank feature weights
rankedGeneList <- mofa_rankFeatList(tidyWeiMat, factor = 1)
# Make Recurrence group positive weights
rankedGeneList <- sort(-rankedGeneList, decreasing = T)

# Do enrichment analysis - Hallmark gene sets (H)
# Retrieve annotated gene sets
msigTabH <- msigdbr(species = 'Homo sapiens', category = 'H', subcategory = NULL) %>%
  dplyr::select(gs_name, human_gene_symbol)
# Run GSEA
gseaResH <- clusterProfiler::GSEA(geneList = rankedGeneList, TERM2GENE = msigTabH,
                                  minGSSize = 10, maxGSSize = 500,
                                  pvalueCutoff = 0.05, pAdjustMethod = 'BH',
                                  by = 'fgsea', eps = 0)

# Plot enrichment analysis results
gseaRes4Plot <- gseaResH@result %>%
  dplyr::select(Description, setSize, NES, pvalue, p.adjust, leading_edge) %>%
  dplyr::arrange(p.adjust) %>%
  dplyr::slice_head(n = 8) %>%
  dplyr::mutate(leading_edge = stringr::str_extract(leading_edge, '\\d\\d'),
                leading_edge = as.numeric(leading_edge),
                Recurrence = ifelse(test = NES > 0, yes = 'Yes', no = 'No')) %>%
  dplyr::rename(gene_ratio = leading_edge)
midGrad <- mean(gseaRes4Plot$p.adjust)

ggplot(gseaRes4Plot, aes(x=NES, y=reorder(Description, NES), fill=p.adjust, col=Recurrence)) +
  geom_bar(stat = 'identity', width = 0.6, size = 2) +
  scale_fill_gradient2(low='#5B1C00', high='#FFFCA5', mid = '#E86900', midpoint = midGrad, name = 'Adjusted pVal') +
  scale_color_manual(values = c(Yes = 'red', No = 'darkblue')) +
  labs(x = 'Normalized Enrichment Score', y = 'Pathway', title = 'MSigDB:H - Factor1') +
  theme_minimal() +
  theme(axis.title = element_text(size = 16, face = 'bold'),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 10, face = 'bold'),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12))
```

```{r include=F, eval=F}
# Compare factors learned from different MOFA models
mofaObj1 <- readRDS('./data/Discovery/mofa/sigFeats_CombDIANTPKlin_DisMDevUntNTL.rds')
mofaObj2 <- readRDS('./data/Discovery/mofa/sigFeats_CombDIANTPKlin_DisMDevUntBPL.rds')

MOFA2::compare_factors(list(NTL = mofaObj1, BPL = mofaObj2))
```




```{r}
sessionInfo()
```






















