---
title: "Single-omics analysis: Untargeted Lipidomics of Discovery cohort"
author: "Qian-Wu Liao"
output:
  workflowr::wflow_html:
    toc: yes
    toc_float: yes
    code_folding: hide
---

<font size='4'> Description: Conduct univariate (t-test) and multivariate (PCA)
analyses on individual datasets, i.e., Tissue and Plasma Untargeted Lipidomics from
AG Hopf (2 datasets), to have overview of data and take initial look at data
power in terms of predicting patient cancer recurrence. </font>

**Metadata variables**\
Patients who got Tissue samples (n = 43):\
Recurrence -> Cancer recurrences, Yes:No = 19:24\
Gender -> Female:Male = 18:25\
Age -> Diagnosis ages ranging from 51 to 76\
Smoking -> Smoker:Ex-smoker:Non-Smoker = 11:26:6\
Stage -> Pathological stages, IB:IIA:IIB = 25:6:12\
Adjuvant -> Adjuvant chemotherapy, True:False = 14:29\
Samples:\
Condition -> Tumor:Normal = 1:1

Patients who got Baseline Plasma samples (n = 63):\
Recurrence -> Cancer recurrences, Yes:No = 24:39\
Gender -> Female:Male = 28:35\
Age -> Diagnosis ages ranging from 40 to 82\
Smoking -> Smoker:Ex-smoker:Non-Smoker = 16:40:7\
Stage -> Pathological stages, IB:IIA:IIB = 39:13:11\
Adjuvant -> Adjuvant chemotherapy, True:False = 20:43

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 8, fig.width = 10, warning = F)
knitr::opts_knit$set(root.dir = '/Users/qianwu/Desktop/SMART-CARE_LungCancer')
```

Load libraries
```{r library loading, message = F}
library('sva')
library('pcaMethods')
library('limma')
library('proDA')
library('pheatmap')
library('ggplotify')
library('ggrepel')
library('SummarizedExperiment')
library('tidyverse')
# Load user-defined functions
source('./code/analysis_pipeline.R')
source('./code/misc.R')

# Set plot theme
th <- theme_bw(base_size = 15) +
    theme(axis.title = element_text(face = 'bold'),
          axis.text = element_text(face = 'bold'),
          axis.ticks = element_line(linewidth = 0.8),
          legend.text = element_text(size = 15))
```

Display associations between cancer recurrence and the other patient metadata variables
to identify potential predictors (or confounders)
```{r}
# Test associations between cancer recurrence and the other patient metadata variables
# Prepare patient metadata
lipTissue <- readRDS('./data/Discovery/AG_Hopf/lipTissueVsn_WBC25.rds')
patientMetadat <- tibble::as_tibble(colData(lipTissue)) %>%
  dplyr::filter(!duplicated(Patient)) %>%
  dplyr::select(Patient, Recurrence, Gender, Age, Smoking, Stage, Adjuvant)
patientRecur <- dplyr::select(patientMetadat, Patient, Recurrence)
patientMetadat <- dplyr::select(patientMetadat, -Recurrence)
testAsso(patientMetadat, patientRecur, cmn_col = 'Patient') %>%
  dplyr::select(-Stat)
```

# Tissue Lipidomics

**Do metadata-assisted quality control**\
After testing associations between PCs and tissue sample conditions, we found that
PC1 and PC2 can significantly separate Tumor and Normal samples, which indicates
decent data quality. Yet, there are some samples misclassified.\
The following 4 samples were misclassified in Krijgsveld Discovery DIA Proteomics:\
Tumor purity: I0HVOL_TU - 87% / 7EAOX7_TU - 70% / MJMTYR_TU - 40% / XFKHGP_TU - 100%\
(Mean of sample tumor cell contents is 74%, ranging from 40% to 100%.)
```{r}
# Load preprocessed data
lipTissue <- readRDS('./data/Discovery/AG_Hopf/lipTissueVsn_WBC25.rds')
# Do metadata-assisted quality control
# Perform single-omics analysis
lipTissueRes <- doSOA(lipTissue, meta_var = 'Condition', pca_method = 'ppca', do_onlyPCA = T)
# Visualize top PCs
pcTab <- lipTissueRes$pcTab %>%
  dplyr::mutate(Label = dplyr::case_when(Sample %in% c('I0HVOL_TU', '7EAOX7_TU',
                                                       'MJMTYR_TU', 'XFKHGP_TU',
                                                       'XFKHGP_NG', 'MJMTYR_NG',
                                                       'I0HVOL_NG', '7EAOX7_NG') ~ Sample))
ggplot(pcTab, aes(x=Condition, y=`PC1 (41.5%)`, color=Condition, fill=Condition, label=Label)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(position = position_jitter(0.3), size = 3, show.legend = F) +
  geom_text_repel(show.legend = F) +
  scale_color_brewer(palette = 'Dark2') +
  scale_fill_brewer(palette = 'Dark2') +
  labs(x = 'Sample condition') +
  th
```

Tumor purity: I0HVOL_TU - 87% / 7EAOX7_TU - 70% / 1NL5BV_TU - 53%\
(Mean of sample tumor cell contents is 74%, ranging from 40% to 100%.)
```{r}
pcTab <- lipTissueRes$pcTab %>%
  dplyr::mutate(Label = dplyr::case_when(Sample %in% c('1NL5BV_TU', '1NL5BV_NG',
                                                       'I0HVOL_TU', 'I0HVOL_NG',
                                                       '7EAOX7_TU', '7EAOX7_NG') ~ Sample))
ggplot(pcTab, aes(x=`PC1 (41.5%)`, y=`PC2 (13.5%)`, col=Condition, label=Label)) +
  geom_point(size = 2.5) +
  geom_text_repel(size = 4, show.legend = F) +
  scale_color_brewer(palette = 'Dark2') +
  th
```
=> Tumor and Normal tissue samples will be analyzed separately for two reasons:
(1) They are characteristically different and might be offsets to each other, and
(2) Normal tissues is always obtainable in case that Tumor tissues were not preserved
after surgery.

## Normal samples

Display significant associations between SVs and sample metadata variables. SVs
(surrogate variables) will be included in linear model for differential analysis
```{r}
# Subset certain samples
lipNormal <- lipTissue[, which(colData(lipTissue)$Condition == 'Normal')]

# Do SVA to identify latent factors that explain unknown variance
# Prepare data matrix and sample metadata table for creating model matrices
datMat <- assay(lipNormal)
# Keep only complete features
keptFeats <- apply(datMat, 1, function(featVec) {
  !any(is.na(featVec))
})
datMat <- datMat[keptFeats,]
metadatTab <- colData(lipNormal) %>%
  tibble::as_tibble(rownames = 'Sample')
# Create full model matrix including both adjustment variables and variables of interest
modelMat <- model.matrix(~ Recurrence, data = metadatTab)
# Create null model matrix including only adjustment variables
modelMat0 <- model.matrix(~ 1, data = metadatTab)

# Perform SVA
# Identify number of latent factors that need to be estimated
numSV = sva::num.sv(datMat, modelMat, method = 'be')
svObj = sva(datMat, modelMat, modelMat0, n.sv = numSV)

# Do association tests between SVs and patient metadata
# Prepare surrogate variable table
svTab <- svObj$sv
rownames(svTab) <- metadatTab$Sample
colnames(svTab) <- paste0('SV', seq_len(ncol(svTab)))
svTab <- tibble::as_tibble(svTab, rownames = 'Sample')
# Prepare patient metadata table
metaTab <- dplyr::select(metadatTab, Sample, Recurrence, Gender, Age, Smoking, Stage,
                         Adjuvant, MissLevel)
# Do association tests
testAsso(svTab, metaTab, cmn_col = 'Sample') %>%
  dplyr::filter(pVal <= 0.05)

# Include surrogate variables into SE object
lipNormalTab <- summExp2df(lipNormal, row_id = 'Feature', col_id = 'Sample')
lipNormalTab <- dplyr::left_join(lipNormalTab, svTab, by = 'Sample')
lipNormal <- df2SummExp(lipNormalTab, row_id = 'Feature', col_id = 'Sample',
                        values = 'Value', row_anno = c('RT', 'CCS', 'MZ'),
                        col_anno = c(colnames(colData(lipNormal)),
                                     paste0('SV', seq_len(ncol(svTab)-1))))
assay(lipNormal) <- as.matrix(assay(lipNormal))
```

```{r message=F}
# Perform single-omics analysis
lipNormalRes <- doSingleOmicsAnalysis(lipNormal, soa_metaVar = 'Recurrence', num_sigFeats = 6,
                                      pca_method = 'ppca', num_PCs = 20, num_PCfeats = 30,
                                      use_proDA = T, pca_metaVar = c('Recurrence', 'Gender', 'Age',
                                                                     'Smoking', 'Stage', 'Adjuvant',
                                                                     'MissLevel'),
                                      show_row = F,
                                      soa_unwantVar = c(paste0('SV', seq_len(6)),
                                                        'MissLevel', 'Stage', 'Gender',
                                                        'Adjuvant', 'Smoking'))
                                      # soa_unwantVar = c('MissLevel', 'Stage', 'Gender',
                                      #                   'Adjuvant', 'Smoking'))
```

Display significant associations between PCs (Var1) and sample metadata variables
(Var2) to overview sources of variance in data
```{r}
lipNormalRes$data.var.source
```

Display significant associations between features (Var1) and cancer recurrence (Var2).
Owing to missing values, there may be situation that only few observations in sample
groups, resulting in unreliable t-statistics. To this end, we used probabilistic
dropout model (proDA) to account for data missingness.
```{r}
lipNormalRes$sig.feat.tab
```

Visualize molecular signatures of significant features in data. Note that features
in rows are ordered by t-statistics.
```{r}
lipNormalRes$sig.feat.heat
```

Visualize top significant features identified in Tumor samples (ns: p > 0.05,
$*$: p <= 0.05, $**$: p <= 0.01)
```{r fig.height=10}
# Prepare data matrix and metadata including both Tumor and Normal samples
datMat <- lipTissueRes$data
smpAnnoTab <- lipTissueRes$smpMetadata
# Prepare significant feature table
featSigAssoTab <- lipNormalRes$sig.feat.tab
num_sigFeats <- 6
gpLevel = c('Yes_Normal', 'No_Normal', 'Yes_Tumor', 'No_Tumor')
condiCol = c('firebrick', 'grey50') #c(Normal, Tumor)
# Extract top significant features and prepare needed information
topSigFeats <- featSigAssoTab[1:num_sigFeats,] %>%
  dplyr::pull(Var1)
topSigFeatDat <- tibble::as_tibble(datMat[topSigFeats,], rownames = 'Var1') %>%
  tidyr::pivot_longer(cols = -'Var1', names_to = 'Sample', values_to = 'Abundance') %>%
  dplyr::left_join(smpAnnoTab, by = 'Sample') %>%
  dplyr::mutate(Var1 = factor(Var1, levels = topSigFeats),
                Recur_Condi = paste0(Recurrence, '_', Condition),
                Recur_Condi = factor(Recur_Condi, levels = gpLevel))
ggplot(topSigFeatDat, aes(x=Recur_Condi, y=Abundance, col=Condition, fill=Recurrence)) +
  geom_boxplot(alpha = 1, outlier.shape = NA, linewidth = 1) +
  geom_jitter(position = position_jitter(0.2), size = 2, show.legend = F) +
  ggpubr::stat_compare_means(method = 't.test', paired = F, method.args = list(var.equal = T),
                             comparisons = list(c('Yes_Normal', 'No_Normal'), c('Yes_Tumor', 'No_Tumor')),
                             label = 'p.signif', tip.length = 0.015, bracket.size = 0.7, size = 4) +
  labs(x = 'Recurrence') +
  scale_color_manual(values = condiCol) +
  scale_fill_manual(values=c('#00BFC4', '#F8766D')) +
  facet_wrap(vars(Var1), scales = 'free') +
  th + theme(strip.text = element_text(size = 13, face = 'bold'),
             axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

Assess data power by p-value histogram
```{r}
lipNormalRes$pval.hist
```

Display significant associations between PCs (Var1) and cancer recurrence (Var2)
```{r}
lipNormalRes$sig.pc.tab
```

Visualize significant PCs and display their top features with greatest absolute loadings
```{r}
lipNormalRes$sig.pc.dist$PC18
lipNormalRes$pc.top.feat.tab$PC18
```

## Tumor samples

Note that tumor purity is also included in linear model for more reliable t-test
results because they are significantly associated with top PCs.
```{r}
# Subset certain samples
lipTumor <- lipTissue[, which(colData(lipTissue)$Condition == 'Tumor')]
# Perform single-omics analysis
lipTumorRes <- doSingleOmicsAnalysis(lipTumor, soa_metaVar = 'Recurrence', num_sigFeats = 6,
                                     pca_method = 'ppca', num_PCs = 20, num_PCfeats = 30,
                                     use_proDA = T, pca_metaVar = c('Recurrence', 'Gender', 'Age',
                                                                    'Smoking', 'Stage', 'Adjuvant',
                                                                    'TumorPurity'),
                                     soa_unwantVar = 'TumorPurity')
```

Display significant associations between PCs (Var1) and sample metadata variables
(Var2) to overview sources of variance in data
```{r}
lipTumorRes$data.var.source
```

Display significant associations between features (Var1) and cancer recurrence (Var2).
```{r}
lipTumorRes$sig.feat.tab
```

Visualize molecular signatures of significant features in data. Note that features
in rows are ordered by t-statistics.
```{r}
lipTumorRes$sig.feat.heat
```

Visualize top significant features identified in Tumor samples (ns: p > 0.05,
$*$: p <= 0.05, $**$: p <= 0.01)
```{r fig.height=10}
# Prepare data matrix and metadata including both Tumor and Normal samples
datMat <- lipTissueRes$data
smpAnnoTab <- lipTissueRes$smpMetadata
# Prepare significant feature table
featSigAssoTab <- lipTumorRes$sig.feat.tab
num_sigFeats <- 6
gpLevel = c('Yes_Tumor', 'No_Tumor', 'Yes_Normal', 'No_Normal')
condiCol = c('grey50', 'firebrick') #c(Normal, Tumor)
# Extract top significant features and prepare needed information
topSigFeats <- featSigAssoTab[1:num_sigFeats,] %>%
  dplyr::pull(Var1)
topSigFeatDat <- tibble::as_tibble(datMat[topSigFeats,], rownames = 'Var1') %>%
  tidyr::pivot_longer(cols = -'Var1', names_to = 'Sample', values_to = 'Abundance') %>%
  dplyr::left_join(smpAnnoTab, by = 'Sample') %>%
  dplyr::mutate(Var1 = factor(Var1, levels = topSigFeats),
                Recur_Condi = paste0(Recurrence, '_', Condition),
                Recur_Condi = factor(Recur_Condi, levels = gpLevel))
ggplot(topSigFeatDat, aes(x=Recur_Condi, y=Abundance, col=Condition, fill=Recurrence)) +
  geom_boxplot(alpha = 1, outlier.shape = NA, linewidth = 1) +
  geom_jitter(position = position_jitter(0.2), size = 2, show.legend = F) +
  ggpubr::stat_compare_means(method = 't.test', paired = F, method.args = list(var.equal = T),
                             comparisons = list(c('Yes_Tumor', 'No_Tumor'), c('Yes_Normal', 'No_Normal')),
                             label = 'p.signif', tip.length = 0.015, bracket.size = 0.7, size = 4) +
  labs(x = 'Recurrence') +
  scale_color_manual(values = condiCol) +
  scale_fill_manual(values=c('#00BFC4', '#F8766D')) +
  facet_wrap(vars(Var1), scales = 'free') +
  th + theme(strip.text = element_text(size = 13, face = 'bold'),
             axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

Assess data power by p-value histogram
```{r}
lipTumorRes$pval.hist
```

There exists no PC significantly associated with cancer recurrence
```{r}
lipTumorRes$sig.pc.tab
```

# Plasma Proteomics
We first focused on analysis of Baseline plasma samples because (1) what we are
most interested in is predicting recurrence, and (2) Baseline samples are less
complex compared to Follow-up samples.

## Baseline samples

Note that plates where samples were located are also included in linear model for
more reliable t-test results because they are significantly associated with top PCs.
```{r}
# Load preprocessed data
lipBase <- readRDS('./data/Discovery/AG_Hopf/lipBaseVsn_B1WBC25.rds')

# Perform single-omics analysis
lipBaseRes <- doSingleOmicsAnalysis(lipBase, soa_metaVar = 'Recurrence', num_sigFeats = 6,
                                    pca_method = 'ppca', num_PCs = 20, num_PCfeats = 30,
                                    use_proDA = T, pca_metaVar = c('Recurrence', 'Gender', 'Age',
                                                                   'Smoking', 'Stage', 'Adjuvant',
                                                                   'Plate'),
                                    soa_unwantVar = 'Plate')
```

Display significant associations between PCs (Var1) and sample metadata variables
(Var2) to overview sources of variance in data
```{r}
lipBaseRes$data.var.source
```

Display significant associations between features (Var1) and cancer recurrence (Var2).
Owing to missing values, there may be situation that only few observations in sample
groups, resulting in unreliable t-statistics. To this end, we used probabilistic
dropout model (proDA) to account for data missingness.
```{r}
lipBaseRes$sig.feat.tab
```

Visualize molecular signatures of significant features in data. Note that features
in rows are ordered by t-statistics.
```{r}
lipBaseRes$sig.feat.heat
```

Visualize top significant features
```{r}
lipBaseRes$top.sig.feat.dist
```

Assess data power by p-value histogram
```{r}
lipBaseRes$pval.hist
```

Display significant associations between PCs (Var1) and cancer recurrence (Var2)
```{r}
lipBaseRes$sig.pc.tab
```

Visualize significant PCs and display their top features with greatest absolute loadings
```{r}
lipBaseRes$sig.pc.dist$PC5
lipBaseRes$pc.top.feat.tab$PC5
```

