---
title: 'Feature selection: Machine learning models'
author: "Qian-Wu Liao"
output:
  workflowr::wflow_html:
    toc: true
    toc_float: true
    code_folding: hide
---

<font size='4'> Description: Do feature selection using machine learning models
(e.g., random forest) trained on combined data of cohorts to identify potential
biomarkers for predicting lung cancer recurrence. </font>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 8, fig.width = 10, warning = F)
knitr::opts_knit$set(root.dir = '/Users/qianwu/Desktop/SMART-CARE_LungCancer')
```

Load libraries
```{r library loading, message = F}
library('spsUtil')
library('randomForest')
library('missForest')
library(xgboost)
library('caret')
library('glmnet')
library('SummarizedExperiment')
library('tidyverse')
# Load user-defined functions
source('./code/misc.R')
source('./code/ml_funcs.R')

# Set plot theme
th <- theme_bw(base_size = 15) +
  theme(axis.title = element_text(face = 'bold'),
        axis.text = element_text(face = 'bold'),
        axis.ticks = element_line(linewidth = 0.8),
        legend.text = element_text(size = 15))
```

# RF

```{r}
getRFRes <- function(se) {
  #' Prepare imputed input data for RF, do initial feature selection, remove highly
  #' correlated features, and train RF
  
  # Prepare input data for RF
  # Retrieve data matrix from SE object
  datMat <- t(assay(se)) %>%
    tibble::as_tibble(rownames = 'Sample')
  # Retrieve patient recurrence annotations
  recurAnno <- tibble::as_tibble(colData(se), rownames = 'Sample') %>%
    dplyr::select(Sample, Recurrence)
  # Combine all needed information
  allFeats <- dplyr::left_join(recurAnno, datMat, by = 'Sample') %>%
    tibble::column_to_rownames('Sample')
  # Do initial feature selection
  # Identify recurrence-related statistically significant features
  soaRes <- doSOA(se, meta_var = 'Recurrence', use_limma = T)
  sigFeats <- dplyr::select(allFeats, c(Recurrence, soaRes$featSigAssoRes$Var1))
  # Cluster highly correlated features and keep only one representative of each cluster
  sigFeatClusters <- rmCorrFeats(t(sigFeats[, -1]), cutoff = 0.8)
  uncorrSigFeats <- dplyr::select(sigFeats, c(Recurrence, rownames(sigFeatClusters$reducedData)))
  
  # Run RF
  # Set random seed for reproducible outcomes
  set.seed(42)
  x <- as.matrix(uncorrSigFeats[, -1])
  y <- uncorrSigFeats[, 1]
  rfRes <- runRF(x, y, targetClass = 'Yes', iter = 1000, split_method = 'random split', save_RF = T)
  
  return(list(rfRes = rfRes, featClusters = sigFeatClusters))
}
```

## NTP (AG Klin.)
Normal Tissue DIA Proteomics from AG Klingm端ller

**Average model performance (Uncorrelated significant features, iter = 1000)**
```{r}
# Load RF training results
rfRes <- readRDS('./data/Discovery/rf/rfResSig_proNormal_randSplit_1000_Klin.rds')

# Report summarized scores
trainedModels <- rfRes$rfRes$rfRes
# Prepare predictions and ground truths
pred <- as.list(rep(NA, length(trainedModels)))
for (i in seq_len(length(trainedModels))) {
  pred[[i]] <- trainedModels[[i]]$test$predicted
}
truth <- rfRes$rfRes$y_test
auc_roc <- rfRes$rfRes$auc_roc
summarizePredPower(pred, truth, auc_roc)
```

**Top important features with highest averaged ranks**
```{r}
# Process feature importance scores
# Rank feature importance scores for all iterations
rankedFeatImpoTab <- apply(rfRes$rfRes$MDG, 2, function(featImpo) {
  orderIdx <- order(featImpo, decreasing = T)
  featImpoRanks <- seq_along(orderIdx)
  names(featImpoRanks) <- orderIdx
  featImpo[as.numeric(names(featImpoRanks))] <- featImpoRanks
  return(featImpo)
})
# Compute mean, SD, and CI of each feature across iterations
medianFeatImpo <- apply(rankedFeatImpoTab, 1, median)
sdFeatImpo <- apply(rankedFeatImpoTab, 1, sd)
lowerCIFeatImpo <- apply(rankedFeatImpoTab, 1, function(featImpoRanks) {
  calcCI(featImpoRanks, bootstrap = T)[1]
})
upperCIFeatImpo <- apply(rankedFeatImpoTab, 1, function(featImpoRanks) {
  calcCI(featImpoRanks, bootstrap = T)[2]
})
# Collect all needed information into a table
Combined_proTissue_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsnBC.rds')
featAnnoTab <- tibble::as_tibble(rowData(Combined_proTissue_Klin), rownames = 'Feature') %>%
  dplyr::rename(Gene = PG.Genes)
medianRankedFeatImpoTab <- data.frame(Feature = names(medianFeatImpo),
                                      ImpoMedian = medianFeatImpo,
                                      ImpoSD = sdFeatImpo,
                                      ImpoLowerCI = lowerCIFeatImpo,
                                      ImpoUpperCI = upperCIFeatImpo) %>%
  dplyr::arrange(ImpoMedian) %>%
  dplyr::left_join(featAnnoTab, by = 'Feature') %>%
  dplyr::mutate(Gene = stringr::str_remove(Gene, ';.+'),
                Feature = stringr::str_remove(Feature, ';.+'))

# Visualize top important features
num_impoFeats <- 15
topFeatImpoTab <- medianRankedFeatImpoTab[seq_len(num_impoFeats),]
ggplot(topFeatImpoTab, aes(x=ImpoMedian, y=factor(Gene, levels = rev(Gene)))) +
  geom_errorbar(aes(xmin=ImpoLowerCI, xmax=ImpoUpperCI)) +
  geom_point(size = 6) +
  labs(x = 'Averaged Rank of Gene Importance', y = 'Gene') +
  th +
  theme(axis.title = element_text(size = 28),
        axis.text = element_text(size = 16))
# ggsave('./output/Discovery/group_meeting/rf_topImpo_combined_proNormal_Klin.png',
#        device = 'png', dpi = 400, height = 8, width = 10)
```

## TTP (AG Klin.)
Tumor Tissue DIA Proteomics from AG Klingm端ller

**Average model performance (Uncorrelated significant features, iter = 1000)**
```{r}
# Load RF training results
rfRes <- readRDS('./data/Discovery/rf/rfResSig_proTumor_randSplit_1000_Klin.rds')

# Report summarized scores
trainedModels <- rfRes$rfRes$rfRes
# Prepare predictions and ground truths
pred <- as.list(rep(NA, length(trainedModels)))
for (i in seq_len(length(trainedModels))) {
  pred[[i]] <- trainedModels[[i]]$test$predicted
}
truth <- rfRes$rfRes$y_test
auc_roc <- rfRes$rfRes$auc_roc
summarizePredPower(pred, truth, auc_roc)
```

**Top important features with highest averaged ranks**
```{r}
# Process feature importance scores
# Rank feature importance scores for all iterations
rankedFeatImpoTab <- apply(rfRes$rfRes$MDG, 2, function(featImpo) {
  orderIdx <- order(featImpo, decreasing = T)
  featImpoRanks <- seq_along(orderIdx)
  names(featImpoRanks) <- orderIdx
  featImpo[as.numeric(names(featImpoRanks))] <- featImpoRanks
  return(featImpo)
})
# Compute mean, SD, and CI of each feature across iterations
meanFeatImpo <- apply(rankedFeatImpoTab, 1, mean)
sdFeatImpo <- apply(rankedFeatImpoTab, 1, sd)
lowerCIFeatImpo <- apply(rankedFeatImpoTab, 1, function(featImpoRanks) {
  calcCI(featImpoRanks, bootstrap = T)[1]
})
upperCIFeatImpo <- apply(rankedFeatImpoTab, 1, function(featImpoRanks) {
  calcCI(featImpoRanks, bootstrap = T)[2]
})
# Collect all needed information into a table
Combined_proTissue_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsnBC.rds')
featAnnoTab <- tibble::as_tibble(rowData(Combined_proTissue_Klin), rownames = 'Feature') %>%
  dplyr::rename(Gene = PG.Genes)
meanRankedFeatImpoTab <- data.frame(Feature = names(meanFeatImpo),
                                    ImpoMean = meanFeatImpo,
                                    ImpoSD = sdFeatImpo,
                                    ImpoLowerCI = lowerCIFeatImpo,
                                    ImpoUpperCI = upperCIFeatImpo) %>%
  dplyr::arrange(ImpoMean) %>%
  dplyr::left_join(featAnnoTab, by = 'Feature') %>%
  dplyr::mutate(Gene = stringr::str_remove(Gene, ';.+'),
                Feature = stringr::str_remove(Feature, ';.+'))

# Visualize top important features
num_impoFeats <- 15
topFeatImpoTab <- meanRankedFeatImpoTab[seq_len(num_impoFeats),]
ggplot(topFeatImpoTab, aes(x=ImpoMean, y=factor(Gene, levels = rev(Gene)))) +
  geom_errorbar(aes(xmin=ImpoLowerCI, xmax=ImpoUpperCI)) +
  geom_point(size = 6) +
  labs(x = 'Averaged Rank of Feature Importance', y = 'Feature') +
  th
```

## NTP (AG Krij.)
Normal Tissue DIA Proteomics from AG Krijgsveld

```{r message=F, eval=F, include=F}
# Prepare input data, which includes imputation, initial feature selection, correlated
# feature removal, and RF training. Take Normal Tissue DIA Proteomics as example

# Load preprocessed data
proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds')
proTissue <- proTissue_Krij
# Subset data
proNormal <- proTissue[, colData(proTissue)$Condition == 'Normal']
proTumor <- proTissue[, colData(proTissue)$Condition == 'Tumor']

# Normal tissues
# Impute missing values using missForest
# dat <- t(assay(proNormal))
# impuDat <- missForest(dat, maxiter = 10, verbose = T)$ximp %>%
#   t()
# assay(proNormal) <- impuDat
# saveRDS(proNormal, './data/Discovery/AG_Krijgsveld/impuByMF/combined_proNormalVsnImpu.rds')

# Load imputed data
proNormal_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/impuByMF/combined_proNormalVsnImpu.rds')
proNormal <- proNormal_Krij

# Prepare input data for RF
# Retrieve data matrix from SE object
datMat <- t(assay(proNormal)) %>%
  tibble::as_tibble(rownames = 'Sample')
# Retrieve patient recurrence annotations
recurAnno <- tibble::as_tibble(colData(proNormal), rownames = 'Sample') %>%
  dplyr::select(Sample, Recurrence)
# Include recurrence information into data matrix
allFeats <- dplyr::left_join(recurAnno, datMat, by = 'Sample') %>%
  tibble::column_to_rownames('Sample')
# Do initial feature selection
# Identify recurrence-related statistically significant features
soaRes <- doSOA(proNormal, meta_var = 'Recurrence', use_limma = T)
sigFeats <- dplyr::select(allFeats, c(Recurrence, soaRes$featSigAssoRes$Var1))
# Cluster highly correlated features and keep only one representative of each cluster
sigFeatClusters <- rmCorrFeats(t(sigFeats[, -1]), cutoff = 0.8)
uncorrSigFeats <- dplyr::select(sigFeats, c(Recurrence, rownames(sigFeatClusters$reducedData)))

# Run RF
# Set random seed for reproducible outcomes
set.seed(42)
x <- as.matrix(uncorrSigFeats[, -1])
y <- uncorrSigFeats[, 1]
spsUtil::quiet(
  rfResSig <- runRF(x, y, targetClass = 'Yes', iter = 1, split_method = 'random split', save_RF = T)
)

# Overview trained RF
# print(rfResSig$rfRes[[1]])
# Play around with thresholds
cutoff <- 0.5
expectedVals <- rfResSig$y_test[[1]]
predictedVals <- as.numeric(rfResSig$rfRes[[1]]$test$votes[, '1'] > cutoff) %>%
  factor()
caret::confusionMatrix(data = predictedVals, reference = expectedVals, positive = '1')
```

**Average model performance (Uncorrelated significant features, iter = 1000)**
```{r}
# Load RF training results
rfResSig <- readRDS('./data/Discovery/rf/rfResSig_proNormal_randSplit_1000_Krij.rds')

# Report summarized scores
trainedModels <- rfRes$rfRes$rfRes
# Prepare predictions and ground truths
pred <- as.list(rep(NA, length(trainedModels)))
for (i in seq_len(length(trainedModels))) {
  pred[[i]] <- trainedModels[[i]]$test$predicted
}
truth <- rfRes$rfRes$y_test
auc_roc <- rfRes$rfRes$auc_roc
summarizePredPower(pred, truth, auc_roc)
```

**Top important features with highest averaged ranks**
```{r}
# Process feature importance scores
# Rank feature importance scores for all iterations
rankedFeatImpoTab <- apply(rfResSig$rfRes$MDG, 2, function(featImpo) {
  orderIdx <- order(featImpo, decreasing = T)
  featImpoRanks <- seq_along(orderIdx)
  names(featImpoRanks) <- orderIdx
  featImpo[as.numeric(names(featImpoRanks))] <- featImpoRanks
  return(featImpo)
})
# Compute mean, SD, and CI of each feature across iterations
meanFeatImpo <- apply(rankedFeatImpoTab, 1, mean)
sdFeatImpo <- apply(rankedFeatImpoTab, 1, sd)
lowerCIFeatImpo <- apply(rankedFeatImpoTab, 1, function(featImpoRanks) {
  calcCI(featImpoRanks, bootstrap = T)[1]
})
upperCIFeatImpo <- apply(rankedFeatImpoTab, 1, function(featImpoRanks) {
  calcCI(featImpoRanks, bootstrap = T)[2]
})
# Collect all needed information into a table
Combined_proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds')
featAnnoTab <- tibble::as_tibble(rowData(Combined_proTissue_Krij), rownames = 'Feature') %>%
  dplyr::select(Feature, Genes) %>%
  dplyr::rename(Gene = Genes)
meanRankedFeatImpoTab <- data.frame(Feature = names(meanFeatImpo),
                                    ImpoMean = meanFeatImpo,
                                    ImpoSD = sdFeatImpo,
                                    ImpoLowerCI = lowerCIFeatImpo,
                                    ImpoUpperCI = upperCIFeatImpo) %>%
  dplyr::arrange(ImpoMean) %>%
  dplyr::left_join(featAnnoTab, by = 'Feature') %>%
  dplyr::mutate(Gene = stringr::str_remove(Gene, ';.+'),
                Feature = stringr::str_remove(Feature, ';.+'))

# Visualize top important features
num_impoFeats <- 15
topFeatImpoTab <- meanRankedFeatImpoTab[seq_len(num_impoFeats),]
ggplot(topFeatImpoTab, aes(x=ImpoMean, y=factor(Gene, levels = rev(Gene)))) +
  geom_errorbar(aes(xmin=ImpoLowerCI, xmax=ImpoUpperCI)) +
  geom_point(size = 6) +
  labs(x = 'Averaged Rank of Feature Importance', y = 'Feature') +
  th
# ggsave('./output/Discovery/group_meeting/combined_rf_featImpo_proNormal_Krij.png',
#        device = 'png', dpi = 400, height = 8, width = 10)
```

```{r eval=F, include=F}
# Visualize abundances of top important features
# Prepare data matrix and metadata including both Tumor and Normal samples
Combined_proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds')
datMat <- assay(Combined_proTissue_Krij)
rownames(datMat) <- stringr::str_remove(rownames(datMat), ';.+')
smpAnnoTab <- tibble::as_tibble(colData(Combined_proTissue_Krij), rownames = 'Sample')

num_impoFeats <- 6
gpLevel = c('Yes_Normal', 'No_Normal', 'Yes_Tumor', 'No_Tumor')
condiCol = c('firebrick', 'grey50') #c(Normal, Tumor)
# Extract top important features and prepare needed information
topImpoFeats <- meanRankedFeatImpoTab[1:num_impoFeats,] %>%
  dplyr::select(Feature, Gene) %>%
  dplyr::mutate(newFeat = paste0(Feature, ' (', Gene, ')'),
                newFeat = factor(newFeat, levels = unique(newFeat)))
topImpoFeatDat <- tibble::as_tibble(datMat[topImpoFeats$Feature,], rownames = 'Feature') %>%
  tidyr::pivot_longer(cols = -'Feature', names_to = 'Sample', values_to = 'Abundance') %>%
  dplyr::left_join(topImpoFeats, by = 'Feature') %>%
  dplyr::left_join(smpAnnoTab, by = 'Sample') %>%
  dplyr::mutate(Recur_Condi = paste0(Recurrence, '_', Condition),
                Recur_Condi = factor(Recur_Condi, levels = gpLevel))
ggplot(topImpoFeatDat, aes(x=Recur_Condi, y=Abundance, col=Condition, fill=Recurrence)) +
  geom_boxplot(alpha = 1, outlier.shape = NA, linewidth = 1) +
  geom_jitter(position = position_jitter(0.2), size = 2, show.legend = F) +
  ggpubr::stat_compare_means(method = 't.test', paired = F, method.args = list(var.equal = T),
                             comparisons = list(c('Yes_Normal', 'No_Normal'), c('Yes_Tumor', 'No_Tumor')),
                             label = 'p.signif', tip.length = 0.015, bracket.size = 0.7, size = 4) +
  labs(x = 'Recurrence') +
  scale_color_manual(values = condiCol) +
  scale_fill_manual(values=c('#00BFC4', '#F8766D')) +
  facet_wrap(vars(newFeat), scales = 'free') +
  th +
  theme(strip.text = element_text(size = 13, face = 'bold'),
             axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

## TTP (AG Krij.)
Tumor Tissue DIA Proteomics from AG Krijgsveld

**Average model performance (Uncorrelated significant features, iter = 1000)**
```{r}
# Load RF training results
rfRes <- readRDS('./data/Discovery/rf/rfResSig_proTumor_randSplit_1000_Krij.rds')

# Report summarized scores
trainedModels <- rfRes$rfRes$rfRes
# Prepare predictions and ground truths
pred <- as.list(rep(NA, length(trainedModels)))
for (i in seq_len(length(trainedModels))) {
  pred[[i]] <- trainedModels[[i]]$test$predicted
}
truth <- rfRes$rfRes$y_test
auc_roc <- rfRes$rfRes$auc_roc
summarizePredPower(pred, truth, auc_roc)
```

**Top important features with highest averaged ranks**
```{r}
# Process feature importance scores
# Rank feature importance scores for all iterations
rankedFeatImpoTab <- apply(rfRes$rfRes$MDG, 2, function(featImpo) {
  orderIdx <- order(featImpo, decreasing = T)
  featImpoRanks <- seq_along(orderIdx)
  names(featImpoRanks) <- orderIdx
  featImpo[as.numeric(names(featImpoRanks))] <- featImpoRanks
  return(featImpo)
})
# Compute mean, SD, and CI of each feature across iterations
meanFeatImpo <- apply(rankedFeatImpoTab, 1, mean)
sdFeatImpo <- apply(rankedFeatImpoTab, 1, sd)
lowerCIFeatImpo <- apply(rankedFeatImpoTab, 1, function(featImpoRanks) {
  calcCI(featImpoRanks, bootstrap = T)[1]
})
upperCIFeatImpo <- apply(rankedFeatImpoTab, 1, function(featImpoRanks) {
  calcCI(featImpoRanks, bootstrap = T)[2]
})
# Collect all needed information into a table
Combined_proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds')
featAnnoTab <- tibble::as_tibble(rowData(Combined_proTissue_Krij), rownames = 'Feature') %>%
  dplyr::select(Feature, Genes) %>%
  dplyr::rename(Gene = Genes)
meanRankedFeatImpoTab <- data.frame(Feature = names(meanFeatImpo),
                                    ImpoMean = meanFeatImpo,
                                    ImpoSD = sdFeatImpo,
                                    ImpoLowerCI = lowerCIFeatImpo,
                                    ImpoUpperCI = upperCIFeatImpo) %>%
  dplyr::arrange(ImpoMean) %>%
  dplyr::left_join(featAnnoTab, by = 'Feature') %>%
  dplyr::mutate(Gene = stringr::str_remove(Gene, ';.+'),
                Feature = stringr::str_remove(Feature, ';.+'))

# Visualize top important features
num_impoFeats <- 15
topFeatImpoTab <- meanRankedFeatImpoTab[seq_len(num_impoFeats),]
ggplot(topFeatImpoTab, aes(x=ImpoMean, y=factor(Gene, levels = rev(Gene)))) +
  geom_errorbar(aes(xmin=ImpoLowerCI, xmax=ImpoUpperCI)) +
  geom_point(size = 6) +
  labs(x = 'Averaged Rank of Feature Importance', y = 'Feature') +
  th
```

# XGBoost

```{r}
iterXGBoost <- function(se, booster = 'gbtree') {
  #' Prepare imputed input data for XGBoost, do initial feature selection, remove
  #' highly correlated features, and train XGBoost
  
  # Do initial feature selection
  # Identify recurrence-related statistically significant features
  soaRes <- doSOA(se, meta_var = 'Recurrence', use_limma = T)
  x <- t(assay(se))
  x <- x[, colnames(x) %in% soaRes$featSigAssoRes$Var1]
  # Cluster highly correlated features and keep only one representative of each cluster
  featClusters <- rmCorrFeats(t(x), cutoff = 0.8)
  x <- t(featClusters$reducedData)
  y <- colData(se)$Recurrence
  
  # Train XGBoost with tree or linear booster
  set.seed(42)
  if (booster == 'gbtree') {
    xgbRes <- runXGBoost(x, y, targetClass = 'Yes', iter = 1000, booster = 'gbtree',
                         nrounds = 100, maxDepth = 6, eta = 0.3, trainSet_ratio = 0.6,
                         plot_ROC = F, save_model = T)
  } else if (booster == 'gblinear') {
    xgbRes <- runXGBoost(x, y, targetClass = 'Yes', iter = 1000, booster = 'gblinear',
                         nrounds = 100, trainSet_ratio = 0.6, plot_ROC = F, save_model = T)
  }
  
  return(list(xgbRes = xgbRes, featClusters = featClusters))
}
```

```{r message=F, eval=F, include=F}
## NTP (AG Krij.)
# Normal Tissue DIA Proteomics from AG Krijgsveld

# XGBoost supports missing values by default. Branch directions for missing values
# are learned during training. Note that 'gblinear' booster treats missing values
# as zeros.

# Load imputed data
proNormal_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/impuByMF/combined_proNormalVsnImpu.rds')
proNormal <- proNormal_Krij
# Do initial feature selection
# Identify recurrence-related statistically significant features
soaRes <- doSOA(proNormal, meta_var = 'Recurrence', use_limma = T)
x <- t(assay(proNormal))
x <- x[, colnames(x) %in% soaRes$featSigAssoRes$Var1]
# Cluster highly correlated features and keep only one representative of each cluster
featClusters <- rmCorrFeats(t(x), cutoff = 0.8)
x <- t(featClusters$reducedData)
y <- colData(proNormal)$Recurrence
y <- ifelse(test = y == 'Yes', yes = 1, no = 0)

# Determine best number of iteration using cross-validation
# allDat <- xgb.DMatrix(data = x, label = y)
# bestIter <- xgb.cv(data = allDat, params = paramsTree, nrounds = 20, nfold = 5, verbose = T)
# ggplot(bestIter$evaluation_log, aes(x=iter, y=train_logloss_mean)) +
#   geom_line(linewidth = 1) +
#   geom_line(aes(y=test_logloss_mean), color = 'firebrick', linewidth = 1) +
#   labs(x = 'Iteration', y = 'Loss') +
#   th
#### => Learning curves look weird...

# Split data into training, validation, and test sets
# Set random seed for reproducible outcomes
set.seed(42)
# trainIdx <- subsetTrainData(x, y, split_method = 'bootstrap', trainSet_ratio = 0.8)
trainIdx <- caret::createDataPartition(y, times = 1, p = 0.6, list = T)
x_train <- x[trainIdx[[1]],, drop = F]
y_train <- y[trainIdx[[1]]]
restIdx <- seq_along(y)[!seq_along(y) %in% trainIdx[[1]]]
ySub <- y[restIdx]
validIdx <- caret::createDataPartition(ySub, times = 1, p = 0.5, list = T)
x_valid <- x[restIdx[validIdx[[1]]],, drop = F]
y_valid <- y[restIdx[validIdx[[1]]]]
x_test <- x[restIdx[-validIdx[[1]]],, drop = F]
y_test <- y[restIdx[-validIdx[[1]]]]
# Collect split data into xgb.DMatrix objects
trainDat <- xgb.DMatrix(data = x_train, label = y_train)
validDat <- xgb.DMatrix(data = x_valid, label = y_valid)
testDat <- xgb.DMatrix(data = x_test, label = y_test)

# Train XGBoost with tree booster
# Prepare parameters for training XGBoost
paramsTree <- list(booster = 'gbtree',
                   objective = 'binary:logistic',
                   eval_metric = 'error',
                   eval_metric = 'logloss',
                   max_depth = 6,
                   eta = 0.3)
# Determine best number of iteration
watchlist <- list(train = trainDat, valid = validDat)
treeBoost <- xgb.train(data = trainDat, params = paramsTree, nrounds = 50,
                       watchlist = watchlist, verbose = 1, early_stopping_rounds = 4)
ggplot(treeBoost$evaluation_log, aes(x=iter, y=train_logloss)) +
  geom_line(linewidth = 1) +
  geom_line(aes(y=valid_logloss), color = 'firebrick', linewidth = 1) +
  labs(x = 'Iteration', y = 'Loss') +
  th
# Retrain XGBoost with combined data including training and validation
# combinedTrainIdx <- c(trainIdx[[1]], restIdx[validIdx[[1]]])
# combinedTrainDat <- xgb.DMatrix(data = x[combinedTrainIdx,, drop = F],
#                                 label = y[combinedTrainIdx])
# treeBoost <- xgb.train(data = combinedTrainDat, params = paramsTree, nrounds = 3, verbose = 1)

# Predict unseen data using trained XGBoost
# Play around with thresholds
cutoff <- 0.5
truth <- getinfo(testDat, 'label') %>%
  factor()
pred <- as.numeric(predict(treeBoost, testDat) > cutoff) %>%
  factor()
caret::confusionMatrix(data = pred, reference = truth, positive = '1')

# Visualize results of trained XGBoost
# Structure of trees
# xgb.plot.tree(model = treeBoost)
# Feature importance
impoTab <- xgb.importance(model = treeBoost)
xgb.ggplot.importance(importance_matrix = impoTab) +
  th
# ggsave('./output/Discovery/group_meeting/combined_xgboost_featImpo_proNormal_Krij.png',
#        device = 'png', dpi = 400, height = 8, width = 10)
```

## NTP (AG Klin.)
Normal Tissue DIA Proteomics from AG Klingm端ller

**Tree booster**
```{r}
# Load imputed data
# proNormal_Klin <- readRDS('./data/Discovery/AG_Klingmueller/impuByMF/proNormalVsnImpu.rds')
# treeBoostRes <- iterXGBoost(proNormal_Klin, booster = 'gbtree')
# saveRDS(treeBoostRes, './data/Discovery/xgboost/tree_xgbResSig_proNormal_randSplit_logloss_1000_Klin.rds')
treeBoostRes <- readRDS('./data/Discovery/xgboost/tree_xgbResSig_proNormal_randSplit_logloss_1000_Klin.rds')

# Report summarized scores
# Prepare predictions and ground truths
pred <- treeBoostRes$xgbRes$y_pred
truth <- treeBoostRes$xgbRes$y_truth
auc_roc <- treeBoostRes$xgbRes$auc_roc
summarizePredPower(pred, truth, auc_roc)

# Take only models with good performance?
bestModels <- which(treeBoostRes$xgbRes$auc_roc == 1)
```

**Linear booster**
```{r}
# Load imputed data
# proNormal_Klin <- readRDS('./data/Discovery/AG_Klingmueller/impuByMF/proNormalVsnImpu.rds')
# linearBoostRes <- iterXGBoost(proNormal_Klin, booster = 'gblinear')
# saveRDS(linearBoostRes, './data/Discovery/xgboost/linear_xgbResSig_proNormal_randSplit_logloss_1000_Klin.rds')
linearBoostRes <- readRDS('./data/Discovery/xgboost/linear_xgbResSig_proNormal_randSplit_error_1000_Klin.rds')

# Report summarized scores
# Prepare predictions and ground truths
pred <- linearBoostRes$xgbRes$y_pred
truth <- linearBoostRes$xgbRes$y_truth
auc_roc <- linearBoostRes$xgbRes$auc_roc
summarizePredPower(pred, truth, auc_roc)

# impoTab <- xgb.importance(model = linearBoost$modelRes[[1]])
# xgb.ggplot.importance(importance_matrix = impoTab,  top_n = 20) +
#   th
```

# Lasso logistic regression

```{r}
iterLogisR <- function(se, iter = 1000, cvFold = 10, cvMeasure = 'auc') {
  #' Prepare imputed input data, do initial feature selection, remove highly correlated
  #' features, and train lasso logistic regression model
  
  # Do initial feature selection
  # Identify recurrence-related statistically significant features
  soaRes <- doSOA(se, meta_var = 'Recurrence', use_limma = T)
  x <- t(assay(se))
  x <- x[, colnames(x) %in% soaRes$featSigAssoRes$Var1]
  # Cluster highly correlated features and keep only one representative of each cluster
  featClusters <- rmCorrFeats(t(x), cutoff = 0.8)
  x <- t(featClusters$reducedData)
  y <- colData(se)$Recurrence
  
  # Train lasso logistic regression model
  set.seed(42)
  lrRes <- runLogisR(x, y, targetClass = 'Yes', iter = iter, regularized_method = 'lasso',
                     cvFold = cvFold, cvMeasure = cvMeasure, used_lambda = 'lambda.min',
                     trainSet_ratio = 0.8, split_method = 'random split',
                     plot_ROC = F, save_model = T)
  
  return(list(lrRes = lrRes, featClusters = featClusters))
}
```

## NTP (AG Klin.)
Normal Tissue DIA Proteomics from AG Klingm端ller

**Models with no any non-zero beta are removed.**
```{r warning=T}
# Load imputed data
# proNormal_Klin <- readRDS('./data/Discovery/AG_Klingmueller/impuByMF/proNormalVsnImpu.rds')
# lrRes <- iterLogisR(proNormal_Klin, iter = 1000, cvFold = 5, cvMeasure = 'deviance')
# saveRDS(lrRes, './data/Discovery/logisR/lrResSig_proNormal_randSplit_1000_5fold_deviance_lmin_Klin.rds')
lrRes <- readRDS('./data/Discovery/logisR/lrResSig_proNormal_randSplit_1000_5fold_deviance_lmin_Klin.rds')

uselessModels <- which(lrRes$lrRes$nNonZero == 0) #higher lambda results in simpler model
# for (i in uselessModels) {
#   pred <- lrRes$lrRes$y_pred[[i]]
#   truth <- lrRes$lrRes$y_truth[[i]]
#   scores <- caret::confusionMatrix(data = pred, reference = truth, positive = '1')
#   cat(scores$overall[['Accuracy']], '\n')
# } # Models that got only intercept make random predictions, 0.5 accuracy and AUC-ROC

# Report summarized scores
# Prepare predictions and ground truths
pred <- lrRes$lrRes$y_pred[-uselessModels]
truth <- lrRes$lrRes$y_truth[-uselessModels]
auc_roc <- lrRes$lrRes$auc_roc[-uselessModels]
suppressWarnings(summarizePredPower(pred, truth, auc_roc)) #warning due to sometimes only pos or neg predictions
```
