---
title: 'Feature selection: Machine learning models (with feature filtering)'
author: "Qian-Wu Liao"
date: "2024-10-08"
output:
  workflowr::wflow_html:
    toc: true
    toc_float: true
    code_folding: hide
---

<font size='4'> Description: Do feature selection with machine learning models trained
on combined data of cohorts (Method Development and Discovery) to identify potential
biomarkers for predicting lung cancer recurrence. Additional feature filtering is
conducted on subdatasets to exclude risk of biases introduced by imputation method
(missForest) used in our workflow, which can impede true but weaker signals contributed
by features more complete. ML models used currently include lasso logistic regression,
random forest, and XGBoost. </font>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 8, fig.width = 10, warning = F)
knitr::opts_knit$set(root.dir = '/Users/qianwu/Desktop/SMART-CARE_LungCancer')
```

Load libraries
```{r library loading, message = F}
library('spsUtil')
library(ggh4x)
library(DT)
library('randomForest')
library('missForest')
library(xgboost)
library('caret')
library('glmnet')
library(ggvenn)
library('SummarizedExperiment')
library('tidyverse')
# Load user-defined functions
source('./code/misc.R')
source('./code/ml_funcs.R')

# Set plot theme
th <- theme_bw(base_size = 15) +
  theme(axis.title = element_text(face = 'bold'),
        axis.text = element_text(face = 'bold'),
        axis.ticks = element_line(linewidth = 0.8),
        legend.text = element_text(size = 15))
```

# TTP (AG Klin.)
Tumor Tissue DIA Proteomics from AG Klingmüller

## Lasso logistic regression

<font size='5'> **Stability selection** </font>

**Performance statistics of 1000 models**
```{r warning=T}
# Load preprocessed data
proTissue_Klin <- readRDS('./data/Discovery/AG_Klingmueller/proTissueVsn.rds')
proTumor_Klin <- proTissue_Klin[, colData(proTissue_Klin)$Condition %in% 'Tumor']
# Load SOA results
proTumorRes_Klin <- readRDS('./data/Discovery/AG_Klingmueller/soaRes/proTumorRes_SVsByProTissueInt.rds')
# Prepare adjusted data
adjProTumor_Klin <- proTumor_Klin
assay(adjProTumor_Klin) <- proTumorRes_Klin$SOA.res$dataCorrect
# Do additional feature filtering to remove features observed in less than 50% of samples
ifRmFeats <- apply(assay(adjProTumor_Klin), 1, function(featVec) {
  sum(is.na(featVec)) / length(featVec) > 0.5
})
adjProTumor_Klin <- adjProTumor_Klin[!ifRmFeats,]

# Load imputed adjusted data
impuAdjProTumor_Klin <- readRDS('./data/Discovery/AG_Klingmueller/impuData/featFilt_impuAdjProTumorVsn_SVsByProTissueInt.rds')

# Train models iteratively
# Prepare significant feature list
sigFeatList <- proTumorRes_Klin$sig.feat.tab$Var1
# lrRes <- iterLogisR(adjProTumor_Klin, doImputation = T, doInitFeatSelection = T, sigFeatList = sigFeatList,
#                     doFeatClustering = T, iter = 1000, cvFold = 5, cvMeasure = 'auc')
# saveRDS(lrRes, './data/Discovery/logisR/featFilt_lrResSOA_SVsByProTissueInt_impuAdjProTumor_0.9r_randSplit_1000_5fold_auc_lmin_Klin.rds')
# saveRDS(lrRes$impuSE, './data/Discovery/AG_Klingmueller/impuData/featFilt_impuAdjProTumorVsn_SVsByProTissueInt.rds')
lrRes <- readRDS('./data/Discovery/logisR/featFilt_lrResSOA_SVsByProTissueInt_impuAdjProTumor_0.9r_randSplit_1000_5fold_auc_lmin_Klin.rds')

# Report summarized scores
# Prepare predictions and ground truths where results from useless models are removed
uselessModels <- which(lrRes$lrRes$nNonZero == 0)
if (length(uselessModels) != 0) {
  pred <- lrRes$lrRes$y_pred[-uselessModels]
  truth <- lrRes$lrRes$y_truth[-uselessModels]
  auc_roc <- lrRes$lrRes$auc_roc[-uselessModels]
} else {
  pred <- lrRes$lrRes$y_pred
  truth <- lrRes$lrRes$y_truth
  auc_roc <- lrRes$lrRes$auc_roc
}
suppressWarnings(summarizePredPower(pred, truth, auc_roc))
```

**Top important features that are most frequently selected**
```{r}
# Visualize top important features
# Prepare feature annotation table
featAnno <- tibble::as_tibble(rowData(proTissue_Klin), rownames = 'Feature')
topImpoFeatViz <- vizTopImpoFeatsLR(lrRes, featAnno = featAnno, num_p1TopFeats = 30)
topImpoFeatViz$pick
```

**Full feature importance table**
```{r}
DT::datatable(topImpoFeatViz$fullImpoFeatTab)
```

<font size='5'> **Systematic training and evaluation** </font>

**Performance of models trained with different numbers of top important features**
```{r}
# Systematically train models with different numbers of top important features
# sysTrainModels <- doSysTrainLogisR(lrRes, impuAdjProTumor_Klin, max_numTopFeats = 50)
# saveRDS(sysTrainModels, './data/Discovery/logisR/featFilt_sysTrain_lrResSOA_SVsByProTissueInt_impuAdjProTumor_randSplit_100_5fold_auc_lmin_Klin.rds')
sysTrainModels <- readRDS('./data/Discovery/logisR/featFilt_sysTrain_lrResSOA_SVsByProTissueInt_impuAdjProTumor_randSplit_100_5fold_auc_lmin_Klin.rds')

# Visualize systematic training
vizSysTrainModelPerf(sysTrainModels, numTopFeats = 30)
# ggsave('./output/Junyan/stabSelec_Stats_proTumor.png', device = 'png', dpi = 400, height = 8, width = 10)
```
=> Top 8 important features are selected, resulting in model with AUC of 0.99 [0.92,1].\
=> Top 14 important features are selected, resulting in model with AUC of 0.988 [0.92,1].\
=> Top 18 important features are selected, resulting in model with AUC of 0.986 [0.92,1].

**Noninformative models that demonstrate effects of top important features**
```{r}
se <- impuAdjProTumor_Klin
# Train models without top important features, aka noninformative models
numTopImpoFeats <- c(5, 10, 15, 20)
# for (num in numTopImpoFeats) {
#   topImpoFeats <- topImpoFeatViz$fullImpoFeatTab$Feature[1:num]
#   # Train lasso logistic regression model
#   reducedData <- lrRes$featClusters$reducedData
#   x <- t(reducedData[!rownames(reducedData) %in% topImpoFeats,])
#   y <- colData(se)$Recurrence
#   lrResNonInfo <- runLogisR(x, y, targetClass = 'Yes', iter = 100, regularized_method = 'lasso',
#                             cvFold = 5, cvMeasure = 'auc', used_lambda = 'lambda.min',
#                             trainSet_ratio = 0.8, split_method = 'random split',
#                             plot_ROC = F, save_model = T)
#   saveRDS(lrResNonInfo, paste0('./data/Discovery/logisR/noninfo/featFilt_NonInfo_rmTop', num,
#                                '_lrResSOA_SVsByProTissueInt_impuAdjProTumor_randSplit_100_5fold_auc_lmin_Klin.rds'))
# }

# Display performance of noninformative models
meanAUC <- c()
lowerAUC <- c()
upperAUC <- c()
uselessModels <- which(lrRes$lrRes$nNonZero == 0)
if (length(uselessModels) != 0) {
  auc_roc <- lrRes$lrRes$auc_roc[-uselessModels]
} else {
  auc_roc <- lrRes$lrRes$auc_roc
}
meanAUC <- c(meanAUC, round(mean(auc_roc), 3))
ci <- calcCI(auc_roc, bootstrap = T)
lowerAUC <- c(lowerAUC, ci[1])
upperAUC <- c(upperAUC, ci[2])
for (num in numTopImpoFeats) {
  lrResNonInfo <- readRDS(paste0('./data/Discovery/logisR/noninfo/featFilt_NonInfo_rmTop', num,
                                 '_lrResSOA_SVsByProTissueInt_impuAdjProTumor_randSplit_100_5fold_auc_lmin_Klin.rds'))
  uselessModels <- which(lrResNonInfo$nNonZero == 0)
  if (length(uselessModels) != 0) {
    auc_roc <- lrResNonInfo$auc_roc[-uselessModels]
  } else {
    auc_roc <- lrResNonInfo$auc_roc
  }
  meanAUC <- c(meanAUC, round(mean(auc_roc), 3))
  ci <- calcCI(auc_roc, bootstrap = T)
  lowerAUC <- c(lowerAUC, ci[1])
  upperAUC <- c(upperAUC, ci[2])
}
aucStatTab <- data.frame(FeatSpace = c('Complete', paste0('-Top ', numTopImpoFeats)),
                         MeanAUC = meanAUC, UpperAUC = upperAUC, LowerAUC = lowerAUC) %>%
  dplyr::mutate(FeatSpace = factor(FeatSpace, levels = FeatSpace))

ggplot(aucStatTab, aes(x=FeatSpace, y=MeanAUC)) +
  geom_bar(stat = 'identity', position = position_dodge(), alpha = 0.9) +
  geom_errorbar(aes(ymin=LowerAUC, ymax=UpperAUC), width = 0.4) +
  labs(x = 'Feature space used to train 100 models', y = 'Mean AUC-ROC',
       title = 'Noninformative models') +
  th
```

<font size='5'> **Investigation of core features** </font>\
*Keep TGFBR3 in mind*

<font size='3'> Imputed data: </font>\
**Heatmap where samples are ordered by linear regression scores**
```{r}
coreFeatPlotsImpu <- vizCoreFeats(lrRes, sysTrainModels, impuAdjProTumor_Klin,
                                  heatmap_numCoreFeats = 8, colFeatAnno = 'Gene',
                                  scatter_twoCoreFeats = c('FABP4', 'FMO5'),
                                  fontsize = 10, fontsize_col = 7, fontsize_row = 11,
                                  cellwidth = 6, cellheight = 18)
coreFeatPlotsImpu$heatmap
# ggsave('./output/Junyan/stabSelec_heatmap_proNormal_impu.png', device = 'png', dpi = 400, height = 8, width = 10)
```

**Scatterplot of two core features**
```{r}
coreFeatPlotsImpu$scatterplot
# ggsave('./output/Junyan/stabSelec_top1&3_proTumor_impu.png', device = 'png', dpi = 400, height = 8, width = 10)
```

<font size='3'> Original data: </font>\
**Heatmap**
```{r}
coreFeatPlots <- vizCoreFeats(lrRes, sysTrainModels, adjProTumor_Klin,
                              heatmap_numCoreFeats = 8, colFeatAnno = 'Gene',
                              scatter_twoCoreFeats = c('FABP4', 'DYNLT3'),
                              fontsize = 10, fontsize_col = 7, fontsize_row = 11,
                              cellwidth = 6, cellheight = 18)
coreFeatPlots$heatmap
# ggsave('./output/Junyan/stabSelec_heatmap_proTumor_ori.png', plot = a, device = 'png', dpi = 400, height = 8, width = 10)
```

**Scatterplot**
```{r}
coreFeatPlots$scatterplot
```

**Missingness and availability of core features**
```{r message=F}
numTopImpoFeats <- 8
# Prepare top important feature list
topImpoFeats <- topImpoFeatViz$fullImpoFeatTab$Feature[1:numTopImpoFeats] %>%
  stringr::str_remove(';.+')
topImpoFeatAnnos <- topImpoFeatViz$fullImpoFeatTab$Annotation[1:numTopImpoFeats] %>%
  stringr::str_remove(';.+')

# Show data missingness
proTumor_Klin <- proTissue_Klin[, colData(proTissue_Klin)$Condition %in% 'Tumor']
featAnnoTab <- tibble::as_tibble(rowData(proTissue_Klin), rownames = 'Feature') %>%
  dplyr::mutate(Feature = stringr::str_remove(Feature, ';.+'),
                Gene = stringr::str_remove(Gene, ';.+'))
featAvailTab <- summExp2df(proTumor_Klin, row_id = 'Feature', col_id = 'Sample') %>%
  dplyr::select(Feature, Value, Recurrence) %>%
  dplyr::mutate(Value = dplyr::case_when(!is.na(Value) ~ 'Observed',
                                         is.na(Value) ~ 'Missing'),
                Feature = stringr::str_remove(Feature, ';.+'),
                Recurrence = dplyr::case_when(Recurrence %in% 'Yes' ~ 'Recurrence',
                                              Recurrence %in% 'No' ~ 'Non-Recurrence')) %>%
  dplyr::filter(Feature %in% topImpoFeats) %>%
  dplyr::group_by(Recurrence, Feature) %>%
  dplyr::mutate(SmpNumber = length(Feature)) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(Recurrence, SmpNumber, Feature, Value) %>%
  dplyr::summarise(Count = dplyr::n()) %>%
  dplyr::ungroup() %>%
  dplyr::left_join(featAnnoTab, by = 'Feature') %>%
  dplyr::mutate(Gene = factor(Gene, levels = topImpoFeatAnnos),
                Recurrence = factor(Recurrence, levels = c('Recurrence', 'Non-Recurrence')),
                Value = factor(Value, levels = c('Missing', 'Observed')),
                Proportion = Count/SmpNumber)

ggplot(featAvailTab, aes(x=Gene, y=Proportion, fill=Value)) +
  geom_col(position = 'stack') +
  facet_wrap(vars(Recurrence), scales = 'free') +
  scale_fill_manual(values = c(Missing = 'grey', Observed = 'black')) +
  labs(title = 'Normal Tissue Proteomics') +
  theme_bw() +
  theme(axis.text.x = element_text(size = 18, face = 'bold', angle = 90, hjust = 1, vjust = 0.5),
        axis.text.y = element_text(size = 12, face = 'bold'),
        axis.title = element_text(size = 20, face = 'bold'),
        legend.title = element_blank(), legend.text = element_text(size = 18),
        strip.text = element_text(size = 18, face = 'bold'),
        plot.title = element_text(size = 22, face = 'bold', hjust = 0.5, vjust = 1.5))


# Show availability of top important features in other datasets
# Prepare feature space of other datasets
# Klingmüller Preprocessed Combined dataset
old_proTissue_Combined_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsn.rds')
featSpace_old_Prepro_Combined_Klin <- rownames(old_proTissue_Combined_Klin) %>%
  stringr::str_remove(';.+')
# Klingmüller Original Combined dataset
featSpace_old_Ori_Combined_Klin <- readr::read_tsv(paste0('./data/MethodDev/AG_Klingmueller/',
                                                          '20230309_103228_SmartCare_TumorFree_Tumor_Yang001_Report.tsv')) %>%
  dplyr::pull(PG.ProteinGroups) %>%
  stringr::str_remove(';.+')
# Krijgsveld Preprocessed Combined dataset
proTissue_Combined_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds')
featSpace_Prepro_Combined_Krij <- rownames(proTissue_Combined_Krij) %>%
  stringr::str_remove(';.+')
# Krijgsveld Original Discovery dataset
featSpace_Ori_Discovery_Krij <- readr::read_tsv(paste0('./data/Discovery/AG_Krijgsveld/',
                                                       '20230815_MarcS_Discovery_Tissue.pg_matrix.tsv')) %>%
  dplyr::pull(Protein.Group) %>%
  stringr::str_remove(';.+')
# Krijgsveld Original Method Development dataset
featSpace_Ori_MethodDev_Krij <- readr::read_delim(paste0('./data/MethodDev/AG_Krijgsveld/',
                                                         '20230726_Thorax_Method_Est_tissue.txt')) %>%
  dplyr::pull(Protein.Group) %>%
  stringr::str_remove(';.+')
# Collect all feature spaces into a list
featSpaceList <- list(Ori_Old_Klin = featSpace_old_Ori_Combined_Klin, Prepro_Old_Klin = featSpace_old_Prepro_Combined_Klin,
                      Ori_MDev_Krij = featSpace_Ori_MethodDev_Krij, Ori_Dis_Krij = featSpace_Ori_Discovery_Krij,
                      Prepro_Comb_Krij = featSpace_Prepro_Combined_Krij)
# Store feature availability in a matrix
featAvailMat <- matrix(nrow = 5, ncol = numTopImpoFeats,
                       dimnames = list(c('Ori_Old_Klin', 'Prepro_Old_Klin', 'Ori_MDev_Krij',
                                         'Ori_Dis_Krij', 'Prepro_Comb_Krij'), topImpoFeats))
for (i in seq_len(nrow(featAvailMat))) {
  for (j in seq_len(ncol(featAvailMat))) {
    if (colnames(featAvailMat)[j] %in% featSpaceList[[rownames(featAvailMat)[i]]]) {
      featAvailMat[i, j] <- 'Yes'
    } else {
      featAvailMat[i, j] <- 'No'
    }
  }
}
# Convert matrix to long data for visualization
featAvailMat <- tibble::as_tibble(featAvailMat, rownames = 'Dataset') %>%
  tidyr::pivot_longer(cols = -'Dataset', names_to = 'Feature', values_to = 'Captured') %>%
  dplyr::mutate(Dataset = factor(Dataset, levels = rev(names(featSpaceList))),
                Captured = factor(Captured, levels = c('Yes', 'No')),
                Feature = plyr::mapvalues(Feature, from = topImpoFeats, to = topImpoFeatAnnos),
                Feature = factor(Feature, levels = topImpoFeatAnnos))

ggplot(featAvailMat, aes(x=Feature, y=Dataset, fill=Captured)) +
  geom_tile() +
  scale_fill_manual(values = c('black', 'grey80')) +
  labs(title = 'Tumor Tissue Proteomics') +
  theme(axis.ticks = element_blank(), panel.grid = element_blank(), panel.background = element_blank(),
        axis.text.x = element_text(size = 18, face = 'bold', angle = 45, hjust = 1, vjust = 1),
        axis.text.y = element_text(size = 18, face = 'bold'),
        axis.title = element_text(size = 20, face = 'bold'),
        legend.title = element_text(size = 20), legend.text = element_text(size = 18),
        plot.title = element_text(size = 28, face = 'bold', hjust = 0.5, vjust = 1.5))
```

## RF

<!-- <font size='5'> **Stability selection** </font> -->

<!-- **Performance statistics of 1000 models** -->
<!-- ```{r} -->
<!-- # Load imputed adjusted data -->
<!-- impuAdjProTumor_Klin <- readRDS('./data/Discovery/AG_Klingmueller/impuAdjProTumorVsn_SVs.rds') -->
<!-- # Load SOA results for retrieving significant feature list for initial feature selection -->
<!-- proTumorRes_Klin <- readRDS('./data/Discovery/AG_Klingmueller/soaRes/proTumorRes_SVs.rds') -->

<!-- # Train models iteratively -->
<!-- # Prepare significant feature list -->
<!-- sigFeatList <- proTumorRes_Klin$sig.feat.tab$Var1 -->
<!-- # rfRes <- iterRF(impuAdjProTumor_Klin, doImputation = F, doInitFeatSelection = T, -->
<!-- #                 sigFeatList = sigFeatList, doFeatClustering = T, iter = 1000) -->
<!-- # saveRDS(rfRes, './data/Discovery/rf/rfResSOA_SVs_impuAdjProTumor_0.9r_randSplit_1000_Klin.rds') -->
<!-- rfRes <- readRDS('./data/Discovery/rf/rfResSOA_SVs_impuAdjProTumor_0.9r_randSplit_1000_Klin.rds') -->

<!-- # Report summarized scores -->
<!-- # Prepare predictions and ground truths -->
<!-- pred <- rfRes$rfRes$y_pred -->
<!-- truth <- rfRes$rfRes$y_truth -->
<!-- auc_roc <- rfRes$rfRes$auc_roc -->
<!-- summarizePredPower(pred, truth, auc_roc) -->
<!-- ``` -->

<!-- **Top important features with highest median ranks** -->
<!-- ```{r} -->
<!-- # Visualize top important features -->
<!-- # Prepare feature annotation table -->
<!-- proTissue_Klin <- readRDS('./data/Discovery/AG_Klingmueller/proTissueVsn.rds') -->
<!-- featAnno <- tibble::as_tibble(rowData(proTissue_Klin), rownames = 'Feature') -->
<!-- topImpoFeatViz <- vizTopImpoFeatsRF(rfRes, featAnno = featAnno, num_p1TopFeats = 30) -->
<!-- topImpoFeatViz$rank -->
<!-- ``` -->

<!-- **Full feature importance table** -->
<!-- ```{r} -->
<!-- DT::datatable(topImpoFeatViz$fullImpoFeatTab) -->
<!-- ``` -->

<!-- <font size='5'> **Systematic training and evaluation** </font> -->

<!-- **Performance of models trained with different numbers of top important features** -->
<!-- ```{r} -->
<!-- # Systematically train models with different numbers of top important features -->
<!-- # sysTrainModels <- doSysTrainRF(rfRes, impuAdjProTumor_Klin, max_numTopFeats = 100) -->
<!-- # saveRDS(sysTrainModels, './data/Discovery/rf/sysTrain_rfResSOA_SVs_impuAdjProTumor_randSplit_100_Klin.rds') -->
<!-- sysTrainModels <- readRDS('./data/Discovery/rf/sysTrain_rfResSOA_SVs_impuAdjProTumor_randSplit_100_Klin.rds') -->

<!-- # Visualize systematic training -->
<!-- vizSysTrainModelPerf(sysTrainModels, numTopFeats = 40) -->
<!-- ``` -->
<!-- => Top 21, resulting in model with AUC of 0.999 [1,1]\ -->
<!-- => Top 28, resulting in model with AUC of 1 [1,1] ???? -->

## XGBoost

<!-- <font size='5'> **Linear booster** </font> -->

<!-- **Models trained with core features selected by lasso logistic regression** -->
<!-- ```{r} -->
<!-- ############### -->
<!-- # Change here # -->
<!-- ################################################################################ -->
<!-- # Load imputed adjusted data -->
<!-- se <- readRDS('./data/Discovery/AG_Klingmueller/impuAdjProTumorVsn_SVs.rds') -->
<!-- # Load trained models -->
<!-- lrRes <- readRDS('./data/Discovery/logisR/lrResSOA_SVs_impuAdjProTumor_0.9r_randSplit_1000_5fold_auc_lmin_Klin.rds') -->
<!-- # Specify number of core features selected by lasso logistic regression -->
<!-- numTopImpoFeats <- 5 -->
<!-- ################################################################################ -->

<!-- # Retrieve core feature list -->
<!-- coreFeats <- vizTopImpoFeatsLR(lrRes)$fullImpoFeatTab$Feature[1:numTopImpoFeats] -->
<!-- # Subset core features from data -->
<!-- coreSE <- se[rownames(se) %in% coreFeats,] -->

<!-- # Train models iteratively -->
<!-- # linearBoostRes <- iterXGBoost(coreSE, booster = 'gblinear', iter = 100, nrounds = 1000) -->
<!-- # saveRDS(linearBoostRes, './data/Discovery/xgboost/linear_xgbRes_SVs_coreImpuAdjProTumor_randSplit_rmse_nR1000_iter100_Klin.rds') -->
<!-- linearBoostRes <- readRDS('./data/Discovery/xgboost/linear_xgbRes_SVs_coreImpuAdjProTumor_randSplit_rmse_nR1000_iter100_Klin.rds') -->

<!-- # Report summarized scores -->
<!-- # Prepare predictions and ground truths -->
<!-- pred <- linearBoostRes$xgbRes$y_pred -->
<!-- truth <- linearBoostRes$xgbRes$y_truth -->
<!-- auc_roc <- linearBoostRes$xgbRes$auc_roc -->
<!-- summarizePredPower(pred, truth, auc_roc) -->
<!-- ``` -->

<!-- **Models trained with grouped recurrence-related features** -->
<!-- ```{r} -->
<!-- ############### -->
<!-- # Change here # -->
<!-- ################################################################################ -->
<!-- # Load imputed adjusted data -->
<!-- se <- readRDS('./data/Discovery/AG_Klingmueller/impuAdjProTumorVsn_SVs.rds') -->
<!-- # Load SOA results for retrieving significant feature list for initial feature selection -->
<!-- soaRes <- readRDS('./data/Discovery/AG_Klingmueller/soaRes/proTumorRes_SVs.rds') -->
<!-- ################################################################################ -->

<!-- # Train models iteratively -->
<!-- # Prepare significant feature list -->
<!-- sigFeatList <- soaRes$sig.feat.tab$Var1 -->
<!-- # linearBoostRes <- iterXGBoost(se, doImputation = F, doInitFeatSelection = T, -->
<!-- #                               sigFeatList = sigFeatList, doFeatClustering = T, -->
<!-- #                               booster = 'gblinear', iter = 100, nrounds = 1000) -->
<!-- # saveRDS(linearBoostRes, './data/Discovery/xgboost/linear_xgbRes_SVs_impuAdjProTumor_randSplit_rmse_nR1000_iter100_Klin.rds') -->
<!-- linearBoostRes <- readRDS('./data/Discovery/xgboost/linear_xgbRes_SVs_impuAdjProTumor_randSplit_rmse_nR1000_iter100_Klin.rds') -->

<!-- # Report summarized scores -->
<!-- # Prepare predictions and ground truths -->
<!-- pred <- linearBoostRes$xgbRes$y_pred -->
<!-- truth <- linearBoostRes$xgbRes$y_truth -->
<!-- auc_roc <- linearBoostRes$xgbRes$auc_roc -->
<!-- summarizePredPower(pred, truth, auc_roc) -->
<!-- ``` -->

<!-- ```{r eval=F, include=F} -->
<!-- # <font size='5'> **Tree booster** </font> -->
<!-- #  -->
<!-- # **Models trained with core features selected by random forest** -->

<!-- ############### -->
<!-- # Change here # -->
<!-- ################################################################################ -->
<!-- # Load imputed adjusted data -->
<!-- se <- readRDS('./data/Discovery/AG_Klingmueller/impuAdjProTumorVsn_SVs.rds') -->
<!-- # Load trained models -->
<!-- rfRes <- readRDS('./data/Discovery/rf/rfResSOA_SVs_impuAdjProTumor_0.9r_randSplit_1000_Klin.rds') -->
<!-- # Specify number of core features selected by random forest -->
<!-- numTopImpoFeats <- 34 -->
<!-- ################################################################################ -->

<!-- # Retrieve core feature list -->
<!-- coreFeats <- vizTopImpoFeatsRF(rfRes)$fullImpoFeatTab$Feature[1:numTopImpoFeats] -->
<!-- # Subset core features from data -->
<!-- coreSE <- se[rownames(se) %in% coreFeats,] -->

<!-- # Train models iteratively -->
<!-- # treeBoostRes <- iterXGBoost(coreSE, booster = 'gbtree', iter = 100, nrounds = 1000) -->
<!-- # saveRDS(treeBoostRes, './data/Discovery/xgboost/tree_xgbRes_SVs_coreImpuAdjProTumor_randSplit_auc_nR1000_iter100_Klin.rds') -->
<!-- treeBoostRes <- readRDS('./data/Discovery/xgboost/tree_xgbRes_SVs_coreImpuAdjProTumor_randSplit_auc_nR1000_iter100_Klin.rds') -->

<!-- # Report summarized scores -->
<!-- # Prepare predictions and ground truths -->
<!-- pred <- treeBoostRes$xgbRes$y_pred -->
<!-- truth <- treeBoostRes$xgbRes$y_truth -->
<!-- auc_roc <- treeBoostRes$xgbRes$auc_roc -->
<!-- summarizePredPower(pred, truth, auc_roc) -->
<!-- ``` -->

<!-- ```{r eval=F, include=F} -->
<!-- # **Models trained with grouped recurrence-related features** -->

<!-- ############### -->
<!-- # Change here # -->
<!-- ################################################################################ -->
<!-- # Load imputed adjusted data -->
<!-- se <- readRDS('./data/Discovery/AG_Klingmueller/impuAdjProTumorVsn_SVs.rds') -->
<!-- # Load SOA results for retrieving significant feature list for initial feature selection -->
<!-- soaRes <- readRDS('./data/Discovery/AG_Klingmueller/soaRes/proTumorRes_SVs.rds') -->
<!-- ################################################################################ -->

<!-- # Train models iteratively -->
<!-- # Prepare significant feature list -->
<!-- sigFeatList <- soaRes$sig.feat.tab$Var1 -->
<!-- # treeBoostRes <- iterXGBoost(se, doImputation = F, doInitFeatSelection = T, -->
<!-- #                             sigFeatList = sigFeatList, doFeatClustering = T, -->
<!-- #                             booster = 'gbtree', iter = 100, nrounds = 1000) -->
<!-- # saveRDS(treeBoostRes, './data/Discovery/xgboost/tree_xgbRes_SVs_impuAdjProTumor_randSplit_auc_nR1000_iter100_Klin.rds') -->
<!-- treeBoostRes <- readRDS('./data/Discovery/xgboost/tree_xgbRes_SVs_impuAdjProTumor_randSplit_auc_nR1000_iter100_Klin.rds') -->

<!-- # Report summarized scores -->
<!-- # Prepare predictions and ground truths -->
<!-- pred <- treeBoostRes$xgbRes$y_pred -->
<!-- truth <- treeBoostRes$xgbRes$y_truth -->
<!-- auc_roc <- treeBoostRes$xgbRes$auc_roc -->
<!-- summarizePredPower(pred, truth, auc_roc) -->
<!-- ``` -->

# NTP (AG Klin.)
Normal Tissue DIA Proteomics from AG Klingmüller

## Lasso logistic regression

<font size='5'> **Stability selection** </font>

**Performance statistics of 1000 models**
```{r warning=T}
# Load preprocessed data
proTissue_Klin <- readRDS('./data/Discovery/AG_Klingmueller/proTissueVsn.rds')
proNormal_Klin <- proTissue_Klin[, colData(proTissue_Klin)$Condition %in% 'Normal']
# # Load SOA results
proNormalRes_Klin <- readRDS('./data/Discovery/AG_Klingmueller/soaRes/proNormalRes_SVsByProTissueInt.rds')
# # Prepare adjusted data
adjProNormal_Klin <- proNormal_Klin
assay(adjProNormal_Klin) <- proNormalRes_Klin$SOA.res$dataCorrect
# Do additional feature filtering to remove features observed in less than 50% of samples
ifRmFeats <- apply(assay(adjProNormal_Klin), 1, function(featVec) {
  sum(is.na(featVec)) / length(featVec) > 0.5
})
adjProNormal_Klin <- adjProNormal_Klin[!ifRmFeats,]

# Load imputed adjusted data
impuAdjProNormal_Klin <- readRDS('./data/Discovery/AG_Klingmueller/impuData/featFilt_impuAdjProNormalVsn_SVsByProTissueInt.rds')

# Train models iteratively
# Prepare significant feature list
sigFeatList <- proNormalRes_Klin$sig.feat.tab$Var1
# 'auc' is not applicable for CV when too few (< 10) observations per fold. Instead,
# 'deviance' will be used.
# lrRes <- iterLogisR(adjProNormal_Klin, doImputation = T, doInitFeatSelection = T, sigFeatList = sigFeatList,
#                     doFeatClustering = T, iter = 1000, cvFold = 5, cvMeasure = 'auc')
# saveRDS(lrRes, './data/Discovery/logisR/featFilt_lrResSOA_SVsByProTissueInt_impuAdjProNormal_0.9r_randSplit_1000_5fold_auc_lmin_Klin.rds')
# saveRDS(lrRes$impuSE, './data/Discovery/AG_Klingmueller/impuData/featFilt_impuAdjProNormalVsn_SVsByProTissueInt.rds')
lrRes <- readRDS('./data/Discovery/logisR/featFilt_lrResSOA_SVsByProTissueInt_impuAdjProNormal_0.9r_randSplit_1000_5fold_auc_lmin_Klin.rds')

# Pinpoint useless models that will be removed, i.e., models with no any non-zero beta
uselessModels <- which(lrRes$lrRes$nNonZero == 0) #higher lambda results in simpler model
# for (i in uselessModels) {
#   pred <- lrRes$lrRes$y_pred[[i]]
#   truth <- lrRes$lrRes$y_truth[[i]]
#   scores <- caret::confusionMatrix(data = pred, reference = truth, positive = '1')
#   cat(scores$overall[['Accuracy']], '\n')
# } # Models that got only intercept make random predictions, 0.5 accuracy and AUC-ROC

# Report summarized scores
# Prepare predictions and ground truths where results from useless models are removed
if (length(uselessModels) != 0) {
  pred <- lrRes$lrRes$y_pred[-uselessModels]
  truth <- lrRes$lrRes$y_truth[-uselessModels]
  auc_roc <- lrRes$lrRes$auc_roc[-uselessModels]
} else {
  pred <- lrRes$lrRes$y_pred
  truth <- lrRes$lrRes$y_truth
  auc_roc <- lrRes$lrRes$auc_roc
}
suppressWarnings(summarizePredPower(pred, truth, auc_roc)) #warning due to sometimes only pos or neg predictions
```

**Top important features that are most frequently selected**
```{r}
# Visualize top important features
# Prepare feature annotation table
featAnno <- tibble::as_tibble(rowData(proTissue_Klin), rownames = 'Feature')
topImpoFeatViz <- vizTopImpoFeatsLR(lrRes, featAnno = featAnno, num_p1TopFeats = 30)
topImpoFeatViz$pick
```

**Full feature importance table**
```{r}
# Display full feature importance table
DT::datatable(topImpoFeatViz$fullImpoFeatTab)
```

<font size='5'> **Systematic training and evaluation** </font>

**Performance of models trained with different numbers of top important features**
```{r}
# Systematically train models with different numbers of top important features
# sysTrainModels <- doSysTrainLogisR(lrRes, impuAdjProNormal_Klin, max_numTopFeats = 50)
# saveRDS(sysTrainModels, './data/Discovery/logisR/featFilt_sysTrain_lrResSOA_SVsByProTissueInt_impuAdjProNormal_randSplit_100_5fold_auc_lmin_Klin.rds')
sysTrainModels <- readRDS('./data/Discovery/logisR/featFilt_sysTrain_lrResSOA_SVsByProTissueInt_impuAdjProNormal_randSplit_100_5fold_auc_lmin_Klin.rds')

# Visualize systematic training
vizSysTrainModelPerf(sysTrainModels, numTopFeats = 30)
# ggsave('./output/Junyan/stabSelec_Stats_proNormal.png', device = 'png', dpi = 400, height = 8, width = 10)
```
=> Top 13 important features are selected, resulting in model with AUC of 0.992 [0.94,1].\
=> Top 21 important features are selected, resulting in model with AUC of 0.994 [0.95,1].\
=> Top 29 important features are selected, resulting in model with AUC of 0.994 [0.95,1].

**Noninformative models that demonstrate effects of top important features**
```{r}
se <- impuAdjProNormal_Klin
# Train models without top important features, aka noninformative models
numTopImpoFeats <- c(5, 10, 15, 20, 25, 30)
# for (num in numTopImpoFeats) {
#   topImpoFeats <- topImpoFeatViz$fullImpoFeatTab$Feature[1:num]
#   # Train lasso logistic regression model
#   reducedData <- lrRes$featClusters$reducedData
#   x <- t(reducedData[!rownames(reducedData) %in% topImpoFeats,])
#   y <- colData(se)$Recurrence
#   lrResNonInfo <- runLogisR(x, y, targetClass = 'Yes', iter = 100, regularized_method = 'lasso',
#                             cvFold = 5, cvMeasure = 'auc', used_lambda = 'lambda.min',
#                             trainSet_ratio = 0.8, split_method = 'random split',
#                             plot_ROC = F, save_model = T)
#   saveRDS(lrResNonInfo, paste0('./data/Discovery/logisR/noninfo/featFilt_NonInfo_rmTop', num,
#                                '_lrResSOA_SVsByProTissueInt_impuAdjProNormal_randSplit_100_5fold_auc_lmin_Klin.rds'))
# }

# Display performance of noninformative models
meanAUC <- c()
lowerAUC <- c()
upperAUC <- c()
uselessModels <- which(lrRes$lrRes$nNonZero == 0)
if (length(uselessModels) != 0) {
  auc_roc <- lrRes$lrRes$auc_roc[-uselessModels]
} else {
  auc_roc <- lrRes$lrRes$auc_roc
}
meanAUC <- c(meanAUC, round(mean(auc_roc), 3))
ci <- calcCI(auc_roc, bootstrap = T)
lowerAUC <- c(lowerAUC, ci[1])
upperAUC <- c(upperAUC, ci[2])
for (num in numTopImpoFeats) {
  lrResNonInfo <- readRDS(paste0('./data/Discovery/logisR/noninfo/featFilt_NonInfo_rmTop', num,
                                 '_lrResSOA_SVsByProTissueInt_impuAdjProNormal_randSplit_100_5fold_auc_lmin_Klin.rds'))
  uselessModels <- which(lrResNonInfo$nNonZero == 0)
  if (length(uselessModels) != 0) {
    auc_roc <- lrResNonInfo$auc_roc[-uselessModels]
  } else {
    auc_roc <- lrResNonInfo$auc_roc
  }
  meanAUC <- c(meanAUC, round(mean(auc_roc), 3))
  ci <- calcCI(auc_roc, bootstrap = T)
  lowerAUC <- c(lowerAUC, ci[1])
  upperAUC <- c(upperAUC, ci[2])
}
aucStatTab <- data.frame(FeatSpace = c('Complete', paste0('-Top ', numTopImpoFeats)),
                         MeanAUC = meanAUC, UpperAUC = upperAUC, LowerAUC = lowerAUC) %>%
  dplyr::mutate(FeatSpace = factor(FeatSpace, levels = FeatSpace))

ggplot(aucStatTab, aes(x=FeatSpace, y=MeanAUC)) +
  geom_bar(stat = 'identity', position = position_dodge(), alpha = 0.9) +
  geom_errorbar(aes(ymin=LowerAUC, ymax=UpperAUC), width = 0.4) +
  labs(x = 'Feature space used to train 100 models', y = 'Mean AUC-ROC',
       title = 'Noninformative models') +
  th
```

<font size='5'> **Investigation of core features** </font>

<font size='3'> Imputed data: </font>\
**Heatmap where samples are ordered by linear regression scores**
```{r}
coreFeatPlotsImpu <- vizCoreFeats(lrRes, sysTrainModels, impuAdjProNormal_Klin,
                                  heatmap_numCoreFeats = 13, colFeatAnno = 'Gene',
                                  scatter_twoCoreFeats = c('OR5K3', 'SEMA4B'),
                                  fontsize = 10, fontsize_col = 7, fontsize_row = 11,
                                  cellwidth = 6, cellheight = 18)
coreFeatPlotsImpu$heatmap
# ggsave('./output/Junyan/stabSelec_heatmap_proNormal_impu.png', device = 'png', dpi = 400, height = 8, width = 10)
```

**Scatterplot of two core features**
```{r}
coreFeatPlotsImpu$scatterplot
# ggsave('./output/Junyan/stabSelec_top2_proNormal_impu.png', device = 'png', dpi = 400, height = 8, width = 10)
```

<font size='3'> Original data: </font>\
**Heatmap**
```{r}
coreFeatPlots <- vizCoreFeats(lrRes, sysTrainModels, adjProNormal_Klin,
                              heatmap_numCoreFeats = 13, colFeatAnno = 'Gene',
                              scatter_twoCoreFeats = c('OR5K3', 'H1-3'),
                              fontsize = 10, fontsize_col = 7, fontsize_row = 11,
                              cellwidth = 6, cellheight = 18)
coreFeatPlots$heatmap
# ggsave('./output/Junyan/stabSelec_heatmap_proNormal_ori.png', device = 'png', dpi = 400, height = 8, width = 10)
```

**Scatterplot**
```{r}
coreFeatPlots$scatterplot
# ggsave('./output/Junyan/stabSelec_top2_proNormal_ori.png', device = 'png', dpi = 400, height = 8, width = 10)
```

**Missingness and availability of core features**
```{r message=F}
numTopImpoFeats <- 13
# Prepare top important feature list
topImpoFeats <- topImpoFeatViz$fullImpoFeatTab$Feature[1:numTopImpoFeats] %>%
  stringr::str_remove(';.+')
topImpoFeatAnnos <- topImpoFeatViz$fullImpoFeatTab$Annotation[1:numTopImpoFeats] %>%
  stringr::str_remove(';.+')

# Show data missingness
proNormal_Klin <- proTissue_Klin[, colData(proTissue_Klin)$Condition %in% 'Normal']
featAnnoTab <- tibble::as_tibble(rowData(proTissue_Klin), rownames = 'Feature') %>%
  dplyr::mutate(Feature = stringr::str_remove(Feature, ';.+'),
                Gene = stringr::str_remove(Gene, ';.+'))
featAvailTab <- summExp2df(proNormal_Klin, row_id = 'Feature', col_id = 'Sample') %>%
  dplyr::select(Feature, Value, Recurrence) %>%
  dplyr::mutate(Value = dplyr::case_when(!is.na(Value) ~ 'Observed',
                                         is.na(Value) ~ 'Missing'),
                Feature = stringr::str_remove(Feature, ';.+'),
                Recurrence = dplyr::case_when(Recurrence %in% 'Yes' ~ 'Recurrence',
                                              Recurrence %in% 'No' ~ 'Non-Recurrence')) %>%
  dplyr::filter(Feature %in% topImpoFeats) %>%
  dplyr::group_by(Recurrence, Feature) %>%
  dplyr::mutate(SmpNumber = length(Feature)) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(Recurrence, SmpNumber, Feature, Value) %>%
  dplyr::summarise(Count = dplyr::n()) %>%
  dplyr::ungroup() %>%
  dplyr::left_join(featAnnoTab, by = 'Feature') %>%
  dplyr::mutate(Gene = factor(Gene, levels = topImpoFeatAnnos),
                Recurrence = factor(Recurrence, levels = c('Recurrence', 'Non-Recurrence')),
                Value = factor(Value, levels = c('Missing', 'Observed')),
                Proportion = Count/SmpNumber)

ggplot(featAvailTab, aes(x=Gene, y=Proportion, fill=Value)) +
  geom_col(position = 'stack') +
  facet_wrap(vars(Recurrence), scales = 'free') +
  scale_fill_manual(values = c(Missing = 'grey', Observed = 'black')) +
  labs(title = 'Normal Tissue Proteomics') +
  theme_bw() +
  theme(axis.text.x = element_text(size = 18, face = 'bold', angle = 90, hjust = 1, vjust = 0.5),
        axis.text.y = element_text(size = 12, face = 'bold'),
        axis.title = element_text(size = 20, face = 'bold'),
        legend.title = element_blank(), legend.text = element_text(size = 18),
        strip.text = element_text(size = 18, face = 'bold'),
        plot.title = element_text(size = 22, face = 'bold', hjust = 0.5, vjust = 1.5))


# Show availability of top important features in other datasets
# Prepare feature space of other datasets
# Klingmüller Preprocessed Combined dataset
old_proTissue_Combined_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsn.rds')
featSpace_old_Prepro_Combined_Klin <- rownames(old_proTissue_Combined_Klin) %>%
  stringr::str_remove(';.+')
# Klingmüller Original Combined dataset
featSpace_old_Ori_Combined_Klin <- readr::read_tsv(paste0('./data/MethodDev/AG_Klingmueller/',
                                                          '20230309_103228_SmartCare_TumorFree_Tumor_Yang001_Report.tsv')) %>%
  dplyr::pull(PG.ProteinGroups) %>%
  stringr::str_remove(';.+')
# Krijgsveld Preprocessed Combined dataset
proTissue_Combined_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds')
featSpace_Prepro_Combined_Krij <- rownames(proTissue_Combined_Krij) %>%
  stringr::str_remove(';.+')
# Krijgsveld Original Discovery dataset
featSpace_Ori_Discovery_Krij <- readr::read_tsv(paste0('./data/Discovery/AG_Krijgsveld/',
                                                       '20230815_MarcS_Discovery_Tissue.pg_matrix.tsv')) %>%
  dplyr::pull(Protein.Group) %>%
  stringr::str_remove(';.+')
# Krijgsveld Original Method Development dataset
featSpace_Ori_MethodDev_Krij <- readr::read_delim(paste0('./data/MethodDev/AG_Krijgsveld/',
                                                         '20230726_Thorax_Method_Est_tissue.txt')) %>%
  dplyr::pull(Protein.Group) %>%
  stringr::str_remove(';.+')
# Collect all feature spaces into a list
featSpaceList <- list(Ori_Old_Klin = featSpace_old_Ori_Combined_Klin, Prepro_Old_Klin = featSpace_old_Prepro_Combined_Klin,
                      Ori_MDev_Krij = featSpace_Ori_MethodDev_Krij, Ori_Dis_Krij = featSpace_Ori_Discovery_Krij,
                      Prepro_Comb_Krij = featSpace_Prepro_Combined_Krij)
# Store feature availability in a matrix
featAvailMat <- matrix(nrow = 5, ncol = numTopImpoFeats,
                       dimnames = list(c('Ori_Old_Klin', 'Prepro_Old_Klin', 'Ori_MDev_Krij',
                                         'Ori_Dis_Krij', 'Prepro_Comb_Krij'), topImpoFeats))
for (i in seq_len(nrow(featAvailMat))) {
  for (j in seq_len(ncol(featAvailMat))) {
    if (colnames(featAvailMat)[j] %in% featSpaceList[[rownames(featAvailMat)[i]]]) {
      featAvailMat[i, j] <- 'Yes'
    } else {
      featAvailMat[i, j] <- 'No'
    }
  }
}
# Convert matrix to long data for visualization
featAvailMat <- tibble::as_tibble(featAvailMat, rownames = 'Dataset') %>%
  tidyr::pivot_longer(cols = -'Dataset', names_to = 'Feature', values_to = 'Captured') %>%
  dplyr::mutate(Dataset = factor(Dataset, levels = rev(names(featSpaceList))),
                Captured = factor(Captured, levels = c('Yes', 'No')),
                Feature = plyr::mapvalues(Feature, from = topImpoFeats, to = topImpoFeatAnnos),
                Feature = factor(Feature, levels = topImpoFeatAnnos))

ggplot(featAvailMat, aes(x=Feature, y=Dataset, fill=Captured)) +
  geom_tile() +
  scale_fill_manual(values = c('black', 'grey80')) +
  labs(title = 'Normal Tissue Proteomics') +
  theme(axis.ticks = element_blank(), panel.grid = element_blank(), panel.background = element_blank(),
        axis.text.x = element_text(size = 18, face = 'bold', angle = 45, hjust = 1, vjust = 1),
        axis.text.y = element_text(size = 18, face = 'bold'),
        axis.title = element_text(size = 20, face = 'bold'),
        legend.title = element_text(size = 20), legend.text = element_text(size = 18),
        plot.title = element_text(size = 28, face = 'bold', hjust = 0.5, vjust = 1.5))
```

## RF

<!-- <font size='5'> **Stability selection** </font> -->

<!-- **Performance statistics of 1000 models** -->
<!-- ```{r} -->
<!-- # Load imputed adjusted data -->
<!-- impuAdjProNormal_Klin <- readRDS('./data/Discovery/AG_Klingmueller/impuAdjProNormalVsn_Stage.rds') -->
<!-- # Load SOA results for retrieving significant feature list for initial feature selection -->
<!-- proNormalRes_Klin <- readRDS('./data/Discovery/AG_Klingmueller/soaRes/proNormalRes_Stage.rds') -->

<!-- # Train models iteratively -->
<!-- # Prepare significant feature list -->
<!-- sigFeatList <- proNormalRes_Klin$sig.feat.tab$Var1 -->
<!-- # rfRes <- iterRF(impuAdjProNormal_Klin, doImputation = F, doInitFeatSelection = T, -->
<!-- #                 sigFeatList = sigFeatList, doFeatClustering = T, iter = 1000) -->
<!-- # saveRDS(rfRes, './data/Discovery/rf/rfResSOA_Stage_impuAdjProNormal_0.9r_randSplit_1000_Klin.rds') -->
<!-- rfRes <- readRDS('./data/Discovery/rf/rfResSOA_Stage_impuAdjProNormal_0.9r_randSplit_1000_Klin.rds') -->

<!-- # Report summarized scores -->
<!-- # Prepare predictions and ground truths -->
<!-- pred <- rfRes$rfRes$y_pred -->
<!-- truth <- rfRes$rfRes$y_truth -->
<!-- auc_roc <- rfRes$rfRes$auc_roc -->
<!-- summarizePredPower(pred, truth, auc_roc) -->
<!-- ``` -->

<!-- **Top important features with highest median ranks** -->
<!-- ```{r} -->
<!-- # Visualize top important features -->
<!-- # Prepare feature annotation table -->
<!-- proTissue_Klin <- readRDS('./data/Discovery/AG_Klingmueller/proTissueVsn.rds') -->
<!-- featAnno <- tibble::as_tibble(rowData(proTissue_Klin), rownames = 'Feature') -->
<!-- topImpoFeatViz <- vizTopImpoFeatsRF(rfRes, featAnno = featAnno, num_p1TopFeats = 30) -->
<!-- topImpoFeatViz$rank + theme(axis.title = element_text(size = 28), -->
<!--                             axis.text = element_text(size = 16)) -->
<!-- # ggsave('./output/Discovery/group_meeting/rf_topImpo_combined_proNormal_Klin.png', -->
<!-- #        device = 'png', dpi = 400, height = 8, width = 10) -->
<!-- ``` -->

<!-- **Full feature importance table** -->
<!-- ```{r} -->
<!-- DT::datatable(topImpoFeatViz$fullImpoFeatTab) -->
<!-- ``` -->

<!-- <font size='5'> **Systematic training and evaluation** </font> -->

<!-- **Performance of models trained with different numbers of top important features** -->
<!-- ```{r} -->
<!-- # Systematically train models with different numbers of top important features -->
<!-- # sysTrainModels <- doSysTrainRF(rfRes, impuAdjProNormal_Klin, max_numTopFeats = 100) -->
<!-- # saveRDS(sysTrainModels, './data/Discovery/rf/sysTrain_rfResSOA_Stage_impuAdjProNormal_randSplit_100_Klin.rds') -->
<!-- sysTrainModels <- readRDS('./data/Discovery/rf/sysTrain_rfResSOA_Stage_impuAdjProNormal_randSplit_100_Klin.rds') -->

<!-- # Visualize systematic training -->
<!-- vizSysTrainModelPerf(sysTrainModels, numTopFeats = 40) -->
<!-- ``` -->
<!-- => Top 24, resulting in model with AUC of 0.934 [0.8,1] -->

<!-- ```{r eval=F, include=F} -->
<!-- # **Noninformative models that demonstrate effects of top important features** -->

<!-- se <- impuAdjProNormal_Klin -->
<!-- # Train models without top important features, aka noninformative models -->
<!-- numTopImpoFeats <- c(10, 50, 100, 200, 300, 400, 500) -->
<!-- # for (num in numTopImpoFeats) { -->
<!-- #   topImpoFeats <- topImpoFeatViz$fullImpoFeatTab$Feature[1:num] -->
<!-- #   # Train lasso logistic regression model -->
<!-- #   reducedData <- rfRes$featClusters$reducedData -->
<!-- #   x <- t(reducedData[!rownames(reducedData) %in% topImpoFeats,]) -->
<!-- #   y <- colData(se)$Recurrence -->
<!-- #   rfResNonInfo <- runRF(x, y, targetClass = 'Yes', iter = 100, split_method = 'random split', -->
<!-- #                         trainSet_ratio = 0.8, ntree = 10000, plot_ROC = F, save_model = T) -->
<!-- #   saveRDS(rfResNonInfo, paste0('./data/Discovery/rf/NonInfo_rmTop', num, '_rfResSOA_Stage_', -->
<!-- #                                'impuAdjProNormal_randSplit_100_Klin.rds')) -->
<!-- # } -->

<!-- # Display performance of noninformative models -->
<!-- meanAUC <- c() -->
<!-- lowerAUC <- c() -->
<!-- upperAUC <- c() -->
<!-- auc_roc <- rfRes$rfRes$auc_roc -->
<!-- meanAUC <- c(meanAUC, round(mean(auc_roc), 3)) -->
<!-- ci <- calcCI(auc_roc, bootstrap = T) -->
<!-- lowerAUC <- c(lowerAUC, ci[1]) -->
<!-- upperAUC <- c(upperAUC, ci[2]) -->
<!-- for (num in numTopImpoFeats) { -->
<!--   lrResNonInfo <- readRDS(paste0('./data/Discovery/rf/NonInfo_rmTop', num, '_rfResSOA_Stage_', -->
<!--                                  'impuAdjProNormal_randSplit_100_Klin.rds')) -->

<!--   auc_roc <- lrResNonInfo$auc_roc -->
<!--   meanAUC <- c(meanAUC, round(mean(auc_roc), 3)) -->
<!--   ci <- calcCI(auc_roc, bootstrap = T) -->
<!--   lowerAUC <- c(lowerAUC, ci[1]) -->
<!--   upperAUC <- c(upperAUC, ci[2]) -->
<!-- } -->
<!-- aucStatTab <- data.frame(FeatSpace = c('Complete', paste0('-Top ', numTopImpoFeats)), -->
<!--                          MeanAUC = meanAUC, UpperAUC = upperAUC, LowerAUC = lowerAUC) %>% -->
<!--   dplyr::mutate(FeatSpace = factor(FeatSpace, levels = FeatSpace)) -->

<!-- ggplot(aucStatTab, aes(x=FeatSpace, y=MeanAUC)) + -->
<!--   geom_bar(stat = 'identity', position = position_dodge(), alpha = 0.9) + -->
<!--   geom_errorbar(aes(ymin=LowerAUC, ymax=UpperAUC), width = 0.4) + -->
<!--   labs(x = 'Feature space used to train 100 models', y = 'Mean AUC-ROC', -->
<!--        title = 'Noninformative models') + -->
<!--   th -->
<!-- ``` -->

## XGBoost

<!-- <font size='5'> **Linear booster** </font> -->

<!-- **Models trained with core features selected by lasso logistic regression** -->
<!-- ```{r} -->
<!-- ############### -->
<!-- # Change here # -->
<!-- ################################################################################ -->
<!-- # Load imputed adjusted data -->
<!-- se <- readRDS('./data/Discovery/AG_Klingmueller/impuAdjProNormalVsn_Stage.rds') -->
<!-- # Load trained models -->
<!-- lrRes <- readRDS('./data/Discovery/logisR/lrResSOA_Stage_impuAdjProNormal_0.9r_randSplit_1000_5fold_auc_lmin_Klin.rds') -->
<!-- # Specify number of core features selected by lasso logistic regression -->
<!-- numTopImpoFeats <- 13 -->
<!-- ################################################################################ -->

<!-- # Retrieve core feature list -->
<!-- coreFeats <- vizTopImpoFeatsLR(lrRes)$fullImpoFeatTab$Feature[1:numTopImpoFeats] -->
<!-- # Subset core features from data -->
<!-- coreSE <- se[rownames(se) %in% coreFeats,] -->

<!-- # Train models iteratively -->
<!-- # linearBoostRes <- iterXGBoost(coreSE, booster = 'gblinear', iter = 100, nrounds = 1000) -->
<!-- # saveRDS(linearBoostRes, './data/Discovery/xgboost/linear_xgbRes_Stage_coreImpuAdjProNormal_randSplit_rmse_nR1000_iter100_Klin.rds') -->
<!-- linearBoostRes <- readRDS('./data/Discovery/xgboost/linear_xgbRes_Stage_coreImpuAdjProNormal_randSplit_rmse_nR1000_iter100_Klin.rds') -->

<!-- # Report summarized scores -->
<!-- # Prepare predictions and ground truths -->
<!-- pred <- linearBoostRes$xgbRes$y_pred -->
<!-- truth <- linearBoostRes$xgbRes$y_truth -->
<!-- auc_roc <- linearBoostRes$xgbRes$auc_roc -->
<!-- summarizePredPower(pred, truth, auc_roc) -->
<!-- ``` -->

<!-- **Models trained with grouped recurrence-related features** -->
<!-- ```{r} -->
<!-- ############### -->
<!-- # Change here # -->
<!-- ################################################################################ -->
<!-- # Load imputed adjusted data -->
<!-- se <- readRDS('./data/Discovery/AG_Klingmueller/impuAdjProNormalVsn_Stage.rds') -->
<!-- # Load SOA results for retrieving significant feature list for initial feature selection -->
<!-- soaRes <- readRDS('./data/Discovery/AG_Klingmueller/soaRes/proNormalRes_Stage.rds') -->
<!-- ################################################################################ -->

<!-- # Train models iteratively -->
<!-- # Prepare significant feature list -->
<!-- sigFeatList <- soaRes$sig.feat.tab$Var1 -->
<!-- # linearBoostRes <- iterXGBoost(se, doImputation = F, doInitFeatSelection = T, -->
<!-- #                               sigFeatList = sigFeatList, doFeatClustering = T, -->
<!-- #                               booster = 'gblinear', iter = 100, nrounds = 1000) -->
<!-- # saveRDS(linearBoostRes, './data/Discovery/xgboost/linear_xgbRes_Stage_impuAdjProNormal_randSplit_rmse_nR1000_iter100_Klin.rds') -->
<!-- linearBoostRes <- readRDS('./data/Discovery/xgboost/linear_xgbRes_Stage_impuAdjProNormal_randSplit_rmse_nR1000_iter100_Klin.rds') -->

<!-- # Report summarized scores -->
<!-- # Prepare predictions and ground truths -->
<!-- pred <- linearBoostRes$xgbRes$y_pred -->
<!-- truth <- linearBoostRes$xgbRes$y_truth -->
<!-- auc_roc <- linearBoostRes$xgbRes$auc_roc -->
<!-- summarizePredPower(pred, truth, auc_roc) -->
<!-- ``` -->

<!-- ```{r eval=F, include=F} -->
<!-- # <font size='5'> **Tree booster** </font> -->
<!-- #  -->
<!-- # **Models trained with core features selected by random forest** -->

<!-- ############### -->
<!-- # Change here # -->
<!-- ################################################################################ -->
<!-- # Load imputed adjusted data -->
<!-- se <- readRDS('./data/Discovery/AG_Klingmueller/impuAdjProNormalVsn_Stage.rds') -->
<!-- # Load trained models -->
<!-- rfRes <- readRDS('./data/Discovery/rf/rfResSOA_Stage_impuAdjProNormal_0.9r_randSplit_1000_Klin.rds') -->
<!-- # Specify number of core features selected by random forest -->
<!-- numTopImpoFeats <- 24 -->
<!-- ################################################################################ -->

<!-- # Retrieve core feature list -->
<!-- coreFeats <- vizTopImpoFeatsRF(rfRes)$fullImpoFeatTab$Feature[1:numTopImpoFeats] -->
<!-- # Subset core features from data -->
<!-- coreSE <- se[rownames(se) %in% coreFeats,] -->

<!-- # Train models iteratively -->
<!-- # treeBoostRes <- iterXGBoost(coreSE, booster = 'gbtree', iter = 100, nrounds = 1000) -->
<!-- # saveRDS(treeBoostRes, './data/Discovery/xgboost/tree_xgbRes_Stage_coreImpuAdjProNormal_randSplit_auc_nR1000_iter100_Klin.rds') -->
<!-- treeBoostRes <- readRDS('./data/Discovery/xgboost/tree_xgbRes_Stage_coreImpuAdjProNormal_randSplit_auc_nR1000_iter100_Klin.rds') -->

<!-- # Report summarized scores -->
<!-- # Prepare predictions and ground truths -->
<!-- pred <- treeBoostRes$xgbRes$y_pred -->
<!-- truth <- treeBoostRes$xgbRes$y_truth -->
<!-- auc_roc <- treeBoostRes$xgbRes$auc_roc -->
<!-- summarizePredPower(pred, truth, auc_roc) -->
<!-- ``` -->

<!-- ```{r eval=F, include=F} -->
<!-- # **Models trained with grouped recurrence-related features** -->

<!-- ############### -->
<!-- # Change here # -->
<!-- ################################################################################ -->
<!-- # Load imputed adjusted data -->
<!-- se <- readRDS('./data/Discovery/AG_Klingmueller/impuAdjProNormalVsn_Stage.rds') -->
<!-- # Load SOA results for retrieving significant feature list for initial feature selection -->
<!-- soaRes <- readRDS('./data/Discovery/AG_Klingmueller/soaRes/proNormalRes_Stage.rds') -->
<!-- ################################################################################ -->

<!-- # Train models iteratively -->
<!-- # Prepare significant feature list -->
<!-- sigFeatList <- soaRes$sig.feat.tab$Var1 -->
<!-- # treeBoostRes <- iterXGBoost(se, doImputation = F, doInitFeatSelection = T, -->
<!-- #                             sigFeatList = sigFeatList, doFeatClustering = T, -->
<!-- #                             booster = 'gbtree', iter = 100, nrounds = 1000) -->
<!-- # saveRDS(treeBoostRes, './data/Discovery/xgboost/tree_xgbRes_Stage_impuAdjProNormal_randSplit_logloss_nR1000_iter100_Klin.rds') -->
<!-- treeBoostRes <- readRDS('./data/Discovery/xgboost/tree_xgbRes_Stage_impuAdjProNormal_randSplit_logloss_nR1000_iter100_Klin.rds') -->

<!-- # Report summarized scores -->
<!-- # Prepare predictions and ground truths -->
<!-- pred <- treeBoostRes$xgbRes$y_pred -->
<!-- truth <- treeBoostRes$xgbRes$y_truth -->
<!-- auc_roc <- treeBoostRes$xgbRes$auc_roc -->
<!-- summarizePredPower(pred, truth, auc_roc) -->
<!-- ``` -->




# TTP (AG Krij.)
Tumor Tissue DIA Proteomics from AG Krijgsveld

## Lasso logistic regression

<font size='5'> **Stability selection** </font>

**Performance statistics of 1000 models**
```{r warning=T}
# Load preprocessed data
proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds')
proTumor_Krij <- proTissue_Krij[, colData(proTissue_Krij)$Condition %in% 'Tumor']
# Load SOA results
proTumorRes_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/soaRes/combined_proTumorRes_SVsByProTissueInt.rds')
# Prepare adjusted data
adjProTumor_Krij <- proTumor_Krij
assay(adjProTumor_Krij) <- proTumorRes_Krij$SOA.res$dataCorrect
# Do additional feature filtering to remove features observed in less than 50% of samples
ifRmFeats <- apply(assay(adjProTumor_Krij), 1, function(featVec) {
  sum(is.na(featVec)) / length(featVec) > 0.5
})
adjProTumor_Krij <- adjProTumor_Krij[!ifRmFeats,]

# Load imputed adjusted data
impuAdjProTumor_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/impuData/featFilt_combined_impuAdjProTumorVsn_SVsByProTissueInt.rds')

# Train models iteratively
# Prepare significant feature list
sigFeatList <- proTumorRes_Krij$sig.feat.tab$Var1
# lrRes <- iterLogisR(adjProTumor_Krij, doImputation = T, doInitFeatSelection = T, sigFeatList = sigFeatList,
#                     doFeatClustering = T, iter = 1000, cvFold = 5, cvMeasure = 'auc')
# saveRDS(lrRes, './data/Discovery/logisR/featFilt_lrResSOA_SVsByProTissueInt_impuAdjProTumor_0.9r_randSplit_1000_5fold_auc_lmin_Krij.rds')
# saveRDS(lrRes$impuSE, './data/Discovery/AG_Krijgsveld/impuData/featFilt_combined_impuAdjProTumorVsn_SVsByProTissueInt.rds')
lrRes <- readRDS('./data/Discovery/logisR/featFilt_lrResSOA_SVsByProTissueInt_impuAdjProTumor_0.9r_randSplit_1000_5fold_auc_lmin_Krij.rds')

# Report summarized scores
# Prepare predictions and ground truths where results from useless models are removed
uselessModels <- which(lrRes$lrRes$nNonZero == 0)
if (length(uselessModels) != 0) {
  pred <- lrRes$lrRes$y_pred[-uselessModels]
  truth <- lrRes$lrRes$y_truth[-uselessModels]
  auc_roc <- lrRes$lrRes$auc_roc[-uselessModels]
} else {
  pred <- lrRes$lrRes$y_pred
  truth <- lrRes$lrRes$y_truth
  auc_roc <- lrRes$lrRes$auc_roc
}
suppressWarnings(summarizePredPower(pred, truth, auc_roc))
```

**Top important features that are most frequently selected**
```{r}
# Visualize top important features
# Prepare feature annotation table
featAnno <- tibble::as_tibble(rowData(proTissue_Krij), rownames = 'Feature') %>%
  dplyr::select(Feature, Genes) %>%
  dplyr::rename(Gene = Genes)
topImpoFeatViz <- vizTopImpoFeatsLR(lrRes, featAnno = featAnno, num_p1TopFeats = 30)
topImpoFeatViz$pick
```

**Full feature importance table**
```{r}
# Display full feature importance table
DT::datatable(topImpoFeatViz$fullImpoFeatTab)
```

<font size='5'> **Systematic training and evaluation** </font>

**Performance of models trained with different numbers of top important features**
```{r}
# Systematically train models with different numbers of top important features
# sysTrainModels <- doSysTrainLogisR(lrRes, impuAdjProTumor_Krij, max_numTopFeats = 50)
# saveRDS(sysTrainModels, './data/Discovery/logisR/featFilt_sysTrain_lrResSOA_SVsByProTissueInt_impuAdjProTumor_randSplit_100_5fold_auc_lmin_Krij.rds')
sysTrainModels <- readRDS('./data/Discovery/logisR/featFilt_sysTrain_lrResSOA_SVsByProTissueInt_impuAdjProTumor_randSplit_100_5fold_auc_lmin_Krij.rds')

# Visualize systematic training
vizSysTrainModelPerf(sysTrainModels, numTopFeats = 30)
```
=> Top 11 important features are selected, resulting in model with AUC of 0.997 [0.98,1].\
=> Top 18 important features are selected, resulting in model with AUC of 0.997 [0.97,1].\
=> Top 21 important features are selected, resulting in model with AUC of 0.997 [0.95,1].

**Noninformative models that demonstrate effects of top important features**
```{r}
se <- impuAdjProTumor_Krij
# Train models without top important features, aka noninformative models
numTopImpoFeats <- c(5, 10, 15, 20, 25)
# for (num in numTopImpoFeats) {
#   topImpoFeats <- topImpoFeatViz$fullImpoFeatTab$Feature[1:num]
#   # Train lasso logistic regression model
#   reducedData <- lrRes$featClusters$reducedData
#   x <- t(reducedData[!rownames(reducedData) %in% topImpoFeats,])
#   y <- colData(se)$Recurrence
#   lrResNonInfo <- runLogisR(x, y, targetClass = 'Yes', iter = 100, regularized_method = 'lasso',
#                             cvFold = 5, cvMeasure = 'auc', used_lambda = 'lambda.min',
#                             trainSet_ratio = 0.8, split_method = 'random split',
#                             plot_ROC = F, save_model = T)
#   saveRDS(lrResNonInfo, paste0('./data/Discovery/logisR/noninfo/featFilt_NonInfo_rmTop', num,
#                                '_lrResSOA_SVsByProTissueInt_impuAdjProTumor_randSplit_100_5fold_auc_lmin_Krij.rds'))
# }

# Display performance of noninformative models
meanAUC <- c()
lowerAUC <- c()
upperAUC <- c()
uselessModels <- which(lrRes$lrRes$nNonZero == 0)
if (length(uselessModels) != 0) {
  auc_roc <- lrRes$lrRes$auc_roc[-uselessModels]
} else {
  auc_roc <- lrRes$lrRes$auc_roc
}
meanAUC <- c(meanAUC, round(mean(auc_roc), 3))
ci <- calcCI(auc_roc, bootstrap = T)
lowerAUC <- c(lowerAUC, ci[1])
upperAUC <- c(upperAUC, ci[2])
for (num in numTopImpoFeats) {
  lrResNonInfo <- readRDS(paste0('./data/Discovery/logisR/noninfo/featFilt_NonInfo_rmTop', num,
                                 '_lrResSOA_SVsByProTissueInt_impuAdjProTumor_randSplit_100_5fold_auc_lmin_Krij.rds'))
  uselessModels <- which(lrResNonInfo$nNonZero == 0)
  if (length(uselessModels) != 0) {
    auc_roc <- lrResNonInfo$auc_roc[-uselessModels]
  } else {
    auc_roc <- lrResNonInfo$auc_roc
  }
  meanAUC <- c(meanAUC, round(mean(auc_roc), 3))
  ci <- calcCI(auc_roc, bootstrap = T)
  lowerAUC <- c(lowerAUC, ci[1])
  upperAUC <- c(upperAUC, ci[2])
}
aucStatTab <- data.frame(FeatSpace = c('Complete', paste0('-Top ', numTopImpoFeats)),
                         MeanAUC = meanAUC, UpperAUC = upperAUC, LowerAUC = lowerAUC) %>%
  dplyr::mutate(FeatSpace = factor(FeatSpace, levels = FeatSpace))

ggplot(aucStatTab, aes(x=FeatSpace, y=MeanAUC)) +
  geom_bar(stat = 'identity', position = position_dodge(), alpha = 0.9) +
  geom_errorbar(aes(ymin=LowerAUC, ymax=UpperAUC), width = 0.4) +
  labs(x = 'Feature space used to train 100 models', y = 'Mean AUC-ROC',
       title = 'Noninformative models') +
  th
```

<font size='5'> **Investigation of core features** </font>

<font size='3'> Imputed data: </font>\
**Heatmap where samples are ordered by linear regression scores**
```{r}
coreFeatPlotsImpu <- vizCoreFeats(lrRes, sysTrainModels, impuAdjProTumor_Krij,
                                  heatmap_numCoreFeats = 11, colFeatAnno = 'Genes',
                                  scatter_twoCoreFeats = c('RETN', 'MX1'),
                                  fontsize = 10, fontsize_col = 7, fontsize_row = 11,
                                  cellwidth = 6, cellheight = 18)
coreFeatPlotsImpu$heatmap
```

**Scatterplot of two core features**
```{r}
coreFeatPlotsImpu$scatterplot
```

<font size='3'> Original data: </font>\
**Heatmap**
```{r}
coreFeatPlots <- vizCoreFeats(lrRes, sysTrainModels, adjProTumor_Krij,
                              heatmap_numCoreFeats = 11, colFeatAnno = 'Genes',
                              scatter_twoCoreFeats = c('RETN', 'MX1'),
                              fontsize = 10, fontsize_col = 7, fontsize_row = 11,
                              cellwidth = 6, cellheight = 18)
coreFeatPlots$heatmap
```

**Scatterplot**
```{r}
coreFeatPlots$scatterplot
```

**Missingness and availability of core features**
```{r message=F}
numTopImpoFeats <- 11
# Prepare top important feature list
topImpoFeats <- topImpoFeatViz$fullImpoFeatTab$Feature[1:numTopImpoFeats] %>%
  stringr::str_remove(';.+')
topImpoFeatAnnos <- topImpoFeatViz$fullImpoFeatTab$Annotation[1:numTopImpoFeats] %>%
  stringr::str_remove(';.+')

# Show data missingness
proTumor_Krij <- proTissue_Krij[, colData(proTissue_Krij)$Condition %in% 'Tumor']
featAnnoTab <- tibble::as_tibble(rowData(proTissue_Krij), rownames = 'Feature') %>%
  dplyr::select(Feature, Genes) %>%
  dplyr::rename(Gene = Genes) %>%
  dplyr::mutate(Feature = stringr::str_remove(Feature, ';.+'),
                Gene = stringr::str_remove(Gene, ';.+'))
featAvailTab <- summExp2df(proTumor_Krij, row_id = 'Feature', col_id = 'Sample') %>%
  dplyr::select(Feature, Value, Recurrence) %>%
  dplyr::mutate(Value = dplyr::case_when(!is.na(Value) ~ 'Observed',
                                         is.na(Value) ~ 'Missing'),
                Feature = stringr::str_remove(Feature, ';.+'),
                Recurrence = dplyr::case_when(Recurrence %in% 'Yes' ~ 'Recurrence',
                                              Recurrence %in% 'No' ~ 'Non-Recurrence')) %>%
  dplyr::filter(Feature %in% topImpoFeats) %>%
  dplyr::group_by(Recurrence, Feature) %>%
  dplyr::mutate(SmpNumber = length(Feature)) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(Recurrence, SmpNumber, Feature, Value) %>%
  dplyr::summarise(Count = dplyr::n()) %>%
  dplyr::ungroup() %>%
  dplyr::left_join(featAnnoTab, by = 'Feature') %>%
  dplyr::mutate(Gene = factor(Gene, levels = topImpoFeatAnnos),
                Recurrence = factor(Recurrence, levels = c('Recurrence', 'Non-Recurrence')),
                Value = factor(Value, levels = c('Missing', 'Observed')),
                Proportion = Count/SmpNumber)

ggplot(featAvailTab, aes(x=Gene, y=Proportion, fill=Value)) +
  geom_col(position = 'stack') +
  facet_wrap(vars(Recurrence), scales = 'free') +
  scale_fill_manual(values = c(Missing = 'grey', Observed = 'black')) +
  labs(title = 'Tumor Tissue Proteomics') +
  theme_bw() +
  theme(axis.text.x = element_text(size = 18, face = 'bold', angle = 90, hjust = 1, vjust = 0.5),
        axis.text.y = element_text(size = 12, face = 'bold'),
        axis.title = element_text(size = 20, face = 'bold'),
        legend.title = element_blank(), legend.text = element_text(size = 18),
        strip.text = element_text(size = 18, face = 'bold'),
        plot.title = element_text(size = 22, face = 'bold', hjust = 0.5, vjust = 1.5))


# Show availability of top important features in other datasets
# Prepare feature space of other datasets
# New Klingmüller Preprocessed Combined dataset
new_proTissue_Klin <- readRDS('./data/Discovery/AG_Klingmueller/proTissueVsn.rds')
featSpace_new_Prepro_Klin <- rownames(new_proTissue_Klin) %>%
  stringr::str_remove(';.+')
# New Klingmüller Original Combined dataset
proPhosTissue_MDevDis <- readRDS('./data/Discovery/AG_Klingmueller/COMP_2696.SMART_Lung_MethodDiscov_ProPhos.rds')
featSpace_new_Ori_Klin <- rowData(proPhosTissue_MDevDis@ExperimentList$Proteome)$UniprotID %>%
  stringr::str_remove(';.+')
# Klingmüller Preprocessed Combined dataset
old_proTissue_Combined_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsn.rds')
featSpace_old_Prepro_Combined_Klin <- rownames(old_proTissue_Combined_Klin) %>%
  stringr::str_remove(';.+')
# Klingmüller Original Combined dataset
featSpace_old_Ori_Combined_Klin <- readr::read_tsv(paste0('./data/MethodDev/AG_Klingmueller/',
                                                          '20230309_103228_SmartCare_TumorFree_Tumor_Yang001_Report.tsv')) %>%
  dplyr::pull(PG.ProteinGroups) %>%
  stringr::str_remove(';.+')
# Collect all feature spaces into a list
featSpaceList <- list(Ori_New_Klin = featSpace_new_Ori_Klin,
                      Prepro_New_Klin = featSpace_new_Prepro_Klin,
                      Ori_Old_Klin = featSpace_old_Ori_Combined_Klin,
                      Prepro_Old_Klin = featSpace_old_Prepro_Combined_Klin)
# Store feature availability in a matrix
featAvailMat <- matrix(nrow = 4, ncol = numTopImpoFeats,
                       dimnames = list(c('Ori_New_Klin', 'Prepro_New_Klin',
                                         'Ori_Old_Klin', 'Prepro_Old_Klin'), topImpoFeats))
for (i in seq_len(nrow(featAvailMat))) {
  for (j in seq_len(ncol(featAvailMat))) {
    if (colnames(featAvailMat)[j] %in% featSpaceList[[rownames(featAvailMat)[i]]]) {
      featAvailMat[i, j] <- 'Yes'
    } else {
      featAvailMat[i, j] <- 'No'
    }
  }
}
# Convert matrix to long data for visualization
featAvailMat <- tibble::as_tibble(featAvailMat, rownames = 'Dataset') %>%
  tidyr::pivot_longer(cols = -'Dataset', names_to = 'Feature', values_to = 'Captured') %>%
  dplyr::mutate(Dataset = factor(Dataset, levels = rev(names(featSpaceList))),
                Captured = factor(Captured, levels = c('Yes', 'No')),
                Feature = plyr::mapvalues(Feature, from = topImpoFeats, to = topImpoFeatAnnos),
                Feature = factor(Feature, levels = topImpoFeatAnnos))

ggplot(featAvailMat, aes(x=Feature, y=Dataset, fill=Captured)) +
  geom_tile() +
  scale_fill_manual(values = c('black', 'grey80')) +
  labs(title = 'Tumor Tissue Proteomics') +
  theme(axis.ticks = element_blank(), panel.grid = element_blank(), panel.background = element_blank(),
        axis.text.x = element_text(size = 18, face = 'bold', angle = 45, hjust = 1, vjust = 1),
        axis.text.y = element_text(size = 18, face = 'bold'),
        axis.title = element_text(size = 20, face = 'bold'),
        legend.title = element_text(size = 20), legend.text = element_text(size = 18),
        plot.title = element_text(size = 28, face = 'bold', hjust = 0.5, vjust = 1.5))
```

## RF

<!-- <font size='5'> **Stability selection** </font> -->

<!-- **Performance statistics of 1000 models** -->
<!-- ```{r} -->
<!-- # Load imputed adjusted data -->
<!-- impuAdjProTumor_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_impuAdjProTumorVsn_SVs.rds') -->
<!-- # Load SOA results for retrieving significant feature list for initial feature selection -->
<!-- proTumorRes_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/soaRes/combined_proTumorRes_SVs.rds') -->

<!-- # Train models iteratively -->
<!-- # Prepare significant feature list -->
<!-- sigFeatList <- proTumorRes_Krij$sig.feat.tab$Var1 -->
<!-- # rfRes <- iterRF(impuAdjProTumor_Krij, doImputation = F, doInitFeatSelection = T, -->
<!-- #                 sigFeatList = sigFeatList, doFeatClustering = T, iter = 1000) -->
<!-- # saveRDS(rfRes, './data/Discovery/rf/rfResSOA_SVs_impuAdjProTumor_0.9r_randSplit_1000_Krij.rds') -->
<!-- rfRes <- readRDS('./data/Discovery/rf/rfResSOA_SVs_impuAdjProTumor_0.9r_randSplit_1000_Krij.rds') -->

<!-- # Report summarized scores -->
<!-- # Prepare predictions and ground truths -->
<!-- pred <- rfRes$rfRes$y_pred -->
<!-- truth <- rfRes$rfRes$y_truth -->
<!-- auc_roc <- rfRes$rfRes$auc_roc -->
<!-- summarizePredPower(pred, truth, auc_roc) -->
<!-- ``` -->

<!-- **Top important features with highest median ranks** -->
<!-- ```{r} -->
<!-- # Visualize top important features -->
<!-- # Prepare feature annotation table -->
<!-- proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds') -->
<!-- featAnno <- tibble::as_tibble(rowData(proTissue_Krij), rownames = 'Feature') %>% -->
<!--   dplyr::select(Feature, Genes) %>% -->
<!--   dplyr::rename(Gene = Genes) -->
<!-- topImpoFeatViz <- vizTopImpoFeatsRF(rfRes, featAnno = featAnno, num_p1TopFeats = 30) -->
<!-- topImpoFeatViz$rank + theme(axis.title = element_text(size = 28), -->
<!--                             axis.text = element_text(size = 16)) -->
<!-- ``` -->

<!-- **Full feature importance table** -->
<!-- ```{r} -->
<!-- DT::datatable(topImpoFeatViz$fullImpoFeatTab) -->
<!-- ``` -->

<!-- <font size='5'> **Systematic training and evaluation** </font> -->

<!-- **Performance of models trained with different numbers of top important features** -->
<!-- ```{r} -->
<!-- # Systematically train models with different numbers of top important features -->
<!-- # sysTrainModels <- doSysTrainRF(rfRes, impuAdjProTumor_Krij, max_numTopFeats = 100) -->
<!-- # saveRDS(sysTrainModels, './data/Discovery/rf/sysTrain_rfResSOA_SVs_impuAdjProTumor_randSplit_100_Krij.rds') -->
<!-- sysTrainModels <- readRDS('./data/Discovery/rf/sysTrain_rfResSOA_SVs_impuAdjProTumor_randSplit_100_Krij.rds') -->

<!-- # Visualize systematic training -->
<!-- vizSysTrainModelPerf(sysTrainModels, numTopFeats = 30) -->
<!-- ``` -->
<!-- => Top 3, resulting in model with AUC of 0.996 [0.97,1]\ -->
<!-- => Top 5, resulting in model with AUC of 0.997 [0.97,1]\ -->
<!-- => Top 9, resulting in model with AUC of 0.998 [0.98,1] -->

# NTP (AG Krij.)
Normal Tissue DIA Proteomics from AG Krijgsveld

## Lasso logistic regression

<font size='5'> **Stability selection** </font>

**Performance statistics of 1000 models**
```{r warning=T}
# Load preprocessed data
proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds')
proNormal_Krij <- proTissue_Krij[, colData(proTissue_Krij)$Condition %in% 'Normal']
# Load SOA results
proNormalRes_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/soaRes/combined_proNormalRes_SVsByProTissueInt.rds')
# Prepare adjusted data
adjProNormal_Krij <- proNormal_Krij
assay(adjProNormal_Krij) <- proNormalRes_Krij$SOA.res$dataCorrect
# Do additional feature filtering to remove features observed in less than 50% of samples
ifRmFeats <- apply(assay(adjProNormal_Krij), 1, function(featVec) {
  sum(is.na(featVec)) / length(featVec) > 0.5
})
adjProNormal_Krij <- adjProNormal_Krij[!ifRmFeats,]

# Load imputed adjusted data
impuAdjProNormal_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/impuData/featFilt_combined_impuAdjProNormalVsn_SVsByProTissueInt.rds')

# Train models iteratively
# Prepare significant feature list
sigFeatList <- proNormalRes_Krij$sig.feat.tab$Var1
# lrRes <- iterLogisR(adjProNormal_Krij, doImputation = T, doInitFeatSelection = T, sigFeatList = sigFeatList,
#                     doFeatClustering = T, iter = 1000, cvFold = 5, cvMeasure = 'auc')
# saveRDS(lrRes, './data/Discovery/logisR/featFilt_lrResSOA_SVsByProTissueInt_impuAdjProNormal_0.9r_randSplit_1000_5fold_auc_lmin_Krij.rds')
# saveRDS(lrRes$impuSE, './data/Discovery/AG_Krijgsveld/impuData/featFilt_combined_impuAdjProNormalVsn_SVsByProTissueInt.rds')
lrRes <- readRDS('./data/Discovery/logisR/featFilt_lrResSOA_SVsByProTissueInt_impuAdjProNormal_0.9r_randSplit_1000_5fold_auc_lmin_Krij.rds')

# Report summarized scores
# Prepare predictions and ground truths where results from useless models are removed
uselessModels <- which(lrRes$lrRes$nNonZero == 0)
if (length(uselessModels) != 0) {
  pred <- lrRes$lrRes$y_pred[-uselessModels]
  truth <- lrRes$lrRes$y_truth[-uselessModels]
  auc_roc <- lrRes$lrRes$auc_roc[-uselessModels]
} else {
  pred <- lrRes$lrRes$y_pred
  truth <- lrRes$lrRes$y_truth
  auc_roc <- lrRes$lrRes$auc_roc
}
suppressWarnings(summarizePredPower(pred, truth, auc_roc))
```

**Top important features that are most frequently selected**
```{r}
# Visualize top important features
# Prepare feature annotation table
featAnno <- tibble::as_tibble(rowData(proTissue_Krij), rownames = 'Feature') %>%
  dplyr::select(Feature, Genes) %>%
  dplyr::rename(Gene = Genes)
topImpoFeatViz <- vizTopImpoFeatsLR(lrRes, featAnno = featAnno, num_p1TopFeats = 30)
topImpoFeatViz$pick
```

**Full feature importance table**
```{r}
# Display full feature importance table
DT::datatable(topImpoFeatViz$fullImpoFeatTab)
```

<font size='5'> **Systematic training and evaluation** </font>

**Performance of models trained with different numbers of top important features**
```{r}
# Systematically train models with different numbers of top important features
# sysTrainModels <- doSysTrainLogisR(lrRes, impuAdjProNormal_Krij, max_numTopFeats = 50)
# saveRDS(sysTrainModels, './data/Discovery/logisR/featFilt_sysTrain_lrResSOA_SVsByProTissueInt_impuAdjProNormal_randSplit_100_5fold_auc_lmin_Krij.rds')
sysTrainModels <- readRDS('./data/Discovery/logisR/featFilt_sysTrain_lrResSOA_SVsByProTissueInt_impuAdjProNormal_randSplit_100_5fold_auc_lmin_Krij.rds')

# Visualize systematic training
vizSysTrainModelPerf(sysTrainModels, numTopFeats = 30)
```
=> Top 5 important features are selected, resulting in model with AUC of 0.991 [0.95,1].\
=> Top 8 important features are selected, resulting in model with AUC of 0.993 [0.95,1].\
=> Top 11 important features are selected, resulting in model with AUC of 0.994 [0.95,1].\
=> Top 20 important features are selected, resulting in model with AUC of 0.994 [0.95,1].

**Noninformative models that demonstrate effects of top important features**
```{r}
se <- impuAdjProNormal_Krij
# Train models without top important features, aka noninformative models
numTopImpoFeats <- c(5, 10, 15, 20)
# for (num in numTopImpoFeats) {
#   topImpoFeats <- topImpoFeatViz$fullImpoFeatTab$Feature[1:num]
#   # Train lasso logistic regression model
#   reducedData <- lrRes$featClusters$reducedData
#   x <- t(reducedData[!rownames(reducedData) %in% topImpoFeats,])
#   y <- colData(se)$Recurrence
#   lrResNonInfo <- runLogisR(x, y, targetClass = 'Yes', iter = 100, regularized_method = 'lasso',
#                             cvFold = 5, cvMeasure = 'auc', used_lambda = 'lambda.min',
#                             trainSet_ratio = 0.8, split_method = 'random split',
#                             plot_ROC = F, save_model = T)
#   saveRDS(lrResNonInfo, paste0('./data/Discovery/logisR/noninfo/featFilt_NonInfo_rmTop', num,
#                                '_lrResSOA_SVsByProTissueInt_impuAdjProNormal_randSplit_100_5fold_auc_lmin_Krij.rds'))
# }

# Display performance of noninformative models
meanAUC <- c()
lowerAUC <- c()
upperAUC <- c()
uselessModels <- which(lrRes$lrRes$nNonZero == 0)
if (length(uselessModels) != 0) {
  auc_roc <- lrRes$lrRes$auc_roc[-uselessModels]
} else {
  auc_roc <- lrRes$lrRes$auc_roc
}
meanAUC <- c(meanAUC, round(mean(auc_roc), 3))
ci <- calcCI(auc_roc, bootstrap = T)
lowerAUC <- c(lowerAUC, ci[1])
upperAUC <- c(upperAUC, ci[2])
for (num in numTopImpoFeats) {
  lrResNonInfo <- readRDS(paste0('./data/Discovery/logisR/noninfo/featFilt_NonInfo_rmTop', num,
                                 '_lrResSOA_SVsByProTissueInt_impuAdjProNormal_randSplit_100_5fold_auc_lmin_Krij.rds'))
  uselessModels <- which(lrResNonInfo$nNonZero == 0)
  if (length(uselessModels) != 0) {
    auc_roc <- lrResNonInfo$auc_roc[-uselessModels]
  } else {
    auc_roc <- lrResNonInfo$auc_roc
  }
  meanAUC <- c(meanAUC, round(mean(auc_roc), 3))
  ci <- calcCI(auc_roc, bootstrap = T)
  lowerAUC <- c(lowerAUC, ci[1])
  upperAUC <- c(upperAUC, ci[2])
}
aucStatTab <- data.frame(FeatSpace = c('Complete', paste0('-Top ', numTopImpoFeats)),
                         MeanAUC = meanAUC, UpperAUC = upperAUC, LowerAUC = lowerAUC) %>%
  dplyr::mutate(FeatSpace = factor(FeatSpace, levels = FeatSpace))

ggplot(aucStatTab, aes(x=FeatSpace, y=MeanAUC)) +
  geom_bar(stat = 'identity', position = position_dodge(), alpha = 0.9) +
  geom_errorbar(aes(ymin=LowerAUC, ymax=UpperAUC), width = 0.4) +
  labs(x = 'Feature space used to train 100 models', y = 'Mean AUC-ROC',
       title = 'Noninformative models') +
  th
```

<font size='5'> **Investigation of core features** </font>

<font size='3'> Imputed data: </font>\
**Heatmap where samples are ordered by linear regression scores**
```{r}
coreFeatPlotsImpu <- vizCoreFeats(lrRes, sysTrainModels, impuAdjProNormal_Krij,
                                  heatmap_numCoreFeats = 5, colFeatAnno = 'Genes',
                                  scatter_twoCoreFeats = c('AQP5', 'WDR55'),
                                  fontsize = 10, fontsize_col = 7, fontsize_row = 11,
                                  cellwidth = 6, cellheight = 18)
coreFeatPlotsImpu$heatmap
```

**Scatterplot of two core features**
```{r}
coreFeatPlotsImpu$scatterplot
```

<font size='3'> Original data: </font>\
**Heatmap**
```{r}
coreFeatPlots <- vizCoreFeats(lrRes, sysTrainModels, adjProNormal_Krij,
                              heatmap_numCoreFeats = 5, colFeatAnno = 'Genes',
                              scatter_twoCoreFeats = c('AQP5', 'PRG4'),
                              fontsize = 10, fontsize_col = 7, fontsize_row = 11,
                              cellwidth = 6, cellheight = 18)
coreFeatPlots$heatmap
```

**Scatterplot**
```{r}
coreFeatPlots$scatterplot
```

**Missingness and availability of core features**
```{r message=F}
numTopImpoFeats <- 5
# Prepare top important feature list
topImpoFeats <- topImpoFeatViz$fullImpoFeatTab$Feature[1:numTopImpoFeats] %>%
  stringr::str_remove(';.+')
topImpoFeatAnnos <- topImpoFeatViz$fullImpoFeatTab$Annotation[1:numTopImpoFeats] %>%
  stringr::str_remove(';.+')

# Show data missingness
proNormal_Krij <- proTissue_Krij[, colData(proTissue_Krij)$Condition %in% 'Normal']
featAnnoTab <- tibble::as_tibble(rowData(proTissue_Krij), rownames = 'Feature') %>%
  dplyr::select(Feature, Genes) %>%
  dplyr::rename(Gene = Genes) %>%
  dplyr::mutate(Feature = stringr::str_remove(Feature, ';.+'),
                Gene = stringr::str_remove(Gene, ';.+'))
featAvailTab <- summExp2df(proNormal_Krij, row_id = 'Feature', col_id = 'Sample') %>%
  dplyr::select(Feature, Value, Recurrence) %>%
  dplyr::mutate(Value = dplyr::case_when(!is.na(Value) ~ 'Observed',
                                         is.na(Value) ~ 'Missing'),
                Feature = stringr::str_remove(Feature, ';.+'),
                Recurrence = dplyr::case_when(Recurrence %in% 'Yes' ~ 'Recurrence',
                                              Recurrence %in% 'No' ~ 'Non-Recurrence')) %>%
  dplyr::filter(Feature %in% topImpoFeats) %>%
  dplyr::group_by(Recurrence, Feature) %>%
  dplyr::mutate(SmpNumber = length(Feature)) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(Recurrence, SmpNumber, Feature, Value) %>%
  dplyr::summarise(Count = dplyr::n()) %>%
  dplyr::ungroup() %>%
  dplyr::left_join(featAnnoTab, by = 'Feature') %>%
  dplyr::mutate(Gene = factor(Gene, levels = topImpoFeatAnnos),
                Recurrence = factor(Recurrence, levels = c('Recurrence', 'Non-Recurrence')),
                Value = factor(Value, levels = c('Missing', 'Observed')),
                Proportion = Count/SmpNumber)

ggplot(featAvailTab, aes(x=Gene, y=Proportion, fill=Value)) +
  geom_col(position = 'stack') +
  facet_wrap(vars(Recurrence), scales = 'free') +
  scale_fill_manual(values = c(Missing = 'grey', Observed = 'black')) +
  labs(title = 'Normal Tissue Proteomics') +
  theme_bw() +
  theme(axis.text.x = element_text(size = 18, face = 'bold', angle = 90, hjust = 1, vjust = 0.5),
        axis.text.y = element_text(size = 12, face = 'bold'),
        axis.title = element_text(size = 20, face = 'bold'),
        legend.title = element_blank(), legend.text = element_text(size = 18),
        strip.text = element_text(size = 18, face = 'bold'),
        plot.title = element_text(size = 22, face = 'bold', hjust = 0.5, vjust = 1.5))


# Show availability of top important features in other datasets
# Prepare feature space of other datasets
# New Klingmüller Preprocessed Combined dataset
new_proTissue_Klin <- readRDS('./data/Discovery/AG_Klingmueller/proTissueVsn.rds')
featSpace_new_Prepro_Klin <- rownames(new_proTissue_Klin) %>%
  stringr::str_remove(';.+')
# New Klingmüller Original Combined dataset
proPhosTissue_MDevDis <- readRDS('./data/Discovery/AG_Klingmueller/COMP_2696.SMART_Lung_MethodDiscov_ProPhos.rds')
featSpace_new_Ori_Klin <- rowData(proPhosTissue_MDevDis@ExperimentList$Proteome)$UniprotID %>%
  stringr::str_remove(';.+')
# Klingmüller Preprocessed Combined dataset
old_proTissue_Combined_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsn.rds')
featSpace_old_Prepro_Combined_Klin <- rownames(old_proTissue_Combined_Klin) %>%
  stringr::str_remove(';.+')
# Klingmüller Original Combined dataset
featSpace_old_Ori_Combined_Klin <- readr::read_tsv(paste0('./data/MethodDev/AG_Klingmueller/',
                                                          '20230309_103228_SmartCare_TumorFree_Tumor_Yang001_Report.tsv')) %>%
  dplyr::pull(PG.ProteinGroups) %>%
  stringr::str_remove(';.+')
# Collect all feature spaces into a list
featSpaceList <- list(Ori_New_Klin = featSpace_new_Ori_Klin,
                      Prepro_New_Klin = featSpace_new_Prepro_Klin,
                      Ori_Old_Klin = featSpace_old_Ori_Combined_Klin,
                      Prepro_Old_Klin = featSpace_old_Prepro_Combined_Klin)
# Store feature availability in a matrix
featAvailMat <- matrix(nrow = 4, ncol = numTopImpoFeats,
                       dimnames = list(c('Ori_New_Klin', 'Prepro_New_Klin',
                                         'Ori_Old_Klin', 'Prepro_Old_Klin'), topImpoFeats))
for (i in seq_len(nrow(featAvailMat))) {
  for (j in seq_len(ncol(featAvailMat))) {
    if (colnames(featAvailMat)[j] %in% featSpaceList[[rownames(featAvailMat)[i]]]) {
      featAvailMat[i, j] <- 'Yes'
    } else {
      featAvailMat[i, j] <- 'No'
    }
  }
}
# Convert matrix to long data for visualization
featAvailMat <- tibble::as_tibble(featAvailMat, rownames = 'Dataset') %>%
  tidyr::pivot_longer(cols = -'Dataset', names_to = 'Feature', values_to = 'Captured') %>%
  dplyr::mutate(Dataset = factor(Dataset, levels = rev(names(featSpaceList))),
                Captured = factor(Captured, levels = c('Yes', 'No')),
                Feature = plyr::mapvalues(Feature, from = topImpoFeats, to = topImpoFeatAnnos),
                Feature = factor(Feature, levels = topImpoFeatAnnos))

ggplot(featAvailMat, aes(x=Feature, y=Dataset, fill=Captured)) +
  geom_tile() +
  scale_fill_manual(values = c('black', 'grey80')) +
  labs(title = 'Normal Tissue Proteomics') +
  theme(axis.ticks = element_blank(), panel.grid = element_blank(), panel.background = element_blank(),
        axis.text.x = element_text(size = 18, face = 'bold', angle = 45, hjust = 1, vjust = 1),
        axis.text.y = element_text(size = 18, face = 'bold'),
        axis.title = element_text(size = 20, face = 'bold'),
        legend.title = element_text(size = 20), legend.text = element_text(size = 18),
        plot.title = element_text(size = 28, face = 'bold', hjust = 0.5, vjust = 1.5))
```

## RF

<!-- <font size='5'> **Stability selection** </font> -->

<!-- **Performance statistics of 1000 models** -->
<!-- ```{r} -->
<!-- # Load imputed adjusted data -->
<!-- impuAdjProNormal_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_impuAdjProNormalVsn_Smoking_Gender.rds') -->
<!-- # Load SOA results for retrieving significant feature list for initial feature selection -->
<!-- proNormalRes_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/soaRes/combined_proNormalRes_Smoking_Gender.rds') -->

<!-- # Train models iteratively -->
<!-- # Prepare significant feature list -->
<!-- sigFeatList <- proNormalRes_Krij$sig.feat.tab$Var1 -->
<!-- # rfRes <- iterRF(impuAdjProNormal_Krij, doImputation = F, doInitFeatSelection = T, -->
<!-- #                 sigFeatList = sigFeatList, doFeatClustering = T, iter = 1000) -->
<!-- # saveRDS(rfRes, './data/Discovery/rf/rfResSOA_Smoking_Gender_impuAdjProNormal_0.9r_randSplit_1000_Krij.rds') -->
<!-- rfRes <- readRDS('./data/Discovery/rf/rfResSOA_Smoking_Gender_impuAdjProNormal_0.9r_randSplit_1000_Krij.rds') -->

<!-- # Report summarized scores -->
<!-- # Prepare predictions and ground truths -->
<!-- pred <- rfRes$rfRes$y_pred -->
<!-- truth <- rfRes$rfRes$y_truth -->
<!-- auc_roc <- rfRes$rfRes$auc_roc -->
<!-- summarizePredPower(pred, truth, auc_roc) -->
<!-- ``` -->

<!-- **Top important features with highest median ranks** -->
<!-- ```{r} -->
<!-- # Visualize top important features -->
<!-- # Prepare feature annotation table -->
<!-- proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds') -->
<!-- featAnno <- tibble::as_tibble(rowData(proTissue_Krij), rownames = 'Feature') %>% -->
<!--   dplyr::select(Feature, Genes) %>% -->
<!--   dplyr::rename(Gene = Genes) -->
<!-- topImpoFeatViz <- vizTopImpoFeatsRF(rfRes, featAnno = featAnno, num_p1TopFeats = 30) -->
<!-- topImpoFeatViz$rank + theme(axis.title = element_text(size = 28), -->
<!--                             axis.text = element_text(size = 16)) -->
<!-- ``` -->

<!-- **Full feature importance table** -->
<!-- ```{r} -->
<!-- DT::datatable(topImpoFeatViz$fullImpoFeatTab) -->
<!-- ``` -->

<!-- <font size='5'> **Systematic training and evaluation** </font> -->

<!-- **Performance of models trained with different numbers of top important features** -->
<!-- ```{r} -->
<!-- # Systematically train models with different numbers of top important features -->
<!-- # sysTrainModels <- doSysTrainRF(rfRes, impuAdjProNormal_Krij, max_numTopFeats = 100) -->
<!-- # saveRDS(sysTrainModels, './data/Discovery/rf/sysTrain_rfResSOA_Smoking_Gender_impuAdjProNormal_randSplit_100_Krij.rds') -->
<!-- sysTrainModels <- readRDS('./data/Discovery/rf/sysTrain_rfResSOA_Smoking_Gender_impuAdjProNormal_randSplit_100_Krij.rds') -->

<!-- # Visualize systematic training -->
<!-- vizSysTrainModelPerf(sysTrainModels, numTopFeats = 40) -->
<!-- ``` -->
<!-- => Top 13, resulting in model with AUC of 0.977 [0.92,1]\ -->
<!-- => Top 23, resulting in model with AUC of 0.977 [0.91,1] -->




<!-- # Unt. Lipid. -->
<!-- Untargeted Lipidomics -->

<!-- ## RF -->

<!-- ```{r} -->
<!-- # NTL -->
<!-- # Load preprocessed data -->
<!-- # Discovery -->
<!-- # dis_lipTissue_Hopf <- readRDS('./data/Discovery/AG_Hopf/lipTissueVsn_WBC25.rds') -->
<!-- # Method development -->
<!-- mDev_lipTissue_Hopf <- readRDS('./data/MethodDev/AG_Hopf/lipTissueVsn.rds') -->
<!-- lipTissue <- mDev_lipTissue_Hopf -->
<!-- # Subset data -->
<!-- lipNormal <- lipTissue[, colData(lipTissue)$Condition == 'Normal'] -->
<!-- # Remove features observed in less than 67% of samples -->
<!-- rmFeats <- apply(assay(lipNormal), 1, function(featVec) { -->
<!--   obProp <- sum(!is.na(featVec)) / length(featVec) -->
<!--   obProp <= 0.67 -->
<!-- }) %>% -->
<!--   which() -->
<!-- if (length(rmFeats) != 0) { -->
<!--   lipNormal <- lipNormal[-rmFeats,] -->
<!-- } -->
<!-- # Impute missing values using missForest -->
<!-- dat <- t(assay(lipNormal)) -->
<!-- impuDat <- missForest(dat, maxiter = 10, verbose = T)$ximp %>% -->
<!--   t() -->
<!-- assay(lipNormal) <- impuDat -->
<!-- # saveRDS(lipNormal, './data/MethodDev/AG_Hopf/impuByMF//lipNormalVsnImpu.rds') -->

<!-- # BPL -->
<!-- # Load preprocessed data -->
<!-- # Discovery -->
<!-- # dis_lipBase_Hopf <- readRDS('./data/Discovery/AG_Hopf/lipBaseVsn_B1WBC25.rds') -->
<!-- # Method development -->
<!-- # Subset data -->
<!-- mDev_lipPlasma_Hopf <- readRDS('./data/MethodDev/AG_Hopf/lipPlasmaVsn.rds') -->
<!-- lipPlasma <- mDev_lipPlasma_Hopf -->
<!-- lipBase <- lipPlasma[, colData(lipPlasma)$TimePoint == 'Baseline'] -->
<!-- # Remove features observed in less than 67% of samples -->
<!-- rmFeats <- apply(assay(lipBase), 1, function(featVec) { -->
<!--   obProp <- sum(!is.na(featVec)) / length(featVec) -->
<!--   obProp <= 0.67 -->
<!-- }) %>% -->
<!--   which() -->
<!-- if (length(rmFeats) != 0) { -->
<!--   lipBase <- lipBase[-rmFeats,] -->
<!-- } -->
<!-- # Impute missing values using missForest -->
<!-- dat <- t(assay(lipBase)) -->
<!-- impuDat <- missForest(dat, maxiter = 10, verbose = T)$ximp %>% -->
<!--   t() -->
<!-- assay(lipBase) <- impuDat -->
<!-- # saveRDS(lipBase, './data/MethodDev/AG_Hopf/impuByMF/lipBaseVsnImpu.rds') -->


<!-- # Train RF -->
<!-- # rfRes <- iterRF(lipBase) -->
<!-- # saveRDS(rfRes, './data/Discovery/rf/untLip/rfResSig_mDev_untLipBase_randSplit_1000') -->
<!-- ``` -->

<!-- ### Dis. NTL -->
<!-- **Average model performance (Uncorrelated significant features, iter = 1000)** -->
<!-- ```{r} -->
<!-- # Load RF training results -->
<!-- rfRes <- readRDS('./data/Discovery/rf/untLip/rfResSig_dis_untLipNormal_randSplit_1000') -->

<!-- # Report summarized scores -->
<!-- trainedModels <- rfRes$rfRes$rfRes -->
<!-- # Prepare predictions and ground truths -->
<!-- pred <- as.list(rep(NA, length(trainedModels))) -->
<!-- for (i in seq_len(length(trainedModels))) { -->
<!--   pred[[i]] <- trainedModels[[i]]$test$predicted -->
<!-- } -->
<!-- truth <- rfRes$rfRes$y_truth -->
<!-- auc_roc <- rfRes$rfRes$auc_roc -->
<!-- summarizePredPower(pred, truth, auc_roc) -->
<!-- ``` -->

<!-- ### Dis. BPL -->
<!-- **Average model performance (Uncorrelated significant features, iter = 1000)** -->
<!-- ```{r} -->
<!-- # Load RF training results -->
<!-- rfRes <- readRDS('./data/Discovery/rf/untLip/rfResSig_dis_untLipBase_randSplit_1000') -->

<!-- # Report summarized scores -->
<!-- trainedModels <- rfRes$rfRes$rfRes -->
<!-- # Prepare predictions and ground truths -->
<!-- pred <- as.list(rep(NA, length(trainedModels))) -->
<!-- for (i in seq_len(length(trainedModels))) { -->
<!--   pred[[i]] <- trainedModels[[i]]$test$predicted -->
<!-- } -->
<!-- truth <- rfRes$rfRes$y_truth -->
<!-- auc_roc <- rfRes$rfRes$auc_roc -->
<!-- summarizePredPower(pred, truth, auc_roc) -->
<!-- ``` -->

<!-- ### MDev. NTL -->
<!-- **Average model performance (Uncorrelated significant features, iter = 1000)** -->
<!-- ```{r} -->
<!-- # Load RF training results -->
<!-- rfRes <- readRDS('./data/Discovery/rf/untLip/rfResSig_mDev_untLipNormal_randSplit_1000') -->

<!-- # Report summarized scores -->
<!-- trainedModels <- rfRes$rfRes$rfRes -->
<!-- # Prepare predictions and ground truths -->
<!-- pred <- as.list(rep(NA, length(trainedModels))) -->
<!-- for (i in seq_len(length(trainedModels))) { -->
<!--   pred[[i]] <- trainedModels[[i]]$test$predicted -->
<!-- } -->
<!-- truth <- rfRes$rfRes$y_truth -->
<!-- auc_roc <- rfRes$rfRes$auc_roc -->
<!-- summarizePredPower(pred, truth, auc_roc) -->
<!-- ``` -->

<!-- ### MDev. BPL -->
<!-- **Average model performance (Uncorrelated significant features, iter = 1000)** -->
<!-- ```{r} -->
<!-- # Load RF training results -->
<!-- rfRes <- readRDS('./data/Discovery/rf/untLip/rfResSig_mDev_untLipBase_randSplit_1000') -->

<!-- # Report summarized scores -->
<!-- trainedModels <- rfRes$rfRes$rfRes -->
<!-- # Prepare predictions and ground truths -->
<!-- pred <- as.list(rep(NA, length(trainedModels))) -->
<!-- for (i in seq_len(length(trainedModels))) { -->
<!--   pred[[i]] <- trainedModels[[i]]$test$predicted -->
<!-- } -->
<!-- truth <- rfRes$rfRes$y_truth -->
<!-- auc_roc <- rfRes$rfRes$auc_roc -->
<!-- summarizePredPower(pred, truth, auc_roc) -->
<!-- ``` -->

