---
title: 'Feature selection: Machine learning models'
author: "Qian-Wu Liao"
output:
  workflowr::wflow_html:
    toc: true
    toc_float: true
    code_folding: hide
---

<font size='4'> Description: Do feature selection with lasso logistic regression
models trained on combined data of cohorts (Method Development and Discovery) to
identify potential biomarkers for predicting lung cancer recurrence. Additional
feature filtering on subdatasets is not conducted, so top important features selected
by models are mostly highly missing. There is risk that imputation method (missForest)
used in our workflow introducing biases and impeding true but weaker signals contributed
by features more complete. On other hand, with brief literature research, those
top important features with high level of missingness seem associated with lung
adenocarcinoma. Moreover, imputed values seem sensible because they are usually of
low abundance, which matches property of MS-based data that missingness is often
MNAR. Next action will be using data imputed by our developing imputation method
(msBayesImpute) to train models to do feature selection and checking result consistency,
which may provides more confidence in our findings. </font>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 8, fig.width = 10, warning = F)
knitr::opts_knit$set(root.dir = '/Users/qianwu/Desktop/SMART-CARE_LungCancer')
```

Load libraries
```{r library loading, message = F}
library('spsUtil')
library(ggh4x)
library(DT)
library('randomForest')
library('missForest')
library(xgboost)
library('caret')
library('glmnet')
library(ggvenn)
library(pheatmap)
library('SummarizedExperiment')
library('tidyverse')
# Load user-defined functions
source('./code/misc.R')
source('./code/ml_funcs.R')

# Set plot theme
th <- theme_bw(base_size = 15) +
  theme(axis.title = element_text(face = 'bold'),
        axis.text = element_text(face = 'bold'),
        axis.ticks = element_line(linewidth = 0.8),
        legend.text = element_text(size = 15))
```

<font size='6'> Lasso logistic regression </font>

# TTP (AG Klin.)
Tumor Tissue DIA Proteomics from AG Klingmüller

<font size='5'> **Stability selection** </font>

**Performance statistics of 1000 models**
```{r warning=T}
# Load preprocessed data
proTissue_Klin <- readRDS('./data/Discovery/AG_Klingmueller/proTissueVsn.rds')
proTumor_Klin <- proTissue_Klin[, colData(proTissue_Klin)$Condition %in% 'Tumor']
# Load SOA results
proTumorRes_Klin <- readRDS('./data/Discovery/AG_Klingmueller/soaRes/proTumorRes_SVsByProTissueInt.rds')
# Prepare adjusted data
adjProTumor_Klin <- proTumor_Klin
assay(adjProTumor_Klin) <- proTumorRes_Klin$SOA.res$dataCorrect

# Load imputed adjusted data
impuAdjProTumor_Klin <- readRDS('./data/Discovery/AG_Klingmueller/impuData/impuAdjProTumorVsn_SVsByProTissueInt.rds')

# Train models iteratively
# Prepare significant feature list
sigFeatList <- proTumorRes_Klin$sig.feat.tab$Var1
# lrRes <- iterLogisR(impuAdjProTumor_Klin, doImputation = F, doInitFeatSelection = T, sigFeatList = sigFeatList,
#                     doFeatClustering = T, iter = 1000, cvFold = 5, cvMeasure = 'auc')
# saveRDS(lrRes, './data/Discovery/logisR/lrResSOA_SVsByProTissueInt_impuAdjProTumor_0.9r_randSplit_1000_5fold_auc_lmin_Klin.rds')
# saveRDS(lrRes$impuSE, './data/Discovery/AG_Klingmueller/impuData/impuAdjProTumorVsn_SVsByProTissueInt.rds')
lrRes <- readRDS('./data/Discovery/logisR/lrResSOA_SVsByProTissueInt_impuAdjProTumor_0.9r_randSplit_1000_5fold_auc_lmin_Klin.rds')

# Report summarized scores
# Prepare predictions and ground truths where results from useless models are removed
uselessModels <- which(lrRes$lrRes$nNonZero == 0)
if (length(uselessModels) != 0) {
  pred <- lrRes$lrRes$y_pred[-uselessModels]
  truth <- lrRes$lrRes$y_truth[-uselessModels]
  auc_roc <- lrRes$lrRes$auc_roc[-uselessModels]
} else {
  pred <- lrRes$lrRes$y_pred
  truth <- lrRes$lrRes$y_truth
  auc_roc <- lrRes$lrRes$auc_roc
}
suppressWarnings(summarizePredPower(pred, truth, auc_roc))
```

**Top important features that are most frequently selected**
```{r}
# Visualize top important features
# Prepare feature annotation table
featAnno <- tibble::as_tibble(rowData(proTissue_Klin), rownames = 'Feature')
topImpoFeatViz <- vizTopImpoFeatsLR(lrRes, featAnno = featAnno, num_p1TopFeats = 30)
topImpoFeatViz$pick
```

**Full feature importance table**
```{r}
DT::datatable(topImpoFeatViz$fullImpoFeatTab)
```

<font size='5'> **Systematic training and evaluation** </font>

**Performance of models trained with different numbers of top important features**
```{r}
# Systematically train models with different numbers of top important features
# sysTrainModels <- doSysTrainLogisR(lrRes, impuAdjProTumor_Klin, max_numTopFeats = 50)
# saveRDS(sysTrainModels, './data/Discovery/logisR/sysTrain_lrResSOA_SVsByProTissueInt_impuAdjProTumor_randSplit_100_5fold_auc_lmin_Klin.rds')
sysTrainModels <- readRDS('./data/Discovery/logisR/sysTrain_lrResSOA_SVsByProTissueInt_impuAdjProTumor_randSplit_100_5fold_auc_lmin_Klin.rds')

# Visualize systematic training
vizSysTrainModelPerf(sysTrainModels, numTopFeats = 30)
# ggsave('./output/Junyan/stabSelec_Stats_proTumor.png', device = 'png', dpi = 400, height = 8, width = 10)
```
=> Top 18 important features are selected, resulting in model with AUC of 0.993 [0.95,1].

**Noninformative models that demonstrate effects of top important features**
```{r}
se <- impuAdjProTumor_Klin
# Train models without top important features, aka noninformative models
numTopImpoFeats <- c(5, 10, 15, 20)
# for (num in numTopImpoFeats) {
#   topImpoFeats <- topImpoFeatViz$fullImpoFeatTab$Feature[1:num]
#   # Train lasso logistic regression model
#   reducedData <- lrRes$featClusters$reducedData
#   x <- t(reducedData[!rownames(reducedData) %in% topImpoFeats,])
#   y <- colData(se)$Recurrence
#   lrResNonInfo <- runLogisR(x, y, targetClass = 'Yes', iter = 100, regularized_method = 'lasso',
#                             cvFold = 5, cvMeasure = 'auc', used_lambda = 'lambda.min',
#                             trainSet_ratio = 0.8, split_method = 'random split',
#                             plot_ROC = F, save_model = T)
#   saveRDS(lrResNonInfo, paste0('./data/Discovery/logisR/noninfo/NonInfo_rmTop', num,
#                                '_lrResSOA_SVsByProTissueInt_impuAdjProTumor_randSplit_100_5fold_auc_lmin_Klin.rds'))
# }

# Display performance of noninformative models
meanAUC <- c()
lowerAUC <- c()
upperAUC <- c()
uselessModels <- which(lrRes$lrRes$nNonZero == 0)
if (length(uselessModels) != 0) {
  auc_roc <- lrRes$lrRes$auc_roc[-uselessModels]
} else {
  auc_roc <- lrRes$lrRes$auc_roc
}
meanAUC <- c(meanAUC, round(mean(auc_roc), 3))
ci <- calcCI(auc_roc, bootstrap = T)
lowerAUC <- c(lowerAUC, ci[1])
upperAUC <- c(upperAUC, ci[2])
for (num in numTopImpoFeats) {
  lrResNonInfo <- readRDS(paste0('./data/Discovery/logisR/noninfo/NonInfo_rmTop', num,
                                 '_lrResSOA_SVsByProTissueInt_impuAdjProTumor_randSplit_100_5fold_auc_lmin_Klin.rds'))
  uselessModels <- which(lrResNonInfo$nNonZero == 0)
  if (length(uselessModels) != 0) {
    auc_roc <- lrResNonInfo$auc_roc[-uselessModels]
  } else {
    auc_roc <- lrResNonInfo$auc_roc
  }
  meanAUC <- c(meanAUC, round(mean(auc_roc), 3))
  ci <- calcCI(auc_roc, bootstrap = T)
  lowerAUC <- c(lowerAUC, ci[1])
  upperAUC <- c(upperAUC, ci[2])
}
aucStatTab <- data.frame(FeatSpace = c('Complete', paste0('-Top ', numTopImpoFeats)),
                         MeanAUC = meanAUC, UpperAUC = upperAUC, LowerAUC = lowerAUC) %>%
  dplyr::mutate(FeatSpace = factor(FeatSpace, levels = FeatSpace))

ggplot(aucStatTab, aes(x=FeatSpace, y=MeanAUC)) +
  geom_bar(stat = 'identity', position = position_dodge(), alpha = 0.9) +
  geom_errorbar(aes(ymin=LowerAUC, ymax=UpperAUC), width = 0.4) +
  labs(x = 'Feature space used to train 100 models', y = 'Mean AUC-ROC',
       title = 'Noninformative models') +
  th
```

<font size='5'> **Investigation of core features** </font>\
*Keep TGFBR3 in mind*

<font size='3'> Imputed data: </font>\
**Heatmap where samples are ordered by linear regression scores**
```{r}
coreFeatPlotsImpu <- vizCoreFeats(lrRes, sysTrainModels, impuAdjProTumor_Klin,
                                  heatmap_numCoreFeats = 18, colFeatAnno = 'Gene',
                                  scatter_twoCoreFeats = c('LDB3', 'FMO5'),
                                  fontsize = 10, fontsize_col = 7, fontsize_row = 11,
                                  cellwidth = 6, cellheight = 18)
coreFeatPlotsImpu$heatmap
# ggsave('./output/Junyan/stabSelec_heatmap_proNormal_impu.png', device = 'png', dpi = 400, height = 8, width = 10)
```

**Scatterplot of two core features**
```{r}
coreFeatPlotsImpu$scatterplot
# ggsave('./output/Junyan/stabSelec_top1&3_proTumor_impu.png', device = 'png', dpi = 400, height = 8, width = 10)
```

<font size='3'> Original data: </font>\
**Heatmap**
```{r}
coreFeatPlots <- vizCoreFeats(lrRes, sysTrainModels, adjProTumor_Klin,
                              heatmap_numCoreFeats = 18, colFeatAnno = 'Gene',
                              scatter_twoCoreFeats = c('LDB3', 'FABP4'),
                              fontsize = 10, fontsize_col = 7, fontsize_row = 11,
                              cellwidth = 6, cellheight = 18)
coreFeatPlots$heatmap
# ggsave('./output/Junyan/stabSelec_heatmap_proTumor_ori.png', plot = a, device = 'png', dpi = 400, height = 8, width = 10)
```

**Scatterplot**
```{r}
coreFeatPlots$scatterplot
```

**Missingness and availability of core features**
```{r message=F}
numTopImpoFeats <- 18
# Prepare top important feature list
topImpoFeats <- topImpoFeatViz$fullImpoFeatTab$Feature[1:numTopImpoFeats] %>%
  stringr::str_remove(';.+')
topImpoFeatAnnos <- topImpoFeatViz$fullImpoFeatTab$Annotation[1:numTopImpoFeats] %>%
  stringr::str_remove(';.+')

# Show data missingness
proTumor_Klin <- proTissue_Klin[, colData(proTissue_Klin)$Condition %in% 'Tumor']
featAnnoTab <- tibble::as_tibble(rowData(proTissue_Klin), rownames = 'Feature') %>%
  dplyr::mutate(Feature = stringr::str_remove(Feature, ';.+'),
                Gene = stringr::str_remove(Gene, ';.+'))
featAvailTab <- summExp2df(proTumor_Klin, row_id = 'Feature', col_id = 'Sample') %>%
  dplyr::select(Feature, Value, Recurrence) %>%
  dplyr::mutate(Value = dplyr::case_when(!is.na(Value) ~ 'Observed',
                                         is.na(Value) ~ 'Missing'),
                Feature = stringr::str_remove(Feature, ';.+'),
                Recurrence = dplyr::case_when(Recurrence %in% 'Yes' ~ 'Recurrence',
                                              Recurrence %in% 'No' ~ 'Non-Recurrence')) %>%
  dplyr::filter(Feature %in% topImpoFeats) %>%
  dplyr::group_by(Recurrence, Feature) %>%
  dplyr::mutate(SmpNumber = length(Feature)) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(Recurrence, SmpNumber, Feature, Value) %>%
  dplyr::summarise(Count = dplyr::n()) %>%
  dplyr::ungroup() %>%
  dplyr::left_join(featAnnoTab, by = 'Feature') %>%
  dplyr::mutate(Gene = factor(Gene, levels = topImpoFeatAnnos),
                Recurrence = factor(Recurrence, levels = c('Recurrence', 'Non-Recurrence')),
                Value = factor(Value, levels = c('Missing', 'Observed')),
                Proportion = Count/SmpNumber)

ggplot(featAvailTab, aes(x=Gene, y=Proportion, fill=Value)) +
  geom_col(position = 'stack') +
  facet_wrap(vars(Recurrence), scales = 'free') +
  scale_fill_manual(values = c(Missing = 'grey', Observed = 'black')) +
  labs(title = 'Normal Tissue Proteomics') +
  theme_bw() +
  theme(axis.text.x = element_text(size = 18, face = 'bold', angle = 90, hjust = 1, vjust = 0.5),
        axis.text.y = element_text(size = 12, face = 'bold'),
        axis.title = element_text(size = 20, face = 'bold'),
        legend.title = element_blank(), legend.text = element_text(size = 18),
        strip.text = element_text(size = 18, face = 'bold'),
        plot.title = element_text(size = 22, face = 'bold', hjust = 0.5, vjust = 1.5))


# Show availability of top important features in other datasets
# Prepare feature space of other datasets
# Klingmüller Preprocessed Combined dataset
old_proTissue_Combined_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsn.rds')
featSpace_old_Prepro_Combined_Klin <- rownames(old_proTissue_Combined_Klin) %>%
  stringr::str_remove(';.+')
# Klingmüller Original Combined dataset
featSpace_old_Ori_Combined_Klin <- readr::read_tsv(paste0('./data/MethodDev/AG_Klingmueller/',
                                                          '20230309_103228_SmartCare_TumorFree_Tumor_Yang001_Report.tsv')) %>%
  dplyr::pull(PG.ProteinGroups) %>%
  stringr::str_remove(';.+')
# Krijgsveld Preprocessed Combined dataset
proTissue_Combined_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds')
featSpace_Prepro_Combined_Krij <- rownames(proTissue_Combined_Krij) %>%
  stringr::str_remove(';.+')
# Krijgsveld Original Discovery dataset
featSpace_Ori_Discovery_Krij <- readr::read_tsv(paste0('./data/Discovery/AG_Krijgsveld/',
                                                       '20230815_MarcS_Discovery_Tissue.pg_matrix.tsv')) %>%
  dplyr::pull(Protein.Group) %>%
  stringr::str_remove(';.+')
# Krijgsveld Original Method Development dataset
featSpace_Ori_MethodDev_Krij <- readr::read_delim(paste0('./data/MethodDev/AG_Krijgsveld/',
                                                         '20230726_Thorax_Method_Est_tissue.txt')) %>%
  dplyr::pull(Protein.Group) %>%
  stringr::str_remove(';.+')
# Collect all feature spaces into a list
featSpaceList <- list(Ori_Old_Klin = featSpace_old_Ori_Combined_Klin, Prepro_Old_Klin = featSpace_old_Prepro_Combined_Klin,
                      Ori_MDev_Krij = featSpace_Ori_MethodDev_Krij, Ori_Dis_Krij = featSpace_Ori_Discovery_Krij,
                      Prepro_Comb_Krij = featSpace_Prepro_Combined_Krij)
# Store feature availability in a matrix
featAvailMat <- matrix(nrow = 5, ncol = numTopImpoFeats,
                       dimnames = list(c('Ori_Old_Klin', 'Prepro_Old_Klin', 'Ori_MDev_Krij',
                                         'Ori_Dis_Krij', 'Prepro_Comb_Krij'), topImpoFeats))
for (i in seq_len(nrow(featAvailMat))) {
  for (j in seq_len(ncol(featAvailMat))) {
    if (colnames(featAvailMat)[j] %in% featSpaceList[[rownames(featAvailMat)[i]]]) {
      featAvailMat[i, j] <- 'Yes'
    } else {
      featAvailMat[i, j] <- 'No'
    }
  }
}
# Convert matrix to long data for visualization
featAvailMat <- tibble::as_tibble(featAvailMat, rownames = 'Dataset') %>%
  tidyr::pivot_longer(cols = -'Dataset', names_to = 'Feature', values_to = 'Captured') %>%
  dplyr::mutate(Dataset = factor(Dataset, levels = rev(names(featSpaceList))),
                Captured = factor(Captured, levels = c('Yes', 'No')),
                Feature = plyr::mapvalues(Feature, from = topImpoFeats, to = topImpoFeatAnnos),
                Feature = factor(Feature, levels = topImpoFeatAnnos))

ggplot(featAvailMat, aes(x=Feature, y=Dataset, fill=Captured)) +
  geom_tile() +
  scale_fill_manual(values = c('black', 'grey80')) +
  labs(title = 'Tumor Tissue Proteomics') +
  theme(axis.ticks = element_blank(), panel.grid = element_blank(), panel.background = element_blank(),
        axis.text.x = element_text(size = 18, face = 'bold', angle = 45, hjust = 1, vjust = 1),
        axis.text.y = element_text(size = 18, face = 'bold'),
        axis.title = element_text(size = 20, face = 'bold'),
        legend.title = element_text(size = 20), legend.text = element_text(size = 18),
        plot.title = element_text(size = 28, face = 'bold', hjust = 0.5, vjust = 1.5))
```

# NTP (AG Klin.)
Normal Tissue DIA Proteomics from AG Klingmüller

<font size='5'> **Stability selection** </font>

**Performance statistics of 1000 models**
```{r warning=T}
# Load preprocessed data
proTissue_Klin <- readRDS('./data/Discovery/AG_Klingmueller/proTissueVsn.rds')
proNormal_Klin <- proTissue_Klin[, colData(proTissue_Klin)$Condition %in% 'Normal']
# # Load SOA results
proNormalRes_Klin <- readRDS('./data/Discovery/AG_Klingmueller/soaRes/proNormalRes_SVsByProTissueInt.rds')
# # Prepare adjusted data
adjProNormal_Klin <- proNormal_Klin
assay(adjProNormal_Klin) <- proNormalRes_Klin$SOA.res$dataCorrect

# Load imputed adjusted data
impuAdjProNormal_Klin <- readRDS('./data/Discovery/AG_Klingmueller/impuData/impuAdjProNormalVsn_SVsByProTissueInt.rds')

# Train models iteratively
# Prepare significant feature list
sigFeatList <- proNormalRes_Klin$sig.feat.tab$Var1
# 'auc' is not applicable for CV when too few (< 10) observations per fold. Instead,
# 'deviance' will be used.
# lrRes <- iterLogisR(impuAdjProNormal_Klin, doImputation = F, doInitFeatSelection = T, sigFeatList = sigFeatList,
#                     doFeatClustering = T, iter = 1000, cvFold = 5, cvMeasure = 'auc')
# saveRDS(lrRes, './data/Discovery/logisR/lrResSOA_SVsByProTissueInt_impuAdjProNormal_0.9r_randSplit_1000_5fold_auc_lmin_Klin.rds')
# saveRDS(lrRes$impuSE, './data/Discovery/AG_Klingmueller/impuData/impuAdjProNormalVsn_SVsByProTissueInt.rds')
lrRes <- readRDS('./data/Discovery/logisR/lrResSOA_SVsByProTissueInt_impuAdjProNormal_0.9r_randSplit_1000_5fold_auc_lmin_Klin.rds')

# Report summarized scores
# Prepare predictions and ground truths where results from useless models are removed
uselessModels <- which(lrRes$lrRes$nNonZero == 0)
if (length(uselessModels) != 0) {
  pred <- lrRes$lrRes$y_pred[-uselessModels]
  truth <- lrRes$lrRes$y_truth[-uselessModels]
  auc_roc <- lrRes$lrRes$auc_roc[-uselessModels]
} else {
  pred <- lrRes$lrRes$y_pred
  truth <- lrRes$lrRes$y_truth
  auc_roc <- lrRes$lrRes$auc_roc
}
suppressWarnings(summarizePredPower(pred, truth, auc_roc))
```

**Top important features that are most frequently selected**
```{r}
# Visualize top important features
# Prepare feature annotation table
featAnno <- tibble::as_tibble(rowData(proTissue_Klin), rownames = 'Feature')
topImpoFeatViz <- vizTopImpoFeatsLR(lrRes, featAnno = featAnno, num_p1TopFeats = 30)
topImpoFeatViz$pick
```

**Full feature importance table**
```{r}
# Display full feature importance table
DT::datatable(topImpoFeatViz$fullImpoFeatTab)
```

<font size='5'> **Systematic training and evaluation** </font>

**Performance of models trained with different numbers of top important features**
```{r}
# Systematically train models with different numbers of top important features
# sysTrainModels <- doSysTrainLogisR(lrRes, impuAdjProNormal_Klin, max_numTopFeats = 30)
# saveRDS(sysTrainModels, './data/Discovery/logisR/sysTrain_lrResSOA_SVsByProTissueInt_impuAdjProNormal_randSplit_100_5fold_auc_lmin_Klin.rds')
sysTrainModels <- readRDS('./data/Discovery/logisR/sysTrain_lrResSOA_SVsByProTissueInt_impuAdjProNormal_randSplit_100_5fold_auc_lmin_Klin.rds')

# Visualize systematic training
vizSysTrainModelPerf(sysTrainModels, numTopFeats = 30)
# ggsave('./output/Junyan/stabSelec_Stats_proNormal.png', device = 'png', dpi = 400, height = 8, width = 10)
```

**Noninformative models that demonstrate effects of top important features**
```{r}
se <- impuAdjProNormal_Klin
# Train models without top important features, aka noninformative models
numTopImpoFeats <- c(2, 3, 4)
# for (num in numTopImpoFeats) {
#   topImpoFeats <- topImpoFeatViz$fullImpoFeatTab$Feature[1:num]
#   # Train lasso logistic regression model
#   reducedData <- lrRes$featClusters$reducedData
#   x <- t(reducedData[!rownames(reducedData) %in% topImpoFeats,])
#   y <- colData(se)$Recurrence
#   lrResNonInfo <- runLogisR(x, y, targetClass = 'Yes', iter = 100, regularized_method = 'lasso',
#                             cvFold = 5, cvMeasure = 'auc', used_lambda = 'lambda.min',
#                             trainSet_ratio = 0.8, split_method = 'random split',
#                             plot_ROC = F, save_model = T)
#   saveRDS(lrResNonInfo, paste0('./data/Discovery/logisR/noninfo/NonInfo_rmTop', num,
#                                '_lrResSOA_SVsByProTissueInt_impuAdjProNormal_randSplit_100_5fold_auc_lmin_Klin.rds'))
# }

# Display performance of noninformative models
meanAUC <- c()
lowerAUC <- c()
upperAUC <- c()
uselessModels <- which(lrRes$lrRes$nNonZero == 0)
if (length(uselessModels) != 0) {
  auc_roc <- lrRes$lrRes$auc_roc[-uselessModels]
} else {
  auc_roc <- lrRes$lrRes$auc_roc
}
meanAUC <- c(meanAUC, round(mean(auc_roc), 3))
ci <- calcCI(auc_roc, bootstrap = T)
lowerAUC <- c(lowerAUC, ci[1])
upperAUC <- c(upperAUC, ci[2])
for (num in numTopImpoFeats) {
  lrResNonInfo <- readRDS(paste0('./data/Discovery/logisR/noninfo/NonInfo_rmTop', num,
                                 '_lrResSOA_SVsByProTissueInt_impuAdjProNormal_randSplit_100_5fold_auc_lmin_Klin.rds'))
  uselessModels <- which(lrResNonInfo$nNonZero == 0)
  if (length(uselessModels) != 0) {
    auc_roc <- lrResNonInfo$auc_roc[-uselessModels]
  } else {
    auc_roc <- lrResNonInfo$auc_roc
  }
  meanAUC <- c(meanAUC, round(mean(auc_roc), 3))
  ci <- calcCI(auc_roc, bootstrap = T)
  lowerAUC <- c(lowerAUC, ci[1])
  upperAUC <- c(upperAUC, ci[2])
}
aucStatTab <- data.frame(FeatSpace = c('Complete', paste0('-Top ', numTopImpoFeats)),
                         MeanAUC = meanAUC, UpperAUC = upperAUC, LowerAUC = lowerAUC) %>%
  dplyr::mutate(FeatSpace = factor(FeatSpace, levels = FeatSpace))

ggplot(aucStatTab, aes(x=FeatSpace, y=MeanAUC)) +
  geom_bar(stat = 'identity', position = position_dodge(), alpha = 0.9) +
  geom_errorbar(aes(ymin=LowerAUC, ymax=UpperAUC), width = 0.4) +
  labs(x = 'Feature space used to train 100 models', y = 'Mean AUC-ROC',
       title = 'Noninformative models') +
  th
```

<font size='5'> **Investigation of core features** </font>

<font size='3'> Imputed data: </font>\
**Heatmap where samples are ordered by linear regression scores**
```{r}
coreFeatPlotsImpu <- vizCoreFeats(lrRes, sysTrainModels, impuAdjProNormal_Klin,
                                  heatmap_numCoreFeats = 4, colFeatAnno = 'Gene',
                                  scatter_twoCoreFeats = c('LILRB4', 'MORF4L2'),
                                  fontsize = 10, fontsize_col = 7, fontsize_row = 11,
                                  cellwidth = 6, cellheight = 18)
coreFeatPlotsImpu$heatmap
# ggsave('./output/Junyan/stabSelec_heatmap_proNormal_impu.png', device = 'png', dpi = 400, height = 8, width = 10)
```

**Scatterplot of two core features**
```{r}
coreFeatPlotsImpu$scatterplot
# ggsave('./output/Junyan/stabSelec_top2_proNormal_impu.png', device = 'png', dpi = 400, height = 8, width = 10)
```

<font size='3'> Original data: </font>\
**Heatmap**
```{r}
coreFeatPlots <- vizCoreFeats(lrRes, sysTrainModels, adjProNormal_Klin,
                              heatmap_numCoreFeats = 4, colFeatAnno = 'Gene',
                              scatter_twoCoreFeats = c('LILRB4', 'MORF4L2'), #'IGHV2-26', 'MAP3K7'
                              fontsize = 10, fontsize_col = 7, fontsize_row = 11,
                              cellwidth = 6, cellheight = 18)
coreFeatPlots$heatmap
# ggsave('./output/Junyan/stabSelec_heatmap_proNormal_ori.png', device = 'png', dpi = 400, height = 8, width = 10)
```

**Scatterplot**
```{r}
coreFeatPlots$scatterplot
# ggsave('./output/Junyan/stabSelec_top2_proNormal_ori.png', device = 'png', dpi = 400, height = 8, width = 10)
```

**Missingness and availability of core features**
```{r message=F}
numTopImpoFeats <- 4
# Prepare top important feature list
topImpoFeats <- topImpoFeatViz$fullImpoFeatTab$Feature[1:numTopImpoFeats] %>%
  stringr::str_remove(';.+')
topImpoFeatAnnos <- topImpoFeatViz$fullImpoFeatTab$Annotation[1:numTopImpoFeats] %>%
  stringr::str_remove(';.+')

# Show data missingness
proNormal_Klin <- proTissue_Klin[, colData(proTissue_Klin)$Condition %in% 'Normal']
featAnnoTab <- tibble::as_tibble(rowData(proTissue_Klin), rownames = 'Feature') %>%
  dplyr::mutate(Feature = stringr::str_remove(Feature, ';.+'),
                Gene = stringr::str_remove(Gene, ';.+'))
featAvailTab <- summExp2df(proNormal_Klin, row_id = 'Feature', col_id = 'Sample') %>%
  dplyr::select(Feature, Value, Recurrence) %>%
  dplyr::mutate(Value = dplyr::case_when(!is.na(Value) ~ 'Observed',
                                         is.na(Value) ~ 'Missing'),
                Feature = stringr::str_remove(Feature, ';.+'),
                Recurrence = dplyr::case_when(Recurrence %in% 'Yes' ~ 'Recurrence',
                                              Recurrence %in% 'No' ~ 'Non-Recurrence')) %>%
  dplyr::filter(Feature %in% topImpoFeats) %>%
  dplyr::group_by(Recurrence, Feature) %>%
  dplyr::mutate(SmpNumber = length(Feature)) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(Recurrence, SmpNumber, Feature, Value) %>%
  dplyr::summarise(Count = dplyr::n()) %>%
  dplyr::ungroup() %>%
  dplyr::left_join(featAnnoTab, by = 'Feature') %>%
  dplyr::mutate(Gene = factor(Gene, levels = topImpoFeatAnnos),
                Recurrence = factor(Recurrence, levels = c('Recurrence', 'Non-Recurrence')),
                Value = factor(Value, levels = c('Missing', 'Observed')),
                Proportion = Count/SmpNumber)

ggplot(featAvailTab, aes(x=Gene, y=Proportion, fill=Value)) +
  geom_col(position = 'stack') +
  facet_wrap(vars(Recurrence), scales = 'free') +
  scale_fill_manual(values = c(Missing = 'grey', Observed = 'black')) +
  labs(title = 'Normal Tissue Proteomics') +
  theme_bw() +
  theme(axis.text.x = element_text(size = 18, face = 'bold', angle = 90, hjust = 1, vjust = 0.5),
        axis.text.y = element_text(size = 12, face = 'bold'),
        axis.title = element_text(size = 20, face = 'bold'),
        legend.title = element_blank(), legend.text = element_text(size = 18),
        strip.text = element_text(size = 18, face = 'bold'),
        plot.title = element_text(size = 22, face = 'bold', hjust = 0.5, vjust = 1.5))


# Show availability of top important features in other datasets
# Prepare feature space of other datasets
# Klingmüller Preprocessed Combined dataset
old_proTissue_Combined_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsn.rds')
featSpace_old_Prepro_Combined_Klin <- rownames(old_proTissue_Combined_Klin) %>%
  stringr::str_remove(';.+')
# Klingmüller Original Combined dataset
featSpace_old_Ori_Combined_Klin <- readr::read_tsv(paste0('./data/MethodDev/AG_Klingmueller/',
                                                          '20230309_103228_SmartCare_TumorFree_Tumor_Yang001_Report.tsv')) %>%
  dplyr::pull(PG.ProteinGroups) %>%
  stringr::str_remove(';.+')
# Krijgsveld Preprocessed Combined dataset
proTissue_Combined_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds')
featSpace_Prepro_Combined_Krij <- rownames(proTissue_Combined_Krij) %>%
  stringr::str_remove(';.+')
# Krijgsveld Original Discovery dataset
featSpace_Ori_Discovery_Krij <- readr::read_tsv(paste0('./data/Discovery/AG_Krijgsveld/',
                                                       '20230815_MarcS_Discovery_Tissue.pg_matrix.tsv')) %>%
  dplyr::pull(Protein.Group) %>%
  stringr::str_remove(';.+')
# Krijgsveld Original Method Development dataset
featSpace_Ori_MethodDev_Krij <- readr::read_delim(paste0('./data/MethodDev/AG_Krijgsveld/',
                                                         '20230726_Thorax_Method_Est_tissue.txt')) %>%
  dplyr::pull(Protein.Group) %>%
  stringr::str_remove(';.+')
# Collect all feature spaces into a list
featSpaceList <- list(Ori_Old_Klin = featSpace_old_Ori_Combined_Klin, Prepro_Old_Klin = featSpace_old_Prepro_Combined_Klin,
                      Ori_MDev_Krij = featSpace_Ori_MethodDev_Krij, Ori_Dis_Krij = featSpace_Ori_Discovery_Krij,
                      Prepro_Comb_Krij = featSpace_Prepro_Combined_Krij)
# Store feature availability in a matrix
featAvailMat <- matrix(nrow = 5, ncol = numTopImpoFeats,
                       dimnames = list(c('Ori_Old_Klin', 'Prepro_Old_Klin', 'Ori_MDev_Krij',
                                         'Ori_Dis_Krij', 'Prepro_Comb_Krij'), topImpoFeats))
for (i in seq_len(nrow(featAvailMat))) {
  for (j in seq_len(ncol(featAvailMat))) {
    if (colnames(featAvailMat)[j] %in% featSpaceList[[rownames(featAvailMat)[i]]]) {
      featAvailMat[i, j] <- 'Yes'
    } else {
      featAvailMat[i, j] <- 'No'
    }
  }
}
# Convert matrix to long data for visualization
featAvailMat <- tibble::as_tibble(featAvailMat, rownames = 'Dataset') %>%
  tidyr::pivot_longer(cols = -'Dataset', names_to = 'Feature', values_to = 'Captured') %>%
  dplyr::mutate(Dataset = factor(Dataset, levels = rev(names(featSpaceList))),
                Captured = factor(Captured, levels = c('Yes', 'No')),
                Feature = plyr::mapvalues(Feature, from = topImpoFeats, to = topImpoFeatAnnos),
                Feature = factor(Feature, levels = topImpoFeatAnnos))

ggplot(featAvailMat, aes(x=Feature, y=Dataset, fill=Captured)) +
  geom_tile() +
  scale_fill_manual(values = c('black', 'grey80')) +
  labs(title = 'Normal Tissue Proteomics') +
  theme(axis.ticks = element_blank(), panel.grid = element_blank(), panel.background = element_blank(),
        axis.text.x = element_text(size = 18, face = 'bold', angle = 45, hjust = 1, vjust = 1),
        axis.text.y = element_text(size = 18, face = 'bold'),
        axis.title = element_text(size = 20, face = 'bold'),
        legend.title = element_text(size = 20), legend.text = element_text(size = 18),
        plot.title = element_text(size = 28, face = 'bold', hjust = 0.5, vjust = 1.5))
```




# TTP (AG Krij.)
Tumor Tissue DIA Proteomics from AG Krijgsveld

<font size='5'> **Stability selection** </font>

**Performance statistics of 1000 models**
```{r warning=T}
# Load preprocessed data
proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds')
proTumor_Krij <- proTissue_Krij[, colData(proTissue_Krij)$Condition %in% 'Tumor']
# Load SOA results
proTumorRes_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/soaRes/combined_proTumorRes_SVsByProTissueInt.rds')
# Prepare adjusted data
adjProTumor_Krij <- proTumor_Krij
assay(adjProTumor_Krij) <- proTumorRes_Krij$SOA.res$dataCorrect

# Load imputed adjusted data
impuAdjProTumor_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/impuData/combined_impuAdjProTumorVsn_SVsByProTissueInt.rds')

# Train models iteratively
# Prepare significant feature list
sigFeatList <- proTumorRes_Krij$sig.feat.tab$Var1
# lrRes <- iterLogisR(impuAdjProTumor_Krij, doImputation = F, doInitFeatSelection = T, sigFeatList = sigFeatList,
#                     doFeatClustering = T, iter = 1000, cvFold = 5, cvMeasure = 'auc')
# saveRDS(lrRes, './data/Discovery/logisR/lrResSOA_SVsByProTissueInt_impuAdjProTumor_0.9r_randSplit_1000_5fold_auc_lmin_Krij.rds')
# saveRDS(lrRes$impuSE, './data/Discovery/AG_Krijgsveld/impuData/combined_impuAdjProTumorVsn_SVsByProTissueInt.rds')
lrRes <- readRDS('./data/Discovery/logisR/lrResSOA_SVsByProTissueInt_impuAdjProTumor_0.9r_randSplit_1000_5fold_auc_lmin_Krij.rds')

# Report summarized scores
# Prepare predictions and ground truths where results from useless models are removed
uselessModels <- which(lrRes$lrRes$nNonZero == 0)
if (length(uselessModels) != 0) {
  pred <- lrRes$lrRes$y_pred[-uselessModels]
  truth <- lrRes$lrRes$y_truth[-uselessModels]
  auc_roc <- lrRes$lrRes$auc_roc[-uselessModels]
} else {
  pred <- lrRes$lrRes$y_pred
  truth <- lrRes$lrRes$y_truth
  auc_roc <- lrRes$lrRes$auc_roc
}
suppressWarnings(summarizePredPower(pred, truth, auc_roc))
```

**Top important features that are most frequently selected**
```{r}
# Visualize top important features
# Prepare feature annotation table
featAnno <- tibble::as_tibble(rowData(proTissue_Krij), rownames = 'Feature') %>%
  dplyr::select(Feature, Genes) %>%
  dplyr::rename(Gene = Genes)
topImpoFeatViz <- vizTopImpoFeatsLR(lrRes, featAnno = featAnno, num_p1TopFeats = 30)
topImpoFeatViz$pick
```

**Full feature importance table**
```{r}
# Display full feature importance table
DT::datatable(topImpoFeatViz$fullImpoFeatTab)
```

<font size='5'> **Systematic training and evaluation** </font>

**Performance of models trained with different numbers of top important features**
```{r}
# Systematically train models with different numbers of top important features
# sysTrainModels <- doSysTrainLogisR(lrRes, impuAdjProTumor_Krij, max_numTopFeats = 50)
# saveRDS(sysTrainModels, './data/Discovery/logisR/sysTrain_lrResSOA_SVsByProTissueInt_impuAdjProTumor_randSplit_100_5fold_auc_lmin_Krij.rds')
sysTrainModels <- readRDS('./data/Discovery/logisR/sysTrain_lrResSOA_SVsByProTissueInt_impuAdjProTumor_randSplit_100_5fold_auc_lmin_Krij.rds')

# Visualize systematic training
vizSysTrainModelPerf(sysTrainModels, numTopFeats = 30)
```
=> Top 8 important features are selected, resulting in model with AUC of 0.993 [0.95,1].\
=> Top 14 important features are selected, resulting in model with AUC of 0.995 [0.95,1].

**Noninformative models that demonstrate effects of top important features**
```{r}
se <- impuAdjProTumor_Krij
# Train models without top important features, aka noninformative models
numTopImpoFeats <- c(5, 10, 15, 20)
# for (num in numTopImpoFeats) {
#   topImpoFeats <- topImpoFeatViz$fullImpoFeatTab$Feature[1:num]
#   # Train lasso logistic regression model
#   reducedData <- lrRes$featClusters$reducedData
#   x <- t(reducedData[!rownames(reducedData) %in% topImpoFeats,])
#   y <- colData(se)$Recurrence
#   lrResNonInfo <- runLogisR(x, y, targetClass = 'Yes', iter = 100, regularized_method = 'lasso',
#                             cvFold = 5, cvMeasure = 'auc', used_lambda = 'lambda.min',
#                             trainSet_ratio = 0.8, split_method = 'random split',
#                             plot_ROC = F, save_model = T)
#   saveRDS(lrResNonInfo, paste0('./data/Discovery/logisR/noninfo/NonInfo_rmTop', num,
#                                '_lrResSOA_SVsByProTissueInt_impuAdjProTumor_randSplit_100_5fold_auc_lmin_Krij.rds'))
# }

# Display performance of noninformative models
meanAUC <- c()
lowerAUC <- c()
upperAUC <- c()
uselessModels <- which(lrRes$lrRes$nNonZero == 0)
if (length(uselessModels) != 0) {
  auc_roc <- lrRes$lrRes$auc_roc[-uselessModels]
} else {
  auc_roc <- lrRes$lrRes$auc_roc
}
meanAUC <- c(meanAUC, round(mean(auc_roc), 3))
ci <- calcCI(auc_roc, bootstrap = T)
lowerAUC <- c(lowerAUC, ci[1])
upperAUC <- c(upperAUC, ci[2])
for (num in numTopImpoFeats) {
  lrResNonInfo <- readRDS(paste0('./data/Discovery/logisR/noninfo/NonInfo_rmTop', num,
                                 '_lrResSOA_SVsByProTissueInt_impuAdjProTumor_randSplit_100_5fold_auc_lmin_Krij.rds'))
  uselessModels <- which(lrResNonInfo$nNonZero == 0)
  if (length(uselessModels) != 0) {
    auc_roc <- lrResNonInfo$auc_roc[-uselessModels]
  } else {
    auc_roc <- lrResNonInfo$auc_roc
  }
  meanAUC <- c(meanAUC, round(mean(auc_roc), 3))
  ci <- calcCI(auc_roc, bootstrap = T)
  lowerAUC <- c(lowerAUC, ci[1])
  upperAUC <- c(upperAUC, ci[2])
}
aucStatTab <- data.frame(FeatSpace = c('Complete', paste0('-Top ', numTopImpoFeats)),
                         MeanAUC = meanAUC, UpperAUC = upperAUC, LowerAUC = lowerAUC) %>%
  dplyr::mutate(FeatSpace = factor(FeatSpace, levels = FeatSpace))

ggplot(aucStatTab, aes(x=FeatSpace, y=MeanAUC)) +
  geom_bar(stat = 'identity', position = position_dodge(), alpha = 0.9) +
  geom_errorbar(aes(ymin=LowerAUC, ymax=UpperAUC), width = 0.4) +
  labs(x = 'Feature space used to train 100 models', y = 'Mean AUC-ROC',
       title = 'Noninformative models') +
  th
```

<font size='5'> **Investigation of core features** </font>

<font size='3'> Imputed data: </font>\
**Heatmap where samples are ordered by linear regression scores**
```{r}
coreFeatPlotsImpu <- vizCoreFeats(lrRes, sysTrainModels, impuAdjProTumor_Krij,
                                  heatmap_numCoreFeats = 8, colFeatAnno = 'Genes',
                                  scatter_twoCoreFeats = c('VSIG2', 'RETN'),
                                  fontsize = 10, fontsize_col = 7, fontsize_row = 11,
                                  cellwidth = 6, cellheight = 18)
coreFeatPlotsImpu$heatmap
```

**Scatterplot of two core features**
```{r}
coreFeatPlotsImpu$scatterplot
```

<font size='3'> Original data: </font>\
**Heatmap**
```{r}
coreFeatPlots <- vizCoreFeats(lrRes, sysTrainModels, adjProTumor_Krij,
                              heatmap_numCoreFeats = 8, colFeatAnno = 'Genes',
                              scatter_twoCoreFeats = c('RETN', 'CES1'),
                              fontsize = 10, fontsize_col = 7, fontsize_row = 11,
                              cellwidth = 6, cellheight = 18)
coreFeatPlots$heatmap
```

**Scatterplot**
```{r}
coreFeatPlots$scatterplot
```

**Missingness and availability of core features**
```{r message=F}
numTopImpoFeats <- 8
# Prepare top important feature list
topImpoFeats <- topImpoFeatViz$fullImpoFeatTab$Feature[1:numTopImpoFeats] %>%
  stringr::str_remove(';.+')
topImpoFeatAnnos <- topImpoFeatViz$fullImpoFeatTab$Annotation[1:numTopImpoFeats] %>%
  stringr::str_remove(';.+')

# Show data missingness
proTumor_Krij <- proTissue_Krij[, colData(proTissue_Krij)$Condition %in% 'Tumor']
featAnnoTab <- tibble::as_tibble(rowData(proTissue_Krij), rownames = 'Feature') %>%
  dplyr::select(Feature, Genes) %>%
  dplyr::rename(Gene = Genes) %>%
  dplyr::mutate(Feature = stringr::str_remove(Feature, ';.+'),
                Gene = stringr::str_remove(Gene, ';.+'))
featAvailTab <- summExp2df(proTumor_Krij, row_id = 'Feature', col_id = 'Sample') %>%
  dplyr::select(Feature, Value, Recurrence) %>%
  dplyr::mutate(Value = dplyr::case_when(!is.na(Value) ~ 'Observed',
                                         is.na(Value) ~ 'Missing'),
                Feature = stringr::str_remove(Feature, ';.+'),
                Recurrence = dplyr::case_when(Recurrence %in% 'Yes' ~ 'Recurrence',
                                              Recurrence %in% 'No' ~ 'Non-Recurrence')) %>%
  dplyr::filter(Feature %in% topImpoFeats) %>%
  dplyr::group_by(Recurrence, Feature) %>%
  dplyr::mutate(SmpNumber = length(Feature)) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(Recurrence, SmpNumber, Feature, Value) %>%
  dplyr::summarise(Count = dplyr::n()) %>%
  dplyr::ungroup() %>%
  dplyr::left_join(featAnnoTab, by = 'Feature') %>%
  dplyr::mutate(Gene = factor(Gene, levels = topImpoFeatAnnos),
                Recurrence = factor(Recurrence, levels = c('Recurrence', 'Non-Recurrence')),
                Value = factor(Value, levels = c('Missing', 'Observed')),
                Proportion = Count/SmpNumber)

ggplot(featAvailTab, aes(x=Gene, y=Proportion, fill=Value)) +
  geom_col(position = 'stack') +
  facet_wrap(vars(Recurrence), scales = 'free') +
  scale_fill_manual(values = c(Missing = 'grey', Observed = 'black')) +
  labs(title = 'Tumor Tissue Proteomics') +
  theme_bw() +
  theme(axis.text.x = element_text(size = 18, face = 'bold', angle = 90, hjust = 1, vjust = 0.5),
        axis.text.y = element_text(size = 12, face = 'bold'),
        axis.title = element_text(size = 20, face = 'bold'),
        legend.title = element_blank(), legend.text = element_text(size = 18),
        strip.text = element_text(size = 18, face = 'bold'),
        plot.title = element_text(size = 22, face = 'bold', hjust = 0.5, vjust = 1.5))


# Show availability of top important features in other datasets
# Prepare feature space of other datasets
# New Klingmüller Preprocessed Combined dataset
new_proTissue_Klin <- readRDS('./data/Discovery/AG_Klingmueller/proTissueVsn.rds')
featSpace_new_Prepro_Klin <- rownames(new_proTissue_Klin) %>%
  stringr::str_remove(';.+')
# New Klingmüller Original Combined dataset
proPhosTissue_MDevDis <- readRDS('./data/Discovery/AG_Klingmueller/COMP_2696.SMART_Lung_MethodDiscov_ProPhos.rds')
featSpace_new_Ori_Klin <- rowData(proPhosTissue_MDevDis@ExperimentList$Proteome)$UniprotID %>%
  stringr::str_remove(';.+')
# Klingmüller Preprocessed Combined dataset
old_proTissue_Combined_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsn.rds')
featSpace_old_Prepro_Combined_Klin <- rownames(old_proTissue_Combined_Klin) %>%
  stringr::str_remove(';.+')
# Klingmüller Original Combined dataset
featSpace_old_Ori_Combined_Klin <- readr::read_tsv(paste0('./data/MethodDev/AG_Klingmueller/',
                                                          '20230309_103228_SmartCare_TumorFree_Tumor_Yang001_Report.tsv')) %>%
  dplyr::pull(PG.ProteinGroups) %>%
  stringr::str_remove(';.+')
# Collect all feature spaces into a list
featSpaceList <- list(Ori_New_Klin = featSpace_new_Ori_Klin,
                      Prepro_New_Klin = featSpace_new_Prepro_Klin,
                      Ori_Old_Klin = featSpace_old_Ori_Combined_Klin,
                      Prepro_Old_Klin = featSpace_old_Prepro_Combined_Klin)
# Store feature availability in a matrix
featAvailMat <- matrix(nrow = 4, ncol = numTopImpoFeats,
                       dimnames = list(c('Ori_New_Klin', 'Prepro_New_Klin',
                                         'Ori_Old_Klin', 'Prepro_Old_Klin'), topImpoFeats))
for (i in seq_len(nrow(featAvailMat))) {
  for (j in seq_len(ncol(featAvailMat))) {
    if (colnames(featAvailMat)[j] %in% featSpaceList[[rownames(featAvailMat)[i]]]) {
      featAvailMat[i, j] <- 'Yes'
    } else {
      featAvailMat[i, j] <- 'No'
    }
  }
}
# Convert matrix to long data for visualization
featAvailMat <- tibble::as_tibble(featAvailMat, rownames = 'Dataset') %>%
  tidyr::pivot_longer(cols = -'Dataset', names_to = 'Feature', values_to = 'Captured') %>%
  dplyr::mutate(Dataset = factor(Dataset, levels = rev(names(featSpaceList))),
                Captured = factor(Captured, levels = c('Yes', 'No')),
                Feature = plyr::mapvalues(Feature, from = topImpoFeats, to = topImpoFeatAnnos),
                Feature = factor(Feature, levels = topImpoFeatAnnos))

ggplot(featAvailMat, aes(x=Feature, y=Dataset, fill=Captured)) +
  geom_tile() +
  scale_fill_manual(values = c('black', 'grey80')) +
  labs(title = 'Tumor Tissue Proteomics') +
  theme(axis.ticks = element_blank(), panel.grid = element_blank(), panel.background = element_blank(),
        axis.text.x = element_text(size = 18, face = 'bold', angle = 45, hjust = 1, vjust = 1),
        axis.text.y = element_text(size = 18, face = 'bold'),
        axis.title = element_text(size = 20, face = 'bold'),
        legend.title = element_text(size = 20), legend.text = element_text(size = 18),
        plot.title = element_text(size = 28, face = 'bold', hjust = 0.5, vjust = 1.5))
```

# NTP (AG Krij.)
Normal Tissue DIA Proteomics from AG Krijgsveld

<font size='5'> **Stability selection** </font>

**Performance statistics of 1000 models**
```{r warning=T}
# Load preprocessed data
proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds')
proNormal_Krij <- proTissue_Krij[, colData(proTissue_Krij)$Condition %in% 'Normal']
# Load SOA results
proNormalRes_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/soaRes/combined_proNormalRes_SVsByProTissueInt.rds')
# Prepare adjusted data
adjProNormal_Krij <- proNormal_Krij
assay(adjProNormal_Krij) <- proNormalRes_Krij$SOA.res$dataCorrect

# Load imputed adjusted data
impuAdjProNormal_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/impuData/combined_impuAdjProNormalVsn_SVsByProTissueInt.rds')

# Train models iteratively
# Prepare significant feature list
sigFeatList <- proNormalRes_Krij$sig.feat.tab$Var1
# lrRes <- iterLogisR(impuAdjProNormal_Krij, doImputation = F, doInitFeatSelection = T, sigFeatList = sigFeatList,
#                     doFeatClustering = T, iter = 1000, cvFold = 5, cvMeasure = 'auc')
# saveRDS(lrRes, './data/Discovery/logisR/lrResSOA_SVsByProTissueInt_impuAdjProNormal_0.9r_randSplit_1000_5fold_auc_lmin_Krij.rds')
# saveRDS(lrRes$impuSE, './data/Discovery/AG_Krijgsveld/impuData/combined_impuAdjProNormalVsn_SVsByProTissueInt.rds')
lrRes <- readRDS('./data/Discovery/logisR/lrResSOA_SVsByProTissueInt_impuAdjProNormal_0.9r_randSplit_1000_5fold_auc_lmin_Krij.rds')

# Report summarized scores
# Prepare predictions and ground truths where results from useless models are removed
uselessModels <- which(lrRes$lrRes$nNonZero == 0)
if (length(uselessModels) != 0) {
  pred <- lrRes$lrRes$y_pred[-uselessModels]
  truth <- lrRes$lrRes$y_truth[-uselessModels]
  auc_roc <- lrRes$lrRes$auc_roc[-uselessModels]
} else {
  pred <- lrRes$lrRes$y_pred
  truth <- lrRes$lrRes$y_truth
  auc_roc <- lrRes$lrRes$auc_roc
}
suppressWarnings(summarizePredPower(pred, truth, auc_roc))
```

**Top important features that are most frequently selected**
```{r}
# Visualize top important features
# Prepare feature annotation table
featAnno <- tibble::as_tibble(rowData(proTissue_Krij), rownames = 'Feature') %>%
  dplyr::select(Feature, Genes) %>%
  dplyr::rename(Gene = Genes)
# Prepare adjusted data
proTissueSVA <- doSVA(proTissue_Krij, wantedVar = c('Condition', 'Recurrence', 'Condition:Recurrence'))
adjProTissue <- proTissue_Krij
assay(adjProTissue) <- proTissueSVA$correctDatMat

topImpoFeatViz <- vizTopImpoFeatsLR(lrRes, featAnno = featAnno, num_p1TopFeats = 30,
                                    fullData = adjProTissue, trainData_smpType = 'Normal', num_p2TopFeats = 1)
topImpoFeatViz$pick
topImpoFeatViz$abun
```

**Full feature importance table**
```{r}
# Display full feature importance table
DT::datatable(topImpoFeatViz$fullImpoFeatTab)
```

<font size='5'> **Investigation of core features** </font>

**Missingness and availability of core features**
```{r}
oriFeat <- assay(adjProNormal_Krij)['O43731',]
impuFeat <- assay(impuAdjProNormal_Krij)['O43731',]
if (identical(names(oriFeat), names(impuFeat))) {
  mat <- matrix(c(oriFeat, impuFeat), nrow = 2, byrow = T,
                dimnames = list(c('KDELR3_Ori', 'KDELR3_Impu'), names(oriFeat)))
}

yPredLinkList <- lrRes$lrRes$y_pred_link
yPredLinkTab <- data.frame(Sample = names(yPredLinkList[[1]]),
                           Score = yPredLinkList[[1]],
                           row.names = NULL)
for (i in 2:length(yPredLinkList)) {
  yPredLink <- yPredLinkList[[i]]
  tmp_yPredLinkTab <- data.frame(Sample = names(yPredLinkList[[i]]),
                                 Score = yPredLinkList[[i]],
                                 row.names = NULL)
  yPredLinkTab <- dplyr::bind_rows(yPredLinkTab, tmp_yPredLinkTab)
}
yPredLinkTab <- dplyr::group_by(yPredLinkTab, Sample) %>%
  dplyr::summarise(Score = mean(Score)) %>%
  dplyr::arrange(dplyr::desc(Score))

smpAnnoTab <- tibble::as_tibble(colData(adjProNormal_Krij), rownames = 'Sample') %>%
    dplyr::select(Sample, Recurrence) %>%
    dplyr::left_join(yPredLinkTab, by = 'Sample') %>%
    tibble::column_to_rownames('Sample')

mat <- mat[, yPredLinkTab$Sample]
smpAnnoCols <- list(Recurrence = c(Yes = '#F8766D', No = '#00BFC4'))
pheatmap(mat, scale = 'row', annotation_col = smpAnnoTab, annotation_colors = smpAnnoCols,
         color = colorRampPalette(c('navy', 'white', 'red'))(100), cluster_rows = F, cluster_cols = F,
         fontsize = 10, fontsize_col = 7, fontsize_row = 11, cellwidth = 6, cellheight = 18)
```

```{r message=F}
numTopImpoFeats <- 1
# Prepare top important feature list
topImpoFeats <- topImpoFeatViz$fullImpoFeatTab$Feature[1:numTopImpoFeats] %>%
  stringr::str_remove(';.+')
topImpoFeatAnnos <- topImpoFeatViz$fullImpoFeatTab$Annotation[1:numTopImpoFeats] %>%
  stringr::str_remove(';.+')

# Show data missingness
proNormal_Krij <- proTissue_Krij[, colData(proTissue_Krij)$Condition %in% 'Normal']
featAnnoTab <- tibble::as_tibble(rowData(proTissue_Krij), rownames = 'Feature') %>%
  dplyr::select(Feature, Genes) %>%
  dplyr::rename(Gene = Genes) %>%
  dplyr::mutate(Feature = stringr::str_remove(Feature, ';.+'),
                Gene = stringr::str_remove(Gene, ';.+'))
featAvailTab <- summExp2df(proNormal_Krij, row_id = 'Feature', col_id = 'Sample') %>%
  dplyr::select(Feature, Value, Recurrence) %>%
  dplyr::mutate(Value = dplyr::case_when(!is.na(Value) ~ 'Observed',
                                         is.na(Value) ~ 'Missing'),
                Feature = stringr::str_remove(Feature, ';.+'),
                Recurrence = dplyr::case_when(Recurrence %in% 'Yes' ~ 'Recurrence',
                                              Recurrence %in% 'No' ~ 'Non-Recurrence')) %>%
  dplyr::filter(Feature %in% topImpoFeats) %>%
  dplyr::group_by(Recurrence, Feature) %>%
  dplyr::mutate(SmpNumber = length(Feature)) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(Recurrence, SmpNumber, Feature, Value) %>%
  dplyr::summarise(Count = dplyr::n()) %>%
  dplyr::ungroup() %>%
  dplyr::left_join(featAnnoTab, by = 'Feature') %>%
  dplyr::mutate(Gene = factor(Gene, levels = topImpoFeatAnnos),
                Recurrence = factor(Recurrence, levels = c('Recurrence', 'Non-Recurrence')),
                Value = factor(Value, levels = c('Missing', 'Observed')),
                Proportion = Count/SmpNumber)

ggplot(featAvailTab, aes(x=Gene, y=Proportion, fill=Value)) +
  geom_col(position = 'stack') +
  facet_wrap(vars(Recurrence), scales = 'free') +
  scale_fill_manual(values = c(Missing = 'grey', Observed = 'black')) +
  labs(title = 'Normal Tissue Proteomics') +
  theme_bw() +
  theme(axis.text.x = element_text(size = 18, face = 'bold'),
        axis.text.y = element_text(size = 12, face = 'bold'),
        axis.title = element_text(size = 20, face = 'bold'),
        legend.title = element_blank(), legend.text = element_text(size = 18),
        strip.text = element_text(size = 18, face = 'bold'),
        plot.title = element_text(size = 22, face = 'bold', hjust = 0.5, vjust = 1.5))


# Show availability of top important features in other datasets
# Prepare feature space of other datasets
# New Klingmüller Preprocessed Combined dataset
new_proTissue_Klin <- readRDS('./data/Discovery/AG_Klingmueller/proTissueVsn.rds')
featSpace_new_Prepro_Klin <- rownames(new_proTissue_Klin) %>%
  stringr::str_remove(';.+')
# New Klingmüller Original Combined dataset
proPhosTissue_MDevDis <- readRDS('./data/Discovery/AG_Klingmueller/COMP_2696.SMART_Lung_MethodDiscov_ProPhos.rds')
featSpace_new_Ori_Klin <- rowData(proPhosTissue_MDevDis@ExperimentList$Proteome)$UniprotID %>%
  stringr::str_remove(';.+')
# Klingmüller Preprocessed Combined dataset
old_proTissue_Combined_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsn.rds')
featSpace_old_Prepro_Combined_Klin <- rownames(old_proTissue_Combined_Klin) %>%
  stringr::str_remove(';.+')
# Klingmüller Original Combined dataset
featSpace_old_Ori_Combined_Klin <- readr::read_tsv(paste0('./data/MethodDev/AG_Klingmueller/',
                                                          '20230309_103228_SmartCare_TumorFree_Tumor_Yang001_Report.tsv')) %>%
  dplyr::pull(PG.ProteinGroups) %>%
  stringr::str_remove(';.+')
# Collect all feature spaces into a list
featSpaceList <- list(Ori_New_Klin = featSpace_new_Ori_Klin,
                      Prepro_New_Klin = featSpace_new_Prepro_Klin,
                      Ori_Old_Klin = featSpace_old_Ori_Combined_Klin,
                      Prepro_Old_Klin = featSpace_old_Prepro_Combined_Klin)
# Store feature availability in a matrix
featAvailMat <- matrix(nrow = 4, ncol = numTopImpoFeats,
                       dimnames = list(c('Ori_New_Klin', 'Prepro_New_Klin',
                                         'Ori_Old_Klin', 'Prepro_Old_Klin'), topImpoFeats))
for (i in seq_len(nrow(featAvailMat))) {
  for (j in seq_len(ncol(featAvailMat))) {
    if (colnames(featAvailMat)[j] %in% featSpaceList[[rownames(featAvailMat)[i]]]) {
      featAvailMat[i, j] <- 'Yes'
    } else {
      featAvailMat[i, j] <- 'No'
    }
  }
}
# Convert matrix to long data for visualization
featAvailMat <- tibble::as_tibble(featAvailMat, rownames = 'Dataset') %>%
  tidyr::pivot_longer(cols = -'Dataset', names_to = 'Feature', values_to = 'Captured') %>%
  dplyr::mutate(Dataset = factor(Dataset, levels = rev(names(featSpaceList))),
                Captured = factor(Captured, levels = c('Yes', 'No')),
                Feature = plyr::mapvalues(Feature, from = topImpoFeats, to = topImpoFeatAnnos),
                Feature = factor(Feature, levels = topImpoFeatAnnos))

ggplot(featAvailMat, aes(x=Feature, y=Dataset, fill=Captured)) +
  geom_tile() +
  scale_fill_manual(values = c('black', 'grey80')) +
  labs(title = 'Normal Tissue Proteomics') +
  theme(axis.ticks = element_blank(), panel.grid = element_blank(), panel.background = element_blank(),
        axis.text.x = element_text(size = 18, face = 'bold'),
        axis.text.y = element_text(size = 18, face = 'bold'),
        axis.title = element_text(size = 20, face = 'bold'),
        legend.title = element_text(size = 20), legend.text = element_text(size = 18),
        plot.title = element_text(size = 28, face = 'bold', hjust = 0.5, vjust = 1.5))
```
