---
title: 'Feature selection: Machine learning models'
author: "Qian-Wu Liao"
output:
  workflowr::wflow_html:
    toc: true
    toc_float: true
    code_folding: hide
---

<font size='4'> Description: Do feature selection using machine learning models
(e.g., random forest) trained on combined data of cohorts to identify potential
biomarkers for predicting lung cancer recurrence. </font>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 8, fig.width = 10, warning = F)
knitr::opts_knit$set(root.dir = '/Users/qianwu/Desktop/SMART-CARE_LungCancer')
```

Load libraries
```{r library loading, message = F}
library('spsUtil')
library('randomForest')
library('missForest')
library('caret')
library('glmnet')
library('SummarizedExperiment')
library('tidyverse')
# Load user-defined functions
source('./code/misc.R')
source('./code/ml_funcs.R')

# Set plot theme
th <- theme_bw(base_size = 15) +
  theme(axis.title = element_text(face = 'bold'),
        axis.text = element_text(face = 'bold'),
        axis.ticks = element_line(linewidth = 0.8),
        legend.text = element_text(size = 15))
```

# RF

## NTP (AG Krij.)
Normal Tissue DIA Proteomics from AG Krijgsveld

```{r message=F, eval=F}
# Prepare input data, which includes imputation, initial feature selection, correlated
# feature removal, and RF training. Take Normal Tissue DIA Proteomics as example

# Load preprocessed data
proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds')
proTissue <- proTissue_Krij
# Subset data
proNormal <- proTissue[, colData(proTissue)$Condition == 'Normal']
proTumor <- proTissue[, colData(proTissue)$Condition == 'Tumor']

# Normal tissues
# Impute missing values using missForest
# dat <- t(assay(proNormal))
# impuDat <- missForest(dat, maxiter = 10, verbose = T)$ximp %>%
#   t()
# assay(proNormal) <- impuDat
# saveRDS(proNormal, './data/Discovery/AG_Krijgsveld/impuByMF/combined_proNormalVsnImpu.rds')

# Load imputed data
proNormal_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/impuByMF/combined_proNormalVsnImpu.rds')
proNormal <- proNormal_Krij

# Prepare input data for RF
# Retrieve data matrix from SE object
datMat <- t(assay(proNormal)) %>%
  tibble::as_tibble(rownames = 'Sample')
# Retrieve patient recurrence annotations
recurAnno <- tibble::as_tibble(colData(proNormal), rownames = 'Sample') %>%
  dplyr::select(Sample, Recurrence)
# Include recurrence information into data matrix
allFeats <- dplyr::left_join(recurAnno, datMat, by = 'Sample') %>%
  tibble::column_to_rownames('Sample')
# Do initial feature selection
# Identify recurrence-related statistically significant features
soaRes <- doSOA(proNormal, meta_var = 'Recurrence', use_limma = T)
sigFeats <- dplyr::select(allFeats, c(Recurrence, soaRes$featSigAssoRes$Var1))
# Cluster highly correlated features and keep only one representative of each cluster
sigFeatClusters <- rmCorrFeats(t(sigFeats[, -1]), cutoff = 0.8)
uncorrSigFeats <- dplyr::select(sigFeats, c(Recurrence, rownames(sigFeatClusters$reducedData)))

# Run RF
# Set random seed for reproducible outcomes
set.seed(42)
x <- as.matrix(uncorrSigFeats[, -1])
y <- uncorrSigFeats[, 1]
spsUtil::quiet(
  rfResSig <- runRF(x, y, targetClass = 'Yes', iter = 1, split_method = 'random split', save_RF = T)
)

# Overview trained RF
# print(rfResSig$rfRes[[1]])
# Play around with thresholds
cutoff <- 0.5
expectedVals <- rfResSig$y_test[[1]]
predictedVals <- as.numeric(rfResSig$rfRes[[1]]$test$votes[, '1'] > cutoff) %>%
  factor()
caret::confusionMatrix(data = predictedVals, reference = expectedVals, positive = '1')
```

**Average model performance (Uncorrelated significant features, iter = 1000)**
```{r}
# Load RF training results
rfResSig <- readRDS('./data/Discovery/rf/rfResSig_proNormal_randSplit_1000_Krij.rds')
# Display mean AUC score
auc <- rfResSig$rfRes$auc_roc
# Compute 95% confidence interval of AUC scores
ci <- calcCI(auc, bootstrap = T)
cat('Mean AUC-ROC score: ', round(mean(auc), 3), '\n',
    '95% CI: [', ci[1], ', ', ci[2], ']', sep = '')
```

**Top important features with highest averaged ranks**
```{r}
# Process feature importance scores
# Rank feature importance scores for all iterations
rankedFeatImpoTab <- apply(rfResSig$rfRes$MDG, 2, function(featImpo) {
  orderIdx <- order(featImpo, decreasing = T)
  featImpoRanks <- seq_along(orderIdx)
  names(featImpoRanks) <- orderIdx
  featImpo[as.numeric(names(featImpoRanks))] <- featImpoRanks
  return(featImpo)
})
# Compute mean, SD, and CI of each feature across iterations
meanFeatImpo <- apply(rankedFeatImpoTab, 1, mean)
sdFeatImpo <- apply(rankedFeatImpoTab, 1, sd)
lowerCIFeatImpo <- apply(rankedFeatImpoTab, 1, function(featImpoRanks) {
  calcCI(featImpoRanks, bootstrap = T)[1]
})
upperCIFeatImpo <- apply(rankedFeatImpoTab, 1, function(featImpoRanks) {
  calcCI(featImpoRanks, bootstrap = T)[2]
})
# Collect all needed information into a table
Combined_proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds')
featAnnoTab <- tibble::as_tibble(rowData(Combined_proTissue_Krij), rownames = 'Feature') %>%
  dplyr::select(Feature, Genes) %>%
  dplyr::rename(Gene = Genes)
meanRankedFeatImpoTab <- data.frame(Feature = names(meanFeatImpo),
                                    ImpoMean = meanFeatImpo,
                                    ImpoSD = sdFeatImpo,
                                    ImpoLowerCI = lowerCIFeatImpo,
                                    ImpoUpperCI = upperCIFeatImpo) %>%
  dplyr::arrange(ImpoMean) %>%
  dplyr::left_join(featAnnoTab, by = 'Feature') %>%
  dplyr::mutate(Gene = stringr::str_remove(Gene, ';.+'),
                Feature = stringr::str_remove(Feature, ';.+'))

# Visualize top important features
num_impoFeats <- 15
topFeatImpoTab <- meanRankedFeatImpoTab[seq_len(num_impoFeats),]
ggplot(topFeatImpoTab, aes(x=ImpoMean, y=factor(Gene, levels = rev(Gene)))) +
  geom_errorbar(aes(xmin=ImpoLowerCI, xmax=ImpoUpperCI)) +
  geom_point(size = 6) +
  labs(x = 'Averaged Rank of Feature Importance', y = 'Feature') +
  th
# ggsave('./output/Discovery/group_meeting/combined_rf_featImpo_proNormal_Krij.png',
#        device = 'png', dpi = 400, height = 8, width = 10)
```

```{r eval=F}
# Visualize abundances of top important features
# Prepare data matrix and metadata including both Tumor and Normal samples
Combined_proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds')
datMat <- assay(Combined_proTissue_Krij)
rownames(datMat) <- stringr::str_remove(rownames(datMat), ';.+')
smpAnnoTab <- tibble::as_tibble(colData(Combined_proTissue_Krij), rownames = 'Sample')

num_impoFeats <- 6
gpLevel = c('Yes_Normal', 'No_Normal', 'Yes_Tumor', 'No_Tumor')
condiCol = c('firebrick', 'grey50') #c(Normal, Tumor)
# Extract top important features and prepare needed information
topImpoFeats <- meanRankedFeatImpoTab[1:num_impoFeats,] %>%
  dplyr::select(Feature, Gene) %>%
  dplyr::mutate(newFeat = paste0(Feature, ' (', Gene, ')'),
                newFeat = factor(newFeat, levels = unique(newFeat)))
topImpoFeatDat <- tibble::as_tibble(datMat[topImpoFeats$Feature,], rownames = 'Feature') %>%
  tidyr::pivot_longer(cols = -'Feature', names_to = 'Sample', values_to = 'Abundance') %>%
  dplyr::left_join(topImpoFeats, by = 'Feature') %>%
  dplyr::left_join(smpAnnoTab, by = 'Sample') %>%
  dplyr::mutate(Recur_Condi = paste0(Recurrence, '_', Condition),
                Recur_Condi = factor(Recur_Condi, levels = gpLevel))
ggplot(topImpoFeatDat, aes(x=Recur_Condi, y=Abundance, col=Condition, fill=Recurrence)) +
  geom_boxplot(alpha = 1, outlier.shape = NA, linewidth = 1) +
  geom_jitter(position = position_jitter(0.2), size = 2, show.legend = F) +
  ggpubr::stat_compare_means(method = 't.test', paired = F, method.args = list(var.equal = T),
                             comparisons = list(c('Yes_Normal', 'No_Normal'), c('Yes_Tumor', 'No_Tumor')),
                             label = 'p.signif', tip.length = 0.015, bracket.size = 0.7, size = 4) +
  labs(x = 'Recurrence') +
  scale_color_manual(values = condiCol) +
  scale_fill_manual(values=c('#00BFC4', '#F8766D')) +
  facet_wrap(vars(newFeat), scales = 'free') +
  th +
  theme(strip.text = element_text(size = 13, face = 'bold'),
             axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

## TTP (AG Krij.)
Tumor Tissue DIA Proteomics from AG Krijgsveld

**Average model performance (Uncorrelated significant features, iter = 1000)**
```{r}
# Load RF training results
rfRes <- readRDS('./data/Discovery/rf/rfResSig_proTumor_randSplit_1000_Krij.rds')
# Display mean AUC score
auc <- rfRes$rfRes$auc
# Compute 95% confidence interval of AUC scores
ci <- calcCI(auc, bootstrap = T)
cat('Mean AUC-ROC score: ', round(mean(auc), 3), '\n',
    '95% CI: [', ci[1], ', ', ci[2], ']', sep = '')
```

**Top important features with highest averaged ranks**
```{r}
# Process feature importance scores
# Rank feature importance scores for all iterations
rankedFeatImpoTab <- apply(rfRes$rfRes$MDG, 2, function(featImpo) {
  orderIdx <- order(featImpo, decreasing = T)
  featImpoRanks <- seq_along(orderIdx)
  names(featImpoRanks) <- orderIdx
  featImpo[as.numeric(names(featImpoRanks))] <- featImpoRanks
  return(featImpo)
})
# Compute mean, SD, and CI of each feature across iterations
meanFeatImpo <- apply(rankedFeatImpoTab, 1, mean)
sdFeatImpo <- apply(rankedFeatImpoTab, 1, sd)
lowerCIFeatImpo <- apply(rankedFeatImpoTab, 1, function(featImpoRanks) {
  calcCI(featImpoRanks, bootstrap = T)[1]
})
upperCIFeatImpo <- apply(rankedFeatImpoTab, 1, function(featImpoRanks) {
  calcCI(featImpoRanks, bootstrap = T)[2]
})
# Collect all needed information into a table
Combined_proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds')
featAnnoTab <- tibble::as_tibble(rowData(Combined_proTissue_Krij), rownames = 'Feature') %>%
  dplyr::select(Feature, Genes) %>%
  dplyr::rename(Gene = Genes)
meanRankedFeatImpoTab <- data.frame(Feature = names(meanFeatImpo),
                                    ImpoMean = meanFeatImpo,
                                    ImpoSD = sdFeatImpo,
                                    ImpoLowerCI = lowerCIFeatImpo,
                                    ImpoUpperCI = upperCIFeatImpo) %>%
  dplyr::arrange(ImpoMean) %>%
  dplyr::left_join(featAnnoTab, by = 'Feature') %>%
  dplyr::mutate(Gene = stringr::str_remove(Gene, ';.+'),
                Feature = stringr::str_remove(Feature, ';.+'))

# Visualize top important features
num_impoFeats <- 15
topFeatImpoTab <- meanRankedFeatImpoTab[seq_len(num_impoFeats),]
ggplot(topFeatImpoTab, aes(x=ImpoMean, y=factor(Gene, levels = rev(Gene)))) +
  geom_errorbar(aes(xmin=ImpoLowerCI, xmax=ImpoUpperCI)) +
  geom_point(size = 6) +
  labs(x = 'Averaged Rank of Feature Importance', y = 'Feature') +
  th
```

## NTP (AG Klin.)
Normal Tissue DIA Proteomics from AG Klingmüller

**Average model performance (Uncorrelated significant features, iter = 1000)**
```{r}
# Load RF training results
rfRes <- readRDS('./data/Discovery/rf/rfResSig_proNormal_randSplit_1000_Klin.rds')
# Display mean AUC score
auc <- rfRes$rfRes$auc
# Compute 95% confidence interval of AUC scores
ci <- calcCI(auc, bootstrap = T)
cat('Mean AUC-ROC score: ', round(mean(auc), 3), '\n',
    '95% CI: [', ci[1], ', ', ci[2], ']', sep = '')
```

**Top important features with highest averaged ranks**
```{r}
# Process feature importance scores
# Rank feature importance scores for all iterations
rankedFeatImpoTab <- apply(rfRes$rfRes$MDG, 2, function(featImpo) {
  orderIdx <- order(featImpo, decreasing = T)
  featImpoRanks <- seq_along(orderIdx)
  names(featImpoRanks) <- orderIdx
  featImpo[as.numeric(names(featImpoRanks))] <- featImpoRanks
  return(featImpo)
})
# Compute mean, SD, and CI of each feature across iterations
meanFeatImpo <- apply(rankedFeatImpoTab, 1, mean)
sdFeatImpo <- apply(rankedFeatImpoTab, 1, sd)
lowerCIFeatImpo <- apply(rankedFeatImpoTab, 1, function(featImpoRanks) {
  calcCI(featImpoRanks, bootstrap = T)[1]
})
upperCIFeatImpo <- apply(rankedFeatImpoTab, 1, function(featImpoRanks) {
  calcCI(featImpoRanks, bootstrap = T)[2]
})
# Collect all needed information into a table
Combined_proTissue_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsnBC.rds')
featAnnoTab <- tibble::as_tibble(rowData(Combined_proTissue_Klin), rownames = 'Feature') %>%
  dplyr::rename(Gene = PG.Genes)
meanRankedFeatImpoTab <- data.frame(Feature = names(meanFeatImpo),
                                    ImpoMean = meanFeatImpo,
                                    ImpoSD = sdFeatImpo,
                                    ImpoLowerCI = lowerCIFeatImpo,
                                    ImpoUpperCI = upperCIFeatImpo) %>%
  dplyr::arrange(ImpoMean) %>%
  dplyr::left_join(featAnnoTab, by = 'Feature') %>%
  dplyr::mutate(Gene = stringr::str_remove(Gene, ';.+'),
                Feature = stringr::str_remove(Feature, ';.+'))

# Visualize top important features
num_impoFeats <- 15
topFeatImpoTab <- meanRankedFeatImpoTab[seq_len(num_impoFeats),]
ggplot(topFeatImpoTab, aes(x=ImpoMean, y=factor(Feature, levels = rev(Feature)))) +
  geom_errorbar(aes(xmin=ImpoLowerCI, xmax=ImpoUpperCI)) +
  geom_point(size = 6) +
  labs(x = 'Averaged Rank of Feature Importance', y = 'Feature') +
  th
```

## TTP (AG Klin.)
Tumor Tissue DIA Proteomics from AG Klingmüller

**Average model performance (Uncorrelated significant features, iter = 1000)**
```{r}
# Load RF training results
rfRes <- readRDS('./data/Discovery/rf/rfResSig_proTumor_randSplit_1000_Klin.rds')
# Display mean AUC score
auc <- rfRes$rfRes$auc
# Compute 95% confidence interval of AUC scores
ci <- calcCI(auc, bootstrap = T)
cat('Mean AUC-ROC score: ', round(mean(auc), 3), '\n',
    '95% CI: [', ci[1], ', ', ci[2], ']', sep = '')
```

**Top important features with highest averaged ranks**
```{r}
# Process feature importance scores
# Rank feature importance scores for all iterations
rankedFeatImpoTab <- apply(rfRes$rfRes$MDG, 2, function(featImpo) {
  orderIdx <- order(featImpo, decreasing = T)
  featImpoRanks <- seq_along(orderIdx)
  names(featImpoRanks) <- orderIdx
  featImpo[as.numeric(names(featImpoRanks))] <- featImpoRanks
  return(featImpo)
})
# Compute mean, SD, and CI of each feature across iterations
meanFeatImpo <- apply(rankedFeatImpoTab, 1, mean)
sdFeatImpo <- apply(rankedFeatImpoTab, 1, sd)
lowerCIFeatImpo <- apply(rankedFeatImpoTab, 1, function(featImpoRanks) {
  calcCI(featImpoRanks, bootstrap = T)[1]
})
upperCIFeatImpo <- apply(rankedFeatImpoTab, 1, function(featImpoRanks) {
  calcCI(featImpoRanks, bootstrap = T)[2]
})
# Collect all needed information into a table
Combined_proTissue_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsnBC.rds')
featAnnoTab <- tibble::as_tibble(rowData(Combined_proTissue_Klin), rownames = 'Feature') %>%
  dplyr::rename(Gene = PG.Genes)
meanRankedFeatImpoTab <- data.frame(Feature = names(meanFeatImpo),
                                    ImpoMean = meanFeatImpo,
                                    ImpoSD = sdFeatImpo,
                                    ImpoLowerCI = lowerCIFeatImpo,
                                    ImpoUpperCI = upperCIFeatImpo) %>%
  dplyr::arrange(ImpoMean) %>%
  dplyr::left_join(featAnnoTab, by = 'Feature') %>%
  dplyr::mutate(Gene = stringr::str_remove(Gene, ';.+'),
                Feature = stringr::str_remove(Feature, ';.+'))

# Visualize top important features
num_impoFeats <- 15
topFeatImpoTab <- meanRankedFeatImpoTab[seq_len(num_impoFeats),]
ggplot(topFeatImpoTab, aes(x=ImpoMean, y=factor(Gene, levels = rev(Gene)))) +
  geom_errorbar(aes(xmin=ImpoLowerCI, xmax=ImpoUpperCI)) +
  geom_point(size = 6) +
  labs(x = 'Averaged Rank of Feature Importance', y = 'Feature') +
  th
```
