---
title: 'MOFA: Downstream analysis'
author: "Qian-Wu Liao"
output:
  workflowr::wflow_html:
    toc: true
    toc_float: true
    code_folding: hide
---

<font size='4'> Description: Do multi-omics integration using MOFA. </font>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 8, fig.width = 10, warning = F)
knitr::opts_knit$set(root.dir = '/Users/qianwu/Desktop/SMART-CARE_LungCancer')
```

Load libraries
```{r library loading, message = F}
library('MOFA2')
library('reticulate')
# use_python('/Users/qianwu/opt/anaconda3/bin/python')
library('AnnotationDbi')
library('org.Hs.eg.db')
# This annotation object is accessed using AnnotationDbi:select()
hs <- org.Hs.eg.db
library('msigdbr')
library('clusterProfiler')
library('ggrepel')
library('SummarizedExperiment')
library('tidyverse')
# Load user-defined functions
source('./code/misc.R')
source('./code/mofa_funcs.R')

# Set plot theme
th <- theme_bw(base_size = 15) +
  theme(axis.title = element_text(face = 'bold'),
        axis.text = element_text(face = 'bold'),
        axis.ticks = element_line(linewidth = 0.8),
        legend.text = element_text(size = 15))
```

```{r message=F}
# Prepare data tables to train MOFA models
# Load normalized datasets
comb_proTissue_Klin <- readRDS('./data/MethodDev/AG_Klingmueller/proTissueVsnBC.rds')
comb_proNormal_Klin <- comb_proTissue_Klin[, colData(comb_proTissue_Klin)$Condition == 'Normal']

comb_proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds')

dis_lipTissue_Hopf <- readRDS('./data/Discovery/AG_Hopf/lipTissueVsn_WBC25.rds')
dis_lipNormal_Hopf <- dis_lipTissue_Hopf[, colData(dis_lipTissue_Hopf)$Condition == 'Normal']

mDev_lipTissue_Hopf <- readRDS('./data/MethodDev/AG_Hopf/lipTissueVsn.rds')
mDev_lipNormal_Hopf <- mDev_lipTissue_Hopf[, colData(mDev_lipTissue_Hopf)$Condition == 'Normal']

dis_lipBase_Hopf <- readRDS('./data/Discovery/AG_Hopf/lipBaseVsn_B1WBC25.rds')

mDev_lipPlasma_Hopf <- readRDS('./data/MethodDev/AG_Hopf/lipPlasmaVsn.rds')
mDev_lipBase_Hopf <- mDev_lipPlasma_Hopf[, colData(mDev_lipPlasma_Hopf)$TimePoint == 'Baseline']

# Sanity check dimensionality of each dataset (Feature x Sample)
dataList <- list(Comb_Tissue_Proteomics_Klin = comb_proTissue_Klin,
                 Comb_Tissue_Proteomics_Krij = comb_proTissue_Krij,
                 Dis_Tissue_Lipidomics_Hopf = dis_lipTissue_Hopf,
                 MDev_Tissue_Lipidomics_Hopf = mDev_lipTissue_Hopf,
                 Dis_Baseline_Lipidomics_Hopf = dis_lipBase_Hopf,
                 MDev_Plasma_Lipidomics_Hopf = mDev_lipPlasma_Hopf)
# lapply(dataList, dim)

# Prepare significant features for feature selection
# soaRes <- readRDS('./data/Discovery/potential_proteins/Combined_proNormalRes_Klin_SV1.rds')
# sigFeats_comb_proNormal_Klin <- soaRes$sig.feat.tab$Var1
# soaRes <- doSOA(dis_lipNormal_Hopf, meta_var = 'Recurrence', use_proDA = T)
# sigFeats_dis_lipNormal_Hopf <- soaRes$featSigAssoRes$Var1
# soaRes <- doSOA(mDev_lipNormal_Hopf, meta_var = 'Recurrence', use_proDA = T)
# sigFeats_mDev_lipNormal_Hopf <- soaRes$featSigAssoRes$Var1
# soaRes <- doSOA(dis_lipBase_Hopf, meta_var = 'Recurrence', use_proDA = T)
# sigFeats_dis_lipBase_Hopf <- soaRes$featSigAssoRes$Var1
# soaRes <- doSOA(mDev_lipBase_Hopf, meta_var = 'Recurrence', use_proDA = T)
# sigFeats_mDev_lipBase_Hopf <- soaRes$featSigAssoRes$Var1

# Convert SE objects to long data for creating MOFA object through 'create_mofa_from_df'
# to include metadata
# comb_proNormalTab_Klin <- mofa_summExp2df(comb_proNormal_Klin, smp_type = 'Tissue',
#                                           data_acqui = 'Combined DIA', data_modal = 'Proteomics') %>%
#   dplyr::filter(feature %in% sigFeats_comb_proNormal_Klin)
# dis_lipNormalTab_Hopf <- mofa_summExp2df(dis_lipNormal_Hopf, smp_type = 'Tissue',
#                                          data_acqui = 'Discovery Untargeted', data_modal = 'Lipidomics') %>%
#   dplyr::filter(feature %in% sigFeats_dis_lipNormal_Hopf)
# mDev_lipNormalTab_Hopf <- mofa_summExp2df(mDev_lipNormal_Hopf, smp_type = 'Tissue',
#                                           data_acqui = 'MethodDev Untargeted', data_modal = 'Lipidomics') %>%
#   dplyr::filter(feature %in% sigFeats_mDev_lipNormal_Hopf)
# dis_lipBaseTab_Hopf <- mofa_summExp2df(dis_lipBase_Hopf, smp_type = 'Plasma',
#                                        data_acqui = ' Discovery Untargeted', data_modal = 'Lipidomics') %>%
#   dplyr::filter(feature %in% sigFeats_dis_lipBase_Hopf)
# mDev_lipBaseTab_Hopf <- mofa_summExp2df(mDev_lipBase_Hopf, smp_type = 'Plasma',
#                                         data_acqui = 'MethodDev Untargeted', data_modal = 'Lipidomics') %>%
#   dplyr::filter(feature %in% sigFeats_mDev_lipBase_Hopf)

# Prepare sample spaces from different datasets to remove unwanted samples later 
smps_CombNTP <- stringr::str_replace(colnames(comb_proNormal_Klin), '_NG', '_B')
smps_DisNTL <- stringr::str_replace(colnames(dis_lipNormal_Hopf), '_NG', '_B')
smps_MDevNTL <- stringr::str_replace(colnames(mDev_lipNormal_Hopf), '_NG', '_B')
smps_DisBPL <- stringr::str_replace(colnames(dis_lipBase_Hopf), '_V1', '_B')
smps_MDevBPL <- colnames(mDev_lipBase_Hopf)
```

# Comb. DIA NTP (Klin.) + All Unt. NTL
Combined DIA Normal Tissue Proteomics (MethodDev + Discovery) from AG Klingmüller +
Discovery and MethodDev Untargeted Normal Tissue Lipidomics\
(since features defined by RT/CCS/MZ have difficulty being combined together from different cohorts)\
**Note that only statistically significant recurrence-related features were used to train a MOFA model**

```{r}
# Train MOFA model
# Remove unwanted samples
# smpViewCounts <- table(c(smps_CombNTP, smps_DisNTL, smps_MDevNTL))
# rmSmps <- names(smpViewCounts)[smpViewCounts == 1]
# tmp_dis_lipNormalTab_Hopf <- dplyr::filter(dis_lipNormalTab_Hopf, !sample %in% rmSmps)
# tmp_mDev_lipNormalTab_Hopf <- dplyr::filter(mDev_lipNormalTab_Hopf, !sample %in% rmSmps)
# 
# mofaObject <- trainMOFA(list(comb_proNormalTab_Klin, tmp_dis_lipNormalTab_Hopf,
#                              tmp_mDev_lipNormalTab_Hopf),
#                         view_data = T, train_mofa = T, num_factors = 10,
#                         save_path = './data/Discovery/mofa/sigFeats_CombDIANTPKlin_DisMDevUntNTL')

# Load trained model
mofaObject <- readRDS('./data/Discovery/mofa/sigFeats_CombDIANTPKlin_DisMDevUntNTL.rds')
```

Overview model\
Note that samples in only one view were removed from model training
```{r}
# Overview model
MOFA2::plot_data_overview(mofaObject)
```

Quantify amount of variance explained by each factor in each data modality
```{r}
# Do sanity check for factor correlations
MOFA2::plot_factor_cor(mofaObject)

# Perform variance dedecomposition analysis (coefficient of determination)
MOFA2::plot_variance_explained(mofaObject, max_r2 = 15)
```

Display and visualize significant recurrence-related factors
```{r}
# Display significant associations between learned factors (Var1) and cancer recurrence
# (Var2) and visualize significant recurrence-related factors
sigFactor <- mofa_vizSigFactor(mofaObject, smpGroup = 'Baseline Group', show_res = F)
sigFactor$sigAssoRes

facTab <- dplyr::mutate(sigFactor$tab4Plot,
                        Recurrence = factor(Recurrence, levels = c('Yes', 'No')),
                        Label = dplyr::case_when(Recurrence %in% 'Yes' & Factor1 > 0 ~ sample,
                                                 Recurrence %in% 'No' & Factor1 < 0 ~ sample))

ggplot(facTab, aes(x=Recurrence, y=Factor1, col=Recurrence, fill=Recurrence, label=Label)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(position = position_jitter(0.2), size = 4, show.legend = F) +
  geom_text_repel(show.legend = F) +
  ggpubr::stat_compare_means(method = 't.test', paired = F, method.args = list(var.equal = T),
                             size = 9, show.legend = F, hjust = 0.4, vjust = 1) +
  scale_color_manual(values=c('#F8766D', '#00BFC4')) +
  scale_fill_manual(values=c('#F8766D', '#00BFC4')) +
  th +
  theme(axis.title = element_text(size = 28),
        axis.text = element_text(size = 16),
        legend.title = element_text(size = 26), legend.text = element_text(size = 24))


ggplot(facTab, aes(x=Recurrence, y=Factor4, col=Recurrence, fill=Recurrence)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(position = position_jitter(0.2), size = 4, show.legend = F) +
  ggpubr::stat_compare_means(method = 't.test', paired = F, method.args = list(var.equal = T),
                             show.legend = F) +
  scale_color_manual(values=c('#F8766D', '#00BFC4')) +
  scale_fill_manual(values=c('#F8766D', '#00BFC4')) +
  th

ggplot(sigFactor$tab4Plot, aes(x=Factor1, y=Factor4, col=Recurrence)) +
  geom_point(size = 4) +
  scale_color_manual(values=c('#00BFC4', '#F8766D')) +
  th

# Compute ROC-AUC score
response <- ifelse(facTab$Recurrence == 'Yes', yes = '1', no = '0') %>%
  factor(levels = c('1', '0'))
predictor <- facTab$Factor1
rocRes <- suppressMessages(
  pROC::roc(response = response, predictor = predictor, plot = T,
            legacy.axes = T, print.auc = T, print.auc.x = 0.4,
            xlab = 'False positive rate', ylab = 'True positive rate',
            main = 'ROC Curve for Factor1',
            cex.lab = 1.2, cex.main = 1.1, col = '#377eb8', lwd = 4,
            direction = '<')
)
```
=> Factor1 is highly correlated with PC1 from Klin. NTP and explains variance also
in MDev. NTL, but in Dis. NTL, which implies that significant features captured
from MDev. and Dis. NTL may provide distinct information probably due to biological
variations in different patient cohorts. In addition, Dis. Tissue Lipidomics got
worse sample or data quality than MDev., which is shown by their metadata-assisted
quality controls (Tumor vs Normal).

Compare Factor1 with recurrence-related PC1 learned from Combined Klin. Normal Tissue Proteomics
```{r}
# Visualize correlation between significant factor (MOFA) and significant PC (PCA)
# Prepare significant factor table
facTab <- sigFactor$tab4Plot %>%
  dplyr::select(sample, Factor1, Recurrence) %>%
  dplyr::mutate(sample = stringr::str_replace(sample, '_B', '_NG'))
# Prepare significant PC table
proNormalRes_Klin <- readRDS('./data/Discovery/potential_proteins/Combined_proNormalRes_Klin_SV1.rds')
# Show significant PCs
# proNormalRes_Klin$sig.pc.tab
pcTab <- proNormalRes_Klin$SOA.res$pcTab %>%
  dplyr::select(Sample, `PC1 (11.1%)`) %>%
  dplyr::rename(sample = Sample)
# Combine all information
combinedTab <- dplyr::left_join(facTab, pcTab, by = 'sample')
# Make a plot and compute correlation
ggplot(combinedTab, aes(x=Factor1, y=`PC1 (11.1%)`)) +
  geom_point(aes(col=Recurrence), size = 4) +
  scale_color_manual(values=c('#00BFC4', '#F8766D')) +
  scale_fill_manual(values=c('#00BFC4', '#F8766D')) +
  ggpubr::stat_cor(aes(label=after_stat(r.label)), method = 'pearson', size = 7) +
  th
```

Visualize enrichment analysis results
```{r message=F}
# Prepare ranked feature list
# Tidy up weight matrix and feature metadata to reduce information loss
proGeneTab <- rowData(comb_proNormal_Klin) %>%
  tibble::as_tibble(rownames = 'Proteins') %>%
  tibble::column_to_rownames('Proteins')
tidyWeiMat <- mofa_keepUniFeats(mofaObject, factor = 1, view = 'Combined DIA Normal Proteomics',
                                feat_anno = proGeneTab, to_genes = T)
# Rank feature weights
rankedGeneList <- mofa_rankFeatList(tidyWeiMat, factor = 1)
# Make Recurrence group positive weights
rankedGeneList <- sort(-rankedGeneList, decreasing = T)

# Do enrichment analysis - Hallmark gene sets (H)
# Retrieve annotated gene sets
msigTabH <- msigdbr(species = 'Homo sapiens', category = 'H', subcategory = NULL) %>%
  dplyr::select(gs_name, human_gene_symbol)
# Run GSEA
gseaResH <- clusterProfiler::GSEA(geneList = rankedGeneList, TERM2GENE = msigTabH,
                                  minGSSize = 10, maxGSSize = 500,
                                  pvalueCutoff = 0.05, pAdjustMethod = 'BH',
                                  by = 'fgsea', eps = 0)

# Plot enrichment analysis results
gseaRes4Plot <- gseaResH@result %>%
  dplyr::select(Description, setSize, NES, pvalue, p.adjust, leading_edge) %>%
  dplyr::arrange(p.adjust) %>%
  dplyr::slice_head(n = 8) %>%
  dplyr::mutate(leading_edge = stringr::str_extract(leading_edge, '\\d\\d'),
                leading_edge = as.numeric(leading_edge),
                Recurrence = ifelse(test = NES > 0, yes = 'Yes', no = 'No')) %>%
  dplyr::rename(gene_ratio = leading_edge)
midGrad <- mean(gseaRes4Plot$p.adjust)

ggplot(gseaRes4Plot, aes(x=NES, y=reorder(Description, NES), fill=p.adjust, col=Recurrence)) +
  geom_bar(stat = 'identity', width = 0.6, size = 2) +
  scale_fill_gradient2(low='#5B1C00', high='#FFFCA5', mid = '#E86900', midpoint = midGrad, name = 'Adjusted pVal') +
  scale_color_manual(values = c(Yes = 'red', No = 'darkblue')) +
  labs(x = 'Normalized Enrichment Score', y = 'Pathway', title = 'MSigDB:H - Factor1') +
  theme_minimal() +
  theme(axis.title = element_text(size = 16, face = 'bold'),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 10, face = 'bold'),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12))

# => Factor4: HALLMARK_MYC_TARGETS_V1
```

```{r include=F, eval=F}
# Compare factors learned from different MOFA models
mofaObj1 <- readRDS('./data/Discovery/mofa/sigFeats_CombDIANTPKlin_DisMDevUntNTL.rds')
mofaObj2 <- readRDS('./data/Discovery/mofa/X/sigFeats_CombDIANTPKlin_DisUntNTL.rds')
mofaObj3 <- readRDS('./data/Discovery/mofa/X/sigFeats_CombDIANTPKlin_MDevUntNTL.rds')

MOFA2::plot_variance_explained(mofaObj2, max_r2 = 15) +
  labs(title = 'DisUntNTL: Factor1&2 Sig.')
MOFA2::plot_variance_explained(mofaObj3, max_r2 = 15) +
  labs(title = 'MDevUntNTL: Factor1&3 Sig.')

MOFA2::compare_factors(list(DisMDevUntNTL = mofaObj1, DisUntNTL = mofaObj2, MDevUntNTL = mofaObj3))
```

# Comb. DIA NTP (Klin.) + All Unt. BPL
Combined DIA Normal Tissue Proteomics (MethodDev + Discovery) from AG Klingmüller +
Discovery and MethodDev Untargeted Baseline Plasma Lipidomics
(since features defined by RT/CCS/MZ have difficulty being combined together from different cohorts)\
**Note that only statistically significant recurrence-related features were used to train a MOFA model**

```{r}
# Train MOFA model
# Remove unwanted samples
# smpViewCounts <- table(c(smps_CombNTP, smps_DisBPL, smps_MDevBPL))
# rmSmps <- names(smpViewCounts)[smpViewCounts == 1]
# tmp_dis_lipBaseTab_Hopf <- dplyr::filter(dis_lipBaseTab_Hopf, !sample %in% rmSmps)
# tmp_mDev_lipBaseTab_Hopf <- dplyr::filter(mDev_lipBaseTab_Hopf, !sample %in% rmSmps)
# 
# mofaObject <- trainMOFA(list(comb_proNormalTab_Klin, tmp_dis_lipBaseTab_Hopf,
#                              tmp_mDev_lipBaseTab_Hopf),
#                         view_data = T, train_mofa = T, num_factors = 10,
#                         save_path = './data/Discovery/mofa/sigFeats_CombDIANTPKlin_DisMDevUntBPL')

# Load trained model
mofaObject <- readRDS('./data/Discovery/mofa/sigFeats_CombDIANTPKlin_DisMDevUntBPL.rds')
# views_names(mofaObject) <- c('Dis. Unt. BPL', 'Comb. DIA NTP', 'MDev. Unt. BPL')
```

Overview model\
Note that samples in only one view were removed from model training
```{r}
# Overview model
MOFA2::plot_data_overview(mofaObject) #+
  # theme(text = element_text(size = 28))
# ggsave('./output/Discovery/2nd_TAC/mofa_data_UntBPL_combined_proNormal_Klin.png',
#        device = 'png', dpi = 400, height = 8, width = 11)
```

Quantify amount of variance explained by each factor in each data modality
```{r}
# Do sanity check for factor correlations
# MOFA2::plot_factor_cor(mofaObject)

# Perform variance dedecomposition analysis (coefficient of determination)
MOFA2::plot_variance_explained(mofaObject, max_r2 = 15) #+
# theme(axis.text.x = element_text(size = 22), axis.text.y = element_text(size =20),
#       legend.title = element_text(size = 22), legend.text = element_text(size = 20))
# ggsave('./output/Discovery/2nd_TAC/mofa_varExplain_UntBPL_combined_proNormal_Klin.png',
       # device = 'png', dpi = 400, height = 8, width = 11)
```

Display and visualize significant recurrence-related factors
```{r}
# Display significant associations between learned factors (Var1) and cancer recurrence
# (Var2) and visualize significant recurrence-related factors
sigFactor <- mofa_vizSigFactor(mofaObject, smpGroup = 'Baseline Group', show_res = F)
sigFactor$sigAssoRes

facTab <- dplyr::mutate(sigFactor$tab4Plot,
                        Recurrence = factor(Recurrence, levels = c('Yes', 'No')),
                        Label = dplyr::case_when(Recurrence %in% 'Yes' & Factor1 > 0 ~ sample,
                                                 Recurrence %in% 'No' & Factor1 < 0 ~ sample))

ggplot(facTab, aes(x=Recurrence, y=Factor1, col=Recurrence, fill=Recurrence, label=Label)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(position = position_jitter(0.2), size = 4, show.legend = F) +
  geom_text_repel(show.legend = F) +
  ggpubr::stat_compare_means(method = 't.test', paired = F, method.args = list(var.equal = T),
                             size = 9, show.legend = F, hjust = 0.4, vjust = 1) +
  scale_color_manual(values=c('#F8766D', '#00BFC4')) +
  scale_fill_manual(values=c('#F8766D', '#00BFC4')) +
  th +
  theme(axis.title = element_text(size = 30), axis.text = element_text(size = 20),
        legend.title = element_text(size = 28), legend.text = element_text(size = 26))
# ggsave('./output/Discovery/2nd_TAC/mofa_fac1_UntBPL_combined_proNormal_Klin.png',
#        device = 'png', dpi = 400, height = 8, width = 11)

# Compute ROC-AUC score
response <- ifelse(facTab$Recurrence == 'Yes', yes = '1', no = '0') %>%
  factor(levels = c('1', '0'))
predictor <- facTab$Factor1
rocRes <- pROC::roc(response = response, predictor = predictor, plot = T,
                    legacy.axes = T, print.auc = T, print.auc.x = 0.4,
                    xlab = substitute(paste(bold('False positive rate'))),
                    ylab = substitute(paste(bold('True positive rate'))),
                    main = 'ROC Curve for Factor1',
                    cex.lab = 1.8, cex.axis = 1.2, cex.main = 1.8, col = '#377eb8',
                    lwd = 6, direction = '<', quiet = T)
# pROC::ci.auc(rocRes)
```

Compare Factor1 with recurrence-related PC1 learned from Combined Klin. Normal Tissue
Proteomics and another Factor1 from model built on Klin. NTP and Unt. NTL
```{r}
# Visualize correlation between significant factors (MOFA) and significant PC (PCA)
# Prepare significant factor table
facTab <- sigFactor$tab4Plot %>%
  dplyr::select(sample, Factor1, Recurrence) %>%
  dplyr::mutate(sample = stringr::str_replace(sample, '_B', '_NG'))
# Prepare significant PC table
proNormalRes_Klin <- readRDS('./data/Discovery/potential_proteins/Combined_proNormalRes_Klin_SV1.rds')
# Show significant PCs
# proNormalRes_Klin$sig.pc.tab
pcTab <- proNormalRes_Klin$SOA.res$pcTab %>%
  dplyr::select(Sample, `PC1 (11.1%)`) %>%
  dplyr::rename(sample = Sample)
# Combine all information
combinedTab <- dplyr::left_join(facTab, pcTab, by = 'sample') %>%
  dplyr::mutate(Recurrence = factor(Recurrence, levels = c('Yes', 'No')))
# Make a plot and compute correlation
ggplot(combinedTab, aes(x=Factor1, y=`PC1 (11.1%)`)) +
  geom_point(aes(col=Recurrence), size = 4) +
  scale_color_manual(values=c('#F8766D', '#00BFC4')) +
  scale_fill_manual(values=c('#F8766D', '#00BFC4')) +
  ggpubr::stat_cor(aes(label=after_stat(r.label)), method = 'pearson', size = 9) +
  th +
  theme(axis.title = element_text(size = 30), axis.text = element_text(size = 20),
        legend.title = element_text(size = 28), legend.text = element_text(size = 26))
# ggsave('./output/Discovery/2nd_TAC/mofa_fac1_vs_pc1_UntBPL_combined_proNormal_Klin.png',
#        device = 'png', dpi = 400, height = 8, width = 11)

# Prepare significant factor tables
facTab <- sigFactor$tab4Plot %>%
  dplyr::select(sample, Factor1, Recurrence) %>%
  dplyr::rename(Factor1_BPL = Factor1)
mofaObj <- readRDS('./data/Discovery/mofa/sigFeats_CombDIANTPKlin_DisMDevUntNTL.rds')
facTab2 <- get_factors(mofaObj)[['Baseline Group']] %>%
  tibble::as_tibble(rownames = 'sample') %>%
  dplyr::select(sample, Factor1) %>%
  dplyr::rename(Factor1_NTL = Factor1)
# Combine all information and remove unwanted samples
combinedTab <- dplyr::left_join(facTab, facTab2, by = 'sample')
# Make a plot and compute correlation
ggplot(combinedTab, aes(x=Factor1_BPL, y=Factor1_NTL)) +
  geom_point(aes(col=Recurrence), size = 4) +
  scale_color_manual(values=c('#00BFC4', '#F8766D')) +
  scale_fill_manual(values=c('#00BFC4', '#F8766D')) +
  ggpubr::stat_cor(aes(label=after_stat(r.label)), method = 'pearson', size = 7) +
  th
```

Visualize enrichment analysis results
```{r message=F}
# Prepare ranked feature list
# Tidy up weight matrix and feature metadata to reduce information loss
proGeneTab <- rowData(comb_proNormal_Klin) %>%
  tibble::as_tibble(rownames = 'Proteins') %>%
  tibble::column_to_rownames('Proteins')
tidyWeiMat <- mofa_keepUniFeats(mofaObject, factor = 1, view = 'Combined DIA Normal Proteomics',
                                feat_anno = proGeneTab, to_genes = T)
# Rank feature weights
rankedGeneList <- mofa_rankFeatList(tidyWeiMat, factor = 1)
# Make Recurrence group positive weights
rankedGeneList <- sort(-rankedGeneList, decreasing = T)

# Do enrichment analysis - Hallmark gene sets (H)
# Retrieve annotated gene sets
msigTabH <- msigdbr(species = 'Homo sapiens', category = 'H', subcategory = NULL) %>%
  dplyr::select(gs_name, human_gene_symbol)
# Run GSEA
gseaResH <- clusterProfiler::GSEA(geneList = rankedGeneList, TERM2GENE = msigTabH,
                                  minGSSize = 10, maxGSSize = 500,
                                  pvalueCutoff = 0.05, pAdjustMethod = 'BH',
                                  by = 'fgsea', eps = 0)

# Plot enrichment analysis results
gseaRes4Plot <- gseaResH@result %>%
  dplyr::select(Description, setSize, NES, pvalue, p.adjust, leading_edge) %>%
  dplyr::arrange(p.adjust) %>%
  dplyr::slice_head(n = 8) %>%
  dplyr::mutate(leading_edge = stringr::str_extract(leading_edge, '\\d\\d'),
                leading_edge = as.numeric(leading_edge),
                Recurrence = ifelse(test = NES > 0, yes = 'Yes', no = 'No')) %>%
  dplyr::rename(gene_ratio = leading_edge)
midGrad <- mean(gseaRes4Plot$p.adjust)

ggplot(gseaRes4Plot, aes(x=NES, y=reorder(Description, NES), fill=p.adjust, col=Recurrence)) +
  geom_bar(stat = 'identity', width = 0.6, size = 2) +
  scale_fill_gradient2(low='#5B1C00', high='#FFFCA5', mid = '#E86900', midpoint = midGrad, name = 'Adjusted pVal') +
  scale_color_manual(values = c(Yes = 'red', No = 'darkblue')) +
  labs(x = 'Normalized Enrichment Score', y = 'Pathway', title = 'MSigDB:H - Factor1') +
  theme_minimal() +
  theme(axis.title = element_text(size = 16, face = 'bold'),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 10, face = 'bold'),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12))
```

```{r include=F, eval=F}
# Compare factors learned from different MOFA models
mofaObj1 <- readRDS('./data/Discovery/mofa/sigFeats_CombDIANTPKlin_DisMDevUntNTL.rds')
mofaObj2 <- readRDS('./data/Discovery/mofa/sigFeats_CombDIANTPKlin_DisMDevUntBPL.rds')

MOFA2::compare_factors(list(NTL = mofaObj1, BPL = mofaObj2))
```
