---
title: "Single-omics analysis: Annotated Untargeted Lipidomics (Discovery + MethodDev)"
author: "Qian-Wu Liao"
output:
  workflowr::wflow_html:
    toc: yes
    toc_float: yes
    code_folding: hide
---

<font size='4'> Description: Conduct univariate (t-test) and multivariate (PCA)
analyses on Method Development, Discovery, and combined annotated Untargeted Tissue
Lipidomics from AG Hopf to have overview of data and take initial look at data power
in terms of predicting patient cancer recurrence. </font>

**Metadata variables**\
Patients who got Tissue samples (n = 85):\
Recurrence -> Cancer recurrences, Yes:No = 29:56\
Gender -> Female:Male = 36:49\
Age -> Diagnosis ages ranging from 51 to 84\
Smoking -> Smoker:Ex-smoker:Non-Smoker = 18:56:11\
Stage -> Pathological stages, IB:IIA:IIB = 50:17:18\
Adjuvant -> Adjuvant chemotherapy, True:False = 26:59\
Samples:\
Condition -> Tumor:Normal = 81:84

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 8, fig.width = 10, warning = F, message = F)
knitr::opts_knit$set(root.dir = '/Users/qianwu/Desktop/SMART-CARE_LungCancer')
```

Load libraries
```{r library loading, message = F}
library('pcaMethods')
library('sva')
library('limma')
library('proDA')
library('pheatmap')
library('ggplotify')
library('ggrepel')
library('ggvenn')
library('AnnotationDbi')
library('org.Hs.eg.db')
# This annotation object is accessed using AnnotationDbi:select()
hs <- org.Hs.eg.db
library('msigdbr')
library('clusterProfiler')
library(spsUtil)
library('SummarizedExperiment')
library('tidyverse')
# Load user-defined functions
source('./code/analysis_pipeline.R')
source('./code/misc.R')

# Set plot theme
th <- theme_bw(base_size = 15) +
    theme(axis.title = element_text(face = 'bold'),
          axis.text = element_text(face = 'bold'),
          axis.ticks = element_line(linewidth = 0.8),
          legend.text = element_text(size = 15))
```

Display associations between cancer recurrence and the other patient metadata variables
to identify potential predictors (or confounders)
```{r}
# Test associations between cancer recurrence and the other patient metadata variables
# Prepare patient metadata
lipTissue <- readRDS(paste0('./data/Discovery/AG_Hopf/partial_annotated/',
                            'Comb_Prepro_lipTissueVsn_Anno_MDevB123.rds'))
patientMetadat <- tibble::as_tibble(colData(lipTissue)) %>%
  dplyr::filter(!duplicated(Patient)) %>%
  dplyr::select(Patient, Recurrence, Gender, Age, Smoking, Stage, Adjuvant)
patientRecur <- dplyr::select(patientMetadat, Patient, Recurrence)
patientMetadat <- dplyr::select(patientMetadat, -Recurrence)
testAsso(patientMetadat, patientRecur, cmn_col = 'Patient') %>%
  dplyr::select(-Stat)
```

# Tissue Lipidomics

## Sample Condition
Tumor vs Normal

### Discovery cohort

<font size='4'> **Preprocessed data without correction** </font>
```{r}
# Load preprocessed data
lipTissueVsn_Dis <- readRDS(paste0('./data/Discovery/AG_Hopf/partial_annotated/',
                                   'Dis_lipTissueVsn_Anno.rds'))
# Subset data by annotated features
lipTissueVsn_Dis <- lipTissueVsn_Dis[!is.na(rowData(lipTissueVsn_Dis)$Lipid),]
# Perform single-omics analysis
# lipTissueRes_Dis <- doSOA(lipTissueVsn_Dis, meta_var = 'Condition', pca_method = 'ppca',
#                           num_PCs = 20, use_proDA = T, level_metaVar = c('Tumor', 'Normal'))
# saveRDS(lipTissueRes_Dis, paste0('./data/Discovery/AG_Hopf/partial_annotated/soaRes/',
#                                  'lipTissueRes_Condi_OnlyAnno_Dis.rds'))
lipTissueRes_Dis <- readRDS(paste0('./data/Discovery/AG_Hopf/partial_annotated/soaRes/',
                                   'lipTissueRes_Condi_OnlyAnno_Dis.rds'))
# Show significant associations between PCs and sample conditions
lipTissueRes_Dis$pcSigAssoRes
# Visualize PCs
pcTab <- lipTissueRes_Dis$pcTab
ggplot(pcTab, aes(x=`PC1 (37.9%)`, y=`PC2 (16%)`, col=Condition)) +
  geom_point(size = 2.5) +
  scale_color_brewer(palette = 'Dark2') +
  th
```

<font size='4'> **Preprocessed data corrected for SVs (unwanted variance)** </font>
```{r}
# Do SVA
lipTissueVsnSVA_Dis <- doSVA(lipTissueVsn_Dis, wantedVar = 'Condition', numSV_method = 'be',
                             asso_metaVar = c('Condition', 'Recurrence', 'Gender', 'Age',
                                              'Smoking', 'Stage', 'Adjuvant', 'TumorPurity'))
# Show significant associations between SVs and sample metadata variables
lipTissueVsnSVA_Dis$sigAssoTab
lipTissueVsnSVs_Dis <- lipTissueVsnSVA_Dis$summExp_SVs

# Perform single-omics analysis
# lipTissueResSVs_Dis <- doSOA(lipTissueVsnSVs_Dis, meta_var = 'Condition', pca_method = 'ppca',
#                              num_PCs = 20, use_proDA = T, level_metaVar = c('Tumor', 'Normal'),
#                              unwantVar = grep('^SV\\d+$', colnames(colData(lipTissueVsnSVs_Dis)), value = T))
# saveRDS(lipTissueResSVs_Dis, paste0('./data/Discovery/AG_Hopf/partial_annotated/soaRes/',
#                                     'lipTissueResSVs_Condi_OnlyAnno_Dis.rds'))
lipTissueResSVs_Dis <- readRDS(paste0('./data/Discovery/AG_Hopf/partial_annotated/soaRes/',
                                      'lipTissueResSVs_Condi_OnlyAnno_Dis.rds'))
# Show significant associations between PCs and sample conditions
lipTissueResSVs_Dis$pcSigAssoRes
# Visualize PCs
pcTab <- lipTissueResSVs_Dis$pcTab %>%
  dplyr::mutate(Label = dplyr::case_when(Sample %in% c('7EAOX7_TU', 'I0HVOL_TU',
                                                       '1NL5BV_TU', 'NG1IWD_TU',
                                                       '1QFJ7P_TU', '125CII_TU',
                                                       '13XAAY_TU', 'PDIR55_NG') ~ Sample))
ggplot(pcTab, aes(x=Condition, y=`PC1 (43.4%)`, col=Condition, fill=Condition, label=Label)) +
  geom_boxplot(alpha = 0.7, outliers = F) +
  geom_jitter(position = position_jitter(0.3), size = 2, show.legend = F) +
  geom_text_repel(show.legend = F) +
  scale_color_brewer(palette = 'Dark2') +
  scale_fill_brewer(palette = 'Dark2') +
  th

# Show tumor purity of misclassified samples
tibble::as_tibble(colData(lipTissueVsn_Dis), rownames = 'Sample') %>%
  dplyr::select(Sample, TumorPurity) %>%
  dplyr::filter(Sample %in% c('7EAOX7_TU', 'I0HVOL_TU', '1NL5BV_TU', 'NG1IWD_TU',
                              '1QFJ7P_TU', '125CII_TU', '13XAAY_TU'))
```

<font size='4'> **Volcano plot of SV-corrected data** </font>
```{r}
# Compute log2(FC) using median normalized data
# Load preprocessed data
lipTissueMedi_Dis <- readRDS(paste0('./data/Discovery/AG_Hopf/partial_annotated/',
                                    'Dis_lipTissueMedi_Anno.rds'))
# Subset data by annotated features
lipTissueMedi_Dis <- lipTissueMedi_Dis[!is.na(rowData(lipTissueMedi_Dis)$Lipid),]
# Do SVA
lipTissueMediSVA_Dis <- doSVA(lipTissueMedi_Dis, wantedVar = 'Condition', numSV_method = 'be',
                              asso_metaVar = c('Condition', 'Recurrence', 'Gender', 'Age',
                                               'Smoking', 'Stage', 'Adjuvant', 'TumorPurity'))
# Show significant associations between SVs and sample metadata variables
lipTissueMediSVA_Dis$sigAssoTab
lipTissueMediSVs_Dis <- lipTissueMediSVA_Dis$summExp_SVs

# Compute log2(FC) using proDA
# Prepare data matrix and sample metadata table for creating model matrices
datMat <- assay(lipTissueMediSVs_Dis)
metadatTab <- tibble::as_tibble(colData(lipTissueMediSVs_Dis), rownames = 'Sample') %>%
  # Make comparison Tumor vs Normal
  dplyr::mutate(Condition = factor(Condition, levels = c('Normal', 'Tumor')))
formu <- paste0('~Condition+', paste0(grep('^SV\\d+$', colnames(colData(lipTissueMediSVs_Dis)),
                                           value = T), collapse = '+')) %>%
  as.formula()
design <- model.matrix(formu, data = metadatTab)

# fit <- proDA(datMat, design = design)
# diffTerm <- proDA::result_names(fit)[2]
# lipTissueLog2FC_SVs_Dis <- proDA::test_diff(fit, contrast = diffTerm)
# saveRDS(lipTissueLog2FC_SVs_Dis, paste0('./data/Discovery/AG_Hopf/partial_annotated/soaRes/',
#                                         'lipTissueLog2FC_SVs_Condi_OnlyAnno_Dis.rds'))
lipTissueLog2FC_SVs_Dis <- readRDS(paste0('./data/Discovery/AG_Hopf/partial_annotated/soaRes/',
                                          'lipTissueLog2FC_SVs_Condi_OnlyAnno_Dis.rds'))


####
# featInterest <- c('PC 34:1', 'PC 34:2', 'PC 32:0', 'PC 38:5', 'PC 36:4', 'PC 36:2',
#                   'PC 38:4', 'PC 36:1', 'PC 30:0', 'PC 34:0', 'PC O-36:5', 'PC 32:1')
# c('PC 32:1', 'PC 36:1', 'PC O-36:5', 'PC 19:2_19:2', 'PC 15:0_15:0', 'PC 17:0_17:0',
#   'PC 16:0_20:4', 'PC 8:0_30:4', 'PC 18:1_18:1', 'PC 16:1_22:4', 'PC 18:0_20:4',
#   'PC 16:0_16:1', 'PC 8:0_28:4', 'PC 20:0_16:4')
featInterest <- c('PC 14:0_18:0', 'PC 16:0_18:1', 'PC 17:1_17:1')

# grep('^PC', volcanoTab$Lipid, value = T)
# grep('^PC \\d+:\\d+$', volcanoTab$Lipid, value = T) %>%
#   stringr::str_remove_all('^PC |:\\d+$')
# grep('^PC \\d+:\\d+_\\d+:\\d+', volcanoTab$Lipid, value = T)
# grep('^PC O-\\d+:\\d+$', volcanoTab$Lipid, value = T)
# grep('^PC\\(.+\\)$', volcanoTab$Lipid, value = T)

# Make volcano plot
# Prepare needed information
pValTab <- lipTissueResSVs_Dis$featAssoRes %>%
  dplyr::select(Var1, pVal, pValAdj) %>%
  dplyr::rename(Feature = Var1)
log2FCTab <- dplyr::select(lipTissueLog2FC_SVs_Dis, name, diff) %>%
  dplyr::rename(Feature = name, log2FC = diff)
featAnnoTab <- tibble::as_tibble(rowData(lipTissueVsn_Dis), rownames = 'Feature') %>%
  dplyr::select(Feature, Lipid, Flag)
# Specify cutoff for log2(FC)
fcCutoff <- 1
# Combine all needed information
volcanoTab <- dplyr::left_join(pValTab, log2FCTab, by = 'Feature') %>%
  dplyr::left_join(featAnnoTab, by = 'Feature') %>%
  # Remove unannotated features
  # dplyr::filter(!is.na(Lipid)) %>%
  # Pinpoint significant features
  dplyr::mutate(DiffExp = dplyr::case_when(pVal <= 0.05 & log2FC >= fcCutoff ~ 'More in Tumor',
                                           pVal <= 0.05 & log2FC <= -fcCutoff ~ 'More in Normal'),
                DiffExp = dplyr::case_when(is.na(DiffExp) ~ 'Not significant',
                                           !is.na(DiffExp) ~ DiffExp),
                DiffExp = factor(DiffExp, levels = c('More in Tumor', 'More in Normal', 'Not significant')),
                Label = dplyr::case_when(Lipid %in% featInterest ~ Lipid)
                # Label = dplyr::case_when(!DiffExp %in% 'Not significant' & grepl('^PC|^PC O', Lipid) ~ Lipid)
  )

ggplot(volcanoTab, aes(x=log2FC, y=-log10(pVal), col=DiffExp, label=Label)) +
  geom_point(size = 2, alpha = 0.9) +
  geom_hline(yintercept = -log10(0.05), linewidth = 0.2) +
  geom_vline(xintercept = c(-fcCutoff, fcCutoff), linewidth = 0.2) +
  geom_text_repel(size = 4, show.legend = F) +
  scale_color_manual(name = '', values = c('red', 'navy', 'grey')) +
  labs(x = 'log2(FC)') +
  th
```

<font size='4'> **Volcano plot of preprocessed data without correction** </font>
```{r}
# Compute log2(FC) using median normalized data
# Load preprocessed data
lipTissueMedi_Dis <- readRDS(paste0('./data/Discovery/AG_Hopf/partial_annotated/',
                                    'Dis_lipTissueMedi_Anno.rds'))
# Subset data by annotated features
lipTissueMedi_Dis <- lipTissueMedi_Dis[!is.na(rowData(lipTissueMedi_Dis)$Lipid),]

# Compute log2(FC) using proDA
# Prepare data matrix and sample metadata table for creating model matrices
datMat <- assay(lipTissueMedi_Dis)
metadatTab <- tibble::as_tibble(colData(lipTissueMedi_Dis), rownames = 'Sample') %>%
  # Make comparison Tumor vs Normal
  dplyr::mutate(Condition = factor(Condition, levels = c('Normal', 'Tumor')))
design <- model.matrix(~ Condition, data = metadatTab)

# fit <- proDA(datMat, design = design)
# diffTerm <- proDA::result_names(fit)[2]
# lipTissueLog2FC_Dis <- proDA::test_diff(fit, contrast = diffTerm)
# saveRDS(lipTissueLog2FC_Dis, paste0('./data/Discovery/AG_Hopf/partial_annotated/soaRes/',
#                                     'lipTissueLog2FC_Condi_OnlyAnno_Dis.rds'))
lipTissueLog2FC_Dis <- readRDS(paste0('./data/Discovery/AG_Hopf/partial_annotated/soaRes/',
                                      'lipTissueLog2FC_Condi_OnlyAnno_Dis.rds'))


####
featInterest <- c('PC 14:0_18:0', 'PC 16:0_18:1', 'PC 17:1_17:1')

# Make volcano plot
# Prepare needed information
pValTab <- lipTissueRes_Dis$featAssoRes %>%
  dplyr::select(Var1, pVal, pValAdj) %>%
  dplyr::rename(Feature = Var1)
log2FCTab <- dplyr::select(lipTissueLog2FC_Dis, name, diff) %>%
  dplyr::rename(Feature = name, log2FC = diff)
featAnnoTab <- tibble::as_tibble(rowData(lipTissueVsn_Dis), rownames = 'Feature') %>%
  dplyr::select(Feature, Lipid, Flag)
# Specify cutoff for log2(FC)
fcCutoff <- 1
# Combine all needed information
volcanoTab <- dplyr::left_join(pValTab, log2FCTab, by = 'Feature') %>%
  dplyr::left_join(featAnnoTab, by = 'Feature') %>%
  # Remove unannotated features
  # dplyr::filter(!is.na(Lipid)) %>%
  # Pinpoint significant features
  dplyr::mutate(DiffExp = dplyr::case_when(pVal <= 0.05 & log2FC >= fcCutoff ~ 'More in Tumor',
                                           pVal <= 0.05 & log2FC <= -fcCutoff ~ 'More in Normal'),
                DiffExp = dplyr::case_when(is.na(DiffExp) ~ 'Not significant',
                                           !is.na(DiffExp) ~ DiffExp),
                DiffExp = factor(DiffExp, levels = c('More in Tumor', 'More in Normal', 'Not significant')),
                Label = dplyr::case_when(Lipid %in% featInterest ~ Lipid)
                # Label = dplyr::case_when(!DiffExp %in% 'Not significant' & grepl('^PC|^PC O', Lipid) ~ Lipid)
  )

ggplot(volcanoTab, aes(x=log2FC, y=-log10(pVal), col=DiffExp, label=Label)) +
  geom_point(size = 2, alpha = 0.9) +
  geom_hline(yintercept = -log10(0.05), linewidth = 0.2) +
  geom_vline(xintercept = c(-fcCutoff, fcCutoff), linewidth = 0.2) +
  geom_text_repel(size = 4, show.legend = F) +
  scale_color_manual(name = '', values = c('red', 'navy', 'grey')) +
  labs(x = 'log2(FC)') +
  th
```

## Patient Recurrence
Recurrence vs Nonrecurrence\
Discovery + MethodDev

<font size='4'> **Metadata-assisted quality control: Tumor vs Normal samples** </font>\
**Data corrected for Cohort using limma::removeBatchEffect**
```{r}
# Load preprocessed data
lipTissue <- readRDS(paste0('./data/Discovery/AG_Hopf/partial_annotated/',
                            'Comb_Prepro_lipTissueVsn_Anno_MDevB123.rds'))
lipTissue_BC <- readRDS(paste0('./data/Discovery/AG_Hopf/partial_annotated/',
                               'Comb_Prepro_lipTissueVsn_Anno_MDevB123_BC.rds'))
# Perform single-omics analysis
lipTissueRes <- doSOA(lipTissue_BC, meta_var = 'Condition', pca_method = 'ppca',
                      num_PCs = 20, do_onlyPCA = T)
# Display significant associations
# lipTissueRes$pcSigAssoRes
# Visualize PCs
pcTab <- lipTissueRes$pcTab %>%
  dplyr::mutate(Cohort = factor(Cohort, levels = c('MethodDev', 'Discovery')))
ggplot(pcTab, aes(x=Condition, y=`PC1 (36.6%)`, col=Condition, fill=Condition)) +
  geom_boxplot(alpha = 0.7, outliers = F) +
  geom_jitter(position = position_jitter(0.3), size = 2, show.legend = F) +
  scale_color_brewer(palette = 'Dark2') +
  scale_fill_brewer(palette = 'Dark2') +
  th

# ggplot(pcTab, aes(x=`PC1 (36.6%)`, y=`PC2 (18.3%)`, col=Cohort, shape=Condition)) +
#   geom_point(size = 4) +
#   scale_color_brewer(palette = 'Set2') +
#   scale_shape_manual(values = c(0, 16)) +
#   th
```

**Cohort-corrected data accounted for SVs identified using sva**\
Perform SVA on cohort-corrected data because SVs identified from uncorrected data
cannot effectively capture sources of unwanted variance (e.g., cohort)
```{r}
# Prepare data accounted for SVs for later visualizations
# Do SVA
lipTissueSVA <- doSVA(lipTissue_BC, wantedVar = c('Condition', 'Recurrence'), numSV_method = 'be',
                      asso_metaVar = c('Cohort', 'Condition', 'Recurrence', 'Gender',
                                       'Age', 'Smoking', 'Stage', 'Adjuvant', 'TumorPurity'))
# Show significant associations between SVs and sample metadata variables
# Sample condition accounts for large variance in data, so perhaps to preserve effects
# of recurrence, some variance contributed from condition is eliminated.
lipTissueSVA$sigAssoTab
# Perform single-omics analysis
SVs <- grep('^SV\\d+$', colnames(colData(lipTissueSVA$summExp_SVs)), value = T)
lipTissueRes4Viz <- doSOA(lipTissueSVA$summExp_SVs, meta_var = 'Condition', pca_method = 'ppca',
                          num_PCs = 20, do_onlyPCA = T, unwantVar = SVs)
# Display significant associations
# lipTissueRes4Viz$pcSigAssoRes
# Visualize PCs
pcTab <- lipTissueRes4Viz$pcTab
ggplot(pcTab, aes(x=Condition, y=`PC1 (50.2%)`, col=Condition, fill=Condition)) +
  geom_boxplot(alpha = 0.7, outliers = F) +
  geom_jitter(position = position_jitter(0.3), size = 2, show.legend = F) +
  scale_color_brewer(palette = 'Dark2') +
  scale_fill_brewer(palette = 'Dark2') +
  th
```

### Normal samples

```{r message=F}
# Performing SOA on uncorrected data and specifying cohort as unwanted variance
# results in insufficient statistical power, so does performing SOA on cohort-corrected data
# Subset certain samples
lipNormal <- lipTissue_BC[, which(colData(lipTissue_BC)$Condition == 'Normal')]
# Prepare feature annotation table
featAnnoTab <- tibble::as_tibble(rowData(lipNormal), rownames = 'Lipid') %>%
  dplyr::select(Lipid, Flag) %>%
  tibble::column_to_rownames('Lipid')

# Append identified SVs from all tissue samples
SVs <- colData(lipTissueSVA$summExp_SVs) %>%
  tibble::as_tibble(rownames = 'Sample') %>%
  dplyr::filter(Condition %in% 'Normal') %>%
  dplyr::select(Sample, contains('SV'))
lipNormalSVs <- summExp2df(lipNormal, row_id = 'Feature', col_id = 'Sample') %>%
  dplyr::left_join(SVs, by = 'Sample') %>%
  df2SummExp(row_id = 'Feature', col_id = 'Sample', values = 'Value', row_anno = c('Formula', 'Flag'),
             col_anno = c('Patient', 'SmpType', 'TimePoint', 'Date', 'Cohort', 'Condition',
                          'Recurrence', 'Gender', 'Age', 'Smoking', 'Stage', 'Adjuvant',
                          'TumorPurity', grep('^SV\\d+$', colnames(SVs), value = T)))

# Perform single-omics analysis
# lipNormalRes <- doSingleOmicsAnalysis(lipNormalSVs, soa_metaVar = 'Recurrence', num_sigFeats = 6,
#                                       pca_method = 'ppca', num_PCs = 20, num_PCfeats = 30,
#                                       use_limma = F, use_proDA = T, feat_anno = featAnnoTab,
#                                       feat_conv = F, plot_title = NULL, show_rownames = T,
#                                       pca_metaVar = c('Recurrence', 'Gender', 'Age',
#                                                       'Smoking', 'Stage', 'Adjuvant',
#                                                       'Cohort'),
#                                       soa_unwantVar = grep('^SV\\d+$', colnames(colData(lipNormalSVs)),
#                                                            value = T),
#                                       level_metaVar = c('Yes', 'No'),
#                                       fontsize_col = 6)
# saveRDS(lipNormalRes, './data/Discovery/AG_Hopf/partial_annotated/soaRes/lipNormalResSVs_Recur_Comb.rds')
lipNormalRes <- readRDS('./data/Discovery/AG_Hopf/partial_annotated/soaRes/lipNormalResSVs_Recur_Comb.rds')
```

Display significant associations between PCs (Var1) and sample metadata variables
(Var2) to overview sources of variance in data
```{r}
lipNormalRes$data.var.source
```

Display significant associations between features (Var1) and cancer recurrence (Var2)
```{r}
dplyr::select(lipNormalRes$sig.feat.tab, -Test)
```

Visualize molecular signatures of significant features in data. Note that features
in rows are ordered by t-statistics.
```{r}
lipNormalRes$sig.feat.heat
```

Visualize top significant features identified in Normal samples (ns: p > 0.05,
$*$: p <= 0.05, $**$: p <= 0.01, $***$: p <= 0.001)
```{r}
# Prepare data matrix and metadata including both Tumor and Normal samples
datMat <- lipTissueRes4Viz$dataCorrect
smpAnnoTab <- lipTissueRes4Viz$smpMetadata
# Prepare significant feature table
featSigAssoTab <- lipNormalRes$sig.feat.tab
num_sigFeats <- 6
# Extract top significant features and prepare needed information
topSigFeats <- featSigAssoTab$Var1[1:num_sigFeats]
topSigFeatDat <- tibble::as_tibble(datMat[topSigFeats,], rownames = 'Feature') %>%
  tidyr::pivot_longer(cols = -'Feature', names_to = 'Sample', values_to = 'Abundance') %>%
  dplyr::left_join(smpAnnoTab, by = 'Sample') %>%
  dplyr::mutate(Feature = factor(Feature, levels = topSigFeats),
                Condition = factor(Condition, levels = c('Normal', 'Tumor')),
                Recurrence = factor(Recurrence, levels = c('Yes', 'No')))
# Visualize top significant features
ggplot(topSigFeatDat, aes(x=Condition, y=Abundance, fill=Recurrence)) +
  geom_boxplot(alpha = 1, linewidth = 1, outlier.color = 'grey50') +
  ggpubr::stat_compare_means(method = 't.test', paired = F, method.args = list(var.equal = T),
                             label = 'p.signif', size = 7, vjust = 0.5) +
  labs(x = 'Sample condition', y = 'Log(Abundance)') +
  scale_fill_manual(values = c('#F8766D', '#00BFC4')) +
  facet_wrap(vars(Feature), scales = 'free') +
  geom_vline(xintercept = 1.5, size = 0.3) +
  th +
  theme(strip.text = element_text(size = 18, face = 'bold'),
        axis.title = element_text(size = 22),
        axis.text.x = element_text(size = 18),
        legend.title = element_text(size = 24), legend.text = element_text(size = 22))
```

Assess data power by p-value histogram
```{r}
lipNormalRes$pval.hist
```

Display significant associations between PCs (Var1) and cancer recurrence (Var2)
```{r}
lipNormalRes$sig.pc.tab
```

Visualize significant PCs and display their top features with greatest absolute loadings
```{r}
lipNormalRes$sig.pc.dist$PC2
lipNormalRes$sig.pc.dist$PC5
```

### Tumor samples

```{r message=F}
# Subset certain samples
lipTumor <- lipTissue_BC[, which(colData(lipTissue_BC)$Condition == 'Tumor')]
# Prepare feature annotation table
featAnnoTab <- tibble::as_tibble(rowData(lipTumor), rownames = 'Lipid') %>%
  dplyr::select(Lipid, Flag) %>%
  tibble::column_to_rownames('Lipid')

# Append identified SVs from all tissue samples
SVs <- colData(lipTissueSVA$summExp_SVs) %>%
  tibble::as_tibble(rownames = 'Sample') %>%
  dplyr::filter(Condition %in% 'Tumor') %>%
  dplyr::select(Sample, contains('SV'))
lipTumorSVs <- summExp2df(lipTumor, row_id = 'Feature', col_id = 'Sample') %>%
  dplyr::left_join(SVs, by = 'Sample') %>%
  df2SummExp(row_id = 'Feature', col_id = 'Sample', values = 'Value', row_anno = c('Formula', 'Flag'),
             col_anno = c('Patient', 'SmpType', 'TimePoint', 'Date', 'Cohort', 'Condition',
                          'Recurrence', 'Gender', 'Age', 'Smoking', 'Stage', 'Adjuvant',
                          'TumorPurity', grep('^SV\\d+$', colnames(SVs), value = T)))

# Perform single-omics analysis
# lipTumorRes <- doSingleOmicsAnalysis(lipTumorSVs, soa_metaVar = 'Recurrence', num_sigFeats = 6,
#                                      pca_method = 'ppca', num_PCs = 20, num_PCfeats = 30,
#                                      use_limma = F, use_proDA = T, feat_anno = featAnnoTab,
#                                      feat_conv = F, plot_title = NULL, show_rownames = T,
#                                      pca_metaVar = c('Recurrence', 'Gender', 'Age',
#                                                      'Smoking', 'Stage', 'Adjuvant',
#                                                      'Cohort'),
#                                      soa_unwantVar = grep('^SV\\d+$', colnames(colData(lipTumorSVs)),
#                                                           value = T),
#                                      level_metaVar = c('Yes', 'No'),
#                                      fontsize_col = 6)
# saveRDS(lipTumorRes, './data/Discovery/AG_Hopf/partial_annotated/soaRes/lipTumorResSVs_Recur_Comb.rds')
lipTumorRes <- readRDS('./data/Discovery/AG_Hopf/partial_annotated/soaRes/lipTumorResSVs_Recur_Comb.rds')
```

Display significant associations between PCs (Var1) and sample metadata variables
(Var2) to overview sources of variance in data
```{r}
lipTumorRes$data.var.source
```

Display significant associations between features (Var1) and cancer recurrence (Var2)
```{r}
dplyr::select(lipTumorRes$sig.feat.tab, -Test)
```

Visualize molecular signatures of significant features in data. Note that features
in rows are ordered by t-statistics.
```{r}
lipTumorRes$sig.feat.heat
```

Visualize top significant features identified in Normal samples (ns: p > 0.05,
$*$: p <= 0.05, $**$: p <= 0.01, $***$: p <= 0.001)
```{r}
# Prepare data matrix and metadata including both Tumor and Normal samples
datMat <- lipTissueRes4Viz$dataCorrect
smpAnnoTab <- lipTissueRes4Viz$smpMetadata
# Prepare significant feature table
featSigAssoTab <- lipTumorRes$sig.feat.tab
num_sigFeats <- 6
# Extract top significant features and prepare needed information
topSigFeats <- featSigAssoTab$Var1[1:num_sigFeats]
topSigFeatDat <- tibble::as_tibble(datMat[topSigFeats,], rownames = 'Feature') %>%
  tidyr::pivot_longer(cols = -'Feature', names_to = 'Sample', values_to = 'Abundance') %>%
  dplyr::left_join(smpAnnoTab, by = 'Sample') %>%
  dplyr::mutate(Feature = factor(Feature, levels = topSigFeats),
                Condition = factor(Condition, levels = c('Tumor', 'Normal')),
                Recurrence = factor(Recurrence, levels = c('Yes', 'No')))
# Visualize top significant features
ggplot(topSigFeatDat, aes(x=Condition, y=Abundance, fill=Recurrence)) +
  geom_boxplot(alpha = 1, linewidth = 1, outlier.color = 'grey50') +
  ggpubr::stat_compare_means(method = 't.test', paired = F, method.args = list(var.equal = T),
                             label = 'p.signif', size = 7, vjust = 0.5) +
  labs(x = 'Sample condition', y = 'Log(Abundance)') +
  scale_fill_manual(values = c('#F8766D', '#00BFC4')) +
  facet_wrap(vars(Feature), scales = 'free') +
  geom_vline(xintercept = 1.5, size = 0.3) +
  th +
  theme(strip.text = element_text(size = 9, face = 'bold'),
        axis.title = element_text(size = 22),
        axis.text.x = element_text(size = 18),
        legend.title = element_text(size = 24), legend.text = element_text(size = 22))
```

Assess data power by p-value histogram
```{r}
lipTumorRes$pval.hist
```

Display significant associations between PCs (Var1) and cancer recurrence (Var2)
```{r}
lipTumorRes$sig.pc.tab
```

Visualize significant PCs and display their top features with greatest absolute loadings
```{r}
lipTumorRes$sig.pc.dist$PC19
lipTumorRes$sig.pc.dist$PC2
```
