---
title: 'Comparison: SOA of Combined DIA Proteomics (Discovery + MethodDev)'
author: "Qian-Wu Liao"
output:
  workflowr::wflow_html:
    toc: yes
    toc_float: yes
    code_folding: hide
---

<font size='4'> Description: Do single-omics analysis of data either unadjusted,
adjusted for SVs constructed by Tumor or Normal subdatasets, or adjusted for SVs
constructed by whole dataset including Tumor and Normal samples. Note that data
has been preprocessed following established preprocessing pipeline. </font>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 8, fig.width = 10, warning = F, message = F)
knitr::opts_knit$set(root.dir = '/Users/qianwu/Desktop/SMART-CARE_LungCancer')
```

Load libraries
```{r library loading, message = F}
library('pcaMethods')
library('sva')
library('limma')
library('proDA')
library('pheatmap')
library('ggplotify')
library('ggrepel')
library('ggvenn')
library('AnnotationDbi')
library('org.Hs.eg.db')
# This annotation object is accessed using AnnotationDbi:select()
hs <- org.Hs.eg.db
library('msigdbr')
library('clusterProfiler')
library(spsUtil)
library('SummarizedExperiment')
library('tidyverse')
# Load user-defined functions
source('./code/analysis_pipeline.R')
source('./code/misc.R')
source('./code/comparison_funcs.R')

# Set plot theme
th <- theme_bw(base_size = 15) +
    theme(axis.title = element_text(face = 'bold'),
          axis.text = element_text(face = 'bold'),
          axis.ticks = element_line(linewidth = 0.8),
          legend.text = element_text(size = 15))
```

# Tissue Proteomics (AG Klingmüller)
Discovery + MethodDev

<font size='4'> **Metadata-assisted quality control: Tumor vs Normal samples** </font>
```{r}
# Load preprocessed data
proTissue_Klin <- readRDS('./data/Discovery/AG_Klingmueller/proTissueVsn.rds')
# Perform single-omics analysis
proTissueRes_Klin <- doSOA(proTissue_Klin, meta_var = 'Condition', pca_method = 'ppca',
                           num_PCs = 20, do_onlyPCA = T)

# Visualize PCs
pcTab <- proTissueRes_Klin$pcTab
ggplot(pcTab, aes(x=Condition, y=`PC1 (33.3%)`, col=Condition, fill=Condition)) +
  geom_boxplot(alpha = 0.7, outliers = F) +
  geom_jitter(position = position_jitter(0.3), size = 2, show.legend = F) +
  scale_color_brewer(palette = 'Dark2') +
  scale_fill_brewer(palette = 'Dark2') +
  th
```

<font size='4'> **Surrogate variable analysis: All Tissue samples** </font>\
Prepare SVs for latter use

Display significant associations between estimated SVs and patient metadata variables
and show data quality after adjusted by SVs\
<font size='3'> **No interaction term - Condition:Recurrence** </font>
```{r}
# Prepare data accounted for SVs for later visualizations
# Do SVA
proTissueSVA_Klin <- doSVA(proTissue_Klin, wantedVar = c('Condition', 'Recurrence'),
                           numSV_method = 'be', asso_metaVar = c('Condition', 'Recurrence',
                                                                 'Gender', 'Age', 'Smoking',
                                                                 'Stage', 'TumorPurity'))
# Show significant associations between SVs and sample metadata variables
# Sample condition accounts for large variance in data, so perhaps to preserve effects
# of recurrence, some variance contributed from condition is eliminated.
proTissueSVA_Klin$sigAssoTab
# Perform single-omics analysis
proTissueSVs_Klin <- proTissueSVA_Klin$summExp_SVs
SVs <- grep('^SV\\d+$', colnames(colData(proTissueSVs_Klin)), value = T)
proTissueRes4Viz_Klin <- doSOA(proTissueSVs_Klin, meta_var = 'Condition', pca_method = 'ppca',
                               num_PCs = 20, do_onlyPCA = T, unwantVar = SVs)

# Visualize PCs
pcTab <- proTissueRes4Viz_Klin$pcTab
ggplot(pcTab, aes(x=Condition, y=`PC1 (57.7%)`, col=Condition, fill=Condition)) +
  geom_boxplot(alpha = 0.7, outliers = F) +
  geom_jitter(position = position_jitter(0.3), size = 2, show.legend = F) +
  scale_color_brewer(palette = 'Dark2') +
  scale_fill_brewer(palette = 'Dark2') +
  th
```

<font size='3'> **With interaction term - Condition:Recurrence** </font>
```{r}
# Prepare data accounted for SVs for later visualizations
# Do SVA
proTissueSVA_Int_Klin <- doSVA(proTissue_Klin, wantedVar = c('Condition', 'Recurrence', 'Condition:Recurrence'),
                               numSV_method = 'be', asso_metaVar = c('Condition', 'Recurrence',
                                                                     'Gender', 'Age', 'Smoking',
                                                                     'Stage', 'TumorPurity'))
# Show significant associations between SVs and sample metadata variables
# Sample condition accounts for large variance in data, so perhaps to preserve effects
# of recurrence, some variance contributed from condition is eliminated.
proTissueSVA_Int_Klin$sigAssoTab
# Perform single-omics analysis
proTissueSVs_Int_Klin <- proTissueSVA_Int_Klin$summExp_SVs
SVs <- grep('^SV\\d+$', colnames(colData(proTissueSVs_Int_Klin)), value = T)
proTissueRes4Viz_Int_Klin <- doSOA(proTissueSVs_Int_Klin, meta_var = 'Condition', pca_method = 'ppca',
                                   num_PCs = 20, do_onlyPCA = T, unwantVar = SVs)

# Visualize PCs
pcTab <- proTissueRes4Viz_Int_Klin$pcTab
ggplot(pcTab, aes(x=Condition, y=`PC1 (58.1%)`, col=Condition, fill=Condition)) +
  geom_boxplot(alpha = 0.7, outliers = F) +
  geom_jitter(position = position_jitter(0.3), size = 2, show.legend = F) +
  scale_color_brewer(palette = 'Dark2') +
  scale_fill_brewer(palette = 'Dark2') +
  th
```
=> Some SVs are significantly associated with Sample Conditions and Patient Tumor
Recurrence (interesting variance). It may be that Condition accounts for major variance
in data, so some associations with constructed SVs are inevitable. For Recurrence,
it may be due to outliler samples observed in molecular signatures, especially of
Normal samples (check Section Normal samples - UNADJUSTED). Those outlier samples
are mostly from Recurrence patient groups, and SVA tries to estimate differences
between outlier and the other samples (uninteresting variance), which makes those
SVs associated with Recurrence.

## Tumor samples

Display associations between cancer recurrence and the other patient metadata variables
from patients included in tumor sample analysis to identify potential confounders (or predictors)
```{r}
# Test associations between cancer recurrence and the other patient metadata variables
# Prepare patient metadata
# Subset certain samples
proTumor_Klin <- proTissue_Klin[, which(colData(proTissue_Klin)$Condition == 'Tumor')]
patientMetadat <- tibble::as_tibble(colData(proTumor_Klin)) %>%
  dplyr::filter(!duplicated(Patient)) %>%
  dplyr::select(Patient, Recurrence, Gender, Age, Smoking, Stage, Adjuvant)
patientRecur <- dplyr::select(patientMetadat, Patient, Recurrence)
patientMetadat <- dplyr::select(patientMetadat, -Recurrence)
testAsso(patientMetadat, patientRecur, cmn_col = 'Patient') %>%
  dplyr::select(-Stat)
```

### UNADJUSTED

```{r message=F}
# Subset certain samples
proTumor_Klin <- proTissue_Klin[, which(colData(proTissue_Klin)$Condition == 'Tumor')]
# Prepare feature annotation table
featAnnoTab <- tibble::as_tibble(rowData(proTumor_Klin), rownames = 'Protein') %>%
  tibble::column_to_rownames('Protein')

# Perform single-omics analysis
# proTumorRes_Klin <- doSingleOmicsAnalysis(proTumor_Klin, soa_metaVar = 'Recurrence', num_sigFeats = 6,
#                                           pca_method = 'ppca', num_PCs = 20, num_PCfeats = 30,
#                                           use_limma = F, use_proDA = T, feat_anno = featAnnoTab,
#                                           feat_conv = T, plot_title = NULL, show_rownames = F,
#                                           pca_metaVar = c('Recurrence', 'Gender', 'Age',
#                                                           'Smoking', 'Stage', 'TumorPurity'),
#                                           level_metaVar = c('Yes', 'No'))
# saveRDS(proTumorRes_Klin, './data/Discovery/AG_Klingmueller/soaRes/trials/proTumorRes.rds')
proTumorRes_Klin <- readRDS('./data/Discovery/AG_Klingmueller/soaRes/trials/proTumorRes.rds')
```

Display significant associations between PCs (Var1) and sample metadata variables
(Var2) to overview sources of variance in data
```{r}
proTumorRes_Klin$data.var.source
```

Display significant associations between features (Var1) and cancer recurrence (Var2)
```{r}
dplyr::select(proTumorRes_Klin$sig.feat.tab, -Test) %>%
  DT::datatable()
```

Assess data power by p-value histogram
```{r}
proTumorRes_Klin$pval.hist
```

Visualize molecular signatures of significant features in data. Note that features
in rows are ordered by t-statistics.
```{r}
proTumorRes_Klin$sig.feat.heat
```

Display significant associations between PCs (Var1) and cancer recurrence (Var2)
```{r}
proTumorRes_Klin$sig.pc.tab
```

### SV-ADJUSTED (PROTUMOR)
**Surrogate variable analysis of Only Tumor samples**

Display associations between estimated surrogate variables and patient metadata variables
```{r message=F}
# Subset certain samples
proTumor_Klin <- proTissue_Klin[, which(colData(proTissue_Klin)$Condition == 'Tumor')]
# Prepare feature annotation table
featAnnoTab <- tibble::as_tibble(rowData(proTumor_Klin), rownames = 'Protein') %>%
  tibble::column_to_rownames('Protein')

# Estimate unwanted variance through surrogate variable analysis
proTumorSVA_Klin <- doSVA(proTumor_Klin, wantedVar = 'Recurrence', numSV_method = 'be',
                          asso_metaVar = c('Recurrence', 'Gender', 'Age', 'Smoking', 'Stage', 'TumorPurity'))
proTumor_Klin <- proTumorSVA_Klin$summExp_SVs
# Show significant associations between SVs and sample metadata variables
proTumorSVA_Klin$sigAssoTab

# Perform single-omics analysis
# proTumorRes_Klin <- doSingleOmicsAnalysis(proTumor_Klin, soa_metaVar = 'Recurrence', num_sigFeats = 6,
#                                           pca_method = 'ppca', num_PCs = 20, num_PCfeats = 30,
#                                           use_limma = F, use_proDA = T, feat_anno = featAnnoTab,
#                                           feat_conv = T, plot_title = NULL, show_rownames = F,
#                                           pca_metaVar = c('Recurrence', 'Gender', 'Age',
#                                                           'Smoking', 'Stage', 'TumorPurity'),
#                                           soa_unwantVar = grep('^SV\\d+$', colnames(colData(proTumor_Klin)),
#                                                                value = T),
#                                           level_metaVar = c('Yes', 'No'))
# saveRDS(proTumorRes_Klin, './data/Discovery/AG_Klingmueller/soaRes/trials/proTumorRes_SVsByProTumor.rds')
proTumorRes_Klin <- readRDS('./data/Discovery/AG_Klingmueller/soaRes/trials/proTumorRes_SVsByProTumor.rds')
```

Display significant associations between features (Var1) and cancer recurrence (Var2)
```{r}
dplyr::select(proTumorRes_Klin$sig.feat.tab, -Test) %>%
  DT::datatable()
```

Assess data power by p-value histogram
```{r}
proTumorRes_Klin$pval.hist
```

Visualize molecular signatures of significant features in data. Note that features
in rows are ordered by t-statistics.
```{r}
proTumorRes_Klin$sig.feat.heat
```

Display significant associations between PCs (Var1) and cancer recurrence (Var2)
```{r}
proTumorRes_Klin$sig.pc.tab
```

Enrichment analysis\
Reference: Hallmark gene sets\
**t-values**
```{r}
# Prepare ranked gene list
statRes_Klin <- proTumorRes_Klin$SOA.res$featAssoRes %>%
  dplyr::select(Var1, Stat) %>%
  tibble::column_to_rownames('Var1')
featAnno_Klin <- as.data.frame(rowData(proTissue_Klin))
rankedGeneList_Klin <- prepRankedFeatList(statRes = statRes_Klin, featAnno = featAnno_Klin, to_gene = T)

# Do GSEA
# Hallmark gene sets
gseaRes_H <- doEA(rankedGeneList_Klin, categoryDB = 'H', numSigGS = 10, rmPrefixGS = '^HALLMARK_', pvalueCutoff = 0.2)
gseaRes_H$plotTopSigGS +
  labs(title = 'Hallmark gene sets')
```

### SV-ADJUSTED (PROTISSUE)
**Surrogate variable analysis of All Tissue samples**

<font size='5'> **No interaction term - Condition:Recurrence** </font>

```{r message=F}
# Subset certain samples
proTumor_Klin <- proTissueSVs_Klin[, which(colData(proTissueSVs_Klin)$Condition == 'Tumor')]
# Prepare feature annotation table
featAnnoTab <- tibble::as_tibble(rowData(proTumor_Klin), rownames = 'Protein') %>%
  tibble::column_to_rownames('Protein')

# Perform single-omics analysis
# proTumorRes_Klin <- doSingleOmicsAnalysis(proTumor_Klin, soa_metaVar = 'Recurrence', num_sigFeats = 6,
#                                           pca_method = 'ppca', num_PCs = 20, num_PCfeats = 30,
#                                           use_limma = F, use_proDA = T, feat_anno = featAnnoTab,
#                                           feat_conv = T, plot_title = NULL, show_rownames = F,
#                                           pca_metaVar = c('Recurrence', 'Gender', 'Age',
#                                                           'Smoking', 'Stage', 'TumorPurity'),
#                                           soa_unwantVar = grep('^SV\\d+$', colnames(colData(proTumor_Klin)),
#                                                                value = T),
#                                           level_metaVar = c('Yes', 'No'))
# saveRDS(proTumorRes_Klin, './data/Discovery/AG_Klingmueller/soaRes/trials/proTumorRes_SVsByProTissue.rds')
proTumorRes_Klin <- readRDS('./data/Discovery/AG_Klingmueller/soaRes/trials/proTumorRes_SVsByProTissue.rds')
```

Display significant associations between features (Var1) and cancer recurrence (Var2)
```{r}
dplyr::select(proTumorRes_Klin$sig.feat.tab, -Test) %>%
  DT::datatable()
```

Assess data power by p-value histogram
```{r}
proTumorRes_Klin$pval.hist
```

Visualize molecular signatures of significant features in data. Note that features
in rows are ordered by t-statistics.
```{r}
proTumorRes_Klin$sig.feat.heat
```

Display significant associations between PCs (Var1) and cancer recurrence (Var2)
```{r}
proTumorRes_Klin$sig.pc.tab
```

Enrichment analysis\
Reference: Hallmark gene sets\
**t-values**
```{r}
# Prepare ranked gene list
statRes_Klin <- proTumorRes_Klin$SOA.res$featAssoRes %>%
  dplyr::select(Var1, Stat) %>%
  tibble::column_to_rownames('Var1')
featAnno_Klin <- as.data.frame(rowData(proTissue_Klin))
rankedGeneList_Klin <- prepRankedFeatList(statRes = statRes_Klin, featAnno = featAnno_Klin, to_gene = T)

# Do GSEA
# Hallmark gene sets
gseaRes_H <- doEA(rankedGeneList_Klin, categoryDB = 'H', numSigGS = 10, rmPrefixGS = '^HALLMARK_', pvalueCutoff = 0.2)
gseaRes_H$plotTopSigGS +
  labs(title = 'Hallmark gene sets')
```

<font size='5'> **With interaction term - Condition:Recurrence** </font>

```{r message=F}
# Subset certain samples
proTumor_Klin <- proTissueSVs_Int_Klin[, which(colData(proTissueSVs_Int_Klin)$Condition == 'Tumor')]
# Prepare feature annotation table
featAnnoTab <- tibble::as_tibble(rowData(proTumor_Klin), rownames = 'Protein') %>%
  tibble::column_to_rownames('Protein')

# Perform single-omics analysis
# proTumorRes_Klin <- doSingleOmicsAnalysis(proTumor_Klin, soa_metaVar = 'Recurrence', num_sigFeats = 6,
#                                           pca_method = 'ppca', num_PCs = 20, num_PCfeats = 30,
#                                           use_limma = F, use_proDA = T, feat_anno = featAnnoTab,
#                                           feat_conv = T, plot_title = NULL, show_rownames = F,
#                                           pca_metaVar = c('Recurrence', 'Gender', 'Age',
#                                                           'Smoking', 'Stage', 'TumorPurity'),
#                                           soa_unwantVar = grep('^SV\\d+$', colnames(colData(proTumor_Klin)),
#                                                                value = T),
#                                           level_metaVar = c('Yes', 'No'))
# saveRDS(proTumorRes_Klin, './data/Discovery/AG_Klingmueller/soaRes/trials/proTumorRes_SVsByProTissueInt.rds')
proTumorRes_Klin <- readRDS('./data/Discovery/AG_Klingmueller/soaRes/trials/proTumorRes_SVsByProTissueInt.rds')
```

Display significant associations between features (Var1) and cancer recurrence (Var2)
```{r}
dplyr::select(proTumorRes_Klin$sig.feat.tab, -Test) %>%
  DT::datatable()
```

Assess data power by p-value histogram
```{r}
proTumorRes_Klin$pval.hist
```

Visualize molecular signatures of significant features in data. Note that features
in rows are ordered by t-statistics.
```{r}
proTumorRes_Klin$sig.feat.heat
```

Display significant associations between PCs (Var1) and cancer recurrence (Var2)
```{r}
proTumorRes_Klin$sig.pc.tab
```

Enrichment analysis\
Reference: Hallmark gene sets\
**t-values**
```{r}
# Prepare ranked gene list
statRes_Klin <- proTumorRes_Klin$SOA.res$featAssoRes %>%
  dplyr::select(Var1, Stat) %>%
  tibble::column_to_rownames('Var1')
featAnno_Klin <- as.data.frame(rowData(proTissue_Klin))
rankedGeneList_Klin <- prepRankedFeatList(statRes = statRes_Klin, featAnno = featAnno_Klin, to_gene = T)

# Do GSEA
# Hallmark gene sets
gseaRes_H <- doEA(rankedGeneList_Klin, categoryDB = 'H', numSigGS = 10, rmPrefixGS = '^HALLMARK_', pvalueCutoff = 0.2)
gseaRes_H$plotTopSigGS +
  labs(title = 'Hallmark gene sets')
```

## Normal samples

Display associations between cancer recurrence and the other patient metadata variables
from patients included in normal sample analysis to identify potential confounders (or predictors)
```{r}
# Test associations between cancer recurrence and the other patient metadata variables
# Prepare patient metadata
# Subset certain samples
proNormal_Klin <- proTissue_Klin[, which(colData(proTissue_Klin)$Condition == 'Normal')]
patientMetadat <- tibble::as_tibble(colData(proNormal_Klin)) %>%
  dplyr::filter(!duplicated(Patient)) %>%
  dplyr::select(Patient, Recurrence, Gender, Age, Smoking, Stage, Adjuvant)
patientRecur <- dplyr::select(patientMetadat, Patient, Recurrence)
patientMetadat <- dplyr::select(patientMetadat, -Recurrence)
testAsso(patientMetadat, patientRecur, cmn_col = 'Patient') %>%
  dplyr::select(-Stat)
```

### UNADJUSTED

```{r message=F}
# Subset certain samples
proNormal_Klin <- proTissue_Klin[, which(colData(proTissue_Klin)$Condition == 'Normal')]
# Prepare feature annotation table
featAnnoTab <- tibble::as_tibble(rowData(proNormal_Klin), rownames = 'Protein') %>%
  tibble::column_to_rownames('Protein')

# Perform single-omics analysis
# proNormalRes_Klin <- doSingleOmicsAnalysis(proNormal_Klin, soa_metaVar = 'Recurrence', num_sigFeats = 6,
#                                            pca_method = 'ppca', num_PCs = 20, num_PCfeats = 30,
#                                            use_limma = F, use_proDA = T, feat_anno = featAnnoTab,
#                                            feat_conv = T, plot_title = NULL, show_rownames = F,
#                                            pca_metaVar = c('Recurrence', 'Gender', 'Age',
#                                                            'Smoking', 'Stage'),
#                                            level_metaVar = c('Yes', 'No'))
                                           # fontsize = 15, fontsize_col = 14
# saveRDS(proNormalRes_Klin, './data/Discovery/AG_Klingmueller/soaRes/trials/proNormalRes.rds')
proNormalRes_Klin <- readRDS('./data/Discovery/AG_Klingmueller/soaRes/trials/proNormalRes.rds')
```

Display significant associations between PCs (Var1) and sample metadata variables
(Var2) to overview sources of variance in data
```{r}
proNormalRes_Klin$data.var.source
```

Display significant associations between features (Var1) and cancer recurrence (Var2)
```{r}
dplyr::select(proNormalRes_Klin$sig.feat.tab, -Test) %>%
  DT::datatable()
```

Assess data power by p-value histogram
```{r}
proNormalRes_Klin$pval.hist
```

Visualize molecular signatures of significant features in data. Note that features
in rows are ordered by t-statistics.
```{r}
proNormalRes_Klin$sig.feat.heat
```
=> There are some samples with much higher or lower abundances (middle) or relatively
higher or lower abundances (right), which can increase effect size, giving rise
to significant results. However, those significant features may be false positives.
It explains why sample clustering looks bad and why much less significant features
are identified with adjusted data. That those samples with distinct behaviors are
due to biological or technical factors can be further investigated.

Display significant associations between PCs (Var1) and cancer recurrence (Var2)
```{r}
proNormalRes_Klin$sig.pc.tab
```

Enrichment analysis\
**t-values**
```{r}
# Prepare ranked gene list
statRes_Klin <- proNormalRes_Klin$SOA.res$featAssoRes %>%
  dplyr::select(Var1, Stat) %>%
  tibble::column_to_rownames('Var1')
featAnno_Klin <- as.data.frame(rowData(proTissue_Klin))
rankedGeneList_Klin <- prepRankedFeatList(statRes = statRes_Klin, featAnno = featAnno_Klin, to_gene = T)

# Do GSEA
# Hallmark gene sets
gseaRes_H <- doEA(rankedGeneList_Klin, categoryDB = 'H', numSigGS = 20, rmPrefixGS = '^HALLMARK_')
gseaRes_H$plotTopSigGS +
  labs(title = 'Hallmark gene sets')
```

### SV-ADJUSTED (PRONORMAL)
**Surrogate variable analysis of Only Normal samples**

Display associations between estimated surrogate variables and patient metadata variables
```{r message=F}
# Subset certain samples
proNormal_Klin <- proTissue_Klin[, which(colData(proTissue_Klin)$Condition == 'Normal')]
# Prepare feature annotation table
featAnnoTab <- tibble::as_tibble(rowData(proNormal_Klin), rownames = 'Protein') %>%
  tibble::column_to_rownames('Protein')

# Estimate unwanted variance through surrogate variable analysis
proNormalSVA_Klin <- doSVA(proNormal_Klin, wantedVar = 'Recurrence', numSV_method = 'be',
                           asso_metaVar = c('Recurrence', 'Gender', 'Age', 'Smoking', 'Stage'))
proNormal_Klin <- proNormalSVA_Klin$summExp_SVs
# Show significant associations between SVs and sample metadata variables
proNormalSVA_Klin$sigAssoTab

# Perform single-omics analysis
# proNormalRes_Klin <- doSingleOmicsAnalysis(proNormal_Klin, soa_metaVar = 'Recurrence', num_sigFeats = 6,
#                                            pca_method = 'ppca', num_PCs = 20, num_PCfeats = 30,
#                                            use_limma = F, use_proDA = T, feat_anno = featAnnoTab,
#                                            feat_conv = T, plot_title = NULL, show_rownames = F,
#                                            pca_metaVar = c('Recurrence', 'Gender', 'Age',
#                                                            'Smoking', 'Stage'),
#                                            soa_unwantVar = grep('^SV\\d+$', colnames(colData(proNormal_Klin)),
#                                                                 value = T),
#                                            level_metaVar = c('Yes', 'No'))
# saveRDS(proNormalRes_Klin, './data/Discovery/AG_Klingmueller/soaRes/trials/proNormalRes_SVsByProNormal.rds')
proNormalRes_Klin <- readRDS('./data/Discovery/AG_Klingmueller/soaRes/trials/proNormalRes_SVsByProNormal.rds')
```

Display significant associations between features (Var1) and cancer recurrence (Var2)
```{r}
dplyr::select(proNormalRes_Klin$sig.feat.tab, -Test) %>%
  DT::datatable()
```

Assess data power by p-value histogram
```{r}
proNormalRes_Klin$pval.hist
```

Visualize molecular signatures of significant features in data. Note that features
in rows are ordered by t-statistics.
```{r}
proNormalRes_Klin$sig.feat.heat
```

Display significant associations between PCs (Var1) and cancer recurrence (Var2)
```{r}
proNormalRes_Klin$sig.pc.tab
```

Enrichment analysis\
**t-values**
```{r}
# Prepare ranked gene list
statRes_Klin <- proNormalRes_Klin$SOA.res$featAssoRes %>%
  dplyr::select(Var1, Stat) %>%
  tibble::column_to_rownames('Var1')
featAnno_Klin <- as.data.frame(rowData(proTissue_Klin))
rankedGeneList_Klin <- prepRankedFeatList(statRes = statRes_Klin, featAnno = featAnno_Klin, to_gene = T)

# Do GSEA
# Hallmark gene sets
gseaRes_H <- doEA(rankedGeneList_Klin, categoryDB = 'H', numSigGS = 20, rmPrefixGS = '^HALLMARK_')
gseaRes_H$plotTopSigGS +
  labs(title = 'Hallmark gene sets')
```

### SV-ADJUSTED (PROTISSUE)
**Surrogate variable analysis of All Tissue samples**

<font size='5'> **No interaction term - Condition:Recurrence** </font>

```{r message=F}
# Subset certain samples
proNormal_Klin <- proTissueSVs_Klin[, which(colData(proTissueSVs_Klin)$Condition == 'Normal')]
# Prepare feature annotation table
featAnnoTab <- tibble::as_tibble(rowData(proNormal_Klin), rownames = 'Protein') %>%
  tibble::column_to_rownames('Protein')

# Perform single-omics analysis
# proNormalRes_Klin <- doSingleOmicsAnalysis(proNormal_Klin, soa_metaVar = 'Recurrence', num_sigFeats = 6,
#                                            pca_method = 'ppca', num_PCs = 20, num_PCfeats = 30,
#                                            use_limma = F, use_proDA = T, feat_anno = featAnnoTab,
#                                            feat_conv = T, plot_title = NULL, show_rownames = F,
#                                            pca_metaVar = c('Recurrence', 'Gender', 'Age',
#                                                            'Smoking', 'Stage'),
#                                            soa_unwantVar = grep('^SV\\d+$', colnames(colData(proNormal_Klin)),
#                                                                 value = T),
#                                            level_metaVar = c('Yes', 'No'))
# saveRDS(proNormalRes_Klin, './data/Discovery/AG_Klingmueller/soaRes/trials/proNormalRes_SVsByProTissue.rds')
proNormalRes_Klin <- readRDS('./data/Discovery/AG_Klingmueller/soaRes/trials/proNormalRes_SVsByProTissue.rds')
```

Display significant associations between features (Var1) and cancer recurrence (Var2)
```{r}
dplyr::select(proNormalRes_Klin$sig.feat.tab, -Test) %>%
  DT::datatable()
```

Assess data power by p-value histogram
```{r}
proNormalRes_Klin$pval.hist
```

Visualize molecular signatures of significant features in data. Note that features
in rows are ordered by t-statistics.
```{r}
proNormalRes_Klin$sig.feat.heat
```

Display significant associations between PCs (Var1) and cancer recurrence (Var2)
```{r}
proNormalRes_Klin$sig.pc.tab
```

Enrichment analysis\
**t-values**
```{r}
# Prepare ranked gene list
statRes_Klin <- proNormalRes_Klin$SOA.res$featAssoRes %>%
  dplyr::select(Var1, Stat) %>%
  tibble::column_to_rownames('Var1')
featAnno_Klin <- as.data.frame(rowData(proTissue_Klin))
rankedGeneList_Klin <- prepRankedFeatList(statRes = statRes_Klin, featAnno = featAnno_Klin, to_gene = T)

# Do GSEA
# Hallmark gene sets
gseaRes_H <- doEA(rankedGeneList_Klin, categoryDB = 'H', numSigGS = 20, rmPrefixGS = '^HALLMARK_')
gseaRes_H$plotTopSigGS +
  labs(title = 'Hallmark gene sets')
```

<font size='5'> **With interaction term - Condition:Recurrence** </font>

```{r message=F}
# Subset certain samples
proNormal_Klin <- proTissueSVs_Int_Klin[, which(colData(proTissueSVs_Int_Klin)$Condition == 'Normal')]
# Prepare feature annotation table
featAnnoTab <- tibble::as_tibble(rowData(proNormal_Klin), rownames = 'Protein') %>%
  tibble::column_to_rownames('Protein')

# Perform single-omics analysis
# proNormalRes_Klin <- doSingleOmicsAnalysis(proNormal_Klin, soa_metaVar = 'Recurrence', num_sigFeats = 6,
#                                            pca_method = 'ppca', num_PCs = 20, num_PCfeats = 30,
#                                            use_limma = F, use_proDA = T, feat_anno = featAnnoTab,
#                                            feat_conv = T, plot_title = NULL, show_rownames = F,
#                                            pca_metaVar = c('Recurrence', 'Gender', 'Age',
#                                                            'Smoking', 'Stage'),
#                                            soa_unwantVar = grep('^SV\\d+$', colnames(colData(proNormal_Klin)),
#                                                                 value = T),
#                                            level_metaVar = c('Yes', 'No'))
# saveRDS(proNormalRes_Klin, './data/Discovery/AG_Klingmueller/soaRes/trials/proNormalRes_SVsByProTissueInt.rds')
proNormalRes_Klin <- readRDS('./data/Discovery/AG_Klingmueller/soaRes/trials/proNormalRes_SVsByProTissueInt.rds')
```

Display significant associations between features (Var1) and cancer recurrence (Var2)
```{r}
dplyr::select(proNormalRes_Klin$sig.feat.tab, -Test) %>%
  DT::datatable()
```

Assess data power by p-value histogram
```{r}
proNormalRes_Klin$pval.hist
```

Visualize molecular signatures of significant features in data. Note that features
in rows are ordered by t-statistics.
```{r}
proNormalRes_Klin$sig.feat.heat
```

Display significant associations between PCs (Var1) and cancer recurrence (Var2)
```{r}
proNormalRes_Klin$sig.pc.tab
```

Enrichment analysis\
**t-values**
```{r}
# Prepare ranked gene list
statRes_Klin <- proNormalRes_Klin$SOA.res$featAssoRes %>%
  dplyr::select(Var1, Stat) %>%
  tibble::column_to_rownames('Var1')
featAnno_Klin <- as.data.frame(rowData(proTissue_Klin))
rankedGeneList_Klin <- prepRankedFeatList(statRes = statRes_Klin, featAnno = featAnno_Klin, to_gene = T)

# Do GSEA
# Hallmark gene sets
gseaRes_H <- doEA(rankedGeneList_Klin, categoryDB = 'H', numSigGS = 20, rmPrefixGS = '^HALLMARK_')
gseaRes_H$plotTopSigGS +
  labs(title = 'Hallmark gene sets')
```




# Tissue Proteomics (AG Krijgsveld)
Discovery + MethodDev

<font size='4'> **Metadata-assisted quality control: Tumor vs Normal samples** </font>
```{r}
# Load preprocessed data
proTissue_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/combined_proTissueVsnBC.rds')
# Perform single-omics analysis
proTissueRes_Krij <- doSOA(proTissue_Krij, meta_var = c('Condition', 'Cohort'),
                           pca_method = 'ppca', num_PCs = 20, do_onlyPCA = T)

# Visualize PCs
pcTab <- proTissueRes_Krij$pcTab
ggplot(pcTab, aes(x=Condition, y=`PC1 (40.7%)`, color=Condition, fill=Condition)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(position = position_jitter(0.3), size = 3, show.legend = F) +
  scale_color_brewer(palette = 'Dark2') +
  scale_fill_brewer(palette = 'Dark2') +
  labs(x = 'Sample condition') +
  th
```

<font size='4'> **Surrogate variable analysis: All Tissue samples** </font>\
Prepare SVs for latter use

Display significant associations between estimated SVs and patient metadata variables
and show data quality after adjusted by SVs\
<font size='3'> **No interaction term - Condition:Recurrence** </font>
```{r}
# Prepare data accounted for SVs for later visualizations
# Do SVA
proTissueSVA_Krij <- doSVA(proTissue_Krij, wantedVar = c('Condition', 'Recurrence'),
                           numSV_method = 'be', asso_metaVar = c('Condition', 'Recurrence',
                                                                 'Gender', 'Age', 'Smoking',
                                                                 'Stage', 'TumorPurity'))
# Show significant associations between SVs and sample metadata variables
# Sample condition accounts for large variance in data, so perhaps to preserve effects
# of recurrence, some variance contributed from condition is eliminated.
proTissueSVA_Krij$sigAssoTab
# Perform single-omics analysis
proTissueSVs_Krij <- proTissueSVA_Krij$summExp_SVs
SVs <- grep('^SV\\d+$', colnames(colData(proTissueSVs_Krij)), value = T)
proTissueRes4Viz_Krij <- doSOA(proTissueSVs_Krij, meta_var = 'Condition', pca_method = 'ppca',
                               num_PCs = 20, do_onlyPCA = T, unwantVar = SVs)

# Visualize PCs
pcTab <- proTissueRes4Viz_Krij$pcTab
ggplot(pcTab, aes(x=Condition, y=`PC1 (63.6%)`, col=Condition, fill=Condition)) +
  geom_boxplot(alpha = 0.7, outliers = F) +
  geom_jitter(position = position_jitter(0.3), size = 2, show.legend = F) +
  scale_color_brewer(palette = 'Dark2') +
  scale_fill_brewer(palette = 'Dark2') +
  th
```

<font size='3'> **With interaction term - Condition:Recurrence** </font>
```{r}
# Prepare data accounted for SVs for later visualizations
# Do SVA
proTissueSVA_Int_Krij <- doSVA(proTissue_Krij, wantedVar = c('Condition', 'Recurrence', 'Condition:Recurrence'),
                               numSV_method = 'be', asso_metaVar = c('Condition', 'Recurrence',
                                                                     'Gender', 'Age', 'Smoking',
                                                                     'Stage', 'TumorPurity'))
# Show significant associations between SVs and sample metadata variables
# Sample condition accounts for large variance in data, so perhaps to preserve effects
# of recurrence, some variance contributed from condition is eliminated.
proTissueSVA_Int_Krij$sigAssoTab
# Perform single-omics analysis
proTissueSVs_Int_Krij <- proTissueSVA_Int_Krij$summExp_SVs
SVs <- grep('^SV\\d+$', colnames(colData(proTissueSVs_Int_Krij)), value = T)
proTissueRes4Viz_Int_Krij <- doSOA(proTissueSVs_Int_Krij, meta_var = 'Condition', pca_method = 'ppca',
                                   num_PCs = 20, do_onlyPCA = T, unwantVar = SVs)

# Visualize PCs
pcTab <- proTissueRes4Viz_Int_Krij$pcTab
ggplot(pcTab, aes(x=Condition, y=`PC1 (64.6%)`, col=Condition, fill=Condition)) +
  geom_boxplot(alpha = 0.7, outliers = F) +
  geom_jitter(position = position_jitter(0.3), size = 2, show.legend = F) +
  scale_color_brewer(palette = 'Dark2') +
  scale_fill_brewer(palette = 'Dark2') +
  th
```

## Tumor samples

Display associations between cancer recurrence and the other patient metadata variables
from patients included in tumor sample analysis to identify potential confounders (or predictors)
```{r}
# Test associations between cancer recurrence and the other patient metadata variables
# Prepare patient metadata
# Subset certain samples
proTumor_Krij <- proTissue_Krij[, which(colData(proTissue_Krij)$Condition == 'Tumor')]
patientMetadat <- tibble::as_tibble(colData(proTumor_Krij)) %>%
  dplyr::filter(!duplicated(Patient)) %>%
  dplyr::select(Patient, Recurrence, Gender, Age, Smoking, Stage, Adjuvant)
patientRecur <- dplyr::select(patientMetadat, Patient, Recurrence)
patientMetadat <- dplyr::select(patientMetadat, -Recurrence)
testAsso(patientMetadat, patientRecur, cmn_col = 'Patient') %>%
  dplyr::select(-Stat)
```

### UNADJUSTED

Display associations between cancer recurrence and the other patient metadata variables
from patients included in tumor sample analysis to identify potential confounders (or predictors)
```{r}
# Subset certain samples
proTumor_Krij <- proTissue_Krij[, which(colData(proTissue_Krij)$Condition == 'Tumor')]
# Prepare feature annotation table
featAnnoTab <- tibble::as_tibble(rowData(proTumor_Krij), rownames = 'Protein') %>%
  dplyr::select(Protein, Genes) %>%
  dplyr::rename(Gene = Genes) %>%
  tibble::column_to_rownames('Protein')

# Perform single-omics analysis
# proTumorRes_Krij <- doSingleOmicsAnalysis(proTumor_Krij, soa_metaVar = 'Recurrence', num_sigFeats = 6,
#                                           pca_method = 'ppca', num_PCs = 20, num_PCfeats = 30,
#                                           use_limma = F, use_proDA = T, feat_anno = featAnnoTab,
#                                           feat_conv = T, plot_title = NULL, show_rownames = F,
#                                           pca_metaVar = c('Recurrence', 'Gender', 'Age',
#                                                           'Smoking', 'Stage', 'TumorPurity'),
#                                           level_metaVar = c('Yes', 'No'))
# saveRDS(proTumorRes_Krij, './data/Discovery/AG_Krijgsveld/soaRes/trials/combined_proTumorRes.rds')
proTumorRes_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/soaRes/trials/combined_proTumorRes.rds')
```

Display significant associations between PCs (Var1) and sample metadata variables
(Var2) to overview sources of variance in data
```{r}
proTumorRes_Krij$data.var.source
```

Display significant associations between features (Var1) and cancer recurrence (Var2).
Owing to missing values, there may be situation that only few observations in sample
groups, resulting in unreliable t-statistics. To this end, we used probabilistic
dropout model (proDA) to account for data missingness.
```{r}
dplyr::select(proTumorRes_Krij$sig.feat.tab, -Test) %>%
  DT::datatable()
```

Assess data power by p-value histogram
```{r}
proTumorRes_Krij$pval.hist
```

Visualize molecular signatures of significant features in data. Note that features
in rows are ordered by t-statistics.
```{r}
proTumorRes_Krij$sig.feat.heat
```

Display significant associations between PCs (Var1) and cancer recurrence (Var2)
```{r}
proTumorRes_Krij$sig.pc.tab
```

### SV-ADJUSTED (PROTUMOR)
**Surrogate variable analysis of Only Tumor samples**

Display associations between estimated surrogate variables and patient metadata variables
```{r}
# Subset certain samples
proTumor_Krij <- proTissue_Krij[, which(colData(proTissue_Krij)$Condition == 'Tumor')]
# Prepare feature annotation table
featAnnoTab <- tibble::as_tibble(rowData(proTumor_Krij), rownames = 'Protein') %>%
  dplyr::select(Protein, Genes) %>%
  dplyr::rename(Gene = Genes) %>%
  tibble::column_to_rownames('Protein')

# Estimate unwanted variance through surrogate variable analysis
proTumorSVA_Krij <- doSVA(proTumor_Krij, wantedVar = 'Recurrence', numSV_method = 'be',
                          asso_metaVar = c('Recurrence', 'Gender', 'Age', 'Smoking', 'Stage', 'TumorPurity'))
proTumor_Krij <- proTumorSVA_Krij$summExp_SVs
# Show significant associations between SVs and sample metadata variables
proTumorSVA_Krij$sigAssoTab

# Perform single-omics analysis
# proTumorRes_Krij <- doSingleOmicsAnalysis(proTumor_Krij, soa_metaVar = 'Recurrence', num_sigFeats = 6,
#                                           pca_method = 'ppca', num_PCs = 20, num_PCfeats = 30,
#                                           use_limma = F, use_proDA = T, feat_anno = featAnnoTab,
#                                           feat_conv = T, plot_title = NULL, show_rownames = F,
#                                           pca_metaVar = c('Recurrence', 'Gender', 'Age',
#                                                           'Smoking', 'Stage', 'TumorPurity'),
#                                           soa_unwantVar = grep('^SV\\d+$', colnames(colData(proTumor_Krij)),
#                                                                value = T),
#                                           level_metaVar = c('Yes', 'No'))
# saveRDS(proTumorRes_Krij, './data/Discovery/AG_Krijgsveld/soaRes/trials/combined_proTumorRes_SVsByProTumor.rds')
proTumorRes_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/soaRes/trials/combined_proTumorRes_SVsByProTumor.rds')
```

Display significant associations between features (Var1) and cancer recurrence (Var2).
Owing to missing values, there may be situation that only few observations in sample
groups, resulting in unreliable t-statistics. To this end, we used probabilistic
dropout model (proDA) to account for data missingness.
```{r}
dplyr::select(proTumorRes_Krij$sig.feat.tab, -Test) %>%
  DT::datatable()
```

Assess data power by p-value histogram
```{r}
proTumorRes_Krij$pval.hist
```

Visualize molecular signatures of significant features in data. Note that features
in rows are ordered by t-statistics.
```{r}
proTumorRes_Krij$sig.feat.heat
```

Display significant associations between PCs (Var1) and cancer recurrence (Var2)
```{r}
proTumorRes_Krij$sig.pc.tab
```

Enrichment analysis\
Reference: Hallmark gene sets\
**t-values**
```{r}
# Prepare ranked gene list
statRes_Krij <- proTumorRes_Krij$SOA.res$featAssoRes %>%
  dplyr::select(Var1, Stat) %>%
  tibble::column_to_rownames('Var1')
featAnno_Krij <- as.data.frame(rowData(proTissue_Krij)) %>%
  dplyr::select(Genes)
rankedGeneList_Krij <- prepRankedFeatList(statRes = statRes_Krij, featAnno = featAnno_Krij, to_gene = T)

# Do GSEA
# Hallmark gene sets
gseaRes_H <- doEA(rankedGeneList_Krij, categoryDB = 'H', numSigGS = 10, rmPrefixGS = '^HALLMARK_')
gseaRes_H$plotTopSigGS +
  labs(title = 'Hallmark gene sets')
```

### SV-ADJUSTED (PROTISSUE)
**Surrogate variable analysis of All Tissue samples**

<font size='5'> **No interaction term - Condition:Recurrence** </font>

```{r}
# Subset certain samples
proTumor_Krij <- proTissueSVs_Krij[, which(colData(proTissueSVs_Krij)$Condition == 'Tumor')]
# Prepare feature annotation table
featAnnoTab <- tibble::as_tibble(rowData(proTumor_Krij), rownames = 'Protein') %>%
  dplyr::select(Protein, Genes) %>%
  dplyr::rename(Gene = Genes) %>%
  tibble::column_to_rownames('Protein')

# Perform single-omics analysis
# proTumorRes_Krij <- doSingleOmicsAnalysis(proTumor_Krij, soa_metaVar = 'Recurrence', num_sigFeats = 6,
#                                           pca_method = 'ppca', num_PCs = 20, num_PCfeats = 30,
#                                           use_limma = F, use_proDA = T, feat_anno = featAnnoTab,
#                                           feat_conv = T, plot_title = NULL, show_rownames = F,
#                                           pca_metaVar = c('Recurrence', 'Gender', 'Age',
#                                                           'Smoking', 'Stage', 'TumorPurity'),
#                                           soa_unwantVar = grep('^SV\\d+$', colnames(colData(proTumor_Krij)),
#                                                                value = T),
#                                           level_metaVar = c('Yes', 'No'))
# saveRDS(proTumorRes_Krij, './data/Discovery/AG_Krijgsveld/soaRes/trials/combined_proTumorRes_SVsByProTissue.rds')
proTumorRes_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/soaRes/trials/combined_proTumorRes_SVsByProTissue.rds')
```

Display significant associations between features (Var1) and cancer recurrence (Var2).
Owing to missing values, there may be situation that only few observations in sample
groups, resulting in unreliable t-statistics. To this end, we used probabilistic
dropout model (proDA) to account for data missingness.
```{r}
dplyr::select(proTumorRes_Krij$sig.feat.tab, -Test) %>%
  DT::datatable()
```

Assess data power by p-value histogram
```{r}
proTumorRes_Krij$pval.hist
```

Visualize molecular signatures of significant features in data. Note that features
in rows are ordered by t-statistics.
```{r}
proTumorRes_Krij$sig.feat.heat
```

Display significant associations between PCs (Var1) and cancer recurrence (Var2)
```{r}
proTumorRes_Krij$sig.pc.tab
```

Enrichment analysis\
Reference: Hallmark gene sets\
**t-values**
```{r}
# Prepare ranked gene list
statRes_Krij <- proTumorRes_Krij$SOA.res$featAssoRes %>%
  dplyr::select(Var1, Stat) %>%
  tibble::column_to_rownames('Var1')
featAnno_Krij <- as.data.frame(rowData(proTissue_Krij)) %>%
  dplyr::select(Genes)
rankedGeneList_Krij <- prepRankedFeatList(statRes = statRes_Krij, featAnno = featAnno_Krij, to_gene = T)

# Do GSEA
# Hallmark gene sets
gseaRes_H <- doEA(rankedGeneList_Krij, categoryDB = 'H', numSigGS = 10, rmPrefixGS = '^HALLMARK_')
gseaRes_H$plotTopSigGS +
  labs(title = 'Hallmark gene sets')
```

<font size='5'> **With interaction term - Condition:Recurrence** </font>

```{r message=F}
# Subset certain samples
proTumor_Krij <- proTissueSVs_Int_Krij[, which(colData(proTissueSVs_Int_Krij)$Condition == 'Tumor')]
# Prepare feature annotation table
featAnnoTab <- tibble::as_tibble(rowData(proTumor_Krij), rownames = 'Protein') %>%
  tibble::column_to_rownames('Protein')

# Perform single-omics analysis
# proTumorRes_Krij <- doSingleOmicsAnalysis(proTumor_Krij, soa_metaVar = 'Recurrence', num_sigFeats = 6,
#                                           pca_method = 'ppca', num_PCs = 20, num_PCfeats = 30,
#                                           use_limma = F, use_proDA = T, feat_anno = featAnnoTab,
#                                           feat_conv = T, plot_title = NULL, show_rownames = F,
#                                           pca_metaVar = c('Recurrence', 'Gender', 'Age',
#                                                           'Smoking', 'Stage', 'TumorPurity'),
#                                           soa_unwantVar = grep('^SV\\d+$', colnames(colData(proTumor_Krij)),
#                                                                value = T),
#                                           level_metaVar = c('Yes', 'No'))
# saveRDS(proTumorRes_Krij, './data/Discovery/AG_Krijgsveld/soaRes/trials/combined_proTumorRes_SVsByProTissueInt.rds')
proTumorRes_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/soaRes/trials/combined_proTumorRes_SVsByProTissueInt.rds')
```

Display significant associations between features (Var1) and cancer recurrence (Var2)
```{r}
dplyr::select(proTumorRes_Krij$sig.feat.tab, -c(Test, Protein.Names)) %>%
  DT::datatable()
```

Assess data power by p-value histogram
```{r}
proTumorRes_Krij$pval.hist
```

Visualize molecular signatures of significant features in data. Note that features
in rows are ordered by t-statistics.
```{r}
proTumorRes_Krij$sig.feat.heat
```

Display significant associations between PCs (Var1) and cancer recurrence (Var2)
```{r}
proTumorRes_Krij$sig.pc.tab
```

Enrichment analysis\
Reference: Hallmark gene sets\
**t-values**
```{r}
# Prepare ranked gene list
statRes_Krij <- proTumorRes_Krij$SOA.res$featAssoRes %>%
  dplyr::select(Var1, Stat) %>%
  tibble::column_to_rownames('Var1')
featAnno_Krij <- as.data.frame(rowData(proTissue_Krij))
rankedGeneList_Krij <- prepRankedFeatList(statRes = statRes_Krij, featAnno = featAnno_Krij, to_gene = T)

# Do GSEA
# Hallmark gene sets
gseaRes_H <- doEA(rankedGeneList_Krij, categoryDB = 'H', numSigGS = 10, rmPrefixGS = '^HALLMARK_', pvalueCutoff = 0.2)
gseaRes_H$plotTopSigGS +
  labs(title = 'Hallmark gene sets')
```

## Normal samples

Display associations between cancer recurrence and the other patient metadata variables
from patients included in normal sample analysis to identify potential confounders (or predictors)
```{r}
# Test associations between cancer recurrence and the other patient metadata variables
# Prepare patient metadata
# Subset certain samples
proNormal_Krij <- proTissue_Krij[, which(colData(proTissue_Krij)$Condition == 'Normal')]
patientMetadat <- tibble::as_tibble(colData(proNormal_Krij)) %>%
  dplyr::filter(!duplicated(Patient)) %>%
  dplyr::select(Patient, Recurrence, Gender, Age, Smoking, Stage, Adjuvant)
patientRecur <- dplyr::select(patientMetadat, Patient, Recurrence)
patientMetadat <- dplyr::select(patientMetadat, -Recurrence)
testAsso(patientMetadat, patientRecur, cmn_col = 'Patient') %>%
  dplyr::select(-Stat)
```

### UNADJUSTED

Display associations between cancer recurrence and the other patient metadata variables
from patients included in normal sample analysis to identify potential confounders (or predictors)
```{r message=F}
# Subset certain samples
proNormal_Krij <- proTissue_Krij[, which(colData(proTissue_Krij)$Condition == 'Normal')]
# Prepare feature annotation table
featAnnoTab <- tibble::as_tibble(rowData(proNormal_Krij), rownames = 'Protein') %>%
  dplyr::select(Protein, Genes) %>%
  dplyr::rename(Gene = Genes) %>%
  tibble::column_to_rownames('Protein')

# Perform single-omics analysis
# proNormalRes_Krij <- doSingleOmicsAnalysis(proNormal_Krij, soa_metaVar = 'Recurrence', num_sigFeats = 6,
#                                            pca_method = 'ppca', num_PCs = 20, num_PCfeats = 30,
#                                            use_limma = F, use_proDA = T, feat_anno = featAnnoTab,
#                                            feat_conv = T, plot_title = NULL, show_rownames = F,
#                                            pca_metaVar = c('Recurrence', 'Gender', 'Age',
#                                                            'Smoking', 'Stage'),
#                                            level_metaVar = c('Yes', 'No'))
# saveRDS(proNormalRes_Krij, './data/Discovery/AG_Krijgsveld/soaRes/trials/combined_proNormalRes.rds')
proNormalRes_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/soaRes/trials/combined_proNormalRes.rds')
```

Display significant associations between PCs (Var1) and sample metadata variables
(Var2) to overview sources of variance in data
```{r}
proNormalRes_Krij$data.var.source
```

Display significant associations between features (Var1) and cancer recurrence (Var2)
```{r}
dplyr::select(proNormalRes_Krij$sig.feat.tab, -Test) %>%
  DT::datatable()
```

Assess data power by p-value histogram
```{r}
proNormalRes_Krij$pval.hist
```

Visualize molecular signatures of significant features in data. Note that features
in rows are ordered by t-statistics.
```{r}
proNormalRes_Krij$sig.feat.heat
```

Display significant associations between PCs (Var1) and cancer recurrence (Var2)
```{r}
proNormalRes_Krij$sig.pc.tab
```

Enrichment analysis\
Reference: Hallmark gene sets\
**t-values**
```{r}
# Prepare ranked gene list
statRes_Krij <- proNormalRes_Krij$SOA.res$featAssoRes %>%
  dplyr::select(Var1, Stat) %>%
  tibble::column_to_rownames('Var1')
featAnno_Krij <- as.data.frame(rowData(proTissue_Krij)) %>%
  dplyr::select(Genes)
rankedGeneList_Krij <- prepRankedFeatList(statRes = statRes_Krij, featAnno = featAnno_Krij, to_gene = T)

# Do GSEA
# Hallmark gene sets
gseaRes_H <- doEA(rankedGeneList_Krij, categoryDB = 'H', numSigGS = 10, rmPrefixGS = '^HALLMARK_', pvalueCutoff = 0.2)
gseaRes_H$plotTopSigGS +
  labs(title = 'Hallmark gene sets')
```

### SV-ADJUSTED (PRONORMAL)
**Surrogate variable analysis of Only Normal samples**

Display associations between estimated surrogate variables and patient metadata variables
```{r message=F}
# Subset certain samples
proNormal_Krij <- proTissue_Krij[, which(colData(proTissue_Krij)$Condition == 'Normal')]
# Prepare feature annotation table
featAnnoTab <- tibble::as_tibble(rowData(proNormal_Krij), rownames = 'Protein') %>%
  dplyr::select(Protein, Genes) %>%
  dplyr::rename(Gene = Genes) %>%
  tibble::column_to_rownames('Protein')

# Estimate unwanted variance through surrogate variable analysis
proNormalSVA_Krij <- doSVA(proNormal_Krij, wantedVar = 'Recurrence', numSV_method = 'be',
                           asso_metaVar = c('Recurrence', 'Gender', 'Age', 'Smoking', 'Stage'))
proNormal_Krij <- proNormalSVA_Krij$summExp_SVs
# Show significant associations between SVs and sample metadata variables
proNormalSVA_Krij$sigAssoTab

# Perform single-omics analysis
# proNormalRes_Krij <- doSingleOmicsAnalysis(proNormal_Krij, soa_metaVar = 'Recurrence', num_sigFeats = 6,
#                                            pca_method = 'ppca', num_PCs = 20, num_PCfeats = 30,
#                                            use_limma = F, use_proDA = T, feat_anno = featAnnoTab,
#                                            feat_conv = T, plot_title = NULL, show_rownames = F,
#                                            pca_metaVar = c('Recurrence', 'Gender', 'Age',
#                                                            'Smoking', 'Stage'),
#                                            soa_unwantVar = grep('^SV\\d+$', colnames(colData(proNormal_Krij)),
#                                                                 value = T),
#                                            level_metaVar = c('Yes', 'No'))
# saveRDS(proNormalRes_Krij, './data/Discovery/AG_Krijgsveld/soaRes/trials/combined_proNormalRes_SVsByProNormal.rds')
proNormalRes_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/soaRes/trials/combined_proNormalRes_SVsByProNormal.rds')
```

Display significant associations between features (Var1) and cancer recurrence (Var2)
```{r}
dplyr::select(proNormalRes_Krij$sig.feat.tab, -Test) %>%
  DT::datatable()
```

Assess data power by p-value histogram
```{r}
proNormalRes_Krij$pval.hist
```

Visualize molecular signatures of significant features in data. Note that features
in rows are ordered by t-statistics.
```{r}
proNormalRes_Krij$sig.feat.heat
```

Display significant associations between PCs (Var1) and cancer recurrence (Var2)
```{r}
proNormalRes_Krij$sig.pc.tab
```

Enrichment analysis\
Reference: Hallmark gene sets\
**t-values**
```{r}
# Prepare ranked gene list
statRes_Krij <- proNormalRes_Krij$SOA.res$featAssoRes %>%
  dplyr::select(Var1, Stat) %>%
  tibble::column_to_rownames('Var1')
featAnno_Krij <- as.data.frame(rowData(proTissue_Krij)) %>%
  dplyr::select(Genes)
rankedGeneList_Krij <- prepRankedFeatList(statRes = statRes_Krij, featAnno = featAnno_Krij, to_gene = T)

# Do GSEA
# Hallmark gene sets
gseaRes_H <- doEA(rankedGeneList_Krij, categoryDB = 'H', numSigGS = 10, rmPrefixGS = '^HALLMARK_', pvalueCutoff = 0.2)
gseaRes_H$plotTopSigGS +
  labs(title = 'Hallmark gene sets')
```

### SV-ADJUSTED (PROTISSUE)
**Surrogate variable analysis of All Tissue samples**

<font size='5'> **No interaction term - Condition:Recurrence** </font>

```{r message=F}
# Subset certain samples
proNormal_Krij <- proTissueSVs_Krij[, which(colData(proTissueSVs_Krij)$Condition == 'Normal')]
# Prepare feature annotation table
featAnnoTab <- tibble::as_tibble(rowData(proNormal_Krij), rownames = 'Protein') %>%
  dplyr::select(Protein, Genes) %>%
  dplyr::rename(Gene = Genes) %>%
  tibble::column_to_rownames('Protein')

# Perform single-omics analysis
# proNormalRes_Krij <- doSingleOmicsAnalysis(proNormal_Krij, soa_metaVar = 'Recurrence', num_sigFeats = 6,
#                                            pca_method = 'ppca', num_PCs = 20, num_PCfeats = 30,
#                                            use_limma = F, use_proDA = T, feat_anno = featAnnoTab,
#                                            feat_conv = T, plot_title = NULL, show_rownames = F,
#                                            pca_metaVar = c('Recurrence', 'Gender', 'Age',
#                                                            'Smoking', 'Stage'),
#                                            soa_unwantVar = grep('^SV\\d+$', colnames(colData(proNormal_Krij)),
#                                                                 value = T),
#                                            level_metaVar = c('Yes', 'No'))
# saveRDS(proNormalRes_Krij, './data/Discovery/AG_Krijgsveld/soaRes/trials/combined_proNormalRes_SVsByProTissue.rds')
proNormalRes_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/soaRes/trials/combined_proNormalRes_SVsByProTissue.rds')
```

Display significant associations between features (Var1) and cancer recurrence (Var2)
```{r}
dplyr::select(proNormalRes_Krij$sig.feat.tab, -Test) %>%
  DT::datatable()
```

Assess data power by p-value histogram
```{r}
proNormalRes_Krij$pval.hist
```

Visualize molecular signatures of significant features in data. Note that features
in rows are ordered by t-statistics.
```{r}
proNormalRes_Krij$sig.feat.heat
```

Display significant associations between PCs (Var1) and cancer recurrence (Var2)
```{r}
proNormalRes_Krij$sig.pc.tab
```

Enrichment analysis\
Reference: Hallmark gene sets\
**t-values**
```{r}
# Prepare ranked gene list
statRes_Krij <- proNormalRes_Krij$SOA.res$featAssoRes %>%
  dplyr::select(Var1, Stat) %>%
  tibble::column_to_rownames('Var1')
featAnno_Krij <- as.data.frame(rowData(proTissue_Krij)) %>%
  dplyr::select(Genes)
rankedGeneList_Krij <- prepRankedFeatList(statRes = statRes_Krij, featAnno = featAnno_Krij, to_gene = T)

# Do GSEA
# Hallmark gene sets
gseaRes_H <- doEA(rankedGeneList_Krij, categoryDB = 'H', numSigGS = 10, rmPrefixGS = '^HALLMARK_', pvalueCutoff = 0.2)
gseaRes_H$plotTopSigGS +
  labs(title = 'Hallmark gene sets')
```

<font size='5'> **With interaction term - Condition:Recurrence** </font>

```{r message=F}
# Subset certain samples
proNormal_Krij <- proTissueSVs_Int_Krij[, which(colData(proTissueSVs_Int_Krij)$Condition == 'Normal')]
# Prepare feature annotation table
featAnnoTab <- tibble::as_tibble(rowData(proNormal_Krij), rownames = 'Protein') %>%
  tibble::column_to_rownames('Protein')

# Perform single-omics analysis
# proNormalRes_Krij <- doSingleOmicsAnalysis(proNormal_Krij, soa_metaVar = 'Recurrence', num_sigFeats = 6,
#                                            pca_method = 'ppca', num_PCs = 20, num_PCfeats = 30,
#                                            use_limma = F, use_proDA = T, feat_anno = featAnnoTab,
#                                            feat_conv = T, plot_title = NULL, show_rownames = F,
#                                            pca_metaVar = c('Recurrence', 'Gender', 'Age',
#                                                            'Smoking', 'Stage'),
#                                            soa_unwantVar = grep('^SV\\d+$', colnames(colData(proNormal_Krij)),
#                                                                 value = T),
#                                            level_metaVar = c('Yes', 'No'))
# saveRDS(proNormalRes_Krij, './data/Discovery/AG_Krijgsveld/soaRes/trials/combined_proNormalRes_SVsByProTissueInt.rds')
proNormalRes_Krij <- readRDS('./data/Discovery/AG_Krijgsveld/soaRes/trials/combined_proNormalRes_SVsByProTissueInt.rds')
```

Display significant associations between features (Var1) and cancer recurrence (Var2)
```{r}
dplyr::select(proNormalRes_Krij$sig.feat.tab, -Test) %>%
  DT::datatable()
```

Assess data power by p-value histogram
```{r}
proNormalRes_Krij$pval.hist
```

Visualize molecular signatures of significant features in data. Note that features
in rows are ordered by t-statistics.
```{r}
proNormalRes_Krij$sig.feat.heat
```

Display significant associations between PCs (Var1) and cancer recurrence (Var2)
```{r}
proNormalRes_Krij$sig.pc.tab
```

Enrichment analysis\
**t-values**
```{r}
# Prepare ranked gene list
statRes_Krij <- proNormalRes_Krij$SOA.res$featAssoRes %>%
  dplyr::select(Var1, Stat) %>%
  tibble::column_to_rownames('Var1')
featAnno_Krij <- as.data.frame(rowData(proTissue_Krij))
rankedGeneList_Krij <- prepRankedFeatList(statRes = statRes_Krij, featAnno = featAnno_Krij, to_gene = T)

# Do GSEA
# Hallmark gene sets
gseaRes_H <- doEA(rankedGeneList_Krij, categoryDB = 'H', numSigGS = 20, rmPrefixGS = '^HALLMARK_')
gseaRes_H$plotTopSigGS +
  labs(title = 'Hallmark gene sets')
```
