---
title: 'Testing: Potential biomarkers'
author: "Qian-Wu Liao"
output:
  workflowr::wflow_html:
    toc: yes
    toc_float: yes
    code_folding: hide
---

<font size='4'> Description: Test predictive ability of potential biomarkers identified
from Method Development cohort in lung Adenocarcinoma recurrence using Discovery cohort. </font>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 8, fig.width = 10, warning = F)
knitr::opts_knit$set(root.dir = '/Users/qianwu/Desktop/SMART-CARE_LungCancer_Discovery')
```

Load libraries
```{r load libraries, message=F}
library(SummarizedExperiment)
library(tidyverse)
# Load user-defined functions
source('./code/misc.R')

# Set plot theme
th <- theme_bw(base_size = 15) +
  theme(axis.title = element_text(face = 'bold'),
        axis.text = element_text(face = 'bold'),
        axis.ticks = element_line(linewidth = 0.8),
        legend.text = element_text(size = 15))
```

# Normal Tissue Proteomics
58 Normal Tissue Proteins were identified in Method Development Normal Tissue DIA
Proteomics from both AG Klingmüller and AG Krijgsveld and also captured in Discovery dataset.

Show predictive ability of these 58 proteins using Discovery Normal Tissue DIA Proteomics
```{r}
# Load normalized data
proTissue <- readRDS('./data/AG_Krijgsveld/proTissueVsn.rds')
# Subset data
smpNormalIdx <- which(colData(proTissue)$Condition == 'Normal')
proNormal <- proTissue[, smpNormalIdx, drop = F]
# Keep first protein as representative if there are more than one
rownames(proNormal) <- stringr::str_remove_all(rownames(proNormal), ';.+')
# Retrieve needed information
smpMetadat <- tibble::as_tibble(colData(proNormal), rownames = 'Sample') %>%
  dplyr::select(Sample, Recurrence)
featMetadat <- tibble::as_tibble(rowData(proNormal), rownames = 'Protein') %>%
  dplyr::select(Protein, Genes) %>%
  dplyr::mutate(Genes = stringr::str_remove_all(Genes, ';.+'))

# Prepare significant features identified from Method Development cohort
methodDev_sigFeats <- readRDS('./data/potential_biomarkers/sigFeats_DIANP_Overlap.rds') %>%
  dplyr::pull(Protein)
methodDev_sigFeats <- intersect(methodDev_sigFeats, rownames(proNormal))
# Do association tests
dat <- t(SummarizedExperiment::assay(proNormal)) %>%
  tibble::as_tibble(rownames = 'Sample') %>%
  dplyr::select(Sample, all_of(methodDev_sigFeats))
proNormalAssoRes <- testAssociation(dat, smpMetadat, cmn_col = 'Sample', use_proDA = T) %>%
  dplyr::select(-c(Var2, Test)) %>%
  dplyr::rename(Protein = Var1, tStat = Stat) %>%
  dplyr::mutate(Gene = plyr::mapvalues(Protein, from = featMetadat$Protein,
                                       to = featMetadat$Genes, warn_missing = F),
                pVal = round(pVal, 4),
                pValAdj = round(pValAdj, 4),
                tStat = round(tStat, 4))
# Show association test results
proNormalAssoRes %>%
  DT::datatable()
```

Visualize predictive ability (TNC, SPRYD4, SPARCL1, ANGPTL2, HK2 are top significant
proteins identified in Method Development cohort, check Method Development analysis
report 3.2)
```{r}
# Prepare data table for plotting
assoRes <- dplyr::select(proNormalAssoRes, Protein, Gene, pVal) %>%
  dplyr::filter(Protein %in% c('P37235', 'Q14515', 'Q9UKU9',
                               'P52789', 'Q8WW59', 'P24821')) %>%
  dplyr::mutate(Gene = paste0(Gene, '\n(p = ', round(pVal, 4), ')'))
proNormalTab4Plot <- SummarizedExperiment::assay(proNormal) %>%
  tibble::as_tibble(rownames = 'Protein') %>%
  tidyr::pivot_longer(cols = -'Protein', names_to = 'Sample', values_to = 'Abundance') %>%
  dplyr::left_join(smpMetadat, by = 'Sample') %>%
  dplyr::left_join(assoRes, by = 'Protein') %>%
  dplyr::filter(Protein %in% assoRes$Protein) %>%
  dplyr::mutate(Gene = factor(Gene, levels = assoRes$Gene))

ggplot(proNormalTab4Plot, aes(x=Recurrence, y=Abundance, col=Recurrence, fill=Recurrence)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(position = position_jitter(0.2), size = 2, show.legend = F) +
  scale_color_manual(values=c('#00BFC4', '#F8766D')) +
  scale_fill_manual(values=c('#00BFC4', '#F8766D')) +
  facet_wrap(vars(Gene), scales = 'free') +
  th + theme(strip.text = element_text(size = 13, face = 'bold'))
```

```{r}
# Do association tests
dat <- t(SummarizedExperiment::assay(proNormal)) %>%
  tibble::as_tibble(rownames = 'Sample')
proNormalAssoRes <- testAssociation(dat, smpMetadat, cmn_col = 'Sample', use_proDA = T) %>%
  dplyr::select(-c(Var2, Test)) %>%
  dplyr::rename(Protein = Var1) %>%
  dplyr::mutate(Gene = plyr::mapvalues(Protein, from = featMetadat$Protein,
                                       to = featMetadat$Genes, warn_missing = F))

# Show association test results
cat(paste('There are', nrow(dplyr::filter(proNormalAssoRes, pVal <= 0.05)), 'significant proteins.'))
hist(proNormalAssoRes$pVal, xlab = 'P-value', main = 'Normal Tissue Proteomics')

# Prepare data table for plotting
assoRes <- dplyr::select(proNormalAssoRes, Protein, Gene, pVal) %>%
  dplyr::mutate(Gene = paste0(Gene, '\n(p = ', round(pVal, 4), ')'))
proNormalTab4Plot <- SummarizedExperiment::assay(proNormal) %>%
  tibble::as_tibble(rownames = 'Protein') %>%
  tidyr::pivot_longer(cols = -'Protein', names_to = 'Sample', values_to = 'Abundance') %>%
  dplyr::left_join(smpMetadat, by = 'Sample') %>%
  dplyr::left_join(assoRes, by = 'Protein') %>%
  dplyr::filter(Protein %in% assoRes$Protein[1:6]) %>%
  dplyr::mutate(Gene = factor(Gene, levels = assoRes$Gene[1:6]))

ggplot(proNormalTab4Plot, aes(x=Recurrence, y=Abundance, col=Recurrence, fill=Recurrence)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(position = position_jitter(0.2), size = 2, show.legend = F) +
  scale_color_manual(values=c('#00BFC4', '#F8766D')) +
  scale_fill_manual(values=c('#00BFC4', '#F8766D')) +
  facet_wrap(vars(Gene), scales = 'free') +
  th + theme(strip.text = element_text(size = 13, face = 'bold'))
```
